### 2023-03-14 owen/01_eda.qmd
add owen initial ideas, commit this

---
title: "Chess EDA"
format: html
bibliography: refs.bib
---

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)

```

## Data

- Confirm the time controls for the Lichess data?

What types of data?
How many games involve titles?

```{r}
#| warning: false

lichess_data <- readRDS("../rdata/lichess_pilot.RData")
```

```{r}
table(lichess_data$Variant)

head(sort(table(lichess_data$TimeControl), decreasing = TRUE))

## this looks to be seconds + incrememnt, makes the most sense 
## could just take common blitz formats, 3 mins and 5 mins,
## would be 50k games still

head(sort(table(lichess_data$Event), decreasing = TRUE))

lichess_data %>% filter(Variant == "Standard") %>% 
  group_by(Event) %>% 
  count() %>% 
  arrange(-n)

```

Get all games from a specific player in a specific format,
to see if `RatingDiff` tracks how
the rating changes.

```{r}
lichess_data %>% filter(TimeControl == "180+0") %>% 
  filter(Event == "Rated Blitz game") %>% 
  arrange(UTCDate, UTCTime) %>% 
  mutate(peng_elo = ifelse(White == "penguingim1", WhiteElo, BlackElo)) %>% 
  mutate(peng_change = ifelse(White == "penguingim1", WhiteRatingDiff,
                              BlackRatingDiff)) %>% 
  select(UTCDate, UTCTime, peng_elo, peng_change) %>%
  mutate(peng_elo = as.numeric(peng_elo)) %>% 
  mutate(peng_change = as.numeric(peng_change)) %>% 
  mutate(ELO_change = peng_elo + peng_change) %>% 
  head(n = 10)

### so RatingDiff gives you the starting ELO for the next game

```

- Casual games don't change the rating, so do we want to remove them from 
the data entirely? Would likely have a different role in winner effects,
etc.

```{r}
## get simpler data

simple_data <- lichess_data %>% 
  filter(Event %in% c("Rated Bullet game", "Rated Blitz game"))

lichess_data %>% 
  filter(Event == "Rated Bullet game") %>% 
  group_by(TimeControl) %>% 
  count() %>% 
  arrange(-n)

bullet_60 <- lichess_data %>% 
  filter(Event == "Rated Bullet game") %>% 
  filter(TimeControl == "60+0")

# bullet_60

# head(sort(table(bullet_60$White), decreasing = TRUE), 20)

## have a few different players with good number of games
## get list of all players, fit some sort of model using all players
## who play enough games

head(sort(table(c(bullet_60$White, bullet_60$Black)), decreasing = TRUE), 25)

## top 25 players possibly enough then, consider all games they are involved
## in

table(bullet_60$Result)

```

## Multilevel Logistic Regression

Similar to @ding2021nhl, we wish to fit a multilevel logistic 
regression to see if we can identify individual 
effects for chess players in this data.

To do this, we will consider the current history of an individual player
as their win ratio in their past $n$ games, where we will investigate
the sensitivity of these results to different choices of $n$.

We will identify focal players as those who were used to query the API,
i.e the `Username` for each game.

```{r, past-n-results}

## given a players data, determine how many of the previous games they won
## do this for a single player first of all
## only need to do this for the potential focal players

curr_user <- "penguingim1"

curr_games <-

```

<!-- ## Streaks -->

<!-- Calculate the length of winning streaks for any given player, -->
<!-- look at how these vary across rankings maybe? -->

```{r}
#| include: false
# get_streaks <- function(games, player_name) {
#   player_games <- games %>% 
#     filter(White == player_name | Black == player_name) 
#   
#   winner <- player_games$
#   ## then get their streaks in here also
# }

peng_games <- bullet_60 %>% 
  filter(White == "penguingim1" | Black == "penguingim1")

peng_games

peng_games$w <- ifelse(peng_games$Username == peng_games$White, 1, 0)

# result column
peng_games$result <- ifelse(peng_games$w == 1 & peng_games$Result == "1-0" | 
                         peng_games$w == 0 & peng_games$Result == "0-1", 1, 0)

peng_games <- peng_games %>% 
  arrange(UTCDate, UTCTime)

peng_games$result

### then count the streaks in this
## win and loss streaks

wins <- which(peng_games$result == 1)

s <- split(wins, cumsum(c(TRUE, diff(wins) != 1)))
win_streaks <- map_dbl(s, length) %>% as.vector()

## now need to get the losing streaks
losses <- which(peng_games$result == 0)
s <- split(losses, cumsum(c(TRUE, diff(losses) != 1)))

loss_streaks <- map_dbl(s, length) %>% as.vector()

sum(loss_streaks) + sum(win_streaks)

nrow(peng_games)
```

<!-- For a given game, get the proportion of games won by the White player -->
<!-- in their last $n$ games, where we can vary $n$ and see how -->
<!-- changing it changes the results. -->

<!-- - Would you want to count wins as white and/or wins as black then? -->

<!-- - General winner effect -->

<!-- - Pick a player as "focal", look at whether they win or lose, -->
<!-- accounting for their colour, their game history, the ranking  -->
<!-- difference to their opponent, and possibly also the  -->
<!-- titles, which could have some sort of psychological impact on the outcome. -->

```{r}
#| include: false

### get past n results

prev_games <- function(player, hist_len) {

}

curr_user <- "penguingim1"

curr_data <- bullet_60 %>% 
  filter(White == curr_user | Black == curr_user)

curr_data$w <- ifelse(curr_data$White == curr_user, 1, 0)

## this doesn't account for draws, gives them equal weight to a loss

## one thing you could do is compare results after a draw with
## results after a win or a loss, see how they differ
## if not effect, and control for everything, then they should be the same

curr_data$result <- ifelse(curr_data$w == 1 & curr_data$Result == "1-0" | 
                         curr_data$w == 0 & curr_data$Result == "0-1", 1, 0)

curr_data %>% 
  select(White, Black, WhiteElo, BlackElo, w, result, UTCDate, UTCTime) %>% 
  arrange(UTCDate, UTCTime)

## then use this to get curr win history for curr_user for each game
```

<!-- ## Hot Hand Fallacy -->

<!-- - Potential different probabilities of success in subsequent games,  -->
<!-- i.e depending on who opponent is, black vs white, etc -->

<!-- - Evidence for some individual effects but -->
<!-- not sure if there is anything out there for -->
<!-- Chess already? -->

<!-- ## Some simple Models -->

<!-- - Take previous 10 games, calculate win proportion in those games -->

<!-- - Use this as a predictor for outcome -->

<!-- - Just predict win probability for White, to control for colour -->

<!-- ```{r} -->
<!-- ## logistic regression for win or less based on previous result -->
<!-- ## controlling for opponents ranking, your ranking, colour -->
<!-- ## previous game -->
<!-- ## time since previous game maybe? -->
<!-- ## -->
<!-- ``` -->

<!-- - The tennis stuff maybe and compare it? -->

<!-- - Does how the game ended make a difference? Don't seem to have this for lichess -->

<!-- - Games in tournaments vs rated, other factors like this -->

<!-- ### References -->

<!-- - The hot hand in the wild -->

<!-- - Surprised by the Hot Hand Fallacy? -->
<!-- A Truth in the Law of Small Numbers -->

### 2023-03-14 owen/hotHand.R

@@ -1,0 +1,47 @@
+rep <- 1e6
+n <- 4
+data <- array(sample(c(0,1), rep*n, replace=TRUE), c(rep,n))
+prob <- rep(NA, rep)
+for (i in 1:rep){
+  heads1 <- data[i,1:(n-1)]==1
+  heads2 <- data[i,2:n]==1
+  prob[i] <- sum(heads1 & heads2)/sum(heads1)
+}
+
+print(mean(prob, na.rm=TRUE))
+
+
+
+## this is the proportion of times you get a head, given you got one
+## the previous time
+
+
+## what about given that you got a streak of 2 heads instead, with n = 10
+
+rep <- 1e6
+n <- 10
+streak_len <- 3
+data <- array(sample(c(0,1), rep*n, replace=TRUE), c(rep,n))
+prob <- rep(NA, rep)
+for (i in 1:rep){
+  ## identify sequences of length 2 
+  heads <- data[i, ] == 1
+  seqs <- sequence(rle(data[i, ])$lengths)
+  length_1_seqs <- which(heads[1:n-1] & seqs[1:n-1] >= streak_len)
+  length_2_seqs <- which(heads[2:n] & seqs[2:n] >= (streak_len + 1))
+  prob[i] <- length(length_2_seqs)/length(length_1_seqs)
+  # 
+  # heads1 <- data[i,1:(n-1)]==1
+  # heads2 <- data[i,2:n]==1
+  # prob[i] <- sum(heads1 & heads2)/sum(heads1)
+}
+
+print(mean(prob, na.rm=TRUE))
+## this calculates it correctly when streak_len = 1
+
+
+sequence(rle(data[1, ])$lengths)
+
+a <- rle(data[1, ])
+
+rep(a$values, a$lengths)

### 2023-03-15 owen/01_eda.qmd
some initial stuff getting win rate

@@
 library(tidyverse)
+library(RcppRoll)
+
+theme_set(theme_bw())
 
@@
 
-- Confirm the time controls for the Lichess data?
-
-What types of data?
-How many games involve titles?
+We first wish to investigate the properties of the scraped data.
 
@@
 ```
+
+We see that the vast majority of games in this dataset 
+are Standard, with the regular rules, etc. Similarly, the most
+commonly played time controls is 180+0,
+which corresponds to 3 minute blitz games with no time increment.
+
+Of standard games, we see that the majority are rated blitz and bullet games.
+These correspond to randomly playing available players on the site 
+at a time. This differs from tournaments, which make up a sizeable proportion
+of standard games also.
 
@@
 Get all games from a specific player in a specific format,
-to see if `RatingDiff` tracks how
-the rating changes.
+to see that `RatingDiff` tracks how
+the rating changes. For the White/Black player based
+on the result of the current game.
 
@@
 etc.
+
+For all current analysis we will restrict to 1 minute rated bullet games,
+the most commonly played format in this dataset, with 13k games compared
+to 12k games of 30 second bullet and 12k games of 3 minute blitz.
 
@@
 ## who play enough games
+```
 
+The top 25 players all play at least 80 games in this format, and across 
+all games in this format we see relatively few draws. This is unsurprising
+as it it hard to end at a draw in these short time controls, when one player
+can lose on time.
+
+```{r}
 head(sort(table(c(bullet_60$White, bullet_60$Black)), decreasing = TRUE), 25)
@@
 We will identify focal players as those who were used to query the API,
-i.e the `Username` for each game.
+i.e the `Username` for each game. We only consider Rated 60 second bullet games
+here.
 
 ```{r, past-n-results}
+#| message: false
 
@@
 curr_user <- "penguingim1"
+curr_n <- 10
 
-curr_games <-
+curr_games <- bullet_60 %>% 
+  filter(White == curr_user | Black == curr_user)
+
+hist_games <- curr_games %>% 
+  arrange(UTCDate, UTCTime) %>% 
+  mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
+  select(White:BlackElo, focal_white)
+
+## need to check all the game types are the same here then
+table(curr_games$Variant)
+table(curr_games$Result)
+
+## then get the result for curr_user for each game, there will be draws here
+
+user_games <- hist_games %>% 
+  mutate(focal_result = case_when(
+    (focal_white == 1 & Result == "1-0") ~ 1,
+    (focal_white == 0 & Result == "0-1") ~ 1,
+    (Result == "1/2-1/2") ~ 0.5,
+    .default = 0
+    )) %>% 
+  mutate(focal_win_prop = c(rep(0, curr_n - 1), roll_mean(focal_result, n = curr_n)))
+
+user_games %>% 
+  ggplot(aes(x = focal_win_prop)) +
+  geom_histogram(bins = 20) +
+  labs(title = "Win Ratio for Previous 10 Games, one Player")
+
+### now just need to get the rolling average of focal result
+### need to pad this output
+
+### could we do this using a map across the players of interest,
+### get their games and then put it all back together afterwards?
+### think that should work, just need to make it into a function then
+
+get_hist <- function(user, games, prev_n) {
+  hist_games <- games %>% 
+    filter(White == user | Black == user) %>% 
+    arrange(UTCDate, UTCTime) %>% 
+    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
+    select(White:BlackElo, focal_white) %>% 
+    mutate(focal_result = case_when(
+      (focal_white == 1 & Result == "1-0") ~ 1,
+      (focal_white == 0 & Result == "0-1") ~ 1,
+      (Result == "1/2-1/2") ~ 0.5,
+      .default = 0
+      )) %>% 
+    mutate(focal_win_prop = c(rep(0, prev_n - 1), 
+                              roll_mean(focal_result, n = prev_n)))
+
+  hist_games
+}
+
+head(get_hist(user = curr_user, games = bullet_60, prev_n = 10))
+
+## this seems to be working, then do it across a vector of
+## players
+
+```
+
+```{r}
+a <- head(sort(table(c(bullet_60$White, bullet_60$Black)), decreasing = TRUE), 25)
+
+top_players <- names(a)
+
+tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 5)
+
+tidy_games %>% 
+  mutate(focal_player = ifelse(focal_white == 1, White, Black)) %>% 
+  group_by(focal_player) %>% 
+  count() %>% 
+  arrange(-n)
+
+## why does this have 17 players rather than 25 then?
+## why are some missing?
 

### 2023-03-17 owen/01_eda.qmd
some initial models

@@
 library(RcppRoll)
+library(cmdstanr)
+library(bayesplot)
 
@@
 ```{r}
-a <- head(sort(table(c(bullet_60$White, bullet_60$Black)), decreasing = TRUE), 25)
+## want to use focal players not all players to get these lists to be the same
+a <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 17)
+## if don't use all focal players here then removing data because they won't be
+## involved
 
@@
 
-tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 5)
+tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 10) %>% 
+  as_tibble()
 
@@
 ```
+
+Now need to figure out how to feed this into a Stan model.
+
+```{r}
+head(tidy_games)
+
+init_data <- tidy_games %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, top_players)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% mutate(ave_prop = focal_win_prop - mean(focal_win_prop))
+
+init_data
+
+### will need
+### number of focal players
+### which focal player is in each row
+### their outcome 0 or 1
+### their colour white (1) or black (0)
+### ranking diff
+
+```
+
+```{r}
+
+stan_data <- list(N = nrow(init_data),
+                  J = length(top_players),
+                  y = init_data$focal_result,
+                  id = init_data$focal_id,
+                  colour = init_data$focal_white,
+                  elo = init_data$elo_diff,
+                  win_prop = init_data$focal_win_prop)
+
+stan_file <- "model1.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit1 <- mod$sample(data = stan_data,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit1$summary()
+
+## but seems to give positive winner effects for all players
+mcmc_hist(fit1$draws("alpha"))
+mcmc_hist(fit1$draws("beta"))
+
+# fit$save_object(file = "simple_multi_logistic.RDS")
+
+```
+
+Instead account for the covariates of colour and elo difference also.
+
+```{r}
+stan_file <- "model2.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit2 <- mod$sample(data = stan_data,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit2$summary()
+
+mcmc_hist(fit2$draws(c("beta", "gamma1", "gamma2")))
+
+# fit$save_object(file = "cov_multi_logistic.RDS")
+
+## these indicate clear positive coefficients for the betas,
+## which are basically winner effects, account for the role
+## of win history
+## elo seems to have a much smaller role
+
+## repeat for a different history, did this using past 5 games, could try
+## others.
+## could also compare to longer games, see if you get the same thing
+## don't put a common prior on the betas, which is what the Cribben paper
+## did. also don't standardize them
+
+```
+
+We can then instead include partial pooling of the $\beta$ coefficients.
+
+```{r}
+stan_file <- "model3.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit3 <- mod$sample(data = stan_data,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit3$summary()
+mcmc_hist(fit3$draws(c("beta", "mu", "tau")))
+
+mcmc_trace(fit3$draws(c("mu", "tau", "beta")))
+
+```
+
+## Repeat using normalized win Ratio
+
+```{r}
+
+stan_data_ave <- list(N = nrow(init_data),
+                  J = length(top_players),
+                  y = init_data$focal_result,
+                  id = init_data$focal_id,
+                  colour = init_data$focal_white,
+                  elo = init_data$elo_diff,
+                  win_prop = init_data$ave_prop)
+
+stan_file <- "model1.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit1_ave <- mod$sample(data = stan_data_ave,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit1_ave$summary()
+
+## but seems to give positive winner effects for all players
+mcmc_hist(fit1_ave$draws("alpha"))
+mcmc_hist(fit1_ave$draws("beta"))
+
+```
+
+```{r}
+stan_file <- "model2.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit2_ave <- mod$sample(data = stan_data_ave,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit2_ave$summary()
+
+mcmc_hist(fit2$draws(c("beta", "gamma1", "gamma2")))
+```
+
+```{r}
+stan_file <- "model3.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit3_ave <- mod$sample(data = stan_data_ave,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit3_ave$summary()
+
+mcmc_hist(fit3$draws(c("beta", "mu", "tau", "gamma1", "gamma2")))
+```
+
+Given this output, we can get the probability a player who is black
+will win against and equally ranked player, when his current win ratio 
+is exactly his average will be $1-\exp(-\alpha_j)$ while if they 
+are playing as white their win probability will be
+$1-\exp(-(\alpha_j + \gamma_1))$, so $\gamma_1$ being positive indicates the win
+probability increases when playing white, which makes sense.
+Playing white makes a player who has 50% chance of winning as black
+have 54% chance of winning with white. Wikipedia says the difference is
+4-6% so similar enough.
 

### 2023-03-17 lichess_model.R

@@ -8,23 +8,28 @@ ldata <- readRDS("rdata/lichess_pilot.Rdata")
 ldata$w <- ifelse(ldata$Username == ldata$White, 1, 0)
 
 # result column
-ldata$result <- ifelse(ldata$w == 1 & ldata$Result == "1-0" | 
+ldata$result <- ifelse(ldata$w == 1 & ldata$Result == "1-0" |
                          ldata$w == 0 & ldata$Result == "0-1", 1, 0)
+
+
 
 # white/black elo adjustments
 ldata$elof <- ifelse(ldata$w == 1, ldata$WhiteElo, ldata$BlackElo)
-ldata$elonf <-ifelse(ldata$w == 0, ldata$WhiteElo, ldata$BlackElo)
+ldata$elonf <- ifelse(ldata$w == 0, ldata$WhiteElo, ldata$BlackElo)
+
+## focal and non focal, if focal is black then nonfocal is white
+
 
 # building a regression model
-plot_data <- ldata[c(1:500),]
+plot_data <- ldata[c(1:500), ]
 
-ggplot(plot_data, aes(x=UTCDate, y=result)) +
+ggplot(plot_data, aes(x = UTCDate, y = result)) +
   geom_point()
 
 ldata$elof <- as.numeric(ldata$elof)
 ldata$elonf <- as.numeric(ldata$elonf)
 ldata$w <- as.numeric(ldata$w)
-ldata1 <- ldata[c(1:1000),]
+ldata1 <- ldata[c(1:1000), ]
 m1 <- glm(result ~ elof + elonf + w + Event, family = binomial(), data = ldata1)
 summary(m1)
 plot(m1)
@@ -35,217 +40,285 @@ options(scipen = 999)
 probabilities <- m1 %>% predict(ldata1, type = "response")
 probabilities <- as.data.frame(probabilities)
 
+## are these used again then?
 
 # fish data WL Effect model fit to this data
-l.data <- data.frame(winner = ifelse(ldata$Result == "1-0", ldata$White, ldata$Black),
-                     loser = ifelse(ldata$Result == "0-1", ldata$White, ldata$Black))
+l.data <- data.frame(
+  winner = ifelse(ldata$Result == "1-0", ldata$White, ldata$Black),
+  loser = ifelse(ldata$Result == "0-1", ldata$White, ldata$Black)
+)
 
 l.data$date <- ldata$UTCDate
 l.data$type <- 0
[…]

### 2023-03-21 owen/01_eda.qmd
tidying up the models and the code

@@
 title: "Chess EDA"
-format: html
+author: Owen G. Ward
+format: 
+  html:
+    toc: true
+    toc-location: left
 bibliography: refs.bib
@@
 
-We first wish to investigate the properties of the scraped data.
+We first wish to investigate the properties of the scraped data. We 
+focus for now on the Lichess data. We see that the vast majority of games in this dataset 
+are Standard Chess, with the regular rules, etc.
 
@@
 
-We see that the vast majority of games in this dataset 
-are Standard, with the regular rules, etc. Similarly, the most
-commonly played time controls is 180+0,
+```{r}
+#| echo: true
+## Variants of Chess in this data
+table(lichess_data$Variant)
+```
+Similarly, the most
+commonly played time control, across all formats of chess, is 180+0,
 which corresponds to 3 minute blitz games with no time increment.
+
+```{r}
+#| echo: true
+head(sort(table(lichess_data$TimeControl), decreasing = TRUE))
+```
 
@@
 These correspond to randomly playing available players on the site 
-at a time. This differs from tournaments, which make up a sizeable proportion
-of standard games also.
+at a given time. This differs from tournaments, which make up a sizeable proportion
+of standard games also (and which might have different dynamics).
 
 ```{r}
-table(lichess_data$Variant)
+#| echo: true
+head(sort(table(lichess_data$Event), decreasing = TRUE))
 
-head(sort(table(lichess_data$TimeControl), decreasing = TRUE))
+```
 
-## this looks to be seconds + incrememnt, makes the most sense 
-## could just take common blitz formats, 3 mins and 5 mins,
-## would be 50k games still
-
-head(sort(table(lichess_data$Event), decreasing = TRUE))
+```{r}
+#| echo: true
 
@@
   count() %>% 
-  arrange(-n)
+  arrange(-n) %>% 
+  head()
 
@@
 
-Get all games from a specific player in a specific format,
-to see that `RatingDiff` tracks how
-the rating changes. For the White/Black player based
-on the result of the current game.
+The variable `ratingDiff` records how the ELO of both the black and
+white player changes based on the outcome of that game.
+Here these ELO scores are across the different event types, which
+may contain multiple time formats. For example, an ELO for Bullet play will 
+change when both 1 minute and 30 second games are played, and also in 
+both tournament and rated games.
+
+<!-- Get all games from a specific player in a specific format, -->
+<!-- to see that `RatingDiff` tracks how -->
+<!-- the rating changes. For the White/Black player based -->
+<!-- on the result of the current game. -->
 
 ```{r}
-lichess_data %>% filter(TimeControl == "180+0") %>% 
+#| echo: false
+#| eval: false
+elo_check <- lichess_data %>% filter(TimeControl == "180+0") %>% 
+  filter(Event == "Rated Blitz game") %>% 
+  arrange(UTCDate, UTCTime) %>% 
+  mutate(peng_elo = ifelse(White == "penguingim1", WhiteElo, BlackElo)) %>% 
+  mutate(peng_change = ifelse(White == "penguingim1", WhiteRatingDiff,
+                              BlackRatingDiff)) %>% 
+  select(UTCDate, UTCTime, peng_elo, peng_change) %>%
+  mutate(peng_elo = as.numeric(peng_elo)) %>% 
+  mutate(peng_change = as.numeric(peng_change)) %>% 
+  mutate(ELO_change = peng_elo + peng_change)
+
+elo_check %>% 
+  mutate(diff = peng_elo - lag(ELO_change))
+
+### so RatingDiff gives you the starting ELO for the next game of the same type
+
+elo_check_2 <- lichess_data %>% 
   filter(Event == "Rated Blitz game") %>% 
@@
   mutate(ELO_change = peng_elo + peng_change) %>% 
-  head(n = 10)
+  mutate(diff = peng_elo - lag(ELO_change))
 
-### so RatingDiff gives you the starting ELO for the next game
+elo_check_2
+
+## so these don't always agree, because could play arena games, etc
 
@@
 
-- Casual games don't change the rating, so do we want to remove them from 
-the data entirely? Would likely have a different role in winner effects,
-etc.
+Casual games don't change the rating, so we should probably treat them
+differently.
+Would likely have a different role in winner effects,
+etc. It is possible that Arena and tournament games also have different 
+dynamics.
 
-For all current analysis we will restrict to 1 minute rated bullet games,
+## Initial Models
+
+For the current initial analysis we
+will restrict to 1 minute rated bullet games,
 the most commonly played format in this dataset, with 13k games compared
@@
 ```{r}
-## get simpler data
-
-simple_data <- lichess_data %>% 
-  filter(Event %in% c("Rated Bullet game", "Rated Blitz game"))
+#| echo: true
+## get initial dataset to use for analysis
 
@@
   filter(TimeControl == "60+0")
-
-# bullet_60
-
-# head(sort(table(bullet_60$White), decreasing = TRUE), 20)
-
-## have a few different players with good number of games
-## get list of all players, fit some sort of model using all players
-## who play enough games
 ```
@@
 ```{r}
+#| echo: true
 head(sort(table(c(bullet_60$White, bullet_60$Black)), decreasing = TRUE), 25)
 
-## top 25 players possibly enough then, consider all games they are involved
-## in
-
+## results of these games
 table(bullet_60$Result)
@@
 
-## Multilevel Logistic Regression
+### Multilevel Logistic Regression
 
@@
 effects for chess players in this data.
+We will incorporate fixed effects for the player effects of the focal
+player in these games, along
 
-To do this, we will consider the current history of an individual player
+We will identify focal players as those who were used to query the API,
+i.e the `Username` for each game. We only consider Rated 60 second bullet games
+here, and use the top 18 players, who all play at least 100 games
+of 1 minute rated bullet chess in this dataset.
+
+We will consider the current history of an individual player
 as their win ratio in their past $n$ games, where we will investigate
@@
 
-We will identify focal players as those who were used to query the API,
-i.e the `Username` for each game. We only consider Rated 60 second bullet games
-here.
+As a small percentage (less than 4%) of these games result in 
+a draw, we will use these draws to get each players current win history,
+before then removing these from the logistic regression model.
+Initially we will choose the previous $$n=10$$ games for each player to
+compute the win ratio, or the average of all current games if
+the historic number of games played is less than 10.
+
+```{r get historic win ratio}
+
+get_hist <- function(user, games, prev_n) {
+  hist_games <- games %>% 
+    filter(White == user | Black == user) %>% 
+    arrange(UTCDate, UTCTime) %>% 
+    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
+    select(White:BlackElo, focal_white) %>% 
+    mutate(focal_result = case_when(
+      (focal_white == 1 & Result == "1-0") ~ 1,
+      (focal_white == 0 & Result == "0-1") ~ 1,
+      (Result == "1/2-1/2") ~ 0.5,
+      .default = 0
+      )) %>% 
+    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
+                              roll_mean(focal_result, n = prev_n)))
+  hist_games
+}
+
+```
 
@@
 #| message: false
+#| eval: false
 
@@
 
+### modified this so that it calculates rolling averarge of previous n
+### games, or all games, when number of previous games is less than n
+
 get_hist <- function(user, games, prev_n) {
@@
       )) %>% 
-    mutate(focal_win_prop = c(rep(0, prev_n - 1), 
+    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                               roll_mean(focal_result, n = prev_n)))
@@
 
+We can see the distribution of the win ratio for these 18 players across all
+games in this sample.
+
 ```{r}
+#| message: false
+
 ## want to use focal players not all players to get these lists to be the same
-a <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 17)
+focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18)
 ## if don't use all focal players here then removing data because they won't be
@@
 
-top_players <- names(a)
+top_players <- names(focal_players)
 
@@
 tidy_games %>% 
-  mutate(focal_player = ifelse(focal_white == 1, White, Black)) %>% 
-  group_by(focal_player) %>% 
-  count() %>% 
-  arrange(-n)
-
-## why does this have 17 players rather than 25 then?
-## why are some missing?
+  ggplot(aes(focal_win_prop)) +
+  geom_histogram() +
+  labs(x = "Win Ratio", title = "Previous 10 Games")
 
@@
 
-Now need to figure out how to feed this into a Stan model.
+Removing the drawn games, we have 12500 games involving these 18 focal players.
 
-```{r}
-head(tidy_games)
+```{r organize data for stan}
 
@@
          focal_win_prop, elo_diff, focal_result) %>% 
-  group_by(focal_id) %>% mutate(ave_prop = focal_win_prop - mean(focal_win_prop))
+  group_by(focal_id) %>% 
+  mutate(ave_prop = focal_win_prop - mean(focal_win_prop)) %>% 
+  filter(focal_result != 0.5)
 
-init_data
-
-### will need
-### number of focal players
-### which focal player is in each row
-### their outcome 0 or 1
-### their colour white (1) or black (0)
-### ranking diff
+nrow(init_data)
 
@@
 
+#### STAN Model
+
+Now we want to fit this multilevel logistic regression. Letting
+$y_{ij}$ be the result of the $i$-th game by the $j$-th focal player,
+then the model we will consider is.
+
+---
+
 ```{r}
+#| eval: false
 
@@
 ```{r}
+#| eval: false
 stan_file <- "model2.stan"
 
-mod <- cmdstan_model(stan_file)
+mod2 <- cmdstan_model(stan_file)
 
-fit2 <- mod$sample(data = stan_data,
+fit2 <- mod2$sample(data = stan_data,
                   seed = 123,
@@
 ```{r}
+#| eval: false
 stan_file <- "model3.stan"
 
-mod <- cmdstan_model(stan_file)
+mod3 <- cmdstan_model(stan_file)
 
-fit3 <- mod$sample(data = stan_data,
+fit3 <- mod3$sample(data = stan_data,
                   seed = 123,
@@
 fit3$summary()
-mcmc_hist(fit3$draws(c("beta", "mu", "tau")))
+mcmc_hist(fit3$draws(c("beta", "mu", "tau", "gamma1", "gamma2")))
 
@@
 ```{r}
-
+#| eval: false
 stan_data_ave <- list(N = nrow(init_data),
@@
 ```{r}
+#| eval: false
 stan_file <- "model2.stan"
@@
 ```{r}
+#| eval: false
 stan_file <- "model3.stan"
@@
 #| include: false
+#| eval: false
 

### 2023-03-22 owen/01_eda.qmd
initial write up

@@
 
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE)
 
@@
 $y_{ij}$ be the result of the $i$-th game by the $j$-th focal player,
-then the model we will consider is.
+then the model we will consider is
 
----
+$$
+P(y_{ij} = 1) = \frac{1}{1 + \exp(-(\alpha_j + \beta_j x_{ij} + \gamma z_{ij}))}
+$$
+
+where:
+
+- $\alpha_j$ is a player level random effect
+- $\beta_j$ is a player level random effect, accounting for the win ratio $x_{ij}$
+- $\gamma$ is the coefficient of game level coefficients
+
+We then partially pool the $\alpha$ and $\beta$ coefficients, so
+
+$$
+\alpha_j \sim \mathcal{N}(\mu_1, \tau_1)
+$$
+$$
+\mu_1 \sim \mathcal{N}(0, 5)
+$$
+$$
+\tau_1 \sim \mathcal{Cauchy}(0, 5)
+$$
+
+and similarly
+
+$$
+\beta_j \sim \mathcal{N}(\mu_2, \tau_2)
+$$
+$$
+\mu_2 \sim \mathcal{N}(0, 5)
+$$
+$$
+\tau_2 \sim \mathcal{Cauchy}(0, 5)
+$$
+
+We rescale the win ratio $x_{ij}$, by the average win ration of
+player $j$. This means $\beta_j$ captures how a players win
+probability changes as their historic performance deviates from 
+the overall average win ratio.
+
+For $z_{ij}$ we incorporate 2 covariates, namely the colour of 
+the focal player and the ELO difference between the focal 
+player and their opponent. So that means we have 
+$\gamma_1$, which tells us how the win probability changes in going
+from black to white,
+and $\gamma_2$ gives how the win probability changes as the focal player
+has a better ELO than their opponent.
+
+```{r fit this model in stan}
+#| eval: false
+
+stan_data_ave <- list(N = nrow(init_data),
+                  J = length(top_players),
+                  y = init_data$focal_result,
+                  id = init_data$focal_id,
+                  colour = init_data$focal_white,
+                  elo = init_data$elo_diff,
+                  win_prop = init_data$ave_prop)
+
+stan_file <- "model3.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit3_ave <- mod$sample(data = stan_data_ave,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit3_ave$save_object("init_model_n10.RDS")
+```
+
+Some of the output of this model is shown below, namely the empirical posteriors
+for each of the parameters.
+
+```{r load in saved fit}
+#| message: false
+
+fit3_ave <- readRDS("init_model_n10.RDS")
+
+fit3_ave$summary()
+
+mcmc_hist(fit3_ave$draws(c("alpha", "mu2", "tau2")))
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
+
+rhat(fit3_ave)
+
+```
+
+#### Interpreting this output
+
+Given this output, we can get the probability a player who is black
+will win against an equally ranked player, when their current win ratio 
+is exactly their average will be $\frac{1}{1+\exp(-\alpha_j)}$.
+So a positive $\alpha_j$ indicates a player favoured to win 
+even when playing black. The overall $\mu_1$ tells us about the 
+global probability of winning as black against an equally ranked
+opponent.
+If a player
+is playing as white their win probability will be
+$\frac{1}{1+\exp(-(\alpha_j + \gamma_1))}$,
+so $\gamma_1$ being positive indicates the win
+probability increases when playing white, which makes sense.
+Playing white makes a player who has 50% chance of winning as black
+have approximately a 54% chance of winning with white. Wikipedia says the increase 
+in win probability from playing white has been estimated to be 
+4-6% so quite similar.
+
+Similarly, $\gamma_2$ tells us about how the ELO of the focal player 
+being larger than their opponent influences their win probability.
+The estimate here indicates that a large ELO difference makes a player very
+likely to win, as would be expected. For example, a player who
+has a 50% chance of winning against an opponent with the same ELO
+will have about a 56% of winning against an opponent with an ELO score
+100 points lower.
+
+### Varying History
+
+We can refit this using varying numbers of games for the
+current win ratio.
+
+```{r 5 prev games}
+#| eval: false
+
+tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 5) %>% 
+  as_tibble()
+
+init_data <- tidy_games %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, top_players)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% 
+  mutate(ave_prop = focal_win_prop - mean(focal_win_prop)) %>% 
+  filter(focal_result != 0.5)
+
+stan_data_ave <- list(N = nrow(init_data),
+                  J = length(top_players),
+                  y = init_data$focal_result,
+                  id = init_data$focal_id,
+                  colour = init_data$focal_white,
+                  elo = init_data$elo_diff,
+                  win_prop = init_data$ave_prop)
+
+stan_file <- "model3.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit3_ave <- mod$sample(data = stan_data_ave,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit3_ave$save_object("init_model_n5.RDS")
+
+```
+
+```{r}
+fit3_ave <- readRDS("init_model_n5.RDS")
+
+fit3_ave$summary()
+
+mcmc_hist(fit3_ave$draws(c("beta", "gamma1", "gamma2")))
+```
+
+```{r 20 prev games}
+#| eval: false
+
+tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 20) %>% 
+  as_tibble()
+
+init_data <- tidy_games %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, top_players)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% 
+  mutate(ave_prop = focal_win_prop - mean(focal_win_prop)) %>% 
+  filter(focal_result != 0.5)
+
+stan_data_ave <- list(N = nrow(init_data),
+                  J = length(top_players),
+                  y = init_data$focal_result,
+                  id = init_data$focal_id,
+                  colour = init_data$focal_white,
+                  elo = init_data$elo_diff,
+                  win_prop = init_data$ave_prop)
+
+stan_file <- "model3.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit3_ave <- mod$sample(data = stan_data_ave,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+## this took 3 minutes
+
+fit3_ave$save_object("init_model_n20.RDS")
+
+```
+
+```{r}
+
+fit3_ave <- readRDS("init_model_n20.RDS")
+
+fit3_ave$summary()
+
+mcmc_hist(fit3_ave$draws(c("beta", "mu1","gamma1", "gamma2")))
+```
+
+```{r prev 50 games}
+#| eval: false
+
+tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 50) %>% 
+  as_tibble()
+
+init_data <- tidy_games %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, top_players)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% 
+  mutate(ave_prop = focal_win_prop - mean(focal_win_prop)) %>% 
+  filter(focal_result != 0.5)
+
+stan_data_ave <- list(N = nrow(init_data),
+                  J = length(top_players),
+                  y = init_data$focal_result,
+                  id = init_data$focal_id,
+                  colour = init_data$focal_white,
+                  elo = init_data$elo_diff,
+                  win_prop = init_data$ave_prop)
+
+stan_file <- "model3.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit3_ave <- mod$sample(data = stan_data_ave,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit3_ave$save_object("init_model_n50.RDS")
+```
+
+```{r}
+#| message: false
+fit3_ave <- readRDS("init_model_n50.RDS")
+
+fit3_ave$summary()
+
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "gamma1", "gamma2")))
+
+```
+
+Using a longer history seems to give a model which fits quicker and
+is more stable. But I'm not sure why that should be the case?
+
+The size of the winner effects does appear to decrease as you look
+at a longer amount of previous games. Unlikely to play more than 
+50 games in a row maybe? Still clear positive effects even 
+using 50 games.
+
+## Comments
+
+- There are some potential concerns with this Stan model and that it could
+be made more efficient and better parameterized. These seem to be a
+larger issue when the number of previous games used for the win ratio
+is smaller, which would make sense as these would depart more from the mean.
+
+- The size of these winner effects decreases as we look at their win ratio over
+more historic games.
+
+- Are there any other variables we should be controlling for here?
 
@@
 
-Instead account for the covariates of colour and elo difference also.
-
 ```{r}
@@
 
-We can then instead include partial pooling of the $\beta$ coefficients.
-
 ```{r}
@@
 ```
-
-## Repeat using normalized win Ratio
 
@@
 
-mcmc_hist(fit3$draws(c("beta", "mu", "tau", "gamma1", "gamma2")))
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
+mcmc_hist(fit3_ave$draws(c("alpha", "mu2", "tau2")))
 ```
-
-Given this output, we can get the probability a player who is black
-will win against and equally ranked player, when his current win ratio 
-is exactly his average will be $1-\exp(-\alpha_j)$ while if they 
-are playing as white their win probability will be
-$1-\exp(-(\alpha_j + \gamma_1))$, so $\gamma_1$ being positive indicates the win
-probability increases when playing white, which makes sense.
-Playing white makes a player who has 50% chance of winning as black
-have 54% chance of winning with white. Wikipedia says the difference is
-4-6% so similar enough.
 

### 2023-03-28 owen/01_eda.qmd
some more stuff for james

@@
 
+We first consider instead using the previous 5 games.
+
 ```{r 5 prev games}
@@
 
-mcmc_hist(fit3_ave$draws(c("beta", "gamma1", "gamma2")))
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "gamma1", "gamma2")))
 ```
+
+We also consider the previous 20 games.
 
@@
 ```
+
+Finally, we consider the previous 50 games also.
 
@@
 
+### Longer Games
+
+In the above example we considered only rated 60 second bullet games.
+We will now instead consider rated 3 minute blitz games, and fit
+the same model. This would indicate how the effect might change as
+we vary the length of the games considered. We have 12k total such blitz games.
+We will use the same players as were considered for the bullet data, 
+of which 16 play blitz games and only 7 play more than 100 blitz games.
+This leads to 9000 games.
+
+```{r get blitz data}
+
+blitz_180 <- lichess_data %>% 
+  filter(Event == "Rated Blitz game") %>% 
+  filter(TimeControl == "180+0")
+
+# blitz_180
+
+blitz_top <- blitz_180 %>% 
+  filter(Username %in% top_players) %>% 
+  group_by(Username) %>% 
+  tally() %>% 
+  arrange(-n) %>% 
+  pull(Username)
+
+## will slightly change the order then compared to the previous data
+
+tidy_games <- map_dfr(blitz_top, get_hist, blitz_180, prev_n = 5) %>% 
+  as_tibble()
+
+```
+
+We first consider using the previous 5 games as the history. We remove
+draws from the outcomes, but use them for the history.
+
+```{r get stan data}
+#| eval: false
+
+init_data <- tidy_games %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, blitz_top)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% 
+  mutate(ave_prop = focal_win_prop - mean(focal_win_prop)) %>% 
+  filter(focal_result != 0.5)
+
+stan_data_ave <- list(N = nrow(init_data),
+                  J = length(blitz_top),
+                  y = init_data$focal_result,
+                  id = init_data$focal_id,
+                  colour = init_data$focal_white,
+                  elo = init_data$elo_diff,
+                  win_prop = init_data$ave_prop)
+
+stan_file <- "model3.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit3_ave <- mod$sample(data = stan_data_ave,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+## save this fit
+
+fit3_ave$save_object("init_model_blitz_n5.RDS")
+
+```
+
+```{r}
+fit3_ave <- readRDS("init_model_blitz_n5.RDS")
+
+fit3_ave$summary()
+
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "gamma1", "gamma2")))
+
+```
+
+We can also repeat this with the previous 20 blitz games as an example also.
+We only consider 13 players now who have more than 20 blitz games from the 
+original selection.
+
+```{r prev 20 blitz}
+#| eval: false
+
+tidy_games <- map_dfr(blitz_top[1:13], get_hist, blitz_180, prev_n = 20) %>% 
+  as_tibble()
+
+init_data <- tidy_games %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, blitz_top)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% 
+  mutate(ave_prop = focal_win_prop - mean(focal_win_prop)) %>% 
+  filter(focal_result != 0.5)
+
+stan_data_ave <- list(N = nrow(init_data),
+                  J = length(blitz_top),
+                  y = init_data$focal_result,
+                  id = init_data$focal_id,
+                  colour = init_data$focal_white,
+                  elo = init_data$elo_diff,
+                  win_prop = init_data$ave_prop)
+
+stan_file <- "model3.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit3_ave <- mod$sample(data = stan_data_ave,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+## save this fit
+
+fit3_ave$save_object("init_model_blitz_n20.RDS")
+
+```
+
+```{r}
+fit3_ave <- readRDS("init_model_blitz_n20.RDS")
+
+fit3_ave$summary()
+
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "gamma1", "gamma2")))
+```
+
+When we consider more historic games the winner effects do appear to shrink,
+as we saw in the bullet data. However, there seems to still be
+very strong evidence for some effect. This model does not fit as 
+well, but this can be fixed.
+
 ## Comments
@@
 
+## Streaks
+
+One of the other interesting components of this data is the performance
+of a player after they draw the previous game, compared to when they
+won or lost the previous game. There are a small number of draws is
+this data set that we can use to investigate this.
+We just get the proportion of results in the next game, for both the 
+bullet and blitz data.
+
 ```{r}
-#| eval: false
+#| echo: true
+prev_data <- bullet_60 %>% 
+  group_by(Username) %>% 
+  arrange(UTCDate, UTCTime) %>% 
+  ## get focal result here also
+  mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
+  mutate(focal_result = case_when(
+    (focal_white == 1 & Result == "1-0") ~ 1,
+    (focal_white == 0 & Result == "0-1") ~ 1,
+    (Result == "1/2-1/2") ~ 0.5,
+    .default = 0
+    )) %>% 
+  mutate(prev_result = lag(focal_result)) %>% 
+  select(White, Black, WhiteElo, BlackElo, focal_white:prev_result)
 
-stan_data <- list(N = nrow(init_data),
-                  J = length(top_players),
-                  y = init_data$focal_result,
-                  id = init_data$focal_id,
-                  colour = init_data$focal_white,
-                  elo = init_data$elo_diff,
-                  win_prop = init_data$focal_win_prop)
+prev_data %>% 
+  drop_na() %>%
+  group_by(prev_result, focal_result) %>% 
+  tally() %>% 
+  group_by(prev_result) %>% 
+  mutate(result_prop = (n/sum(n)))
 
-stan_file <- "model1.stan"
+## get some confidence intervals for these differences, might be enough
+## evidence for a winner effect after winning, maybe not for a loser effect 
+## after losing
 
-mod <- cmdstan_model(stan_file)
+## can also repeat this for the blitz data
 
-fit1 <- mod$sample(data = stan_data,
-                  seed = 123,
-                  chains = 4,
-                  parallel_chains = 4,
-                  refresh = 100)
+prev_blitz <- blitz_180 %>% group_by(Username) %>% 
+  arrange(UTCDate, UTCTime) %>% 
+  ## get focal result here also
+  mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
+  mutate(focal_result = case_when(
+    (focal_white == 1 & Result == "1-0") ~ 1,
+    (focal_white == 0 & Result == "0-1") ~ 1,
+    (Result == "1/2-1/2") ~ 0.5,
+    .default = 0
+    )) %>% 
+  mutate(prev_result = lag(focal_result)) %>% 
+  select(White, Black, WhiteElo, BlackElo, focal_white:prev_result)
 
-fit1$summary()
+prev_blitz %>% 
+  drop_na() %>%
+  group_by(prev_result, focal_result) %>% 
+  tally() %>% 
+  group_by(prev_result) %>% 
+  mutate(result_prop = (n/sum(n)))
 
-## but seems to give positive winner effects for all players
-mcmc_hist(fit1$draws("alpha"))
-mcmc_hist(fit1$draws("beta"))
-
-# fit$save_object(file = "simple_multi_logistic.RDS")
+## see something similar with the blitz data, maybe not as strong but
+## seems to be there
 
 ```
-
-```{r}
-#| eval: false
-stan_file <- "model2.stan"
-
-mod2 <- cmdstan_model(stan_file)
-
-fit2 <- mod2$sample(data = stan_data,
-                  seed = 123,
-                  chains = 4,
-                  parallel_chains = 4,
-                  refresh = 100)
-
-fit2$summary()
-
-mcmc_hist(fit2$draws(c("beta", "gamma1", "gamma2")))
-
-# fit$save_object(file = "cov_multi_logistic.RDS")
-
-## these indicate clear positive coefficients for the betas,
-## which are basically winner effects, account for the role
-## of win history
-## elo seems to have a much smaller role
-
-## repeat for a different history, did this using past 5 games, could try
-## others.
-## could also compare to longer games, see if you get the same thing
-## don't put a common prior on the betas, which is what the Cribben paper
-## did. also don't standardize them
-
-```
-
-```{r}
-#| eval: false
-stan_file <- "model3.stan"
-
-mod3 <- cmdstan_model(stan_file)
-
-fit3 <- mod3$sample(data = stan_data,
-                  seed = 123,
-                  chains = 4,
-                  parallel_chains = 4,
-                  refresh = 100)
-
-fit3$summary()
-mcmc_hist(fit3$draws(c("beta", "mu", "tau", "gamma1", "gamma2")))
-
-mcmc_trace(fit3$draws(c("mu", "tau", "beta")))
-
-```
-
-```{r}
-#| eval: false
-stan_data_ave <- list(N = nrow(init_data),
-                  J = length(top_players),
-                  y = init_data$focal_result,
-                  id = init_data$focal_id,
-                  colour = init_data$focal_white,
-                  elo = init_data$elo_diff,
-                  win_prop = init_data$ave_prop)
-
-stan_file <- "model1.stan"
-
-mod <- cmdstan_model(stan_file)
-
-fit1_ave <- mod$sample(data = stan_data_ave,
-                  seed = 123,
-                  chains = 4,
-                  parallel_chains = 4,
-                  refresh = 100)
-
-fit1_ave$summary()
-
-## but seems to give positive winner effects for all players
-mcmc_hist(fit1_ave$draws("alpha"))
-mcmc_hist(fit1_ave$draws("beta"))
-
-```
-
-```{r}
-#| eval: false
-stan_file <- "model2.stan"
-
-mod <- cmdstan_model(stan_file)
-
-fit2_ave <- mod$sample(data = stan_data_ave,
-                  seed = 123,
-                  chains = 4,
-                  parallel_chains = 4,
-                  refresh = 100)
-
-fit2_ave$summary()
-
-mcmc_hist(fit2$draws(c("beta", "gamma1", "gamma2")))
-```
-
-```{r}
-#| eval: false
-stan_file <- "model3.stan"
-
-mod <- cmdstan_model(stan_file)
-
-fit3_ave <- mod$sample(data = stan_data_ave,
-                  seed = 123,
-                  chains = 4,
-                  parallel_chains = 4,
-                  refresh = 100)
-
-fit3_ave$summary()
-
-mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
-mcmc_hist(fit3_ave$draws(c("alpha", "mu2", "tau2")))
-```
-
-<!-- ## Streaks -->
 

### 2023-04-17 owen/Write Ups/01_eda.qmd
fixing a bug in the model where the previous n games included the current result we are predicting. This explains the problem with all the effects looking the same. No gives more reasonable results, permuting the data gives clear indications for no effects then also

--- 2023-03-28 owen/01_eda.qmd
@@
 format: 
-  html:
-    toc: true
-    toc-location: left
+  # html:
+  #   toc: true
+  #   toc-location: left
+  pdf
 bibliography: refs.bib
@@
 
-lichess_data <- readRDS("../rdata/lichess_pilot.RData")
+lichess_data <- readRDS("../../rdata/lichess_pilot.RData")
 ```
@@
 effects for chess players in this data.
-We will incorporate fixed effects for the player effects of the focal
-player in these games, along
+We will incorporate random effects for the player effects of the focal
+player in these games, along with random player level winner effects. 
+We take common fixed effects for the effect of colour and 
+the ELO difference.
 
@@
 i.e the `Username` for each game. We only consider Rated 60 second bullet games
-here, and use the top 18 players, who all play at least 100 games
+here initially, and use the top 18 players, who all play at least 100 games
 of 1 minute rated bullet chess in this dataset.
@@
   group_by(focal_id) %>% 
-  mutate(ave_prop = focal_win_prop - mean(focal_win_prop)) %>% 
+  mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_win_prop)) %>% 
   filter(focal_result != 0.5)
@@
 $$
-\alpha_j \sim \mathcal{N}(\mu_1, \tau_1)
+\alpha_j \sim \mathcal{N}(\mu_2, \tau_2)
 $$
@@
 $$
-\beta_j \sim \mathcal{N}(\mu_2, \tau_2)
+\beta_j \sim \mathcal{N}(\mu_1, \tau_1)
 $$
@@
 
-stan_file <- "model3.stan"
+stan_file <- "../model3.stan"
 
@@
 
-fit3_ave$save_object("init_model_n10.RDS")
+fit3_ave$save_object("../model_fits/init_model_n10.RDS")
 ```
@@
 
-fit3_ave <- readRDS("init_model_n10.RDS")
+fit3_ave <- readRDS("../model_fits/init_model_n10.RDS")
 
@@
 has a 50% chance of winning against an opponent with the same ELO
-will have about a 56% of winning against an opponent with an ELO score
+will have about a 61% chance of winning against an opponent with an ELO score
 100 points lower.
@@
 
-fit3_ave$save_object("init_model_n5.RDS")
+fit3_ave$save_object("../model_fits/init_model_n5.RDS")
 
@@
 ```{r}
-fit3_ave <- readRDS("init_model_n5.RDS")
+fit3_ave <- readRDS("../model_fits/init_model_n5.RDS")
 
@@
 
-fit3_ave$save_object("init_model_n20.RDS")
+fit3_ave$save_object("../model_fits/init_model_n20.RDS")
 
@@
 
-fit3_ave <- readRDS("init_model_n20.RDS")
+fit3_ave <- readRDS("../model_fits/init_model_n20.RDS")
 
@@
 
-fit3_ave$save_object("init_model_n50.RDS")
+fit3_ave$save_object("../model_fits/init_model_n50.RDS")
 ```
@@
 #| message: false
-fit3_ave <- readRDS("init_model_n50.RDS")
+fit3_ave <- readRDS("../model_fits/init_model_n50.RDS")
 
@@
 ```
-
-Using a longer history seems to give a model which fits quicker and
-is more stable. But I'm not sure why that should be the case?
 
@@
 
-stan_file <- "model3.stan"
+# stan_file <- "../model3.stan"
 
@@
 
-fit3_ave$save_object("init_model_blitz_n5.RDS")
+fit3_ave$save_object("../model_fits/init_model_blitz_n5.RDS")
 
@@
 ```{r}
-fit3_ave <- readRDS("init_model_blitz_n5.RDS")
+fit3_ave <- readRDS("../model_fits/init_model_blitz_n5.RDS")
 
@@
 
-stan_file <- "model3.stan"
-
-mod <- cmdstan_model(stan_file)
+# stan_file <- "model3.stan"
+# 
+# mod <- cmdstan_model(stan_file)
 
@@
 
-fit3_ave$save_object("init_model_blitz_n20.RDS")
+fit3_ave$save_object("../model_fits/init_model_blitz_n20.RDS")
 
@@
 ```{r}
-fit3_ave <- readRDS("init_model_blitz_n20.RDS")
+fit3_ave <- readRDS("../model_fits/init_model_blitz_n20.RDS")
 
@@
 
-mcmc_hist(fit3_ave$draws(c("beta", "mu1", "gamma1", "gamma2")))
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1","gamma1", "gamma2")))
 ```
@@
 
-- There are some potential concerns with this Stan model and that it could
-be made more efficient and better parameterized. These seem to be a
-larger issue when the number of previous games used for the win ratio
-is smaller, which would make sense as these would depart more from the mean.
-
 - The size of these winner effects decreases as we look at their win ratio over
@@
 - Are there any other variables we should be controlling for here?
+
+- How data dependent is this result? Should investigate the
+performance across different players, the strength of the winner
+effect as a function of the ELO of a player, etc.
 
@@
 ```{r}
-#| echo: true
+
 prev_data <- bullet_60 %>% 

### 2023-04-17 owen/Write Ups/02_more_eda.qmd
fixing a bug in the model where the previous n games included the current result we are predicting. This explains the problem with all the effects looking the same. No gives more reasonable results, permuting the data gives clear indications for no effects then also

---
title: "Model Checking"
author: Owen G. Ward
date: 04/17/2023
format: 
  # html:
  #   toc: true
  #   toc-location: left
  pdf
bibliography: refs.bib
---

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(RcppRoll)
library(cmdstanr)
library(bayesplot)

theme_set(theme_bw())

```

## Permuting Data

A quick sanity check on the existing model is to see if there are 
still strong winner and loser effects present when we permute the 
ordering of the events. This can be checked quite easily.

```{r clean lichess data}
#| warning: false

lichess_data <- readRDS("../../rdata/lichess_pilot.RData")

bullet_60 <- lichess_data %>% 
  filter(Event == "Rated Bullet game") %>% 
  filter(TimeControl == "60+0")

get_hist <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
      )) %>% 
    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(focal_result, n = prev_n)))
  hist_games
}

focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18)
## if don't use all focal players here then removing data because they won't be
## involved

top_players <- names(focal_players)

tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 10) %>% 
  as_tibble()

```

```{r permute results}

## need to do this inside the `get_hist` function

## permute the results of all games for a given player before then
## computing the current win proportion

get_hist_perm <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
      )) %>% 
    mutate(perm_result = sample(focal_result)) %>% 
    mutate(perm_win_prop = c(cumsum(perm_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(perm_result, n = prev_n)))
  hist_games
}

perm_games <- map_dfr(top_players, get_hist_perm, bullet_60, prev_n = 10) %>% 
  as_tibble()
```

```{r}
tidy_games %>% 
  ggplot(aes(focal_win_prop)) + geom_histogram() +
  labs(title = "True Win Proportions")

perm_games %>% 
  ggplot(aes(perm_win_prop)) + geom_histogram() +
  labs(title = "Permuted Win Proportions")

## there is some sort of difference here, but not sure how big it is

```

Can also plot the current previous $n$ win proportion over time. 
Under the permuted model this should look more like random noise.

```{r}

perm_games %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  group_by(focal_user) %>% 
  mutate(index = row_number()) %>% 
  ggplot(aes(index, perm_win_prop, colour = focal_user)) +
  geom_line() +
  facet_wrap(~focal_user, scales = "free_x") +
  labs(title = "Permuted Games")

tidy_games %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  group_by(focal_user) %>% 
  mutate(index = row_number()) %>% 
  ggplot(aes(index, focal_win_prop, colour = focal_user)) +
  geom_line() +
  facet_wrap(~focal_user, scales = "free_x") +
  labs(title = "Original Data")

## these look more random anyway

```

Now let's fit the stan model to the permuted data and investigate what
the effects are.

```{r get stan data permuted}

perm_init_data <- perm_games %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         perm_win_prop, elo_diff, perm_result) %>% 
  group_by(focal_id) %>% 
  mutate(ave_prop = lag(perm_win_prop, default = 0) - mean(perm_win_prop)) %>% 
  filter(perm_result != 0.5)

perm_init_data

stan_data_ave <- list(N = nrow(perm_init_data),
                  J = length(top_players),
                  y = perm_init_data$perm_result,
                  id = perm_init_data$focal_id,
                  colour = perm_init_data$focal_white,
                  elo = perm_init_data$elo_diff,
                  win_prop = perm_init_data$ave_prop)

```

```{r fit stan model}
#| eval: false

stan_file <- "../model3.stan"

mod <- cmdstan_model(stan_file)

fit3_ave <- mod$sample(data = stan_data_ave,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit3_ave$save_object("../model_fits/perm_model_n10.RDS")
```

Then we want to compare the estimates from this permuted model to the true 
data.

```{r}

fit3_ave <- readRDS("../model_fits/perm_model_n10.RDS")
fit3_orig <- readRDS("../model_fits/init_model_n10.RDS")

fit3_ave$summary()

mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))

mcmc_hist(fit3_orig$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
```

So, when we permute the data in this way, get the same winner/loser effects.
This means its not a real effect. What could be causing this?

- Weird data sample?

- Not controlling for some other important predictors?

- The magnitude of these effects is the same also, which indicates its a problem

- Did I permute the data the right way then?

What about if the history is simply uniform on the [-1, 1]. What sort of
effects do you see then?

```{r}
#| eval: false
stan_data_rand <- list(N = nrow(perm_init_data),
                  J = length(top_players),
                  y = perm_init_data$perm_result,
                  id = perm_init_data$focal_id,
                  colour = perm_init_data$focal_white,
                  elo = perm_init_data$elo_diff,
                  win_prop = perm_init_data$ave_prop)

random_win_prop <- runif(n = stan_data_rand$N, min = -1)

stan_data_rand$win_prop <- random_win_prop

fit3_rand <- mod$sample(data = stan_data_rand,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit3_rand$save_object(file = "../model_fits/unif_hist.RDS")

```

```{r}

fit3_rand <- readRDS("../model_fits/unif_hist.RDS")

mcmc_hist(fit3_rand$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))

fit3_rand$summary()

```

### 2023-04-18 owen/Write Ups/01_eda.qmd
some small fixes

@@
 So a positive $\alpha_j$ indicates a player favoured to win 
-even when playing black. The overall $\mu_1$ tells us about the 
+even when playing black. The overall $\mu_2$ tells us about the 
 global probability of winning as black against an equally ranked
-opponent.
+opponent, which is just below 0, indicating a slightly less than 50% 
+chance of winning.
 If a player
@@
 
-mcmc_hist(fit3_ave$draws(c("beta", "mu1", "gamma1", "gamma2")))
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
 ```
 
+The posterior of these effects does seem to change somewhat 
+as we vary the history. This also seems to vary from player to player.
 We also consider the previous 20 games.
@@
 
-mcmc_hist(fit3_ave$draws(c("beta", "mu1","gamma1", "gamma2")))
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1","gamma1", "gamma2")))
 ```
@@
 
-mcmc_hist(fit3_ave$draws(c("beta", "mu1", "gamma1", "gamma2")))
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1","gamma1", "gamma2")))
 
@@
 at a longer amount of previous games. Unlikely to play more than 
-50 games in a row maybe? Still clear positive effects even 
-using 50 games.
+50 games in a row maybe? It does appear that the mean
+for $\tau_2$ increases as we consider longer histories.
 
@@
 
-mcmc_hist(fit3_ave$draws(c("beta", "mu1", "gamma1", "gamma2")))
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
 
 ```
+
+When we instead consider blitz games there seems to be considerably less
+evidence for these winner effects, with many of the $\beta$ parameters
+centered around 0 now. This seems reasonable, indicates
+smaller effects in longer games.
 
@@
 
-mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1","gamma1", "gamma2")))
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
 ```
 
-When we consider more historic games the winner effects do appear to shrink,
-as we saw in the bullet data. However, there seems to still be
-very strong evidence for some effect. This model does not fit as 
-well, but this can be fixed.
+This seems to mirror the smaller number of blitz games quite closely.
 
 ## Comments
-
-- The size of these winner effects decreases as we look at their win ratio over
-more historic games.
 

### 2023-04-25 owen/Write Ups/01_eda.qmd
trying out some new things, fitting some new models

@@
 format: 
-  # html:
-  #   toc: true
-  #   toc-location: left
-  pdf
+  html:
+    toc: true
+    toc-location: left
+  # pdf
 bibliography: refs.bib
@@
 library(bayesplot)
+library(posterior)
 
@@
 - $\beta_j$ is a player level random effect, accounting for the win ratio $x_{ij}$
-- $\gamma$ is the coefficient of game level coefficients
+- $\gamma$ is a fixed effect of game level coefficients
 
@@
 $$
+\mu_2 \sim \mathcal{N}(0, 5)
+$$
+$$
+\tau_2 \sim \mathcal{Cauchy}(0, 5)
+$$
+
+and similarly
+
+$$
+\beta_j \sim \mathcal{N}(\mu_1, \tau_1)
+$$
+$$
 \mu_1 \sim \mathcal{N}(0, 5)
@@
 \tau_1 \sim \mathcal{Cauchy}(0, 5)
-$$
-
-and similarly
-
-$$
-\beta_j \sim \mathcal{N}(\mu_1, \tau_1)
-$$
-$$
-\mu_2 \sim \mathcal{N}(0, 5)
-$$
-$$
-\tau_2 \sim \mathcal{Cauchy}(0, 5)
 $$
@@
 
-fit3_ave$summary()
+# fit3_ave$summary()
 
@@
 mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
-
-rhat(fit3_ave)
-
+```
+
+The following metric (R-hat) indicates if the Bayesian model is
+converging. We would like values of smaller than 1.01 here, with values
+less than 1 possible in a sample. 
+```{r compute r-hat}
+bayesplot::rhat(fit3_ave)
 ```
@@
 
+### How big are these effects
+
+We then want to quantify how much these estimated effects change the 
+expected win probability.
+We have seen the size of the effects above for the non history parameters.
+The largest posterior mean for a $\beta\approx 1.56$. Since the normalized win
+proportion is between -0.6 and 0.6, that means that for a game
+where this player has a 50% chance of winning with a normalized average win
+proportion of 0, if their normalized win proportion is 0.5, the
+win probability will increase by $\approx 19\%$.
+This would occur, for example, if they win 50% of their total games but have
+won all 10 of their last 10. This is a similar size effect to the 
+increase in win probability which would come from playing an 
+opponent with an ELO ranking that is 170 points lower.
+
+Similarly, a normalized win history 0.25 
+above average corresponds to a difference 
+of approximately 80 in ELO scores.
+
+The underlying parameter $\mu_1$ has posterior mean of 0.606. 
+A normalised win history of 0.25 with a winner effect of this size
+would increase the win probability by about 4%, a similar effect
+as playing as white instead of black, all other covariates being
+equal.
+
+### Probability of Positive Winner Effects
+
+For each of these players, we can use the posterior draws to
+estimate the posterior probability that $\beta >0$, i.e
+that there is a positive effect.
+
+```{r estimate prob beta positive}
+
+fit3_ave <- readRDS("../model_fits/init_model_n10.RDS")
+
+# stan_draws <- fit3_ave$draws()
+# 
+# 
+# all_draws <- as_draws_df(stan_draws)
+# 
+# ndraws <- nrow(all_draws)
+# 
+# as_tibble(all_draws) %>% 
+#   select(starts_with("beta")) %>% 
+#   apply(2, function(x) sum(x>0)/ndraws)
+#
+
+prob_positive <- function(stan_draws) {
+  all_draws <- as_draws_df(stan_draws)
+  ndraws <- nrow(all_draws)
+  as_tibble(all_draws) %>% 
+  select(starts_with("beta")) %>% 
+  apply(2, function(x) sum(x>0)/ndraws)
+}
+
+prob_positive(fit3_ave$draws())
+
+```
+
+We can then use this metric to compare how these winner effects change 
+as we change both the length of the history and the type of games used.
+
 ### Varying History
@@
 
-fit3_ave$summary()
+# fit3_ave$summary()
 
 mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
+
+prob_positive(fit3_ave$draws())
 ```
@@
 
-fit3_ave$summary()
+# fit3_ave$summary()
 
 mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1","gamma1", "gamma2")))
+
+prob_positive(fit3_ave$draws())
 ```
@@
 
+prob_positive(fit3_ave$draws())
 ```

### 2023-04-25 owen/Write Ups/02_more_eda.qmd
trying out some new things, fitting some new models

@@
 library(cmdstanr)
+library(posterior)
 library(bayesplot)
@@
 
+prob_positive <- function(stan_draws, param = "beta") {
+  all_draws <- as_draws_df(stan_draws)
+  ndraws <- nrow(all_draws)
+  as_tibble(all_draws) %>% 
+  select(starts_with(param)) %>% 
+  apply(2, function(x) sum(x>0)/ndraws)
+}
+
 ```
+
+## Identifying Problems
+
+An issue was identified with the current implementation,
+where the `previous_win_history` used, which was supposed to
+be the proportion of the past $n$ games won compared to the
+players overall win percentage, contained the result of the current game
+(the outcome). When this is corrected the strong equal positive winner
+effects are removed, and we now see a variable effect, 
+with some players still displaying positive winner effects. However,
+several of the players no longer show such an effect.
 
@@
 
-Now let's fit the stan model to the permuted data and investigate what
+Now we fit the stan model to the permuted data and investigate what
 the effects are.
@@
 
-perm_init_data
+# perm_init_data
 
@@
 
-fit3_ave$summary()
+# fit3_ave$summary()
 
-mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2"))) +
+  labs(title = "Permuted Data")
 
-mcmc_hist(fit3_orig$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
+mcmc_hist(fit3_orig$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2"))) +
+  labs(title = "True Data")
+
 ```
 
-So, when we permute the data in this way, get the same winner/loser effects.
-This means its not a real effect. What could be causing this?
+```{r}
+#| echo: true
 
-- Weird data sample?
+## for the permuted data
+prob_positive(fit3_ave)
 
-- Not controlling for some other important predictors?
+## for the real data
+prob_positive(fit3_orig)
 
-- The magnitude of these effects is the same also, which indicates its a problem
+```
 
-- Did I permute the data the right way then?
+Permuted data gives effects which are all zero on average, which is reasonable.
+This would be expected as permuting the ordering of the games should
+remove any possible winner effects. Similarly, the 
+probability of positive effects is closer to 0.5, which
+would be expected under a model with no effects.
 
-What about if the history is simply uniform on the [-1, 1]. What sort of
-effects do you see then?
+Another way to confirm that having no history should lead to no 
+winner effects is shown below.
+We now replace the historic win proportion with uniform 
+draws on the interval $[-1, 1]$. The estimated winner effects 
+under such a history are shown below.
 
@@
 
+We also see that the probability of these effects being positive is
+close to 0.5, as would be expected under this model.
+
 ```{r}
@@
 
-fit3_rand$summary()
+# fit3_rand$summary()
+
+prob_positive(fit3_rand)
 
 ```
+
+## Separating out history
+
+We want to next look at the history as both the current win 
+streak, up to but not including the previous game, along with the result 
+of the previous game, to see if we can identify how potential winner 
+effects can be decomposed between these two components.
+
+We do this with a model of the form
+
+$$
+P(y_{ij} = 1) = \frac{1}{1 + \exp(-(\alpha_j + \beta_j x_{ij} + 
+\delta_j x_{ij}^{*} + 
+\gamma z_{ij}))},
+$$
+
+where now $x_{ij}^{*}$ is the result of the __previous__ game
+played by player $j$, and $x_{ij}$ is their normalised 
+win ratio over the previous $n$ games, not including the single
+previous game.
+
+```{r get data for winner effects}
+
+hist_data_init <- tidy_games %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, top_players)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% 
+  mutate(ave_prop = lag(focal_win_prop, n = 2, default = 0) - 
+           mean(focal_win_prop),
+         prev_game = lag(focal_result, default = 0)) %>% 
+  filter(focal_result != 0.5)
+
+stan_data_hist <- list(N = nrow(hist_data_init),
+                  J = length(top_players),
+                  y = hist_data_init$focal_result,
+                  id = hist_data_init$focal_id,
+                  colour = hist_data_init$focal_white,
+                  elo = hist_data_init$elo_diff,
+                  win_prop = hist_data_init$ave_prop,
+                  prev_game = hist_data_init$prev_game)
+
+```
+
+```{r fit the stan model}
+#| eval: false
+
+stan_file <- "../model4.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit4 <- mod$sample(data = stan_data_hist,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit4$summary()
+
+fit4$save_object(file = "../model_fits/model4.RDS")
+
+```
+
+```{r load the stan fit}
+
+fit4 <- readRDS("../model_fits/model4.RDS")
+
+mcmc_hist(fit4$draws(c("delta"))) +
+  labs(title = "Effect of Previous Game")
+
+mcmc_hist(fit4$draws(c("beta"))) +
+  labs(title = "Effect of Previous n games, excluding most recent game")
+
+```
+
+This model seems to fit well, with no potential concerns also.
+
+We can then again compare these estimated effects. This
+
+```{r check whether positive effects}
+#| echo: true
+
+## the beta parameters (previous n, not last game)
+prob_positive(fit4)
+
+## the delta parameters (single game) 
+prob_positive(fit4, param = "delta")
+```
+
+This seems to indicate that the previous game on its own does not have as clear
+an effect, but that the streak of previous games shows stronger evidence
+for a winner effect.
+
+## Consider all bullet games
+
+Previously we considered only 60 second bullet games. We wish to 
+consider if there is evidence for this effect across all bullet games.
+
+```{r get all bullet games}
+
+all_bullet <- lichess_data %>% 
+  filter(Event == "Rated Bullet game")
+
+tidy_games <- map_dfr(top_players, get_hist, all_bullet, prev_n = 10) %>% 
+  as_tibble()
+
+tidy_games
+
+stan_data <- tidy_games %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, top_players)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% 
+  mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_win_prop)) %>% 
+  filter(focal_result != 0.5)
+
+```
+
+```{r refit model 3 for all bullet}
+#| eval: false
+
+stan_data_ave <- list(N = nrow(stan_data),
+                  J = length(top_players),
+                  y = stan_data$focal_result,
+                  id = stan_data$focal_id,
+                  colour = stan_data$focal_white,
+                  elo = stan_data$elo_diff,
+                  win_prop = stan_data$ave_prop)
+
+stan_file <- "../model3.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit3_bullet <- mod$sample(data = stan_data_ave,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit3_bullet$save_object(file = "../model_fits/model3_n10_all_bullet.RDS")
+```
+
+```{r}
+fit3_bullet <- readRDS("../model_fits/model3_n10_all_bullet.RDS")
+
+mcmc_hist(fit3_bullet$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
+
+prob_positive(fit3_bullet)
+```
+
+How do these estimates compare to those with only the 60 second games?
+The model with all bullet games has all posterior means greater than 0.
+There is not much of an overall change between the estimated mean
+effect sizes between the two datasets.
+Using all bullet games also leads to 
+smaller posterior standard deviations for the
+winner effect size for all players.
+
+```{r refit model 4 for all bullet games}
+#| eval: false
+
+hist_data_init <- tidy_games %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, top_players)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% 
+  mutate(ave_prop = lag(focal_win_prop, n = 2, default = 0) - 
+           mean(focal_win_prop),
+         prev_game = lag(focal_result, default = 0)) %>% 
+  filter(focal_result != 0.5)
+
+stan_data_hist <- list(N = nrow(hist_data_init),
+                  J = length(top_players),
+                  y = hist_data_init$focal_result,
+                  id = hist_data_init$focal_id,
+                  colour = hist_data_init$focal_white,
+                  elo = hist_data_init$elo_diff,
+                  win_prop = hist_data_init$ave_prop,
+                  prev_game = hist_data_init$prev_game)
+
+stan_file <- "../model4.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit4 <- mod$sample(data = stan_data_hist,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit4$summary()
+
+fit4$save_object(file = "../model_fits/model4_n10_all_bullet.RDS")
+
+```
+
+```{r examine this model fit}
+fit4 <- readRDS("../model_fits/model4_n10_all_bullet.RDS")
+
+mcmc_hist(fit4$draws(variables = c("beta", "mu1", "tau1", "gamma1", "gamma2")))
+
+mcmc_hist(fit4$draws(variables = c("delta", "mu3", "tau3")))
+
+prob_positive(fit4)
+prob_positive(fit4, param = "delta")
+```
+
+## Just previous game
+
+We could also only use the previous game as a predictor, 
+without using the rest of the history.
+
+```{r}
+#| eval: false
+
+stan_data_hist <- list(N = nrow(hist_data_init),
+                  J = length(top_players),
+                  y = hist_data_init$focal_result,
+                  id = hist_data_init$focal_id,
+                  colour = hist_data_init$focal_white,
+                  elo = hist_data_init$elo_diff,
+                  prev_game = hist_data_init$prev_game)
+
+stan_file <- "../model5.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit5 <- mod$sample(data = stan_data_hist,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit5$save_object(file = "../model_fits/model5_n10_all_blitz.RDS")
+
+```
+
+```{r}
+fit5 <- readRDS("../model_fits/model5_n10_all_blitz.RDS")
+
+mcmc_hist(fit5$draws(variables = c("delta")))
+
+prob_positive(fit5, param = "delta")
+
+```
+
+Fitting using just the previous game as a predictor seems to give a reasonable
+model, with a similar interpretation as when we include the history also.
+
+For this model, a posterior mean value for delta of 0.25 corresponds to 7%
+increase in the win probability having won the previous game compared 
+to having lost it.
+
+Not clear if this difference is due to excluding the win
+history or due to using a larger/different dataset.
+
+### Next Steps
+
+- Can we compare the performance of these models, using loo-cv 
+or similar?
+
+- Could this model be parameterized better?
+
+- Any other possible problems with this?
+
+- What about some sort of mixture of the previous game and the others before 
+that?
+
+ - Some sort of decaying effect?

### 2023-04-28 owen/Write Ups/02_more_eda.qmd
updating the models and a bit more writing up on it also

@@
 library(bayesplot)
+library(loo)
 
@@
 
-## Permuting Data
+### Permuting Data
 
@@
 
+## Initial Model
+
+For completeness, we include the initial model output here again, to 
+show the results.
+
+$$
+P(y_{ij} = 1) = \frac{1}{1 + \exp(-(\alpha_j + \beta_j x_{ij} + \gamma z_{ij}))}
+$$
+
+where:
+
+- $\alpha_j$ is a player level random effect
+- $\beta_j$ is a player level random effect, accounting for the win ratio $x_{ij}$
+- $\gamma$ is a fixed effect of game level coefficients
+
+We then partially pool the $\alpha$ and $\beta$ coefficients, so
+
+$$
+\alpha_j \sim \mathcal{N}(\mu_2, \tau_2)
+$$
+$$
+\mu_2 \sim \mathcal{N}(0, 5)
+$$
+$$
+\tau_2 \sim \mathcal{Cauchy}(0, 5)
+$$
+
+and similarly
+
+$$
+\beta_j \sim \mathcal{N}(\mu_1, \tau_1)
+$$
+$$
+\mu_1 \sim \mathcal{N}(0, 5)
+$$
+$$
+\tau_1 \sim \mathcal{Cauchy}(0, 5)
+$$
+
+We rescale the win ratio $x_{ij}$, by the average win ration of
+player $j$. This means $\beta_j$ captures how a players win
+probability changes as their historic performance deviates from 
+the overall average win ratio.
+
+For $z_{ij}$ we incorporate 2 covariates, namely the colour of 
+the focal player and the ELO difference between the focal 
+player and their opponent. So that means we have 
+$\gamma_1$, which tells us how the win probability changes in going
+from black to white,
+and $\gamma_2$ gives how the win probability changes as the focal player
+has a better ELO than their opponent.
+
+We show the posterior estimates for these parameters below.
+
+```{r load in initial model here}
+
+fit3_ave <- readRDS("../model_fits/init_model_n10.RDS")
+
+# fit3_ave$summary()
+
+mcmc_hist(fit3_ave$draws(c("alpha", "mu2", "tau2")))
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
+```
+
+#### Interpreting this output
+
+Given this output, we can get the probability a player who is black
+will win against an equally ranked player, when their current win ratio 
+is exactly their average will be $\frac{1}{1+\exp(-\alpha_j)}$.
+So a positive $\alpha_j$ indicates a player favoured to win 
+even when playing black. The overall $\mu_2$ tells us about the 
+global probability of winning as black against an equally ranked
+opponent, which is just below 0, indicating a slightly less than 50% 
+chance of winning.
+If a player
+is playing as white their win probability will be
+$\frac{1}{1+\exp(-(\alpha_j + \gamma_1))}$,
+so $\gamma_1$ being positive indicates the win
+probability increases when playing white, which makes sense.
+Playing white makes a player who has 50% chance of winning as black
+have approximately a 54% chance of winning with white. Wikipedia says the increase 
+in win probability from playing white has been estimated to be 
+4-6% so quite similar.
+
+Similarly, $\gamma_2$ tells us about how the ELO of the focal player 
+being larger than their opponent influences their win probability.
+The estimate here indicates that a large ELO difference makes a player very
+likely to win, as would be expected. For example, a player who
+has a 50% chance of winning against an opponent with the same ELO
+will have about a 61% chance of winning against an opponent with an ELO score
+100 points lower.
+
+#### How big are these effects
+
+We then want to quantify how much these estimated effects change the 
+expected win probability.
+We have seen the size of the effects above for the non history parameters.
+The largest posterior mean for a $\beta\approx 1.56$. Since the normalized win
+proportion is between -0.6 and 0.6, that means that for a game
+where this player has a 50% chance of winning with a normalized average win
+proportion of 0, if their normalized win proportion is 0.5, the
+win probability will increase by $\approx 19\%$.
+This would occur, for example, if they win 50% of their total games but have
+won all 10 of their last 10. This is a similar size effect to the 
+increase in win probability which would come from playing an 
+opponent with an ELO ranking that is 170 points lower.
+
+Similarly, a normalized win history 0.25 
+above average corresponds to a difference 
+of approximately 80 in ELO scores.
+
+The underlying parameter $\mu_1$ has posterior mean of 0.606. 
+A normalised win history of 0.25 with a winner effect of this size
+would increase the win probability by about 4%, a similar effect
+as playing as white instead of black, all other covariates being
+equal.
+
 ## Separating out history
@@
 win ratio over the previous $n$ games, not including the single
-previous game.
+previous game. Note that $x_{ij}^{*}$ is not normalised,
+so can take on values $0,0.5,1$.
 
@@
 
-```{r fit the stan model}
+```{r fit model 4}
 #| eval: false
@@
 
+When we fit this model we can see the posterior distributions 
+of these $\delta$ parameters, which correspond to the 
+role the __single previous__ game has in the 
+win probability for the next game.
+
 ```{r load the stan fit}
@@
 
-This model seems to fit well, with no potential concerns also.
+We can then again compare these estimated effects. Most of these $\delta$
+have posterior means between $[-0.2, 0.2]$. 
+$\delta=0.2$ corresponds to winning the previous game leading to an 
+increase in the win probability of 5%, while holding all 
+other variables fixed.
 
-We can then again compare these estimated effects. This
+It is maybe not immediate how to compare the effect sizes for $\beta$ and 
+$\delta$ here.
 
@@
 for a winner effect.
+
+We will also fit another model, using only
+the previous game as history, to this same dataset.
+
+```{r model 5 bullet 60}
+#| eval: false
+
+hist_data_init <- tidy_games %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, top_players)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% 
+  mutate(ave_prop = lag(focal_win_prop, n = 2, default = 0) - 
+           mean(focal_win_prop),
+         prev_game = lag(focal_result, default = 0)) %>% 
+  filter(focal_result != 0.5)
+
+stan_data_hist <- list(N = nrow(hist_data_init),
+                  J = length(top_players),
+                  y = hist_data_init$focal_result,
+                  id = hist_data_init$focal_id,
+                  colour = hist_data_init$focal_white,
+                  elo = hist_data_init$elo_diff,
+                  prev_game = hist_data_init$prev_game)
+
+stan_file <- "../model5.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit5 <- mod$sample(data = stan_data_hist,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit5$save_object(file = "../model_fits/model5.RDS")
+
+```
+
+We can see the estimated effects from this model below.
+
+```{r}
+fit5 <- readRDS("../model_fits/model5.RDS")
+
+mcmc_hist(fit5$draws(variables = c("delta", "mu3", "tau3", "gamma1", "gamma2")))
+prob_positive(fit5$draws(), param = "delta")
+```
+
+The estimated effect sizes for $\delta$ are very similar to those estimated
+under the previous model, indicating that including the 
+longer history provides an improvement over just the previous game 
+(particularly given the model comparison below).
 
@@
 
-## Just previous game
+#### Just previous game
 
@@
 
-### Next Steps
+## Model Checking
 
-- Can we compare the performance of these models, using loo-cv 
-or similar?
+We now have a bunch of variations on the original model. We have fit these
+on both only 60 second games and all bullet games.
+We do some initial model checking, using only the
+smaller data initially (to speed up the computation).
 
-- Could this model be parameterized better?
+As a model check, we use each of the posterior draws to
+generate corresponding simulated outcomes, for each
+of the games, given the covariates.
+We can then use this to get a posterior predictive distribution
+for the number of games won by each of the players, which 
+we show in the histograms below. Each 
+red vertical line corresponds to the true
+number of games won by a player.
 
-- Any other possible problems with this?
+We see that the true values are very plausible under each
+of these models and provide one reasonable sanity check.
+
+```{r get data for this section}
+tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 10) %>% 
+  as_tibble()
+## For model 3
+
+init_data <- tidy_games %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, top_players)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% 
+  mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_win_prop)) %>% 
+  filter(focal_result != 0.5)
+
+stan_data_ave <- list(N = nrow(init_data),
+                  J = length(top_players),
+                  y = init_data$focal_result,
+                  id = init_data$focal_id,
+                  colour = init_data$focal_white,
+                  elo = init_data$elo_diff,
+                  win_prop = init_data$ave_prop)
+
+## For model 4 and model 5
+
+hist_data_init <- tidy_games %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, top_players)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% 
+  mutate(ave_prop = lag(focal_win_prop, n = 2, default = 0) - 
+           mean(focal_win_prop),
+         prev_game = lag(focal_result, default = 0)) %>% 
+  filter(focal_result != 0.5)
+
+stan_data_hist <- list(N = nrow(hist_data_init),
+                  J = length(top_players),
+                  y = hist_data_init$focal_result,
+                  id = hist_data_init$focal_id,
+                  colour = hist_data_init$focal_white,
+                  elo = hist_data_init$elo_diff,
+                  win_prop = hist_data_init$ave_prop,
+                  prev_game = hist_data_init$prev_game)
+
+```
+
+```{r load single model at a time for model checking}
+fit3_ave <- readRDS("../model_fits/init_model_n10.RDS")
+
+fit3_samples <- as_draws_df(fit3_ave$draws())
+
+fit3_samp <- fit3_samples %>% 
+  select(!starts_with(c("log_lik", "yrep")))
+
+y_rep <- fit3_samples %>% select(starts_with("y_rep"))
+
+rm(fit3_samples)
+
+### then want to join this with the correct input, i.e player, etc
+
+y_rep_mod <- y_rep %>% 
+  rownames_to_column() %>%  
+  pivot_longer(-rowname) %>% 
+  pivot_wider(names_from=rowname, values_from=value) %>% 
+  ## remove the draws and chains here
+  mutate(focal_id = stan_data_ave$id)
+
+## what is the check here? number of games won by a player for each of
+## the posterior samples, compared to the number of games the actually won
+
+games_won <- y_rep_mod %>% 
+  pivot_longer(cols = `1`:`4000`, names_to = "draw", values_to = "y") %>% 
+  group_by(focal_id, draw) %>% 
+  summarise(games_won = sum(y))
+
+## then stick the true number of games on top of this and see how
+## it looks
+
+orig_data <- tibble(outcome = stan_data_ave$y,
+                    focal_id = stan_data_ave$id)
+
+orig_games_won <- orig_data %>% 
+  group_by(focal_id) %>% 
+  summarise(games_won = sum(outcome))
+
+games_won %>% 
+  ggplot(aes(x = games_won)) +
+  geom_histogram() +
+  facet_wrap(~focal_id, scales = "free") +
+  geom_vline(data = orig_games_won, 
+             mapping = aes(xintercept = games_won), col = "red") +
+  labs(title = "Model 3")
+
+## could get posterior predictive p-values here but they should all 
+## be fine, looks pretty decent,
+## smallest p-value is 0.06 here for id = 16
+
+```
+
+```{r repeat for model 4}
+
+fit4 <- readRDS("../model_fits/model4.RDS")
+
+fit4_samples <- as_draws_df(fit4$draws())
+
+y_rep <- fit4_samples %>% select(starts_with("y_rep"))
+
+rm(fit4_samples)
+
+y_rep_mod <- y_rep %>% 
+  rownames_to_column() %>%  
+  pivot_longer(-rowname) %>% 
+  pivot_wider(names_from=rowname, values_from=value) %>% 
+  ## remove the draws and chains here
+  mutate(focal_id = stan_data_hist$id)
+
+games_won <- y_rep_mod %>% 
+  pivot_longer(cols = `1`:`4000`, names_to = "draw", values_to = "y") %>% 
+  group_by(focal_id, draw) %>% 
+  summarise(games_won = sum(y))
+
+games_won %>% 
+  ggplot(aes(x = games_won)) +
+  geom_histogram() +
+  facet_wrap(~focal_id, scales = "free") +
+  geom_vline(data = orig_games_won, 
+             mapping = aes(xintercept = games_won), col = "red") +
+  labs(title = "Model 4")
+
+## this model looks event better than model 3
+
+```
+
+```{r repeat this posterior check for model 5}
+
+fit5 <- readRDS("../model_fits/model5.RDS")
+
+fit5_samples <- as_draws_df(fit5$draws())
+
+y_rep <- fit5_samples %>% select(starts_with("y_rep"))
+
+rm(fit5_samples)
+
+y_rep_mod <- y_rep %>% 
+  rownames_to_column() %>%  
+  pivot_longer(-rowname) %>% 
+  pivot_wider(names_from=rowname, values_from=value) %>% 
+  ## remove the draws and chains here
+  mutate(focal_id = stan_data_hist$id)
+
+games_won <- y_rep_mod %>% 
+  pivot_longer(cols = `1`:`4000`, names_to = "draw", values_to = "y") %>% 
+  group_by(focal_id, draw) %>% 
+  summarise(games_won = sum(y))
+
+games_won %>% 
+  ggplot(aes(x = games_won)) +
+  geom_histogram() +
+  facet_wrap(~focal_id, scales = "free") +
+  geom_vline(data = orig_games_won, 
+             mapping = aes(xintercept = games_won), col = "red") +
+  labs(title = "Model 5")
+
+```
+
+Model 5 appears to look as good as model 4, in terms
+of posterior predictive distribution for the output.
+
+### Model comparison
+
+Finally, we also consider some model comparison
+between these 3 proposed models. Here we
+compute the leave one out cross validation.
+This allows us to compare across the models, with
+the general rule of thumb being a difference 
+of more than 2 standard errors indicates a better model.
+
+```{r}
+#| cache: true
+comp <- loo_compare(fit3_ave$loo(), fit4$loo(), fit5$loo())
+
+comp
+
+## this indicated model 2 slightly better than model 4, clearly better than
+## model 3
+
+```
+
+Here we see that the second model, including both the previous game
+and
+the $n$ games before that, is an improvement over the other 
+two models. The model which considers only the single previous game
+is closer.
+
+## Further Ideas
 
@@
 - Some sort of decaying effect?
+
+- Is there an interaction between the winner effect and the ability
+of the player, in terms of their ELO score?

### 2023-05-23 owen/Write Ups/02_more_eda.qmd
writing some code to get the versions of the models James shared somewhat up and running, have it ready to fit them on the cluster also

@@
 format: 
-  html:
-    toc: true
-    toc-location: left
-  # pdf
+  # html:
+  #   toc: true
+  #   toc-location: left
+  pdf
 bibliography: refs.bib
@@
 
+```{r prob positive fit3 beta}
+#| echo: true
+prob_positive(fit3_ave$draws())
+```
+
 #### Interpreting this output
@@
 mcmc_hist(fit5$draws(variables = c("delta", "mu3", "tau3", "gamma1", "gamma2")))
+
+```
+
+```{r}
+#| echo: true
 prob_positive(fit5$draws(), param = "delta")
@@
 ```{r get all bullet games}
-
 all_bullet <- lichess_data %>%
@@
 
+#### First Model all Bullet
+
 ```{r}
@@
 winner effect size for all players.
+
+#### Separating out History
 

### 2023-05-23 owen/Write Ups/03_existing_models.qmd
writing some code to get the versions of the models James shared somewhat up and running, have it ready to fit them on the cluster also

---
title: "Model Checking"
author: Owen G. Ward
date: 04/17/2023
format: 
  # html:
  #   toc: true
  #   toc-location: left
  pdf
bibliography: refs.bib
---

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(RcppRoll)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(loo)

options(mc.cores = parallel::detectCores())

theme_set(theme_bw())

prob_positive <- function(stan_draws, param = "beta") {
  all_draws <- as_draws_df(stan_draws)
  ndraws <- nrow(all_draws)
  as_tibble(all_draws) %>% 
  select(starts_with(param)) %>% 
  apply(2, function(x) sum(x>0)/ndraws)
}

```

## Model Comparison

We want to compare these potential models, which is very similar
to the model we fit previously.
The only change is that we now need the history for
both players, not just the focal player.

```{r load in the initial data}
#| cache: true

lichess_data <- readRDS("../../rdata/lichess_pilot.RData")

bullet_60 <- lichess_data %>% 
  filter(Event == "Rated Bullet game") %>% 
  filter(TimeControl == "60+0") %>% 
  mutate(game_id = row_number())

get_hist <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white, game_id) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
      )) %>% 
    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(focal_result, n = prev_n)))s
  hist_games
}

focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18)
## if don't use all focal players here then removing data because they won't be
## involved

top_players <- names(focal_players)

## then get the opponents of these players

full_opponents <- bullet_60 %>% 
  filter(Username %in% top_players) %>% 
  select(White, Black)

all_opps <- c(full_opponents$White, full_opponents$Black)

opponents <- unique(all_opps)

## then need to get the history of these players, but maybe need
## to store game id
## can just match using date and time I guess?
## need to modify the function then to account for this

get_hist_opp <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(opp_white = ifelse(user == White, 1, 0)) %>% 
    select(White:BlackElo, opp_white, game_id) %>% 
    mutate(opp_result = case_when(
      (opp_white == 1 & Result == "1-0") ~ 1,
      (opp_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
      ))

  ngames <- nrow(hist_games)

  if(ngames < prev_n) {
    hist_games <- hist_games %>% 
    mutate(opp_win_prop = opp_result/(1:ngames))
  }
  else{
    hist_games <- hist_games %>% 
      mutate(opp_win_prop = c(cumsum(opp_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(opp_result, n = prev_n)))
  }
  hist_games$player <- user
  hist_games
}
## so now the history is the history corresponding to the name
## in the "player" column

remaining_opponents <- opponents[!(opponents %in% top_players)]

opp_hist <- map_dfr(remaining_opponents, get_hist_opp, bullet_60, prev_n = 5) %>% 
  as_tibble()

## so this gives the history of the opponents for each of their 
## games, can use this to then first get their previous result, etc,
## then join this with the other results

prev_opp <- opp_hist %>% 
  group_by(player) %>% 
  arrange(UTCDate, UTCTime) %>% 
  mutate(prev_opp_result = lag(opp_result, default = 0), 
         prev_opp_prop = lag(opp_win_prop, default = 0)) %>%
  select(game_id, player, prev_opp_result, prev_opp_prop)

tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 5) %>% 
  as_tibble()

all <- tidy_games %>% 
  left_join(prev_opp, by = "game_id") %>% 
  mutate(focal = ifelse(focal_white == 1, White, Black))
## this looks more reasonable now
## this has the history of the opponent included, not the history 
## of the focal player yet

final_data <- all %>% 
  group_by(focal) %>% 
  arrange(UTCDate, UTCTime) %>% 
  mutate(prev_focal_result = lag(focal_result, default = 0),
         prev_focal_prop = lag(focal_win_prop, default = 0)) %>% 
  ungroup()

## then tidy this up for stan

```

This object `prev_game_both` should have the previous result of both players in
each game, if that game is in the dataset.
Similarly, `final_prev` has the games played by the focal players, along 
with their previous result and their opponents previous result, if that
game was in the original dataset used (all 60 second bullet games here).

Then convert this into data for stan.

```{r get stan data for simple model}

final_data

stan_prev <- final_data %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>%
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         prev_focal_result, prev_opp_result, focal_result)

stan_data <- list(N = nrow(stan_prev),
                  J = length(top_players),
                  y = stan_prev$focal_result,
                  id = stan_prev$focal_id,
                  focal_prev = stan_prev$prev_focal_result,
                  opp_prev = stan_prev$prev_opp_result)

```

```{r fit the simple stan model}
#| eval: false

## remember no effect for non focal players in this at all 
## at the moment

stan_file <- "../Comp1.stan"

mod <- cmdstan_model(stan_file)

fit1_simple <- mod$sample(data = stan_data,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

## this fits at least now...

fit1_simple$summary()
mcmc_hist(fit1_simple$draws(c("alpha", "beta")))

```

## Notes on these models

- How does the choice of focal player make a difference in these models?

```{r thinking about the choice of focal}

table(bullet_60$Username)

tidy_bullet <- bullet_60 %>% 
  select(White:BlackElo, Username, game_id) %>% 
  mutate(opponent = ifelse(Username == White, Black, White),
         focal_white = ifelse(Username == White, 1, 0))

all_players <- c(tidy_bullet$Username, tidy_bullet$opponent)
players <- unique(all_players)

## then just get all the history for this
all_hist <- map_dfr(players, get_hist_opp, bullet_60, prev_n = 5) %>% 
  as_tibble() %>% 
  select(game_id:player)

## this should have each game appear twice, for each of the two players
## involved, so need to figure out how to handle that best
## left join with tidy games matching player to focal and then
## to opponent, rename the columns as needed

tidy_stan <- tidy_bullet %>% 
  left_join(all_hist, by = c("game_id", "Username" = "player")) %>% 
  arrange(UTCDate, UTCTime) %>% 
  rename(focal_result = opp_result, focal_win_prop = opp_win_prop) %>% 
  left_join(all_hist, by = c("game_id", "opponent" = "player")) %>% 
  as_tibble()

## this is all good, just need to lag both and add them in,
## and also get the ids for the stan code

## getting the overall win probabilities for each player also

player_avg <- all_hist %>% 
  group_by(player) %>% 
  summarise(overall_avg = mean(opp_result))

## then just join this in also to use it

```

```{r getting the final data for potential stan model}

final_data <- tidy_stan %>% 
  left_join(player_avg, by = c("Username" = "player")) %>% 
  rename(focal_avg = overall_avg) %>% 
  left_join(player_avg, by = c("opponent" = "player")) %>% 
  rename(opp_avg = overall_avg) %>% 
  group_by(Username) %>% 
  arrange(UTCDate, UTCTime) %>% 
  mutate(focal_prev_result = lag(focal_result, default = 0)) %>% 
  ungroup() %>% 
  group_by(opponent) %>% 
  arrange(UTCDate, UTCTime) %>% 
  mutate(opp_prev_result = lag(opp_result, default = 0)) %>% 
  ungroup() %>% 
  mutate(focal_id = match(Username, players),
         opp_id = match(opponent, players)) %>% 
  filter(focal_result != 0.5)

```

```{r}

final_data

stan_data_all <- list(N = nrow(final_data),
                      J = length(players),
                      y = final_data$focal_result,
                      focal_id = final_data$focal_id,
                      opp_id = final_data$opp_id,
                      focal_prev = final_data$focal_prev_result,
                      opp_prev = final_data$opp_prev_result,
                      focal_avg = final_data$focal_avg,
                      opp_avg = final_data$opp_avg)

saveRDS(stan_data_all, file = "../cluster_scripts/stan_data.RDS")

stan_file <- "../Comp2.stan"

mod2 <- cmdstan_model(stan_file)

fit2 <- mod2$sample(data = stan_data_all,
                    seed = 123,
                    chains = 4,
                    parallel_chains = 4,
                    refresh = 100,
                    iter_sampling = 500,
                    iter_warmup = 100)

```

### 2023-05-23 owen/cluster_scripts/Fit_All.R

@@ -1,0 +1,27 @@
+### May 23rd, 2023,
+## to fit simple w/l effect model for large chess data on cluster
+## using cmdstanr and see what the output looks like
+
+
+library(cmdstanr)
+
+library(tidyverse)
+library(here)
+
+
+stan_data_all <- readRDS(here("owen", "cluster_scripts", "stan_data.RDS"))
+
+
+stan_file <- here("owen","Comp2.stan")
+
+mod2 <- cmdstan_model(stan_file)
+
+fit2 <- mod2$sample(data = stan_data_all,
+                    seed = 123,
+                    chains = 4,
+                    parallel_chains = 4,
+                    refresh = 100)
+
+fit2$summary()
+
+fit2$save_object(file = here("owen", "cluster_scripts", "Cluster_stan.RDS"))

### 2023-05-23 owen/cluster_scripts/Fit_All.R
Update to the fit file writing some code to get the versions of the models James shared somewhat up and running, have it ready to fit them on the cluster also

@@ -1,0 +1,29 @@
+### May 23rd, 2023,
+## to fit simple w/l effect model for large chess data on cluster
+## using cmdstanr and see what the output looks like
+
+
+library(cmdstanr)
+
+library(tidyverse)
+library(here)
+
+
+stan_data_all <- readRDS(here("owen", "cluster_scripts", "stan_data.RDS"))
+
+
+stan_file <- here("owen","Comp2.stan")
+
+mod2 <- cmdstan_model(stan_file)
+
+fit2 <- mod2$sample(data = stan_data_all,
+                    seed = 123,
+                    chains = 4,
+                    parallel_chains = 4,
+                    refresh = 100,
+		    iter_warmup = 2000, 
+		    iter_sampling = 2000)
+
+fit2$summary()
+
+fit2$save_object(file = here("owen", "cluster_scripts", "Cluster_stan.RDS"))

### 2023-05-25 owen/Write Ups/03_existing_models.qmd
update the data to use normalised running average

@@
 library(loo)
+library(here)
 
@@
     mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
-                              roll_mean(focal_result, n = prev_n)))s
+                              roll_mean(focal_result, n = prev_n)))
   hist_games
@@
   as_tibble() %>% 
-  select(game_id:player)
+  select(game_id:player, UTCDate, UTCTime)
+
+min_hist <- all_hist %>% select(game_id:player)
 
@@
 tidy_stan <- tidy_bullet %>% 
-  left_join(all_hist, by = c("game_id", "Username" = "player")) %>% 
+  left_join(min_hist, by = c("game_id", "Username" = "player")) %>% 
   arrange(UTCDate, UTCTime) %>% 
   rename(focal_result = opp_result, focal_win_prop = opp_win_prop) %>% 
-  left_join(all_hist, by = c("game_id", "opponent" = "player")) %>% 
+  left_join(min_hist, by = c("game_id", "opponent" = "player")) %>% 
   as_tibble()
@@
 
-## getting the overall win probabilities for each player also
+## getting the overall current win probabilities for each player also, 
+## up to and including the current game
 
@@
   group_by(player) %>% 
-  summarise(overall_avg = mean(opp_result))
+  arrange(UTCDate, UTCTime, .by_group = TRUE) %>%
+  mutate(nrow = row_number()) %>% 
+  mutate(curr_avg = cumsum(opp_result)/nrow) %>% 
+  select(game_id, player, curr_avg)
+
+global_avg <- all_hist %>% 
+  group_by(player) %>% 
+  summarise(total_avg = mean(opp_result))
 
@@
 final_data <- tidy_stan %>% 
-  left_join(player_avg, by = c("Username" = "player")) %>% 
-  rename(focal_avg = overall_avg) %>% 
-  left_join(player_avg, by = c("opponent" = "player")) %>% 
-  rename(opp_avg = overall_avg) %>% 
+  left_join(player_avg, by = c("Username" = "player", "game_id")) %>% 
+  rename(focal_avg = curr_avg) %>% 
+  left_join(player_avg, by = c("opponent" = "player", "game_id")) %>% 
+  rename(opp_avg = curr_avg) %>% 
   group_by(Username) %>% 
   arrange(UTCDate, UTCTime) %>% 
-  mutate(focal_prev_result = lag(focal_result, default = 0)) %>% 
+  mutate(focal_prev_result = lag(focal_result, default = 0),
+         focal_runn_avg = lag(focal_avg, n = 2, default = 0)) %>% 
   ungroup() %>% 
@@
   arrange(UTCDate, UTCTime) %>% 
-  mutate(opp_prev_result = lag(opp_result, default = 0)) %>% 
+  mutate(opp_prev_result = lag(opp_result, default = 0),
+         opp_runn_avg = lag(opp_avg, n = 2, default = 0)) %>% 
   ungroup() %>% 
@@
          opp_id = match(opponent, players)) %>% 
-  filter(focal_result != 0.5)
+  filter(focal_result != 0.5) %>% 
+  select(-focal_avg, - opp_avg, - opp_result, 
+         - focal_win_prop, - opp_win_prop)
+
+final_data
+## remove some unused columns
+
+norm_data <- final_data %>% 
+  left_join(global_avg, by = c("Username" = "player")) %>% 
+  rename(focal_global_avg = total_avg) %>% 
+  left_join(global_avg, by = c("opponent" = "player")) %>% 
+  rename(opp_global_avg = total_avg) %>% 
+  mutate(focal_prev_result = focal_prev_result - focal_global_avg,
+         focal_runn_avg = focal_runn_avg - focal_global_avg,
+         opp_prev_result = opp_prev_result - opp_global_avg,
+         opp_runn_avg = opp_runn_avg - opp_global_avg)
 
@@
 
+Now this matches nicely, contains their historic winning average up to 2 games
+previously, along with their result in the previous game.
+
+Note that this has not been standardized currently.
+
 ```{r}
 
-final_data
+norm_data
 
-stan_data_all <- list(N = nrow(final_data),
+stan_data_all <- list(N = nrow(norm_data),
                       J = length(players),
-                      y = final_data$focal_result,
-                      focal_id = final_data$focal_id,
-                      opp_id = final_data$opp_id,
-                      focal_prev = final_data$focal_prev_result,
-                      opp_prev = final_data$opp_prev_result,
-                      focal_avg = final_data$focal_avg,
-                      opp_avg = final_data$opp_avg)
+                      y = norm_data$focal_result,
+                      focal_id = norm_data$focal_id,
+                      opp_id = norm_data$opp_id,
+                      focal_prev = norm_data$focal_prev_result,
+                      opp_prev = norm_data$opp_prev_result,
+                      focal_avg = norm_data$focal_runn_avg,
+                      opp_avg = norm_data$opp_runn_avg)
 
@@
 ```
+
+## Fitting this to all data
+
+Fit this model to all data on the cluster, just to see the results.
+The traceplots and R-hat for this model indicates it doesn't work well,
+does not converge.
+
+## Comparing models on smaller data
+
+Is there a small subset of players who play each other often that we can
+use in the dataset that can be used to compare these models?
+
+```{r}
+
+bullet_60 %>% 
+  filter(White %in% focal_players) %>% 
+  filter(Black %in% focal_players)
+
+## no games between any of the players used to scrape the data
+## would maybe need to get more players then, to use this for a comparison
+
+```
+
+```{r}
+library(chessR)
+
+chessdotcom_leaders <- chessdotcom_leaderboard(game_type = "live_bullet")
+
+top_bullet <- chessdotcom_leaders %>% top_n(10, wt = -rank) %>% 
+  select(player_id, username)
+
+chessdotcom_hikaru_recent <- get_raw_chessdotcom(usernames = "Hikaru",
+                                                 year_month = c(202201:202205))
+
+jan_games <- top_bullet$username %>% 
+  map_dfr(~get_raw_chessdotcom(.x, year_month = 202201))
+
+top_games <- jan_games %>% 
+  filter(White %in% top_bullet$username) %>% 
+  filter(Black %in% top_bullet$username)
+
+## get 500 games here then for 1 month, probably enough to fit an initial model
+## most of these games bullet too, so could use that
+
+top_games %>% 
+  group_by(White, Black, TimeClass) %>% 
+  tally() %>% 
+  arrange(-n)
+
+top_games %>% 
+  filter(TimeClass == "blitz")
+
+```

### 2023-05-25 owen/cluster_scripts/Fit_All.R

@@ -5,6 +5,8 @@
 
 library(cmdstanr)
 
+options(mc.cores = parallel::detectCores())
+
 library(tidyverse)
 library(here)
 
@@ -20,7 +22,7 @@ fit2 <- mod2$sample(data = stan_data_all,
                     seed = 123,
                     chains = 4,
                     parallel_chains = 4,
-                    refresh = 100)
+                    refresh = 100, iter)
 
 fit2$summary()
 

### 2023-05-26 owen/Write Ups/03_existing_models.qmd
identifying potential problems with the model

@@
 
+```{r}
+stan_fit <- readRDS("../cluster_scripts/Cluster_stan.RDS")
+
+mcmc_hist(stan_fit$draws(c("alpha[1]", "alpha[2]")))
+
+r_hat <- rhat(stan_fit)
+
+summary(r_hat)
+
+which(r_hat > 1.15)
+## just problems with mu and tau for convergence here, otherwise good?
+
+```
+
 ## Comparing models on smaller data
@@
 
-chessdotcom_hikaru_recent <- get_raw_chessdotcom(usernames = "Hikaru",
-                                                 year_month = c(202201:202205))
+# chessdotcom_hikaru_recent <- get_raw_chessdotcom(usernames = "Hikaru",
+#                                                  year_month = c(202201:202205))
 
-jan_games <- top_bullet$username %>% 
-  map_dfr(~get_raw_chessdotcom(.x, year_month = 202201))
+games_22 <- top_bullet$username %>% 
+  map_dfr(~get_raw_chessdotcom(.x, year_month = c(202201:202203)))
 
-top_games <- jan_games %>% 
+top_games <- games_22 %>% 
   filter(White %in% top_bullet$username) %>% 
@@
 
-top_games %>% 
-  filter(TimeClass == "blitz")
+bullet_games <- top_games %>% 
+  filter(TimeClass == "bullet")
+
+bullet_top10 <- saveRDS(bullet_games, file = "chess_com_bullet_10.RDS")
 

### 2023-05-26 owen/cluster_scripts/Fit_All.R
update cluster script so that can run more complex models, add updated version of the data so that it is an exact comparison to the data I used for my models

@@ -9,9 +9,11 @@ options(mc.cores = parallel::detectCores())
 
 library(tidyverse)
 library(here)
+library(loo)
+library(ggplot2)
 
 
-stan_data_all <- readRDS(here("owen", "cluster_scripts", "stan_data.RDS"))
+stan_data_all <- readRDS(here("owen", "cluster_scripts", "stan_data_ppt_n10.RDS"))
 
 
 stan_file <- here("owen","Comp2.stan")
@@ -29,6 +31,65 @@ fit2 <- mod2$sample(data = stan_data_all,
 
 fit2$summary()
 
-fit2$save_object(file = here("owen", "cluster_scripts", "Cluster_stan.RDS"))
+fit2$save_object(file = here("owen", "cluster_scripts", "Cluster_stan_ppt_n10.RDS"))
 
 
+## Do some model checking here using the draws to get 
+## posterior predictive distributions
+
+fit_samples <- as_draws_df(fit2$draws())
+
+
+fit_samp <- fit_samples %>% 
+  select(!starts_with(c("log_lik", "yrep")))
+
+y_rep <- fit_samples %>% select(starts_with("y_rep"))
+
+
+
+### then want to join this with the correct input, i.e player, etc
+
+
+y_rep_mod <- y_rep %>% 
+  rownames_to_column() %>%  
+  pivot_longer(-rowname) %>% 
+  pivot_wider(names_from=rowname, values_from=value) %>% 
+  ## remove the draws and chains here
+  mutate(focal_id = stan_data_all$focal_id)
+
+
+## what is the check here? number of games won by a player for each of
+## the posterior samples, compared to the number of games the actually won
+
+
+games_won <- y_rep_mod %>% 
+  pivot_longer(cols = `1`:`8000`, names_to = "draw", values_to = "y") %>% 
+  group_by(focal_id, draw) %>% 
+  summarise(games_won = sum(y)) 
+
+
+orig_data <- tibble(outcome = stan_data_all$y,
+                    focal_id = stan_data_all$focal_id)
+
+orig_games_won <- orig_data %>% 
+  group_by(focal_id) %>% 
+  summarise(games_won = sum(outcome))
+
+
+p1 <- games_won %>% 
+  ggplot(aes(x = games_won)) +
+  geom_histogram() +
+  facet_wrap(~focal_id, scales = "free") +
+  geom_vline(data = orig_games_won, 
+             mapping = aes(xintercept = games_won), col = "red") +
[…]

### 2023-05-27 owen/cluster_scripts/Fit_All.R
runs

@@ -11,7 +11,7 @@ library(tidyverse)
 library(here)
 library(loo)
 library(ggplot2)
-
+library(posterior)
 
 stan_data_all <- readRDS(here("owen", "cluster_scripts", "stan_data_ppt_n10.RDS"))
 
@@ -20,19 +20,20 @@ stan_file <- here("owen","Comp2.stan")
 
 mod2 <- cmdstan_model(stan_file)
 
-fit2 <- mod2$sample(data = stan_data_all,
-                    seed = 123,
-                    chains = 4,
-                    parallel_chains = 4,
-                    refresh = 100,
-		                iter_warmup = 2000, 
-		                iter_sampling = 2000)
+#fit2 <- mod2$sample(data = stan_data_all,
+#                    seed = 123,
+#                    chains = 4,
+#                    parallel_chains = 4,
+#                    refresh = 100,
+#		                iter_warmup = 2000, 
+#		                iter_sampling = 2000)
 
 
-fit2$summary()
+#fit2$summary()
 
-fit2$save_object(file = here("owen", "cluster_scripts", "Cluster_stan_ppt_n10.RDS"))
+#fit2$save_object(file = here("owen", "cluster_scripts", "Cluster_stan_ppt_n10.RDS"))
 
+fit2 <- readRDS(here("owen", "cluster_scripts", "Cluster_stan_ppt_n10.RDS"))
 
 ## Do some model checking here using the draws to get 
 ## posterior predictive distributions
@@ -90,6 +91,6 @@ ggsave(filename = here("owen", "cluster_scripts", "ppd_model.png"), plot = p1)
 
 fit4 <- readRDS(here("owen", "model_fits", "model4.RDS"))
 
-comp <- loo_compare(fit2, fit4)
+comp <- loo_compare(fit2$loo(), fit4$loo())
 
 print(comp)

### 2023-05-27 owen/cluster_scripts/Fit_All.R
beginning to tidy

@@ -11,7 +11,7 @@ library(tidyverse)
 library(here)
 library(loo)
 library(ggplot2)
-
+library(posterior)
 
 stan_data_all <- readRDS(here("owen", "cluster_scripts", "stan_data_ppt_n10.RDS"))
 
@@ -31,7 +31,8 @@ fit2 <- mod2$sample(data = stan_data_all,
 
 fit2$summary()
 
-fit2$save_object(file = here("owen", "cluster_scripts", "Cluster_stan_ppt_n10.RDS"))
+fit2$save_object(file = here("owen", "cluster_scripts",
+                             "Cluster_stan_ppt_2_n10.RDS"))
 
 
 ## Do some model checking here using the draws to get 
@@ -84,7 +85,7 @@ p1 <- games_won %>%
              mapping = aes(xintercept = games_won), col = "red") +
   labs(title = "PPT Model, No Covariates")
 
-ggsave(filename = here("owen", "cluster_scripts", "ppd_model.png"), plot = p1)
+ggsave(filename = here("owen", "cluster_scripts", "ppd_model2.png"), plot = p1)
 
 ## then load in the competing model and compare them using loo
 

### 2023-05-27 owen/Write Ups/03_existing_models.qmd

--- 2023-05-23 owen/Write Ups/03_existing_models.qmd
@@
 author: Owen G. Ward
-date: 04/17/2023
+date: 05/26/2023
 format: 
@@
 
+This document will summarise what has been done so far.
+
 ```{r setup}
@@
 
-knitr::opts_chunk$set(echo = FALSE, message = FALSE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 8)
 
@@
 
-## Model Comparison
+## Initial Models Fit
 
-We want to compare these potential models, which is very similar
-to the model we fit previously.
-The only change is that we now need the history for
-both players, not just the focal player.
+Based on @ding2021nhl, an initial model proposed was of the form
 
-```{r load in the initial data}
-#| cache: true
+$$
+P(y_{ij} = 1) = \frac{1}{1 + \exp(-(\alpha_j + \beta_j x_{ij} + \gamma z_{ij}))}
+$$
 
-lichess_data <- readRDS("../../rdata/lichess_pilot.RData")
+where this is the probability of the **j-th focal player** 
+winning their **i-th game**. Here:
 
-bullet_60 <- lichess_data %>% 
-  filter(Event == "Rated Bullet game") %>% 
-  filter(TimeControl == "60+0") %>% 
-  mutate(game_id = row_number())
+- $\alpha_j$ is a player level random effect
+- $\beta_j$ is a player level random effect, accounting for the win ratio $x_{ij}$
+of the focal player over their previous $n$ games
+- $\gamma$ is a fixed effect of game level coefficients
 
-get_hist <- function(user, games, prev_n) {
-  hist_games <- games %>% 
-    filter(White == user | Black == user) %>% 
-    arrange(UTCDate, UTCTime) %>% 
-    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
-    select(White:BlackElo, focal_white, game_id) %>% 
-    mutate(focal_result = case_when(
-      (focal_white == 1 & Result == "1-0") ~ 1,
-      (focal_white == 0 & Result == "0-1") ~ 1,
-      (Result == "1/2-1/2") ~ 0.5,
-      .default = 0
-      )) %>% 
-    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
-                              roll_mean(focal_result, n = prev_n)))s
-  hist_games
-}
+We then partially pool the $\alpha$ and $\beta$ coefficients, so
 
-focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18)
-## if don't use all focal players here then removing data because they won't be
-## involved
+$$
+\alpha_j \sim \mathcal{N}(\mu_2, \tau_2)
+$$
+$$
+\mu_2 \sim \mathcal{N}(0, 5)
+$$
+$$
+\tau_2 \sim \mathcal{Cauchy}(0, 5)
+$$
 
-top_players <- names(focal_players)
+and similarly
 
-## then get the opponents of these players
+$$
+\beta_j \sim \mathcal{N}(\mu_1, \tau_1)
+$$
+$$
+\mu_1 \sim \mathcal{N}(0, 5)
+$$
+$$
+\tau_1 \sim \mathcal{Cauchy}(0, 5)
+$$
 
-full_opponents <- bullet_60 %>% 
-  filter(Username %in% top_players) %>% 
-  select(White, Black)
+We rescale the win ratio $x_{ij}$, by the average win ratio of
+player $j$ across all 
+games they play in the dataset.
+This means $\beta_j$ captures how a players win
+probability changes as their historic performance deviates from 
+the overall average win ratio.
 
-all_opps <- c(full_opponents$White, full_opponents$Black)
+For $z_{ij}$ we incorporate 2 covariates, namely the colour of 
+the focal player and the ELO difference between the focal 
+player and their opponent. So that means we have 
+$\gamma_1$, which tells us how the win probability changes in going
+from black to white,
+and $\gamma_2$ gives how the win probability changes as the focal player
+has a better ELO than their opponent.
 
-opponents <- unique(all_opps)
+We can show this result, taking the original lichess data
+and the 18 players with the most games as focal players.
 
-## then need to get the history of these players, but maybe need
-## to store game id
-## can just match using date and time I guess?
-## need to modify the function then to account for this
+```{r load in initial model here}
+
+fit3_ave <- readRDS("../model_fits/init_model_n10.RDS")
+
+# fit3_ave$summary()
+
+mcmc_hist(fit3_ave$draws(c("alpha", "mu2", "tau2")))
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
+```
+
+#### Interpreting this output
+
+Given this output, we can get the probability a player who is black
+will win against an equally ranked player, when their current win ratio 
+is exactly their average will be $\frac{1}{1+\exp(-\alpha_j)}$.
+So a positive $\alpha_j$ indicates a player favoured to win 
+even when playing black. The overall $\mu_2$ tells us about the 
+global probability of winning as black against an equally ranked
+opponent, which is just below 0, indicating a slightly less than 50% 
+chance of winning.
+If a player
+is playing as white their win probability will be
+$\frac{1}{1+\exp(-(\alpha_j + \gamma_1))}$,
+so $\gamma_1$ being positive indicates the win
+probability increases when playing white, which makes sense.
+Playing white makes a player who has 50% chance of winning as black
+have approximately a 54% chance of winning with white. Wikipedia says the increase 
+in win probability from playing white has been estimated to be 
+4-6% so quite similar.
+
+Similarly, $\gamma_2$ tells us about how the ELO of the focal player 
+being larger than their opponent influences their win probability.
+The estimate here indicates that a large ELO difference makes a player very
+likely to win, as would be expected. For example, a player who
+has a 50% chance of winning against an opponent with the same ELO
+will have about a 61% chance of winning against an opponent with an ELO score
+100 points lower.
+
+#### How big are these effects
+
+We then want to quantify how much these estimated effects change the 
+expected win probability.
+We have seen the size of the effects above for the non history parameters.
+The largest posterior mean for a $\beta\approx 1.56$. Since the normalized win
+proportion is between -0.6 and 0.6, that means that for a game
+where this player has a 50% chance of winning with a normalized average win
+proportion of 0, if their normalized win proportion is 0.5, the
+win probability will increase by $\approx 19\%$.
+This would occur, for example, if they win 50% of their total games but have
+won all 10 of their last 10. This is a similar size effect to the 
+increase in win probability which would come from playing an 
+opponent with an ELO ranking that is 170 points lower.
+
+Similarly, a normalized win history 0.25 
+above average corresponds to a difference 
+of approximately 80 in ELO scores.
+
+The underlying parameter $\mu_1$ has posterior mean of 0.606. 
+A normalised win history of 0.25 with a winner effect of this size
+would increase the win probability by about 4%, a similar effect
+as playing as white instead of black, all other covariates being
+equal.
+
+#### Winner Effects here
+
+Here, the presence of positive estimates for $\beta$ would indicate the
+presence of a __winner/loser__ effect, where a player is performing
+better (worse) than their overall performance is more (less)
+likely to win their subsequent games. As some of these distributions
+take both positive and negative values, we can compute the 
+posterior probability that $\beta>0$ for each of the focal players.
+
+```{r prob positive fit3 beta}
+#| echo: true
+prob_positive(fit3_ave$draws())
+```
+
+This indicates that most of these $\beta$ parameters are strongly seen 
+as positive.
+
+### Updated Model
+
+An extension of this model was then to consider the "history" 
+as being composed of the previous game and a more general "winning streak",
+which is fit by the following model.
+
+$$
+P(y_{ij} = 1) = \frac{1}{1 + \exp(-(\alpha_j + \beta_j x_{ij} + 
+\delta_j x_{ij}^{*} + 
+\gamma z_{ij}))},
+$$
+
+where now $x_{ij}^{*}$ is the result of the __previous__ game
+played by player $j$, and $x_{ij}$ is their normalised 
+win ratio over the previous $n$ games, not including the single
+previous game. Note that $x_{ij}^{*}$ is not normalised,
+so can take on values $0,0.5,1$.
+
+```{r load the stan fit}
+#| fig-width: 8
+#| fig-height: 6
+
+fit4 <- readRDS("../model_fits/model4.RDS")
+
+mcmc_hist(fit4$draws(c("delta", "mu3", "tau3"))) +
+  labs(title = "Effect of Previous Game")
+
+mcmc_hist(fit4$draws(c("beta", "mu1", "tau1"))) +
+  labs(title = "Effect of Previous n games, excluding most recent game")
+
+```
+
+The $\delta$ parameters then indicate the effect that the previous
+game alone has on the next game, while $\beta$ indicates the
+effect of the overall current streak (excluding the previous game).
+
+**Should the previous game be normalised to give deviation from average 
+performance?**
+
+### Model Checking Comparison
+
+We can use the posterior draws to do some model checking, generating
+synthetic data based on the posterior draws. These indicate that 
+our models are reasonable and able to describe this data well.
+Similarly we can do some model checking which indicates the model
+which separates out the history performs substantially better 
+than the other model.
+
+## Model from Powerpoint
+
+The slides shared give a similar form of model for this type
+of data to estimate winner/loser effects. 
+Written in the format shown here, that model would look like
+
+$$
+P(y_{Fi} = 1) =\frac{1}{1 + exp(-(\alpha_F + \alpha_{opp_i} + \beta_j\hat{w}_{Fi})
++ \beta_{opp_i}\hat{w}_{Oi} + \delta_j(w_{F,i-1} - \hat{w}_{Fi}) +
+\delta_{opp_i}(w_{O,i-1}-\hat{w}_{Oi})))}.
+$$
+
+Here we have:
+
+- $\hat{w}_{Fi}, \hat{w}_{Oi}$ is some measure of the long run performance
+focal player and their opponent
+- $w_{F,i-1} - \hat{w}_{Fi}$ measures the deviation between both players 
+previous result and their more long run performance.
+
+One difference with this model is that we now need to consider the 
+history of all opponents played by focal players also.
+For the chess example where many players are involved, this leads to thousands
+of opponents, many who plan only a small number of games.
+
+Covariates could also be included here, as described before.
+This model can be fit to the same initial dataset we considered above.
+
+```{r get same data as before for this example}
+#| eval: false
 
@@
 }
-## so now the history is the history corresponding to the name
-## in the "player" column
 
-remaining_opponents <- opponents[!(opponents %in% top_players)]
+focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18)
+## if don't use all focal players here then removing data because they won't be
+## involved
 
-opp_hist <- map_dfr(remaining_opponents, get_hist_opp, bullet_60, prev_n = 5) %>% 
-  as_tibble()
+top_players <- names(focal_players)
 
-## so this gives the history of the opponents for each of their 
-## games, can use this to then first get their previous result, etc,
-## then join this with the other results
-
-prev_opp <- opp_hist %>% 
-  group_by(player) %>% 
-  arrange(UTCDate, UTCTime) %>% 
-  mutate(prev_opp_result = lag(opp_result, default = 0), 
-         prev_opp_prop = lag(opp_win_prop, default = 0)) %>%
-  select(game_id, player, prev_opp_result, prev_opp_prop)
-
-tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 5) %>% 
-  as_tibble()
-
-all <- tidy_games %>% 
-  left_join(prev_opp, by = "game_id") %>% 
-  mutate(focal = ifelse(focal_white == 1, White, Black))
-## this looks more reasonable now
-## this has the history of the opponent included, not the history 
-## of the focal player yet
-
-final_data <- all %>% 
-  group_by(focal) %>% 
-  arrange(UTCDate, UTCTime) %>% 
-  mutate(prev_focal_result = lag(focal_result, default = 0),
-         prev_focal_prop = lag(focal_win_prop, default = 0)) %>% 
-  ungroup()
-
-## then tidy this up for stan
-
-```
-
-This object `prev_game_both` should have the previous result of both players in
-each game, if that game is in the dataset.
-Similarly, `final_prev` has the games played by the focal players, along 
-with their previous result and their opponents previous result, if that
-game was in the original dataset used (all 60 second bullet games here).
-
-Then convert this into data for stan.
-
-```{r get stan data for simple model}
-
-final_data
-
-stan_prev <- final_data %>% 
-  mutate(WhiteElo = as.numeric(WhiteElo), 
-         BlackElo = as.numeric(BlackElo)) %>% 
-  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>%
-  mutate(elo_diff = ifelse(focal_white == 1, 
-                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
-  mutate(focal_id = match(focal_user, top_players)) %>% 
-  select(focal_user, focal_id, focal_white, 
-         prev_focal_result, prev_opp_result, focal_result)
-
-stan_data <- list(N = nrow(stan_prev),
-                  J = length(top_players),
-                  y = stan_prev$focal_result,
-                  id = stan_prev$focal_id,
-                  focal_prev = stan_prev$prev_focal_result,
-                  opp_prev = stan_prev$prev_opp_result)
-
-```
-
-```{r fit the simple stan model}
-#| eval: false
-
-## remember no effect for non focal players in this at all 
-## at the moment
-
-stan_file <- "../Comp1.stan"
-
-mod <- cmdstan_model(stan_file)
-
-fit1_simple <- mod$sample(data = stan_data,
-                  seed = 123,
-                  chains = 4,
-                  parallel_chains = 4,
-                  refresh = 100)
-
-## this fits at least now...
-
-fit1_simple$summary()
-mcmc_hist(fit1_simple$draws(c("alpha", "beta")))
-
-```
-
-## Notes on these models
-
-- How does the choice of focal player make a difference in these models?
-
-```{r thinking about the choice of focal}
-
-table(bullet_60$Username)
-
-tidy_bullet <- bullet_60 %>% 
+kept_games <- bullet_60 %>% 
+  filter(Username %in% top_players) %>% 
   select(White:BlackElo, Username, game_id) %>%
@@
 
-all_players <- c(tidy_bullet$Username, tidy_bullet$opponent)
-players <- unique(all_players)
+curr_players <- unique(c(kept_games$White, kept_games$Black))
 
-## then just get all the history for this
-all_hist <- map_dfr(players, get_hist_opp, bullet_60, prev_n = 5) %>% 
+all_hist <- map_dfr(curr_players, get_hist_opp, bullet_60, prev_n = 10) %>%
   as_tibble() %>%
-  select(game_id:player)
+  select(game_id:player, UTCDate, UTCTime)
 
-## this should have each game appear twice, for each of the two players
-## involved, so need to figure out how to handle that best
-## left join with tidy games matching player to focal and then
-## to opponent, rename the columns as needed
+min_hist <- all_hist %>% select(game_id:player)
 
-tidy_stan <- tidy_bullet %>% 
-  left_join(all_hist, by = c("game_id", "Username" = "player")) %>% 
+tidy_stan <- kept_games %>%
+  left_join(min_hist, by = c("game_id", "Username" = "player")) %>%
   arrange(UTCDate, UTCTime) %>%
   rename(focal_result = opp_result, focal_win_prop = opp_win_prop) %>%
-  left_join(all_hist, by = c("game_id", "opponent" = "player")) %>% 
+  left_join(min_hist, by = c("game_id", "opponent" = "player")) %>%
   as_tibble()
-
-## this is all good, just need to lag both and add them in,
-## and also get the ids for the stan code
-
-## getting the overall win probabilities for each player also
 
@@
   group_by(player) %>%
-  summarise(overall_avg = mean(opp_result))
+  arrange(UTCDate, UTCTime, .by_group = TRUE) %>%
+  mutate(nrow = row_number()) %>%
+  mutate(curr_avg = cumsum(opp_result)/nrow) %>%
+  select(game_id, player, curr_avg)
 
-## then just join this in also to use it
+global_avg <- all_hist %>%
+  group_by(player) %>%
+  summarise(total_avg = mean(opp_result))
 
@@
 
-```{r getting the final data for potential stan model}
+```{r prep this for the stan model}
+#| eval: false
 
 final_data <- tidy_stan %>%
-  left_join(player_avg, by = c("Username" = "player")) %>% 
-  rename(focal_avg = overall_avg) %>% 
-  left_join(player_avg, by = c("opponent" = "player")) %>% 
-  rename(opp_avg = overall_avg) %>% 
+  left_join(player_avg, by = c("Username" = "player", "game_id")) %>%
+  rename(focal_avg = curr_avg) %>%
+  left_join(player_avg, by = c("opponent" = "player", "game_id")) %>%
+  rename(opp_avg = curr_avg) %>%
   group_by(Username) %>%
-  arrange(UTCDate, UTCTime) %>% 
-  mutate(focal_prev_result = lag(focal_result, default = 0)) %>% 
+  arrange(UTCDate, UTCTime, .by_group = TRUE) %>%
+  mutate(focal_prev_result = lag(focal_result, default = 0),
+         focal_runn_avg = lag(focal_win_prop, n = 2, default = 0)) %>%
   ungroup() %>%
   group_by(opponent) %>%
-  arrange(UTCDate, UTCTime) %>% 
-  mutate(opp_prev_result = lag(opp_result, default = 0)) %>% 
+  arrange(UTCDate, UTCTime, .by_group = TRUE) %>%
+  mutate(opp_prev_result = lag(opp_result, default = 0),
+         opp_runn_avg = lag(opp_win_prop, n = 2, default = 0)) %>%
   ungroup() %>%
-  mutate(focal_id = match(Username, players),
-         opp_id = match(opponent, players)) %>% 
-  filter(focal_result != 0.5)
+  mutate(focal_id = match(Username, curr_players),
+         opp_id = match(opponent, curr_players)) %>%
+  filter(focal_result != 0.5) %>%
+  select(-focal_avg, - opp_avg, - opp_result,
+         - focal_win_prop, - opp_win_prop) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, as.numeric(WhiteElo) - 
+                             as.numeric(BlackElo), 
+                           as.numeric(BlackElo) - as.numeric(WhiteElo)))
+
+final_data
+
+## now running average is based on the previous 10 games, not all previous games
 
@@
 
-```{r}
-
-final_data
+```{r process for stan and fit}
+#| eval: false
 
 stan_data_all <- list(N = nrow(final_data),
-                      J = length(players),
+                      J = length(curr_players),
                       y = final_data$focal_result,
@@
                       opp_prev = final_data$opp_prev_result,
-                      focal_avg = final_data$focal_avg,
-                      opp_avg = final_data$opp_avg)
+                      focal_avg = final_data$focal_runn_avg,
+                      opp_avg = final_data$opp_runn_avg,
+                      elo_diff = final_data$elo_diff,
+                      colour = final_data$focal_white)
 
-saveRDS(stan_data_all, file = "../cluster_scripts/stan_data.RDS")
+saveRDS(stan_data_all, file = "../cluster_scripts/stan_data_ppt_n10.RDS")
 
@@
 ```
+
+```{r plot some parts of this model}
+
+## estimated effects for the focal players
+## posterior predictive checks
+
+```
+
+Comparing the predictive performance of the model from 
+the slides (**without covariates**) it is unsurprisingly 
+not comparable to the current iteration of the model we have here.
+
+## Questions, Comments
+
+- The choice of the focal player for these types of models. Is there a
+way this should be done?
+
+- I think that you probably do need to include the history of 
+both players in a match. Otherwise could identify a winner effect for the
+focal player which is actually a loser effect for the opponent. However,
+it may not be practical to have a separate effect for all players in a
+given dataset.
+
+- Need to think some more about what to normalize and how important
+that is for models of this form.
+
+-----
+
+<!-- <!-- ## Model Comparison -->
+
+<!-- We want to compare these potential models, which is very similar -->
+<!-- to the model we fit previously. -->
+<!-- The only change is that we now need the history for -->
+<!-- both players, not just the focal player. -->
+
+<!-- ```{r load in the initial data} -->
+<!-- #| cache: true -->
+
+<!-- lichess_data <- readRDS("../../rdata/lichess_pilot.RData") -->
+
+<!-- bullet_60 <- lichess_data %>% -->
+<!--   filter(Event == "Rated Bullet game") %>% -->
+<!--   filter(TimeControl == "60+0") %>% -->
+<!--   mutate(game_id = row_number()) -->
+
+<!-- get_hist <- function(user, games, prev_n) { -->
+<!--   hist_games <- games %>% -->
+<!--     filter(White == user | Black == user) %>% -->
+<!--     arrange(UTCDate, UTCTime) %>% -->
+<!--     mutate(focal_white = ifelse(Username == White, 1, 0)) %>% -->
+<!--     select(White:BlackElo, focal_white, game_id) %>% -->
+<!--     mutate(focal_result = case_when( -->
+<!--       (focal_white == 1 & Result == "1-0") ~ 1, -->
+<!--       (focal_white == 0 & Result == "0-1") ~ 1, -->
+<!--       (Result == "1/2-1/2") ~ 0.5, -->
+<!--       .default = 0 -->
+<!--       )) %>% -->
+<!--     mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), -->
+<!--                               roll_mean(focal_result, n = prev_n))) -->
+<!--   hist_games -->
+<!-- } -->
+
+<!-- focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18) -->
+<!-- ## if don't use all focal players here then removing data because they won't be -->
+<!-- ## involved -->
+
+<!-- top_players <- names(focal_players) -->
+
+<!-- ## then get the opponents of these players -->
+
+<!-- full_opponents <- bullet_60 %>% -->
+<!--   filter(Username %in% top_players) %>% -->
+<!--   select(White, Black) -->
+
+<!-- all_opps <- c(full_opponents$White, full_opponents$Black) -->
+
+<!-- opponents <- unique(all_opps) -->
+
+<!-- ## then need to get the history of these players, but maybe need -->
+<!-- ## to store game id -->
+<!-- ## can just match using date and time I guess? -->
+<!-- ## need to modify the function then to account for this -->
+
+<!-- get_hist_opp <- function(user, games, prev_n) { -->
+<!--   hist_games <- games %>% -->
+<!--     filter(White == user | Black == user) %>% -->
+<!--     arrange(UTCDate, UTCTime) %>% -->
+<!--     mutate(opp_white = ifelse(user == White, 1, 0)) %>% -->
+<!--     select(White:BlackElo, opp_white, game_id) %>% -->
+<!--     mutate(opp_result = case_when( -->
+<!--       (opp_white == 1 & Result == "1-0") ~ 1, -->
+<!--       (opp_white == 0 & Result == "0-1") ~ 1, -->
+<!--       (Result == "1/2-1/2") ~ 0.5, -->
+<!--       .default = 0 -->
+<!--       )) -->
+
+<!--   ngames <- nrow(hist_games) -->
+
+<!--   if(ngames < prev_n) { -->
+<!--     hist_games <- hist_games %>% -->
+<!--     mutate(opp_win_prop = opp_result/(1:ngames)) -->
+<!--   } -->
+<!--   else{ -->
+<!--     hist_games <- hist_games %>% -->
+<!--       mutate(opp_win_prop = c(cumsum(opp_result[1:(prev_n - 1)])/(1:(prev_n -1)), -->
+<!--                               roll_mean(opp_result, n = prev_n))) -->
+<!--   } -->
+<!--   hist_games$player <- user -->
+<!--   hist_games -->
+<!-- } -->
+<!-- ## so now the history is the history corresponding to the name -->
+<!-- ## in the "player" column -->
+
+<!-- remaining_opponents <- opponents[!(opponents %in% top_players)] -->
+
+<!-- opp_hist <- map_dfr(remaining_opponents, get_hist_opp, bullet_60, prev_n = 5) %>% -->
+<!--   as_tibble() -->
+
+<!-- ## so this gives the history of the opponents for each of their -->
+<!-- ## games, can use this to then first get their previous result, etc, -->
+<!-- ## then join this with the other results -->
+
+<!-- prev_opp <- opp_hist %>% -->
+<!--   group_by(player) %>% -->
+<!--   arrange(UTCDate, UTCTime) %>% -->
+<!--   mutate(prev_opp_result = lag(opp_result, default = 0), -->
+<!--          prev_opp_prop = lag(opp_win_prop, default = 0)) %>% -->
+<!--   select(game_id, player, prev_opp_result, prev_opp_prop) -->
+
+<!-- tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 5) %>% -->
+<!--   as_tibble() -->
+
+<!-- all <- tidy_games %>% -->
+<!--   left_join(prev_opp, by = "game_id") %>% -->
+<!--   mutate(focal = ifelse(focal_white == 1, White, Black)) -->
+<!-- ## this looks more reasonable now -->
+<!-- ## this has the history of the opponent included, not the history -->
+<!-- ## of the focal player yet -->
+
+<!-- final_data <- all %>% -->
+<!--   group_by(focal) %>% -->
+<!--   arrange(UTCDate, UTCTime) %>% -->
+<!--   mutate(prev_focal_result = lag(focal_result, default = 0), -->
+<!--          prev_focal_prop = lag(focal_win_prop, default = 0)) %>% -->
+<!--   ungroup() -->
+
+<!-- ## then tidy this up for stan -->
+
+<!-- ``` -->
+
+<!-- This object `prev_game_both` should have the previous result of both players in -->
+<!-- each game, if that game is in the dataset. -->
+<!-- Similarly, `final_prev` has the games played by the focal players, along -->
+<!-- with their previous result and their opponents previous result, if that -->
+<!-- game was in the original dataset used (all 60 second bullet games here). -->
+
+<!-- Then convert this into data for stan. -->
+
+<!-- ```{r get stan data for simple model} -->
+
+<!-- final_data -->
+
+<!-- stan_prev <- final_data %>% -->
+<!--   mutate(WhiteElo = as.numeric(WhiteElo), -->
+<!--          BlackElo = as.numeric(BlackElo)) %>% -->
+<!--   mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% -->
+<!--   mutate(elo_diff = ifelse(focal_white == 1, -->
+<!--                            WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% -->
+<!--   mutate(focal_id = match(focal_user, top_players)) %>% -->
+<!--   select(focal_user, focal_id, focal_white, -->
+<!--          prev_focal_result, prev_opp_result, focal_result) -->
+
+<!-- stan_data <- list(N = nrow(stan_prev), -->
+<!--                   J = length(top_players), -->
+<!--                   y = stan_prev$focal_result, -->
+<!--                   id = stan_prev$focal_id, -->
+<!--                   focal_prev = stan_prev$prev_focal_result, -->
+<!--                   opp_prev = stan_prev$prev_opp_result) -->
+
+<!-- ``` -->
+
+<!-- ```{r fit the simple stan model} -->
+<!-- #| eval: false -->
+
+<!-- ## remember no effect for non focal players in this at all -->
+<!-- ## at the moment -->
+
+<!-- stan_file <- "../Comp1.stan" -->
+
+<!-- mod <- cmdstan_model(stan_file) -->
+
+<!-- fit1_simple <- mod$sample(data = stan_data, -->
+<!--                   seed = 123, -->
+<!--                   chains = 4, -->
+<!--                   parallel_chains = 4, -->
+<!--                   refresh = 100) -->
+
+<!-- ## this fits at least now... -->
+
+<!-- fit1_simple$summary() -->
+<!-- mcmc_hist(fit1_simple$draws(c("alpha", "beta"))) -->
+
+<!-- ``` -->
+
+<!-- ## Notes on these models -->
+
+<!-- - How does the choice of focal player make a difference in these models? -->
+
+<!-- ```{r thinking about the choice of focal} -->
+
+<!-- table(bullet_60$Username) -->
+
+<!-- tidy_bullet <- bullet_60 %>% -->
+<!--   select(White:BlackElo, Username, game_id) %>% -->
+<!--   mutate(opponent = ifelse(Username == White, Black, White), -->
+<!--          focal_white = ifelse(Username == White, 1, 0)) -->
+
+<!-- all_players <- c(tidy_bullet$Username, tidy_bullet$opponent) -->
+<!-- players <- unique(all_players) -->
+
+<!-- ## then just get all the history for this -->
+<!-- all_hist <- map_dfr(players, get_hist_opp, bullet_60, prev_n = 5) %>% -->
+<!--   as_tibble() %>% -->
+<!--   select(game_id:player, UTCDate, UTCTime) -->
+
+<!-- min_hist <- all_hist %>% select(game_id:player) -->
+
+<!-- ## this should have each game appear twice, for each of the two players -->
+<!-- ## involved, so need to figure out how to handle that best -->
+<!-- ## left join with tidy games matching player to focal and then -->
+<!-- ## to opponent, rename the columns as needed -->
+
+<!-- tidy_stan <- tidy_bullet %>% -->
+<!--   left_join(min_hist, by = c("game_id", "Username" = "player")) %>% -->
+<!--   arrange(UTCDate, UTCTime) %>% -->
+<!--   rename(focal_result = opp_result, focal_win_prop = opp_win_prop) %>% -->
+<!--   left_join(min_hist, by = c("game_id", "opponent" = "player")) %>% -->
+<!--   as_tibble() -->
+
+<!-- ## this is all good, just need to lag both and add them in, -->
+<!-- ## and also get the ids for the stan code -->
+
+<!-- ## getting the overall current win probabilities for each player also, -->
+<!-- ## up to and including the current game -->
+
+<!-- player_avg <- all_hist %>% -->
+<!--   group_by(player) %>% -->
+<!--   arrange(UTCDate, UTCTime, .by_group = TRUE) %>% -->
+<!--   mutate(nrow = row_number()) %>% -->
+<!--   mutate(curr_avg = cumsum(opp_result)/nrow) %>% -->
+<!--   select(game_id, player, curr_avg) -->
+
+<!-- global_avg <- all_hist %>% -->
+<!--   group_by(player) %>% -->
+<!--   summarise(total_avg = mean(opp_result)) -->
+
+<!-- ## then just join this in also to use it -->
+
+<!-- ``` -->
+
+<!-- ```{r getting the final data for potential stan model} -->
+
+<!-- final_data <- tidy_stan %>% -->
+<!--   left_join(player_avg, by = c("Username" = "player", "game_id")) %>% -->
+<!--   rename(focal_avg = curr_avg) %>% -->
+<!--   left_join(player_avg, by = c("opponent" = "player", "game_id")) %>% -->
+<!--   rename(opp_avg = curr_avg) %>% -->
+<!--   group_by(Username) %>% -->
+<!--   arrange(UTCDate, UTCTime) %>% -->
+<!--   mutate(focal_prev_result = lag(focal_result, default = 0), -->
+<!--          focal_runn_avg = lag(focal_avg, n = 2, default = 0)) %>% -->
+<!--   ungroup() %>% -->
+<!--   group_by(opponent) %>% -->
+<!--   arrange(UTCDate, UTCTime) %>% -->
+<!--   mutate(opp_prev_result = lag(opp_result, default = 0), -->
+<!--          opp_runn_avg = lag(opp_avg, n = 2, default = 0)) %>% -->
+<!--   ungroup() %>% -->
+<!--   mutate(focal_id = match(Username, players), -->
+<!--          opp_id = match(opponent, players)) %>% -->
+<!--   filter(focal_result != 0.5) %>% -->
+<!--   select(-focal_avg, - opp_avg, - opp_result, -->
+<!--          - focal_win_prop, - opp_win_prop) -->
+
+<!-- final_data -->
+<!-- ## remove some unused columns -->
+
+<!-- norm_data <- final_data %>% -->
+<!--   left_join(global_avg, by = c("Username" = "player")) %>% -->
+<!--   rename(focal_global_avg = total_avg) %>% -->
+<!--   left_join(global_avg, by = c("opponent" = "player")) %>% -->
+<!--   rename(opp_global_avg = total_avg) %>% -->
+<!--   mutate(focal_prev_result = focal_prev_result - focal_global_avg, -->
+<!--          focal_runn_avg = focal_runn_avg - focal_global_avg, -->
+<!--          opp_prev_result = opp_prev_result - opp_global_avg, -->
+<!--          opp_runn_avg = opp_runn_avg - opp_global_avg) -->
+
+<!-- ``` -->
+
+<!-- Now this matches nicely, contains their historic winning average up to 2 games -->
+<!-- previously, along with their result in the previous game. -->
+
+<!-- Note that this has not been standardized currently. -->
+
+<!-- ```{r} -->
+
+<!-- norm_data -->
+
+<!-- stan_data_all <- list(N = nrow(norm_data), -->
+<!--                       J = length(players), -->
+<!--                       y = norm_data$focal_result, -->
+<!--                       focal_id = norm_data$focal_id, -->
+<!--                       opp_id = norm_data$opp_id, -->
+<!--                       focal_prev = norm_data$focal_prev_result, -->
+<!--                       opp_prev = norm_data$opp_prev_result, -->
+<!--                       focal_avg = norm_data$focal_runn_avg, -->
+<!--                       opp_avg = norm_data$opp_runn_avg) -->
+
+<!-- saveRDS(stan_data_all, file = "../cluster_scripts/stan_data.RDS") -->
+
+<!-- stan_file <- "../Comp2.stan" -->
+
+<!-- mod2 <- cmdstan_model(stan_file) -->
+
+<!-- fit2 <- mod2$sample(data = stan_data_all, -->
+<!--                     seed = 123, -->
+<!--                     chains = 4, -->
+<!--                     parallel_chains = 4, -->
+<!--                     refresh = 100, -->
+<!--                     iter_sampling = 500, -->
+<!--                     iter_warmup = 100) -->
+
+<!-- ``` -->
+
+<!-- ## Fitting this to all data -->
+
+<!-- Fit this model to all data on the cluster, just to see the results. -->
+<!-- The traceplots and R-hat for this model indicates it doesn't work well, -->
+<!-- does not converge. -->
+
+<!-- ```{r} -->
+<!-- stan_fit <- readRDS("../cluster_scripts/Cluster_stan.RDS") -->
+
+<!-- mcmc_hist(stan_fit$draws(c("alpha[1]", "alpha[2]"))) -->
+
+<!-- r_hat <- rhat(stan_fit) -->
+
+<!-- summary(r_hat) -->
+
+<!-- which(r_hat > 1.15) -->
+<!-- ## just problems with mu and tau for convergence here, otherwise good? -->
+
+<!-- ``` -->
+
+<!-- ## Comparing models on smaller data -->
+
+<!-- Is there a small subset of players who play each other often that we can -->
+<!-- use in the dataset that can be used to compare these models? -->
+
+<!-- ```{r} -->
+
+<!-- bullet_60 %>% -->
+<!--   filter(White %in% focal_players) %>% -->
+<!--   filter(Black %in% focal_players) -->
+
+<!-- ## no games between any of the players used to scrape the data -->
+<!-- ## would maybe need to get more players then, to use this for a comparison -->
+
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- library(chessR) -->
+
+<!-- chessdotcom_leaders <- chessdotcom_leaderboard(game_type = "live_bullet") -->
+
+<!-- top_bullet <- chessdotcom_leaders %>% top_n(10, wt = -rank) %>% -->
+<!--   select(player_id, username) -->
+
+<!-- # chessdotcom_hikaru_recent <- get_raw_chessdotcom(usernames = "Hikaru", -->
+<!-- #                                                  year_month = c(202201:202205)) -->
+
+<!-- games_22 <- top_bullet$username %>% -->
+<!--   map_dfr(~get_raw_chessdotcom(.x, year_month = c(202201:202203))) -->
+
+<!-- top_games <- games_22 %>% -->
+<!--   filter(White %in% top_bullet$username) %>% -->
+<!--   filter(Black %in% top_bullet$username) -->
+
+<!-- ## get 500 games here then for 1 month, probably enough to fit an initial model -->
+<!-- ## most of these games bullet too, so could use that -->
+
+<!-- top_games %>% -->
+<!--   group_by(White, Black, TimeClass) %>% -->
+<!--   tally() %>% -->
+<!--   arrange(-n) -->
+
+<!-- bullet_games <- top_games %>% -->
+<!--   filter(TimeClass == "bullet") -->
+
+<!-- bullet_top10 <- saveRDS(bullet_games, file = "chess_com_bullet_10.RDS") -->
+
+<!-- ``` -->

### 2023-05-27 owen/cluster_scripts/Fit_All.R
run new model

@@ -16,7 +16,7 @@ library(posterior)
 stan_data_all <- readRDS(here("owen", "cluster_scripts", "stan_data_ppt_n10.RDS"))
 
 
-stan_file <- here("owen","Comp2.stan")
+stan_file <- here("owen","Comp3.stan")
 
 mod2 <- cmdstan_model(stan_file)
 

### 2023-05-27 owen/Write Ups/03_existing_models.qmd
finish off the write up

@@
 ```{r plot some parts of this model}
+#| fig-height: 8
 
@@
 
+fit_ppt <- readRDS(here("owen", "cluster_scripts", "Cluster_stan_ppt_2_n10.RDS"))
+
+focal <- paste0("beta[", c(1, 127, 1207, 1263, 1519, 1635, 
+                           1789, 2084, 2139, 2441, 2503, 2561, 
+                           2679, 2707, 2781, 2836, 3533, 3714), "]")
+
+mcmc_hist(fit_ppt$draws(focal))
+
+focal_gamma <- paste0("gamma[", c(1, 127, 1207, 1263, 1519, 1635, 
+                           1789, 2084, 2139, 2441, 2503, 2561, 
+                           2679, 2707, 2781, 2836, 3533, 3714), "]")
+
+mcmc_hist(fit_ppt$draws(focal_gamma))
 ```
 
-Comparing the predictive performance of the model from 
-the slides (**without covariates**) it is unsurprisingly 
-not comparable to the current iteration of the model we have here.
+![Predictive Checks](../cluster_scripts/ppd_model2.png)
+
+This model gives similar predictive performance to the previously proposed
+model.
 

### 2023-05-29 owen/Write Ups/03_existing_models.qmd
add comments to make self contained

@@
 
+```{r process data for this model}
+#| eval: false
+
+lichess_data <- readRDS("../../rdata/lichess_pilot.RData")
+
+lichess_data %>% 
+  filter(Event == "Rated Bullet game") %>% 
+  group_by(TimeControl) %>% 
+  count() %>% 
+  arrange(-n)
+
+bullet_60 <- lichess_data %>% 
+  filter(Event == "Rated Bullet game") %>% 
+  filter(TimeControl == "60+0")
+
+get_hist <- function(user, games, prev_n) {
+  hist_games <- games %>% 
+    filter(White == user | Black == user) %>% 
+    arrange(UTCDate, UTCTime) %>% 
+    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
+    select(White:BlackElo, focal_white) %>% 
+    mutate(focal_result = case_when(
+      (focal_white == 1 & Result == "1-0") ~ 1,
+      (focal_white == 0 & Result == "0-1") ~ 1,
+      (Result == "1/2-1/2") ~ 0.5,
+      .default = 0
+      )) %>% 
+    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
+                              roll_mean(focal_result, n = prev_n)))
+
+  hist_games
+}
+
+focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18)
+
+top_players <- names(focal_players)
+
+tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 10) %>% 
+  as_tibble()
+
+init_data <- tidy_games %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, top_players)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% 
+  mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_win_prop)) %>% 
+  filter(focal_result != 0.5)
+
+```
+
+```{r fit stan model}
+#| eval: false
+
+## takes approx 3 minutes on m1 pro mac
+
+stan_data_ave <- list(N = nrow(init_data),
+                  J = length(top_players),
+                  y = init_data$focal_result,
+                  id = init_data$focal_id,
+                  colour = init_data$focal_white,
+                  elo = init_data$elo_diff,
+                  win_prop = init_data$ave_prop)
+
+stan_file <- "../model3.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit3_ave <- mod$sample(data = stan_data_ave,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit3_ave$save_object("../model_fits/init_model_n10.RDS")
+```
+
 ```{r load in initial model here}
@@
 so can take on values $0,0.5,1$.
+
+```{r data for model 2}
+#| eval: false
+
+hist_data_init <- tidy_games %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, top_players)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% 
+  mutate(ave_prop = lag(focal_win_prop, n = 2, default = 0) - 
+           mean(focal_win_prop),
+         prev_game = lag(focal_result, default = 0)) %>% 
+  filter(focal_result != 0.5)
+
+stan_data_hist <- list(N = nrow(hist_data_init),
+                  J = length(top_players),
+                  y = hist_data_init$focal_result,
+                  id = hist_data_init$focal_id,
+                  colour = hist_data_init$focal_white,
+                  elo = hist_data_init$elo_diff,
+                  win_prop = hist_data_init$ave_prop,
+                  prev_game = hist_data_init$prev_game)
+
+```
+
+```{r fit the stan model}
+#| eval: false
+
+## takes approx 4 minutes on m1 pro mac
+
+stan_file <- "../model4.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit4 <- mod$sample(data = stan_data_hist,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+
+fit4$summary()
+
+fit4$save_object(file = "../model_fits/model4.RDS")
+```
 
@@
 
+### this takes several hours, needs to be run on a cluster
+
 stan_file <- "../Comp2.stan"
@@
 This model gives similar predictive performance to the previously proposed
-model.
+model. However, there are some minor concerns with model fitting 
+which should be addressed (convergence issues).
 

### 2023-06-06 owen/Write Ups/03_existing_models.qmd
update the stan models to constrain scale params, even though forced to already by model
tidy up the current write up to include the fixes to the model, explore the relationship between posterior mean estimates and overall elo for players

@@
 ```{r process data for this model}
-#| eval: false
 
@@
   filter(Event == "Rated Bullet game") %>% 
-  filter(TimeControl == "60+0")
+  filter(TimeControl == "60+0") %>% 
+  mutate(game_id = row_number())
+
+```
+
+```{r}
+#| eval: false
 
@@
 
+stan_data_all <- readRDS("../cluster_scripts/stan_data_ppt_n10.RDS")
 ### this takes several hours, needs to be run on a cluster
 
-stan_file <- "../Comp2.stan"
+stan_file <- "../Comp3.stan"
 
@@
 ```{r plot some parts of this model}
-#| fig-height: 8
+#| fig-height: 5
+#| fig-cap: Player effects for the 18 players considered by the previous model.
 
@@
 
-fit_ppt <- readRDS(here("owen", "cluster_scripts", "Cluster_stan_ppt_2_n10.RDS"))
+fit_ppt <- readRDS(here("owen", "cluster_scripts",
+                        "Cluster_stan_ppt_2_n10_June5.RDS"))
+
+focal <- paste0("alpha[", c(1, 127, 1207, 1263, 1519, 1635, 
+                           1789, 2084, 2139, 2441, 2503, 2561, 
+                           2679, 2707, 2781, 2836, 3533, 3714), "]")
+
+mcmc_hist(fit_ppt$draws(focal))
+
+## most of these alphas centered around 0, 1/2 slightly positive and
+## 1/2 slightly negative. these player effects
+
+```
+
+```{r}
+#| fig-height: 5
+#| fig-cap: Longer run (10 game) winner effects for those 18 players.
 
@@
 
+## some of these betas not 0 really, this is more interesting maybe
+## beta corresponds to the long run effect (prev 10 games)
+
+```
+
+```{r}
+#| fig-height: 5
+#| fig-cap: Effect of deviation in previous game from previous 10.
+
 focal_gamma <- paste0("gamma[", c(1, 127, 1207, 1263, 1519, 1635, 
@@
 mcmc_hist(fit_ppt$draws(focal_gamma))
+
+## some of these slightly positive or negative, but not much
+## this corresponds to previous game deviation from prev 10 games
+
 ```
 
-![Predictive Checks](../cluster_scripts/ppd_model2.png)
+```{r}
+#| fig-height: 5
+#| fig-cap: The global parameters for this model.
+
+mcmc_hist(fit_ppt$draws(c("mu1", "mu2", "mu3",
+                          "tau1", "tau2", "tau3")))
+
+```
+
+![Predictive Checks](../cluster_scripts/ppd_model2_June5.png)
 
 This model gives similar predictive performance to the previously proposed
-model. However, there are some minor concerns with model fitting 
-which should be addressed (convergence issues).
+model.
 
@@
 
------
+- How do these winner effects vary with the rating of players,
+and how their rating changes over time?
 
-<!-- <!-- ## Model Comparison -->
+```{r}
 
-<!-- We want to compare these potential models, which is very similar -->
-<!-- to the model we fit previously. -->
-<!-- The only change is that we now need the history for -->
-<!-- both players, not just the focal player. -->
+focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18)
+## if don't use all focal players here then removing data because they won't be
+## involved
 
-<!-- ```{r load in the initial data} -->
-<!-- #| cache: true -->
+top_players <- names(focal_players)
 
-<!-- lichess_data <- readRDS("../../rdata/lichess_pilot.RData") -->
+kept_games <- bullet_60 %>% 
+  filter(Username %in% top_players) %>% 
+  select(White:BlackElo, Username, game_id) %>%
+  mutate(opponent = ifelse(Username == White, Black, White),
+         focal_white = ifelse(Username == White, 1, 0))
 
-<!-- bullet_60 <- lichess_data %>% -->
-<!--   filter(Event == "Rated Bullet game") %>% -->
-<!--   filter(TimeControl == "60+0") %>% -->
-<!--   mutate(game_id = row_number()) -->
+curr_players <- unique(c(kept_games$White, kept_games$Black))
 
-<!-- get_hist <- function(user, games, prev_n) { -->
-<!--   hist_games <- games %>% -->
-<!--     filter(White == user | Black == user) %>% -->
-<!--     arrange(UTCDate, UTCTime) %>% -->
-<!--     mutate(focal_white = ifelse(Username == White, 1, 0)) %>% -->
-<!--     select(White:BlackElo, focal_white, game_id) %>% -->
-<!--     mutate(focal_result = case_when( -->
-<!--       (focal_white == 1 & Result == "1-0") ~ 1, -->
-<!--       (focal_white == 0 & Result == "0-1") ~ 1, -->
-<!--       (Result == "1/2-1/2") ~ 0.5, -->
-<!--       .default = 0 -->
-<!--       )) %>% -->
-<!--     mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), -->
-<!--                               roll_mean(focal_result, n = prev_n))) -->
-<!--   hist_games -->
-<!-- } -->
+## note that rhat(fit_ppt) should work, just too big here
 
-<!-- focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18) -->
-<!-- ## if don't use all focal players here then removing data because they won't be -->
-<!-- ## involved -->
+white <- bullet_60 %>% 
+  select(game_id, White, WhiteElo) %>% 
+  rename(Player = White, ELO = WhiteElo)
 
-<!-- top_players <- names(focal_players) -->
+black <- bullet_60 %>% 
+  select(game_id, Black, BlackElo) %>% 
+  rename(Player = Black, ELO = BlackElo)
 
-<!-- ## then get the opponents of these players -->
+game_elo <- bind_rows(white, black)
 
-<!-- full_opponents <- bullet_60 %>% -->
-<!--   filter(Username %in% top_players) %>% -->
-<!--   select(White, Black) -->
+avg_ELO <- game_elo %>% 
+  mutate(ELO = as.numeric(ELO)) %>% 
+  group_by(Player) %>% 
+  summarise(mean_rating = mean(ELO))
 
-<!-- all_opps <- c(full_opponents$White, full_opponents$Black) -->
+## need to order these correctly, based on the indexing used for stan
 
-<!-- opponents <- unique(all_opps) -->
+player_info <- tibble(Player = curr_players) %>% 
+  mutate(id = row_number())
 
-<!-- ## then need to get the history of these players, but maybe need -->
-<!-- ## to store game id -->
-<!-- ## can just match using date and time I guess? -->
-<!-- ## need to modify the function then to account for this -->
+avg_ELO <- avg_ELO %>% 
+  left_join(player_info, by = "Player") %>% 
+  drop_na()
 
-<!-- get_hist_opp <- function(user, games, prev_n) { -->
-<!--   hist_games <- games %>% -->
-<!--     filter(White == user | Black == user) %>% -->
-<!--     arrange(UTCDate, UTCTime) %>% -->
-<!--     mutate(opp_white = ifelse(user == White, 1, 0)) %>% -->
-<!--     select(White:BlackElo, opp_white, game_id) %>% -->
-<!--     mutate(opp_result = case_when( -->
-<!--       (opp_white == 1 & Result == "1-0") ~ 1, -->
-<!--       (opp_white == 0 & Result == "0-1") ~ 1, -->
-<!--       (Result == "1/2-1/2") ~ 0.5, -->
-<!--       .default = 0 -->
-<!--       )) -->
+## then get the posterior means for each of the parameters for this.
 
-<!--   ngames <- nrow(hist_games) -->
+alpha_draws <- as_draws_df(fit_ppt$draws(c("alpha")))
+beta_draws <- as_draws_df(fit_ppt$draws(c("beta")))
+gamma_draws <- as_draws_df(fit_ppt$draws(c("gamma")))
 
-<!--   if(ngames < prev_n) { -->
-<!--     hist_games <- hist_games %>% -->
-<!--     mutate(opp_win_prop = opp_result/(1:ngames)) -->
-<!--   } -->
-<!--   else{ -->
-<!--     hist_games <- hist_games %>% -->
-<!--       mutate(opp_win_prop = c(cumsum(opp_result[1:(prev_n - 1)])/(1:(prev_n -1)), -->
-<!--                               roll_mean(opp_result, n = prev_n))) -->
-<!--   } -->
-<!--   hist_games$player <- user -->
-<!--   hist_games -->
-<!-- } -->
-<!-- ## so now the history is the history corresponding to the name -->
-<!-- ## in the "player" column -->
+alpha_mean <- apply(alpha_draws, 2, mean)
+beta_mean <- apply(beta_draws, 2, mean)
+gamma_mean <- apply(gamma_draws, 2, mean)
+## then get the id's from this, based on the names
+names(alpha_mean)
 
-<!-- remaining_opponents <- opponents[!(opponents %in% top_players)] -->
+coef_info <- tibble(mean_alpha = alpha_mean, 
+                    mean_beta = beta_mean,
+                    mean_gamma = gamma_mean) %>% 
+  mutate(id = row_number())
 
-<!-- opp_hist <- map_dfr(remaining_opponents, get_hist_opp, bullet_60, prev_n = 5) %>% -->
-<!--   as_tibble() -->
+avg_ELO %>% 
+  left_join(coef_info, by = "id") %>% 
+  ggplot(aes(mean_alpha, mean_rating)) +
+  geom_point() + 
+  labs(x = "Posterior Mean for Alpha", y = "Mean ELO")
 
-<!-- ## so this gives the history of the opponents for each of their -->
-<!-- ## games, can use this to then first get their previous result, etc, -->
-<!-- ## then join this with the other results -->
+avg_ELO %>% 
+  left_join(coef_info, by = "id") %>% 
+  ggplot(aes(mean_beta, mean_rating)) +
+  geom_point() + 
+  labs(x = "Posterior Mean for Beta", y = "Mean ELO")
 
-<!-- prev_opp <- opp_hist %>% -->
-<!--   group_by(player) %>% -->
-<!--   arrange(UTCDate, UTCTime) %>% -->
-<!--   mutate(prev_opp_result = lag(opp_result, default = 0), -->
-<!--          prev_opp_prop = lag(opp_win_prop, default = 0)) %>% -->
-<!--   select(game_id, player, prev_opp_result, prev_opp_prop) -->
+avg_ELO %>% 
+  left_join(coef_info, by = "id") %>% 
+  ggplot(aes(mean_gamma, mean_rating)) +
+  geom_point() + 
+  labs(x = "Posterior Mean for Gamma", y = "Mean ELO")
 
-<!-- tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 5) %>% -->
-<!--   as_tibble() -->
+top_id <- c(1, 127, 1207, 1263, 1519, 1635, 
+                           1789, 2084, 2139, 2441, 2503, 2561, 
+                           2679, 2707, 2781, 2836, 3533, 3714)
 
-<!-- all <- tidy_games %>% -->
-<!--   left_join(prev_opp, by = "game_id") %>% -->
-<!--   mutate(focal = ifelse(focal_white == 1, White, Black)) -->
-<!-- ## this looks more reasonable now -->
-<!-- ## this has the history of the opponent included, not the history -->
-<!-- ## of the focal player yet -->
+avg_ELO %>% 
+  left_join(coef_info, by = "id") %>% 
+  mutate(top = ifelse(id %in% top_id, "TOP", "REST")) %>% 
+  ggplot(aes(mean_alpha, mean_rating, colour = top)) +
+  geom_point() + 
+  labs(x = "Posterior Mean for Alpha", y = "Mean ELO")
 
-<!-- final_data <- all %>% -->
-<!--   group_by(focal) %>% -->
-<!--   arrange(UTCDate, UTCTime) %>% -->
-<!--   mutate(prev_focal_result = lag(focal_result, default = 0), -->
-<!--          prev_focal_prop = lag(focal_win_prop, default = 0)) %>% -->
-<!--   ungroup() -->
+avg_ELO %>% 
+  left_join(coef_info, by = "id") %>% 
+  mutate(top = ifelse(id %in% top_id, "TOP", "REST")) %>% 
+  ggplot(aes(mean_beta, mean_rating, colour = top)) +
+  geom_point() + 
+  labs(x = "Posterior Mean for Beta", y = "Mean ELO")
 
-<!-- ## then tidy this up for stan -->
+avg_ELO %>% 
+  left_join(coef_info, by = "id") %>% 
+  mutate(top = ifelse(id %in% top_id, "TOP", "REST")) %>% 
+  ggplot(aes(mean_gamma, mean_rating, colour = top)) +
+  geom_point() + 
+  labs(x = "Posterior Mean for Gamma", y = "Mean ELO")
 
-<!-- ``` -->
+## this just gives you the players who play few games all lumped together 
+## basically, which makes sense
 
-<!-- This object `prev_game_both` should have the previous result of both players in -->
-<!-- each game, if that game is in the dataset. -->
-<!-- Similarly, `final_prev` has the games played by the focal players, along -->
-<!-- with their previous result and their opponents previous result, if that -->
-<!-- game was in the original dataset used (all 60 second bullet games here). -->
-
-<!-- Then convert this into data for stan. -->
-
-<!-- ```{r get stan data for simple model} -->
-
-<!-- final_data -->
-
-<!-- stan_prev <- final_data %>% -->
-<!--   mutate(WhiteElo = as.numeric(WhiteElo), -->
-<!--          BlackElo = as.numeric(BlackElo)) %>% -->
-<!--   mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% -->
-<!--   mutate(elo_diff = ifelse(focal_white == 1, -->
-<!--                            WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% -->
-<!--   mutate(focal_id = match(focal_user, top_players)) %>% -->
-<!--   select(focal_user, focal_id, focal_white, -->
-<!--          prev_focal_result, prev_opp_result, focal_result) -->
-
-<!-- stan_data <- list(N = nrow(stan_prev), -->
-<!--                   J = length(top_players), -->
-<!--                   y = stan_prev$focal_result, -->
-<!--                   id = stan_prev$focal_id, -->
-<!--                   focal_prev = stan_prev$prev_focal_result, -->
-<!--                   opp_prev = stan_prev$prev_opp_result) -->
-
-<!-- ``` -->
-
-<!-- ```{r fit the simple stan model} -->
-<!-- #| eval: false -->
-
-<!-- ## remember no effect for non focal players in this at all -->
-<!-- ## at the moment -->
-
-<!-- stan_file <- "../Comp1.stan" -->
-
-<!-- mod <- cmdstan_model(stan_file) -->
-
-<!-- fit1_simple <- mod$sample(data = stan_data, -->
-<!--                   seed = 123, -->
-<!--                   chains = 4, -->
-<!--                   parallel_chains = 4, -->
-<!--                   refresh = 100) -->
-
-<!-- ## this fits at least now... -->
-
-<!-- fit1_simple$summary() -->
-<!-- mcmc_hist(fit1_simple$draws(c("alpha", "beta"))) -->
-
-<!-- ``` -->
-
-<!-- ## Notes on these models -->
-
-<!-- - How does the choice of focal player make a difference in these models? -->
-
-<!-- ```{r thinking about the choice of focal} -->
-
-<!-- table(bullet_60$Username) -->
-
-<!-- tidy_bullet <- bullet_60 %>% -->
-<!--   select(White:BlackElo, Username, game_id) %>% -->
-<!--   mutate(opponent = ifelse(Username == White, Black, White), -->
-<!--          focal_white = ifelse(Username == White, 1, 0)) -->
-
-<!-- all_players <- c(tidy_bullet$Username, tidy_bullet$opponent) -->
-<!-- players <- unique(all_players) -->
-
-<!-- ## then just get all the history for this -->
-<!-- all_hist <- map_dfr(players, get_hist_opp, bullet_60, prev_n = 5) %>% -->
-<!--   as_tibble() %>% -->
-<!--   select(game_id:player, UTCDate, UTCTime) -->
-
-<!-- min_hist <- all_hist %>% select(game_id:player) -->
-
-<!-- ## this should have each game appear twice, for each of the two players -->
-<!-- ## involved, so need to figure out how to handle that best -->
-<!-- ## left join with tidy games matching player to focal and then -->
-<!-- ## to opponent, rename the columns as needed -->
-
-<!-- tidy_stan <- tidy_bullet %>% -->
-<!--   left_join(min_hist, by = c("game_id", "Username" = "player")) %>% -->
-<!--   arrange(UTCDate, UTCTime) %>% -->
-<!--   rename(focal_result = opp_result, focal_win_prop = opp_win_prop) %>% -->
-<!--   left_join(min_hist, by = c("game_id", "opponent" = "player")) %>% -->
-<!--   as_tibble() -->
-
-<!-- ## this is all good, just need to lag both and add them in, -->
-<!-- ## and also get the ids for the stan code -->
-
-<!-- ## getting the overall current win probabilities for each player also, -->
-<!-- ## up to and including the current game -->
-
-<!-- player_avg <- all_hist %>% -->
-<!--   group_by(player) %>% -->
-<!--   arrange(UTCDate, UTCTime, .by_group = TRUE) %>% -->
-<!--   mutate(nrow = row_number()) %>% -->
-<!--   mutate(curr_avg = cumsum(opp_result)/nrow) %>% -->
-<!--   select(game_id, player, curr_avg) -->
-
-<!-- global_avg <- all_hist %>% -->
-<!--   group_by(player) %>% -->
-<!--   summarise(total_avg = mean(opp_result)) -->
-
-<!-- ## then just join this in also to use it -->
-
-<!-- ``` -->
-
-<!-- ```{r getting the final data for potential stan model} -->
-
-<!-- final_data <- tidy_stan %>% -->
-<!--   left_join(player_avg, by = c("Username" = "player", "game_id")) %>% -->
-<!--   rename(focal_avg = curr_avg) %>% -->
-<!--   left_join(player_avg, by = c("opponent" = "player", "game_id")) %>% -->
-<!--   rename(opp_avg = curr_avg) %>% -->
-<!--   group_by(Username) %>% -->
-<!--   arrange(UTCDate, UTCTime) %>% -->
-<!--   mutate(focal_prev_result = lag(focal_result, default = 0), -->
-<!--          focal_runn_avg = lag(focal_avg, n = 2, default = 0)) %>% -->
-<!--   ungroup() %>% -->
-<!--   group_by(opponent) %>% -->
-<!--   arrange(UTCDate, UTCTime) %>% -->
-<!--   mutate(opp_prev_result = lag(opp_result, default = 0), -->
-<!--          opp_runn_avg = lag(opp_avg, n = 2, default = 0)) %>% -->
-<!--   ungroup() %>% -->
-<!--   mutate(focal_id = match(Username, players), -->
-<!--          opp_id = match(opponent, players)) %>% -->
-<!--   filter(focal_result != 0.5) %>% -->
-<!--   select(-focal_avg, - opp_avg, - opp_result, -->
-<!--          - focal_win_prop, - opp_win_prop) -->
-
-<!-- final_data -->
-<!-- ## remove some unused columns -->
-
-<!-- norm_data <- final_data %>% -->
-<!--   left_join(global_avg, by = c("Username" = "player")) %>% -->
-<!--   rename(focal_global_avg = total_avg) %>% -->
-<!--   left_join(global_avg, by = c("opponent" = "player")) %>% -->
-<!--   rename(opp_global_avg = total_avg) %>% -->
-<!--   mutate(focal_prev_result = focal_prev_result - focal_global_avg, -->
-<!--          focal_runn_avg = focal_runn_avg - focal_global_avg, -->
-<!--          opp_prev_result = opp_prev_result - opp_global_avg, -->
-<!--          opp_runn_avg = opp_runn_avg - opp_global_avg) -->
-
-<!-- ``` -->
-
-<!-- Now this matches nicely, contains their historic winning average up to 2 games -->
-<!-- previously, along with their result in the previous game. -->
-
-<!-- Note that this has not been standardized currently. -->
-
-<!-- ```{r} -->
-
-<!-- norm_data -->
-
-<!-- stan_data_all <- list(N = nrow(norm_data), -->
-<!--                       J = length(players), -->
-<!--                       y = norm_data$focal_result, -->
-<!--                       focal_id = norm_data$focal_id, -->
-<!--                       opp_id = norm_data$opp_id, -->
-<!--                       focal_prev = norm_data$focal_prev_result, -->
-<!--                       opp_prev = norm_data$opp_prev_result, -->
-<!--                       focal_avg = norm_data$focal_runn_avg, -->
-<!--                       opp_avg = norm_data$opp_runn_avg) -->
-
-<!-- saveRDS(stan_data_all, file = "../cluster_scripts/stan_data.RDS") -->
-
-<!-- stan_file <- "../Comp2.stan" -->
-
-<!-- mod2 <- cmdstan_model(stan_file) -->
-
-<!-- fit2 <- mod2$sample(data = stan_data_all, -->
-<!--                     seed = 123, -->
-<!--                     chains = 4, -->
-<!--                     parallel_chains = 4, -->
-<!--                     refresh = 100, -->
-<!--                     iter_sampling = 500, -->
-<!--                     iter_warmup = 100) -->
-
-<!-- ``` -->
-
-<!-- ## Fitting this to all data -->
-
-<!-- Fit this model to all data on the cluster, just to see the results. -->
-<!-- The traceplots and R-hat for this model indicates it doesn't work well, -->
-<!-- does not converge. -->
-
-<!-- ```{r} -->
-<!-- stan_fit <- readRDS("../cluster_scripts/Cluster_stan.RDS") -->
-
-<!-- mcmc_hist(stan_fit$draws(c("alpha[1]", "alpha[2]"))) -->
-
-<!-- r_hat <- rhat(stan_fit) -->
-
-<!-- summary(r_hat) -->
-
-<!-- which(r_hat > 1.15) -->
-<!-- ## just problems with mu and tau for convergence here, otherwise good? -->
-
-<!-- ``` -->
-
-<!-- ## Comparing models on smaller data -->
-
-<!-- Is there a small subset of players who play each other often that we can -->
-<!-- use in the dataset that can be used to compare these models? -->
-
-<!-- ```{r} -->
-
-<!-- bullet_60 %>% -->
-<!--   filter(White %in% focal_players) %>% -->
-<!--   filter(Black %in% focal_players) -->
-
-<!-- ## no games between any of the players used to scrape the data -->
-<!-- ## would maybe need to get more players then, to use this for a comparison -->
-
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- library(chessR) -->
-
-<!-- chessdotcom_leaders <- chessdotcom_leaderboard(game_type = "live_bullet") -->
-
-<!-- top_bullet <- chessdotcom_leaders %>% top_n(10, wt = -rank) %>% -->
-<!--   select(player_id, username) -->
-
-<!-- # chessdotcom_hikaru_recent <- get_raw_chessdotcom(usernames = "Hikaru", -->
-<!-- #                                                  year_month = c(202201:202205)) -->
-
-<!-- games_22 <- top_bullet$username %>% -->
-<!--   map_dfr(~get_raw_chessdotcom(.x, year_month = c(202201:202203))) -->
-
-<!-- top_games <- games_22 %>% -->
-<!--   filter(White %in% top_bullet$username) %>% -->
-<!--   filter(Black %in% top_bullet$username) -->
-
-<!-- ## get 500 games here then for 1 month, probably enough to fit an initial model -->
-<!-- ## most of these games bullet too, so could use that -->
-
-<!-- top_games %>% -->
-<!--   group_by(White, Black, TimeClass) %>% -->
-<!--   tally() %>% -->
-<!--   arrange(-n) -->
-
-<!-- bullet_games <- top_games %>% -->
-<!--   filter(TimeClass == "bullet") -->
-
-<!-- bullet_top10 <- saveRDS(bullet_games, file = "chess_com_bullet_10.RDS") -->
-
-<!-- ``` -->
+```

### 2023-07-21 owen/Write Ups/03_existing_models.qmd
quick update to instead of fitting to every game, construct the windows, only take the final game in each window, use that rather than all games. plotted the parameter estimates and did posterior predictive checking. seems to work reasonably well still

@@
 
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 8)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8)
 
@@
 ## then get the id's from this, based on the names
-names(alpha_mean)
+# names(alpha_mean)
 
@@
 ```
+
+## Updates July
+
+Only consider a single game in each window?
+
+```{r}
+
+get_hist_mod <- function(user, games, prev_n) {
+  hist_games <- games %>% 
+    filter(White == user | Black == user) %>% 
+    arrange(UTCDate, UTCTime) %>% 
+    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
+    select(White:BlackElo, focal_white) %>% 
+    mutate(focal_result = case_when(
+      (focal_white == 1 & Result == "1-0") ~ 1,
+      (focal_white == 0 & Result == "0-1") ~ 1,
+      (Result == "1/2-1/2") ~ 0.5,
+      .default = 0
+      )) %>% 
+    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
+                              roll_mean(focal_result, n = prev_n)),
+           window_id = row_number()/prev_n)
+
+  hist_games %>% filter(window_id - round(window_id) == 0)
+}
+
+focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18)
+
+top_players <- names(focal_players)
+
+tidy_games_mod <- map_dfr(top_players, get_hist_mod, bullet_60, prev_n = 10) %>% 
+  as_tibble()
+
+## so this takes windows of size 10, so keep 10th game, 
+## use history from the previous 9 games each time
+
+tidy_games_mod
+
+init_data_mod <- tidy_games_mod %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, top_players)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% 
+  mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_win_prop)) %>% 
+  filter(focal_result != 0.5)
+
+```
+
+Then quickly fit model to this data.
+
+```{r}
+stan_data_mod <- list(N = nrow(init_data_mod),
+                  J = length(top_players),
+                  y = init_data_mod$focal_result,
+                  id = init_data_mod$focal_id,
+                  colour = init_data_mod$focal_white,
+                  elo = init_data_mod$elo_diff,
+                  win_prop = init_data_mod$ave_prop)
+
+stan_file <- "../model3.stan"
+
+mod <- cmdstan_model(stan_file)
+
+fit3_mod <- mod$sample(data = stan_data_mod,
+                  seed = 123,
+                  chains = 4,
+                  parallel_chains = 4,
+                  refresh = 100)
+```
+
+```{r}
+fit3_mod$summary()
+
+mcmc_hist(fit3_mod$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
+```
+
+Get the posterior predictive checks.
+
+```{r}
+fit_samples <- as_draws_df(fit3_mod$draws())
+
+fit_samp <- fit_samples %>% 
+  select(!starts_with(c("log_lik", "yrep")))
+
+y_rep <- fit_samples %>% select(starts_with("y_rep"))
+
+rm(fit_samples)
+rm(fit_samp)
+
+### then want to join this with the correct input, i.e player, etc
+
+y_rep_mod <- y_rep %>% 
+  rownames_to_column() %>%  
+  pivot_longer(-rowname) %>% 
+  pivot_wider(names_from=rowname, values_from=value) %>% 
+  ## remove the draws and chains here
+  mutate(focal_id = stan_data_mod$id)
+
+games_won <- y_rep_mod %>% 
+  pivot_longer(cols = `1`:`4000`, names_to = "draw", values_to = "y") %>% 
+  group_by(focal_id, draw) %>% 
+  summarise(games_won = sum(y))
+
+orig_data <- tibble(outcome = stan_data_mod$y,
+                    focal_id = stan_data_mod$id)
+
+orig_games_won <- orig_data %>% 
+  group_by(focal_id) %>% 
+  summarise(games_won = sum(outcome))
+
+games_won %>% 
+  ggplot(aes(x = games_won)) +
+  geom_histogram(bins = 20) +
+  facet_wrap(~focal_id, scales = "free") +
+  geom_vline(data = orig_games_won, 
+             mapping = aes(xintercept = games_won), col = "red") +
+  labs(title = "Final Outcome in Window Only")
+
+```

### 2023-09-05 owen/Write Ups/03_existing_models.qmd
add quick version for fitting model to permuted data

@@
 ```
+
+### Robustness to Correlation
+
+A valid concern here is the large correlation between the data used at
+each step. For example, when using the previous 10 games as the history,
+subsequent games will use an average with 9 of the 10 terms in the average being
+identical. This leads to severe correlation.
+
+One potential way to deal with this (mentioned by the authors of the Deng et al 
+paper) is to test for this effect. One way to do this would be to consider
+a permutation test. The results can be randomly permuted, which allows the
+construction of an empirical set of estimates under the permutation.
+This would need to be considered for this data also.

### 2023-09-05 owen/cluster_scripts/perm_results.R

@@ -1,0 +1,105 @@
+#### September 5th 2023
+
+## Considering a permutation test for evaluating the robustness of
+## the results to serial correlations in the data
+
+
+## We consider this first for the subset of data consisting of 18 focal players
+
+library(cmdstanr)
+
+options(mc.cores = parallel::detectCores())
+
+library(tidyverse)
+library(here)
+library(loo)
+library(ggplot2)
+library(posterior)
+
+
+jobid <- Sys.getenv("SLURM_ARRAY_TASK_ID")
+jobid <- as.numeric(jobid)
+sim_id <- jobid
+
+lichess_data <- readRDS("rdata/lichess_pilot.RData")
+
+lichess_data %>% 
+  filter(Event == "Rated Bullet game") %>% 
+  group_by(TimeControl) %>% 
+  count() %>% 
+  arrange(-n)
+
+
+bullet_60 <- lichess_data %>% 
+  filter(Event == "Rated Bullet game") %>% 
+  filter(TimeControl == "60+0") %>% 
+  mutate(game_id = row_number()) 
+
+
+
+focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18)
+
+
+top_players <- names(focal_players)
+
+
+## take only the games involving focal players, confirm this
+## this identical to original data
+## then permute the results
+
+focal_games <- bullet_60 %>% 
+  filter(White %in% top_players | Black %in% top_players) %>% 
+  mutate(Result = sample(Result))
+
+
+
+tidy_games <- map_dfr(top_players, get_hist, focal_games, prev_n = 10) %>% 
+  as_tibble()
+
+
+init_data <- tidy_games %>% 
+  mutate(WhiteElo = as.numeric(WhiteElo), 
+         BlackElo = as.numeric(BlackElo)) %>% 
+  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
+  mutate(elo_diff = ifelse(focal_white == 1, 
+                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
+  mutate(focal_id = match(focal_user, top_players)) %>% 
+  select(focal_user, focal_id, focal_white, 
+         focal_win_prop, elo_diff, focal_result) %>% 
+  group_by(focal_id) %>% 
+  mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_win_prop)) %>% 
+  filter(focal_result != 0.5)
+
+
+
[…]

### 2023-09-05 owen/cluster_scripts/perm_submit.sh

@@ -1,0 +1,20 @@
+#! /bin/bash
+#
+#SBATCH --mem-per-cpu 8000
+#SBATCH -c 8
+#SBATCH -t 300:00
+#SBATCH -a 1-10
+#SBATCH --mail-user=oward@sfu.ca
+##SBATCH --mail-type=ALL
+
+module load gcc r/4.2.2
+
+echo "Launching R"
+date
+
+Rscript perm_results.R
+
+echo "Completed"
+date
+
+# end of script

### 2023-09-05 owen/cluster_scripts/perm_results.R
something with the file paths

@@ -21,7 +21,7 @@ jobid <- Sys.getenv("SLURM_ARRAY_TASK_ID")
 jobid <- as.numeric(jobid)
 sim_id <- jobid
 
-lichess_data <- readRDS("rdata/lichess_pilot.RData")
+lichess_data <- readRDS(here("rdata", "lichess_pilot.RData"))
 
 lichess_data %>% 
   filter(Event == "Rated Bullet game") %>% 

### 2023-09-05 owen/cluster_scripts/perm_results.R
add missing function

@@ -51,7 +51,23 @@ focal_games <- bullet_60 %>%
   filter(White %in% top_players | Black %in% top_players) %>% 
   mutate(Result = sample(Result))
 
-
+get_hist <- function(user, games, prev_n) {
+  hist_games <- games %>% 
+    filter(White == user | Black == user) %>% 
+    arrange(UTCDate, UTCTime) %>% 
+    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
+    select(White:BlackElo, focal_white) %>% 
+    mutate(focal_result = case_when(
+      (focal_white == 1 & Result == "1-0") ~ 1,
+      (focal_white == 0 & Result == "0-1") ~ 1,
+      (Result == "1/2-1/2") ~ 0.5,
+      .default = 0
+    )) %>% 
+    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
+                              roll_mean(focal_result, n = prev_n)))
+  
+  hist_games
+}
 
 tidy_games <- map_dfr(top_players, get_hist, focal_games, prev_n = 10) %>% 
   as_tibble()

### 2023-09-05 owen/cluster_scripts/perm_results.R
add missing lib call

@@ -15,7 +15,7 @@ library(here)
 library(loo)
 library(ggplot2)
 library(posterior)
-
+library(RcppRoll)
 
 jobid <- Sys.getenv("SLURM_ARRAY_TASK_ID")
 jobid <- as.numeric(jobid)

### 2023-09-05 owen/cluster_scripts/perm_results.R
more stuff

@@ -97,11 +97,11 @@ stan_data_ave <- list(N = nrow(init_data),
                       win_prop = init_data$ave_prop)
 
 
-stan_file <- "owen/model3.stan"
+stan_file <- here("owen", "model3.stan")
 
 mod <- cmdstan_model(stan_file)
 
-fit3_ave <- mod$sample(data = stan_data_ave,
+fit3 <- mod$sample(data = stan_data_ave,
                        seed = 123,
                        chains = 4,
                        parallel_chains = 4,
@@ -112,7 +112,7 @@ fit3_ave <- mod$sample(data = stan_data_ave,
 ## to the simulation number
 
 
-par_ests <- fit3_ave$summary(c("alpha", "beta", "gamma1", "gamma2",
+par_ests <- fit3$summary(c("alpha", "beta", "gamma1", "gamma2",
                    "mu1", "mu2", "tau1", "tau2")) %>% 
   mutate(sim_id = sim_id)
 

### 2023-09-07 owen/Write Ups/03_existing_models.qmd
checking some stuff about the permutations

@@
 
-```{r}
+```{r, message=FALSE}
 stan_data_mod <- list(N = nrow(init_data_mod),
@@
 ```{r}
-fit3_mod$summary()
+# fit3_mod$summary()
 
@@
 construction of an empirical set of estimates under the permutation.
-This would need to be considered for this data also.
+
+This can be implemented for the simple initial model, considering
+only games involving the 18 most frequent players. We randomly permute the 
+results from the corresponding matches and refit the above model 100 times.
+
+```{r}
+perm_path <- here("owen", "cluster_scripts", "perm_outputs", "perm_outputs")
+
+sims <- list.files(path = perm_path)
+
+perm_results <- sims %>% 
+  map_dfr(~readRDS(here("owen", "cluster_scripts",
+                        "perm_outputs", "perm_outputs",
+                        .x))) %>% 
+  select(variable, mean, median, sim_id)
+
+## global parameters here
+
+fit3_ave <- readRDS("../model_fits/init_model_n10.RDS")
+
+orig <- fit3_ave$summary(c("alpha", "beta", "gamma1", "gamma2", "mu1", "mu2",
+                   "tau1", "tau2"))
+
+rm(fit3_ave)
+
+perm_results %>% 
+  filter(variable == "mu1") %>% 
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  labs(title = "Permutations of Mu_1",
+       subtitle = "True data appears different",
+       x = "mu1") +
+  geom_vline(data = orig %>% 
+               filter(variable == "mu1"),
+             aes(xintercept = mean), col = "red")
+
+perm_results %>% 
+  filter(variable == "mu2") %>% 
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  labs(title = "Permutations of Mu_2",
+       subtitle = "True data appears different",
+       x = "mu2") +
+  geom_vline(data = orig %>% 
+               filter(variable == "mu2"),
+             aes(xintercept = mean), col = "red")
+
+## try it also for some of the winner effects
+
+perm_results %>% 
+  filter(variable == "gamma1") %>% 
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  labs(title = "Permutations of gamma_1",
+       subtitle = "True data appears different",
+       x = "gamma_1") +
+  geom_vline(data = orig %>% 
+               filter(variable == "gamma1"),
+             aes(xintercept = mean), col = "red")
+
+perm_results %>% 
+  filter(variable == "gamma2") %>% 
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  labs(title = "Permutations of gamma_2",
+       subtitle = "True data appears different",
+       x = "gamma_2") +
+  geom_vline(data = orig %>% 
+               filter(variable == "gamma2"),
+             aes(xintercept = mean), col = "red")
+
+perm_results %>% 
+  filter(variable == "tau1") %>% 
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  labs(title = "Permutations of tau_1",
+       subtitle = "True data appears different",
+       x = "tau1") +
+  geom_vline(data = orig %>% 
+               filter(variable == "tau1"),
+             aes(xintercept = mean), col = "red")
+
+perm_results %>% 
+  filter(variable == "tau2") %>% 
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  labs(title = "Permutations of tau_2",
+       subtitle = "True data appears different",
+       x = "tau_2") +
+  geom_vline(data = orig %>% 
+             filter(variable == "tau2"),
+             aes(xintercept = mean), col = "red")
+
+```
+
+This quick permutations seem to indicate that the is evidence for these 
+effects, which is not present when the results are randomly permuted 
+instead.
+
+```{r repeat for all alpha}
+
+alpha_var <- paste0("alpha[", 1:18, "]")
+
+# alpha_var
+
+for(i in seq_along(alpha_var)){
+  plot_curr <- perm_results %>% 
+  filter(variable == alpha_var[i]) %>% 
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  labs(title = paste0("Permutations of ", alpha_var[i]),
+       subtitle = "True data appears different",
+       x = alpha_var[i]) +
+  geom_vline(data = orig %>% 
+             filter(variable == alpha_var[i]),
+             aes(xintercept = mean), col = "red")
+  print(plot_curr)
+}
+
+```
+
+```{r repeat for all beta}
+
+beta_var <- paste0("beta[", 1:18, "]")
+
+# alpha_var
+
+for(i in seq_along(beta_var)){
+  plot_curr <- perm_results %>% 
+  filter(variable == beta_var[i]) %>% 
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  labs(title = paste0("Permutations of ", beta_var[i]),
+       subtitle = "True data appears different",
+       x = alpha_var[i]) +
+  geom_vline(data = orig %>% 
+             filter(variable == beta_var[i]),
+             aes(xintercept = mean), col = "red")
+  print(plot_curr)
+}
+
+```
+
+Are the estimates that are not centered around 0 explained by the 
+overall proportion of wins in the dataset?
+
+```{r}
+
+init_data
+
+mean(init_data$focal_result)
+
+## not that different to 1/(1 + exp(-mu_2)), 
+## i.e just the proportion of games won by the focal player
+
+## is it something similar for gamma1, larger number of games as white
+## by the focal players?
+
+mean(init_data$focal_white)
+
+init_data %>% 
+  group_by(focal_white, focal_result) %>% 
+  tally() %>% 
+  mutate(prop = n /sum(n))
+
+init_data %>% 
+  group_by(focal_result) %>% 
+  tally() %>% 
+  mutate(prop = n /sum(n))
+
+```

### 2023-10-05 owen/Write Ups/03_existing_models.qmd
adding indiv level permutations also

@@
 ```{r}
-#| eval: false
+#| eval: true
 
@@
 For the chess example where many players are involved, this leads to thousands
-of opponents, many who plan only a small number of games.
+of opponents, many who play only a small number of games.
 
@@
 Only consider a single game in each window?
+This would be done by only using the final game in each subsequent
+window of size $n$, reducing the number of observations in 
+the data by a factor of $n$ essentially. Initially it was
+thought that this was being considered in the Deng et al. paper,
+but that was incorrect.
 
@@
 
-tidy_games_mod
+# tidy_games_mod
 
@@
 
-```{r, message=FALSE}
+```{r, message=FALSE, }
 stan_data_mod <- list(N = nrow(init_data_mod),
@@
                   chains = 4,
-                  parallel_chains = 4,
-                  refresh = 100)
+                  parallel_chains = 4)
 ```
@@
 
-This quick permutations seem to indicate that the is evidence for these 
+This quick permutation seem to indicate that there is evidence for these 
 effects, which is not present when the results are randomly permuted 
@@
        subtitle = "True data appears different",
-       x = alpha_var[i]) +
+       x = beta_var[i]) +
   geom_vline(data = orig %>% 
@@
 ```
+
+#### Checking Permuting within Players only
+
+```{r read the data in}
+perm_path <- here("owen", "cluster_scripts",
+                  "perm_outputs", "perm_outputs_indiv")
+
+sims <- list.files(path = perm_path)
+
+perm_results <- sims %>% 
+  map_dfr(~readRDS(here("owen", "cluster_scripts",
+                        "perm_outputs", "perm_outputs_indiv",
+                        .x))) %>% 
+  select(variable, mean, median, sim_id)
+
+```
+
+```{r plot global pars}
+
+## have orig from above
+
+## mean winner effect
+
+perm_results %>% 
+  filter(variable == "mu1") %>% 
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  labs(title = "Permutations of Mu_1",
+       subtitle = "True data appears different",
+       x = "mu1") +
+  geom_vline(data = orig %>% 
+               filter(variable == "mu1"),
+             aes(xintercept = mean), col = "red")
+
+perm_results %>% 
+  filter(variable == "mu2") %>% 
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  labs(title = "Permutations of Mu_2",
+       subtitle = "This is the mean of the player effects",
+       x = "mu2") +
+  geom_vline(data = orig %>% 
+               filter(variable == "mu2"),
+             aes(xintercept = mean), col = "red")
+
+## effect of playing as white
+
+perm_results %>% 
+  filter(variable == "gamma1") %>% 
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  labs(title = "Permutations of gamma_1",
+       subtitle = "True data appears similar",
+       x = "gamma_1") +
+  geom_vline(data = orig %>% 
+               filter(variable == "gamma1"),
+             aes(xintercept = mean), col = "red")
+
+## effect of elo difference
+
+perm_results %>% 
+  filter(variable == "gamma2") %>% 
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  labs(title = "Permutations of gamma_2",
+       subtitle = "True data appears different",
+       x = "gamma_2") +
+  geom_vline(data = orig %>% 
+               filter(variable == "gamma2"),
+             aes(xintercept = mean), col = "red")
+
+```
+
+```{r check alpha first}
+
+alpha_var <- paste0("alpha[", 1:18, "]")
+
+# alpha_var
+
+for(i in seq_along(alpha_var)){
+  plot_curr <- perm_results %>% 
+  filter(variable == alpha_var[i]) %>% 
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  labs(title = paste0("Permutations of ", alpha_var[i]),
+       subtitle = "True data appears different",
+       x = alpha_var[i]) +
+  geom_vline(data = orig %>% 
+             filter(variable == alpha_var[i]),
+             aes(xintercept = mean), col = "red")
+  print(plot_curr)
+}
+
+## some of these still significant here, need to check it
+
+```
+
+```{r check beta also}
+
+beta_var <- paste0("beta[", 1:18, "]")
+
+# alpha_var
+
+for(i in seq_along(beta_var)){
+  plot_curr <- perm_results %>% 
+  filter(variable == beta_var[i]) %>% 
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  labs(title = paste0("Permutations of ", beta_var[i]),
+       subtitle = "True data appears different",
+       x = beta_var[i]) +
+  geom_vline(data = orig %>% 
+             filter(variable == beta_var[i]),
+             aes(xintercept = mean), col = "red")
+  print(plot_curr)
+}
+
+```

### 2023-10-05 owen/cluster_scripts/perm_results.R

@@ -98,7 +98,7 @@ init_data <- tidy_games %>%
 
 ## to check the win proportions
 
-# init_data %>% 
+# init_data %>%
 #   group_by(focal_white, focal_result) %>%
 #   tally() %>%
 #   mutate(prop = n/sum(n))

### 2023-11-27 owen/cluster_scripts/perm_submit.sh
small changes

@@ -3,7 +3,7 @@
 #SBATCH --mem-per-cpu 8000
 #SBATCH -c 8
 #SBATCH -t 300:00
-#SBATCH -a 1-10
+#SBATCH -a 6-100
 #SBATCH --mail-user=oward@sfu.ca
 ##SBATCH --mail-type=ALL
 

### 2023-11-27 owen/cluster_scripts/submit.sh

@@ -1,8 +1,8 @@
 #! /bin/bash
 #
-#SBATCH --mem-per-cpu 8000
+#SBATCH --mem-per-cpu 12000
 #SBATCH -c 8
-#SBATCH -t 300:00
+#SBATCH -t 600:00
 #SBATCH --mail-user=oward@sfu.ca
 #SBATCH --mail-type=ALL
 
@@ -11,7 +11,7 @@ module load gcc r/4.2.2
 echo "Launching R"
 date
 
-Rscript Fit_All.R
+Rscript Large_Fit.R
 
 echo "Completed"
 date

### 2023-11-27 owen/cluster_scripts/Large_Fit.R
update to use all rated bullet games

@@ -79,8 +79,8 @@ small_data <- lichess_data %>%
   # filter(Event == "Rated Bullet game") %>% 
   # filter(TimeControl == "60+0") %>% 
   filter(Variant == "Standard") %>% 
-  filter(grepl("Rated", Event)) %>% 
-  filter(!grepl("Classical|Correspondence", Event))
+  filter(grepl("Rated Bullet game", Event)) #%>% 
+  # filter(!grepl("Classical|Correspondence", Event))
 
 users <- unique(small_data$Username)
 
@@ -128,6 +128,11 @@ fit3_ave <- mod$sample(data = stan_data_ave,
                        refresh = 100)
 
 
+## save the stan fit as not actually that large here
+
+fit3_ave$save_object(file = here(save_path, "all_rated_bullet.RDS"))
+
+
 ## create some summary plots of these results
 
 players <- users
@@ -138,14 +143,14 @@ player_labels <- as_labeller(players)
 mcmc_hist(fit3_ave$draws(c("mu2", "mu1", "tau2", "tau1", "gamma1", "gamma2")),
           facet_args = list(scales = "free"))
 
-ggsave(filename = paste0(save_path, "/global_pars.png"),
+ggsave(filename = paste0(save_path, "/global_pars_all_rated_bullet.png"),
                          width = 8, height = 8, units = "in")
 
 
 mcmc_hist(fit3_ave$draws("beta"),
           facet_args = list(labeller = player_labels)) 
 
-ggsave(filename = paste0(save_path, "/winner_pars.png"),
+ggsave(filename = paste0(save_path, "/winner_pars_all_rated_bullet.png"),
        width = 8, height = 8, units = "in")
 
 names(players) <- paste0("alpha[", 1:length(users), "]")
@@ -153,6 +158,6 @@ player_labels <- as_labeller(players)
 mcmc_hist(fit3_ave$draws("alpha"),
           facet_args = list(labeller = player_labels))
 
-ggsave(filename = paste0(save_path, "/indiv_pars.png"),
+ggsave(filename = paste0(save_path, "/indiv_pars_all_rated_bullet.png"),
        width = 8, height = 8, units = "in")
 

### 2023-11-28 owen/Write Ups/03_existing_models.qmd
rename this file to avoid any confusion
some small changes
rerun some of this analysis, with the simpler model here

@@
 ---
-title: "Model Checking"
+title: "Existing Models"
 author: Owen G. Ward
@@
 
-fit3_ave$save_object("../model_fits/init_model_n10.RDS")
+# fit3_ave$save_object("../model_fits/init_model_n10.RDS")
 ```
@@
 
-fit3_ave <- readRDS("../model_fits/init_model_n10.RDS")
+# fit3_ave <- readRDS("../model_fits/init_model_n10.RDS")
 
@@
 
-mcmc_hist(fit3_ave$draws(c("alpha", "mu2", "tau2")))
-mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
+mcmc_hist(fit3_ave$draws(c("alpha", "mu2")))
+mcmc_hist(fit3_ave$draws(c("beta", "mu1", "gamma1", "gamma2")))
 ```
@@
 Playing white makes a player who has 50% chance of winning as black
-have approximately a 54% chance of winning with white. Wikipedia says the increase 
+have approximately a 54% chance of winning with white. Wikipedia says
+the increase 
 in win probability from playing white has been estimated to be 
@@
 
-fit4$save_object(file = "../model_fits/model4.RDS")
+# fit4$save_object(file = "../model_fits/model4.RDS")
 ```
@@
 
-fit4 <- readRDS("../model_fits/model4.RDS")
-
-mcmc_hist(fit4$draws(c("delta", "mu3", "tau3"))) +
+# fit4 <- readRDS("../model_fits/model4.RDS")
+
+mcmc_hist(fit4$draws(c("delta", "mu3"))) +
   labs(title = "Effect of Previous Game")
 
-mcmc_hist(fit4$draws(c("beta", "mu1", "tau1"))) +
+mcmc_hist(fit4$draws(c("beta", "mu1"))) +
   labs(title = "Effect of Previous n games, excluding most recent game")

### 2023-11-28 owen/cluster_scripts/Large_Fit.R
fitting a smaller sample locally for this data

@@ -28,7 +28,7 @@ save_path <- here("results/lichess1700-1900/")
 
 
 files <- list.files(data_path)
-files <- files[1:10]
+# files <- files[1:3]
 
 ## need to write a function to process them separately then I think
 

### 2023-11-29 owen/cluster_scripts/Large_Fit.R
local changes

@@ -28,7 +28,7 @@ save_path <- here("results/lichess1700-1900/")
 
 
 files <- list.files(data_path)
-files <- files[1:10]
+files <- files[1:15]
 
 ## need to write a function to process them separately then I think
 

### 2023-12-07 owen/cluster_scripts/Large_Fit.R
add another folder

@@ -22,14 +22,16 @@ theme_set(theme_bw())
 ### load in the data to use
 
 ## rerun for 2000-2200 next
-# data_path <- here("box_data/lichess1700-1900/")
-data_path <- here("box_data/lichess2000-2200/")
-# save_path <- here("results/lichess1700-1900/")
-save_path <- here("results/lichess2000-2200/")
+##data_path <- here("box_data/lichess1700-1900/")
+##data_path <- here("box_data/lichess2000-2200/")
+data_path <- here("box_data/lichess2300-2500/")
+##save_path <- here("results/lichess1700-1900/")
+##save_path <- here("results/lichess2000-2200/")
+save_path <- here("results/lichess2300-2500/")
 dir.create(save_path, showWarnings = FALSE)
 
 files <- list.files(data_path)
-files <- files[1:25]
+files <- files[1:8]
 
 ## need to write a function to process them separately then I think
 
@@ -141,7 +143,7 @@ names(players) <- paste0("beta[", 1:length(users), "]")
 
 player_labels <- as_labeller(players)
 
-mcmc_hist(fit3_ave$draws(c("mu2", "mu1", "tau2", "tau1", "gamma1", "gamma2")),
+mcmc_hist(fit3_ave$draws(c("mu2", "mu1",  "gamma1", "gamma2")),
           facet_args = list(scales = "free"))
 
 ggsave(filename = paste0(save_path, "/global_pars_all_rated_bullet_model.png"),

### 2023-12-08 owen/Write Ups/04_model_checking.qmd
update, add this file to github

---
title: "Model Checking"
author: Owen G. Ward
date: today
format: 
  # html:
  #   toc: true
  #   toc-location: left
  pdf
bibliography: refs.bib
---

The aim of this document is to summarise the current version 
of the model and describe some of the ways we have assessed the
suitability of this model to our data. The
aim is for this document to be self contained.

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 8)

library(tidyverse)
library(RcppRoll)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(loo)
library(here)

options(mc.cores = parallel::detectCores())

theme_set(theme_bw())

prob_positive <- function(stan_draws, param = "beta") {
  all_draws <- as_draws_df(stan_draws)
  ndraws <- nrow(all_draws)
  as_tibble(all_draws) %>% 
  select(starts_with(param)) %>% 
  apply(2, function(x) sum(x>0)/ndraws)
}

```

## The current model

Based on @ding2021nhl, an initial model proposed was of the form

$$
P(y_{ij} = 1) = \frac{1}{1 + \exp(-(\alpha_j + \beta_j x_{ij} + \gamma z_{ij}))}
$$

where this is the probability of the **j-th focal player** 
winning their **i-th game**. Here:

- $\alpha_j$ is a player level random effect
- $\beta_j$ is a player level random effect, accounting for the win ratio $x_{ij}$
of the focal player over their previous $n$ games
- $\gamma$ is a fixed effect of game level coefficients (colour, difference
in ELO ranking)

We then partially pool the $\alpha$ and $\beta$ coefficients, so

$$
\alpha_j \sim \mathcal{N}(\mu_2, 1)
$$
$$
\mu_2 \sim \mathcal{N}(0, 1)
$$
$$
\tau_2 \sim \mathcal{N}(0, 1)
$$

and similarly

$$
\beta_j \sim \mathcal{N}(\mu_1, 1)
$$
$$
\mu_1 \sim \mathcal{N}(0, 1)
$$
$$
\tau_1 \sim \mathcal{N}(0, 1)
$$

We rescale the win ratio $x_{ij}$, by the average win ratio of
player $j$ across all 
games they play in the dataset.
This means $\beta_j$ captures how a players win
probability changes as their historic performance deviates from 
their overall average win ratio.

For $z_{ij}$ we incorporate 2 covariates, namely the colour of 
the focal player and the ELO difference between the focal 
player and their opponent. So that means we have 
$\gamma_1$, which tells us how the win probability changes in going
from black to white,
and $\gamma_2$ gives how the win probability changes as the focal player
has a better ELO than their opponent.

__These predictors are not currently rescaled__.

## Fitting this to Lichess Data

We fit this to some of the large lichess data. We consider only a subset 
of the players in a given rating range, due to time constraints.

### 1700-1900 Players

We fit this to games from 25 players in the 1700-1900 range. 
We looked only at standard rated bullet games (no tournaments, etc),
which was 445k games, ranging from 500 - 50k games per player.
Fitting this model take over 12 hours on a computing cluster.

This results in the following estimated parameters.

```{r}
fit <- readRDS(here("results/", "lichess1700-1900/", 
                    "all_rated_bullet_model3.RDS"))

mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
  labs(title = "Global Parameters",
       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

```

- Here the mean for $\mu_2$ is -0.0651, corresponding to a 48% chance of 
winning as black against an equally ranked opponent. This parameter
is not clearly different from 0 here. Posterior SD is large.

- $\gamma_1$ is estimated to be about 0.115, indicating that playing 
as white increases the win probability by 3%, holding other parameters 
fixed. Posterior SD is very small.

- $\gamma_2$ indicates that a 10 point ELO difference increases the win 
probability by about 1%, while a 100 ELO score difference 
increases the win probability by about 9%. Posterior SD is very small.

- The global estimate for $\mu_1$, the global winner effect, is 
around 0. Posterior SD is large here.

```{r get names 1700-1900}
#| cache: true

## need to get user names here, from the data again

data_path <- here("box_data/lichess1700-1900/")

files <- list.files(data_path)
files <- files[1:25]

read_player <- function(path, file){
  dat <- read_csv(file = paste0(data_path, file),
                  col_types = cols(UTCDate = col_date("%Y.%m.%d"),
                                   WhiteTitle = col_character(),
                                   BlackTitle = col_character(),
                                   WhiteElo = col_character(),
                                   BlackElo = col_character(),
                                   FEN = col_character())) %>% 
    select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
           WhiteElo, BlackElo, Variant, TimeControl, Termination) %>% 
    mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
           BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo)))
  dat
}

get_hist <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
    )) %>% 
    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(focal_result, n = prev_n)))

  hist_games
}

lichess_data <- files %>% 
  map_dfr(~read_player(data_path, .x))

## restrict to rated rapid and shorter here
## this also removes the NAs, which makes sense

small_data <- lichess_data %>% 
  # filter(Event == "Rated Bullet game") %>% 
  # filter(TimeControl == "60+0") %>% 
  filter(Variant == "Standard") %>% 
  filter(grepl("Rated Bullet game", Event)) #%>% 
  # filter(!grepl("Classical|Correspondence", Event))

users <- unique(small_data$Username)

rm(lichess_data)
rm(small_data)

```

```{r winner-effects-17-19}
#| fig.height = 7

players <- users
names(players) <- paste0("beta[", 1:length(users), "]")

player_labels <- as_labeller(players)

mcmc_hist(fit$draws("beta"),
          facet_args = list(labeller = player_labels)) +
  labs(title = "Player Winner Effects")
```

Here we see that there is a range of plausible winner effects for 
different players.

- Some have posterior distributions which are only positive, some 
are only negative and some are centered around or near 0.

-The largest posterior mean here is 1.32. This is saying that if on
average you win 50% of your games, but have instead won 10 in a row, your
probability of winning the next would increase by ~16% over that baseline.
However most estimated effects are small, and most of these posterior
distributions contain 0.

```{r indiv-effects-17-19}
#| fig.height = 6

names(players) <- paste0("alpha[", 1:length(users), "]")
player_labels <- as_labeller(players)
mcmc_hist(fit$draws("alpha"),
          facet_args = list(labeller = player_labels)) +
  labs(title = "Individual Player Effects")
```

- Most of the individual player effects are somewhat negative but
close to 0.

- Many are quite similar to the global parameter here.

### 2000-2200 Players

We fit this model also to 10 players rated between 2000 and 2200. 
With the same restrictions on the games this gives us 285k games.
We have between 1200 and 89k games for each of these 10 players.

```{r}

fit2 <- readRDS(here("results/", "lichess2000-2200/", 
                    "all_rated_bullet_model.RDS"))

mcmc_hist(fit2$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
  labs(title = "Global Parameters",
       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

```

```{r get-player-names}
#| cache: true

data_path <- here("box_data/lichess2000-2200/")

files <- list.files(data_path)
files <- files[1:10]

read_player <- function(path, file){
  dat <- read_csv(file = paste0(data_path, file),
                  col_types = cols(UTCDate = col_date("%Y.%m.%d"),
                                   WhiteTitle = col_character(),
                                   BlackTitle = col_character(),
                                   WhiteElo = col_character(),
                                   BlackElo = col_character(),
                                   FEN = col_character())) %>% 
    select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
           WhiteElo, BlackElo, Variant, TimeControl, Termination) %>% 
    mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
           BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo)))
  dat
}

get_hist <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
    )) %>% 
    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(focal_result, n = prev_n)))

  hist_games
}

lichess_data <- files %>% 
  map_dfr(~read_player(data_path, .x))

## restrict to rated rapid and shorter here
## this also removes the NAs, which makes sense

small_data <- lichess_data %>% 
  # filter(Event == "Rated Bullet game") %>% 
  # filter(TimeControl == "60+0") %>% 
  filter(Variant == "Standard") %>% 
  filter(grepl("Rated Bullet game", Event)) #%>% 
  # filter(!grepl("Classical|Correspondence", Event))

users_20_22 <- unique(small_data$Username)

rm(lichess_data)
rm(small_data)

```

```{r}

players <- users_20_22
names(players) <- paste0("beta[", 1:length(users_20_22), "]")

player_labels <- as_labeller(players)

mcmc_hist(fit2$draws("beta"),
          facet_args = list(labeller = player_labels)) +
  labs(title = "Player Winner Effects")

```

```{r}
names(players) <- paste0("alpha[", 1:length(users_20_22), "]")
player_labels <- as_labeller(players)
mcmc_hist(fit2$draws("alpha"),
          facet_args = list(labeller = player_labels)) +
  labs(title = "Individual Player Effects")
```

- Here the effect of a difference in ranking appears to be slightly larger.
This leads to a bigger change in the win probability when the difference
in ranking between the two players is larger.

- The posterior standard deviations of the $\mu$ parameters is slightly
larger in this model, but this could be explained by the smaller number of 
games.

- A smaller sample, but half of these posterior distributions 
of the winner effects contain 0.

### 2300-2500 Players

We look at 8 players rated between 2300 and 2500. This gives us 160k games,
with players playing 4k-58k games.

```{r}
fit3 <- readRDS(here("results/", "lichess2300-2500/", 
                    "all_rated_bullet_model.RDS"))

mcmc_hist(fit2$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
  labs(title = "Global Parameters",
       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")
```

```{r get users 2300-2500}
#| cache: true

data_path <- here("box_data/lichess2300-2500/")

files <- list.files(data_path)
files <- files[1:8]

lichess_data <- files %>% 
  map_dfr(~read_player(data_path, .x))

## restrict to rated rapid and shorter here
## this also removes the NAs, which makes sense

small_data <- lichess_data %>% 
  # filter(Event == "Rated Bullet game") %>% 
  # filter(TimeControl == "60+0") %>% 
  filter(Variant == "Standard") %>% 
  filter(grepl("Rated Bullet game", Event)) #%>% 
  # filter(!grepl("Classical|Correspondence", Event))

users_23_25 <- unique(small_data$Username)

rm(lichess_data)
rm(small_data)

```

```{r}

players <- users_23_25
names(players) <- paste0("beta[", 1:length(users_23_25), "]")

player_labels <- as_labeller(players)

mcmc_hist(fit3$draws("beta"),
          facet_args = list(labeller = player_labels)) +
  labs(title = "Player Winner Effects")

```

```{r}
names(players) <- paste0("alpha[", 1:length(users_23_25), "]")
player_labels <- as_labeller(players)
mcmc_hist(fit3$draws("alpha"),
          facet_args = list(labeller = player_labels)) +
  labs(title = "Individual Player Effects")
```

## Other Questions

Do focal players play each other in this data?

```{r}
data_path <- here("box_data/lichess1700-1900/")

files <- list.files(data_path)
files <- files[1:25]

read_player <- function(path, file){
  dat <- read_csv(file = paste0(data_path, file),
                  col_types = cols(UTCDate = col_date("%Y.%m.%d"),
                                   WhiteTitle = col_character(),
                                   BlackTitle = col_character(),
                                   WhiteElo = col_character(),
                                   BlackElo = col_character(),
                                   FEN = col_character())) %>% 
    select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
           WhiteElo, BlackElo, Variant, TimeControl, Termination) %>% 
    mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
           BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo)))
  dat
}

lichess_data <- files %>% 
  map_dfr(~read_player(data_path, .x))

users <- unique(lichess_data$Username)

overlap_data <- lichess_data %>% 
  filter(White %in% users & Black %in% users) %>% 
  filter(Variant == "Standard") %>% 
  filter(grepl("Rated Bullet game", Event))

rm(lichess_data)

overlap_data %>% group_by(Username) %>% 
  tally() %>% 
  arrange(-n)
```

Certain players have a large amount of overlap, looking only at the games we
consider in the analysis. How should these games be accounted for?

```{r check for 2000-2200 games also}
#| eval: false

## happens here also, but not as often for this sample

data_path <- here("box_data/lichess2000-2200/")

files <- list.files(data_path)
files <- files[1:10]

read_player <- function(path, file){
  dat <- read_csv(file = paste0(data_path, file),
                  col_types = cols(UTCDate = col_date("%Y.%m.%d"),
                                   WhiteTitle = col_character(),
                                   BlackTitle = col_character(),
                                   WhiteElo = col_character(),
                                   BlackElo = col_character(),
                                   FEN = col_character())) %>% 
    select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
           WhiteElo, BlackElo, Variant, TimeControl, Termination) %>% 
    mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
           BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo)))
  dat
}

lichess_data <- files %>% 
  map_dfr(~read_player(data_path, .x))

users <- unique(lichess_data$Username)

overlap_data <- lichess_data %>% 
  filter(White %in% users & Black %in% users) %>% 
  filter(Variant == "Standard") %>% 
  filter(grepl("Rated Bullet game", Event))

rm(lichess_data)

overlap_data %>% group_by(Username) %>% 
  tally() %>% 
  arrange(-n)

```

### Other Comments

- Model checking is hard for these large datasets

- Not easy to do the permutation fits for all games

- Hard to simulate fake data also

- Are we ignoring any covariates which could influence the model here?

- How do these effects change if we only look at games which were played 
directly before?

<!-- Below some ideas about posterior predictive checking, hard to currently  -->
<!-- consider for the larger datasets we're working with. -->

<!-- ## Posterior Predictive Checking -->

<!-- ## Permuting the Original Data -->

<!-- We can also use permutations of the data to examine how -->
<!-- robust these results are. -->

<!-- ### Global Permutation -->

<!-- The simplest permutation of our model would be to simply permute -->
<!-- all results at random and then process the data and refit this model. -->

<!-- ```{r} -->
<!-- perm_path <- here("owen", "cluster_scripts", "perm_outputs", "perm_outputs") -->

<!-- sims <- list.files(path = perm_path) -->

<!-- perm_results <- sims %>%  -->
<!--   map_dfr(~readRDS(here("owen", "cluster_scripts", -->
<!--                         "perm_outputs", "perm_outputs", -->
<!--                         .x))) %>%  -->
<!--   select(variable, mean, median, sim_id) -->

<!-- ## global parameters here -->

<!-- fit3_ave <- readRDS("../model_fits/init_model_n10.RDS") -->

<!-- orig <- fit3_ave$summary(c("alpha", "beta", "gamma1", "gamma2", "mu1", "mu2", -->
<!--                    "tau1", "tau2"))  -->

<!-- rm(fit3_ave)  -->

<!-- perm_results %>%  -->
<!--   filter(variable == "mu1") %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = "Permutations of Mu_1", -->
<!--        subtitle = "True data appears different", -->
<!--        x = "mu1") + -->
<!--   geom_vline(data = orig %>%  -->
<!--                filter(variable == "mu1"), -->
<!--              aes(xintercept = mean), col = "red") -->

<!-- perm_results %>%  -->
<!--   filter(variable == "mu2") %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = "Permutations of Mu_2", -->
<!--        subtitle = "True data appears different", -->
<!--        x = "mu2") + -->
<!--   geom_vline(data = orig %>%  -->
<!--                filter(variable == "mu2"), -->
<!--              aes(xintercept = mean), col = "red") -->

<!-- ## try it also for some of the winner effects -->

<!-- perm_results %>%  -->
<!--   filter(variable == "gamma1") %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = "Permutations of gamma_1", -->
<!--        subtitle = "True data appears different", -->
<!--        x = "gamma_1") + -->
<!--   geom_vline(data = orig %>%  -->
<!--                filter(variable == "gamma1"), -->
<!--              aes(xintercept = mean), col = "red") -->

<!-- perm_results %>%  -->
<!--   filter(variable == "gamma2") %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = "Permutations of gamma_2", -->
<!--        subtitle = "True data appears different", -->
<!--        x = "gamma_2") + -->
<!--   geom_vline(data = orig %>%  -->
<!--                filter(variable == "gamma2"), -->
<!--              aes(xintercept = mean), col = "red") -->

<!-- perm_results %>%  -->
<!--   filter(variable == "tau1") %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = "Permutations of tau_1", -->
<!--        subtitle = "True data appears different", -->
<!--        x = "tau1") + -->
<!--   geom_vline(data = orig %>%  -->
<!--                filter(variable == "tau1"), -->
<!--              aes(xintercept = mean), col = "red") -->

<!-- perm_results %>%  -->
<!--   filter(variable == "tau2") %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = "Permutations of tau_2", -->
<!--        subtitle = "True data appears different", -->
<!--        x = "tau_2") + -->
<!--   geom_vline(data = orig %>%  -->
<!--              filter(variable == "tau2"), -->
<!--              aes(xintercept = mean), col = "red") -->

<!-- ``` -->

<!-- This quick permutation seem to indicate that there is evidence for these  -->
<!-- effects, which is not present when the results are randomly permuted  -->
<!-- instead. -->

<!-- Now we can look at the permutation distributions of $\alpha$. These are -->
<!-- centered at $\mu_2$ with standard deviation $\tau_2$. -->
<!-- As expected, all permutation distributions are essentially  -->
<!-- draws from the the global mean player effect, which corresponds to some  -->
<!-- overall number of games won. -->
<!-- Most of the estimates fall in these permutation distributions, essentially -->
<!-- indicating that the individual effect is the same as the global effect for these -->
<!-- players.  -->
<!-- For the players with a smaller effect, they are less likely to win -->
<!-- then the overall average, while players with a larger effect are more likely to -->
<!-- win. -->

<!-- I suspect the jumpy nature of these permutation distributions would -->
<!-- disappear if more permutations were considered. -->

<!-- ```{r repeat for all alpha} -->

<!-- alpha_var <- paste0("alpha[", 1:18, "]") -->

<!-- # alpha_var -->

<!-- for(i in seq_along(alpha_var)){ -->
<!--   plot_curr <- perm_results %>%  -->
<!--   filter(variable == alpha_var[i]) %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = paste0("Permutations of ", alpha_var[i]), -->
<!--        subtitle = "True data appears different", -->
<!--        x = alpha_var[i]) + -->
<!--   geom_vline(data = orig %>%  -->
<!--              filter(variable == alpha_var[i]), -->
<!--              aes(xintercept = mean), col = "red") -->
<!--   print(plot_curr) -->
<!-- } -->

<!-- ``` -->

<!-- Given this global permutation of all game results, if we look at the  -->
<!-- winner effect, we should expect to not see any evidence for these under -->
<!-- the permutation distribution. The individual effects have mean $\mu_1$  -->
<!-- and sd $\tau_1$. Note that these $\beta$ estimates under the  -->
<!-- original data, which correspond to to winner and loser effects, appear -->
<!-- to have stronger evidence under this permutation. Almost all of these are  -->
<!-- positive and the one negative estimate is likely not significant, based -->
<!-- on the posterior distribution anyway. -->

<!-- ```{r repeat for all beta} -->

<!-- beta_var <- paste0("beta[", 1:18, "]") -->

<!-- # alpha_var -->

<!-- for(i in seq_along(beta_var)){ -->
<!--   plot_curr <- perm_results %>%  -->
<!--   filter(variable == beta_var[i]) %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = paste0("Permutations of ", beta_var[i]), -->
<!--        subtitle = "True data appears different", -->
<!--        x = beta_var[i]) + -->
<!--   geom_vline(data = orig %>%  -->
<!--              filter(variable == beta_var[i]), -->
<!--              aes(xintercept = mean), col = "red") -->
<!--   print(plot_curr) -->
<!-- } -->

<!-- ``` -->

<!-- Are the estimates that are not centered around 0 explained by the  -->
<!-- overall proportion of wins in the dataset? When you permute all the results -->
<!-- you essentially get that 48% of games are won by a focal player playing -->
<!-- as black. This is basically $1/(1+\exp(-1(-0.1)))$. Note that the  -->
<!-- proportion is higher in the original data. -->

<!-- ```{r} -->
<!-- #| eval: true -->
<!-- #| echo: false -->

<!-- lichess_data <- readRDS("../../rdata/lichess_pilot.RData") -->

<!-- bullet_60 <- lichess_data %>%  -->
<!--   filter(Event == "Rated Bullet game") %>%  -->
<!--   filter(TimeControl == "60+0") %>%  -->
<!--   mutate(game_id = row_number())  -->

<!-- get_hist <- function(user, games, prev_n) { -->
<!--   hist_games <- games %>%  -->
<!--     filter(White == user | Black == user) %>%  -->
<!--     arrange(UTCDate, UTCTime) %>%  -->
<!--     mutate(focal_white = ifelse(Username == White, 1, 0)) %>%  -->
<!--     select(White:BlackElo, focal_white) %>%  -->
<!--     mutate(focal_result = case_when( -->
<!--       (focal_white == 1 & Result == "1-0") ~ 1, -->
<!--       (focal_white == 0 & Result == "0-1") ~ 1, -->
<!--       (Result == "1/2-1/2") ~ 0.5, -->
<!--       .default = 0 -->
<!--       )) %>%  -->
<!--     mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)),  -->
<!--                               roll_mean(focal_result, n = prev_n))) -->

<!--   hist_games -->
<!-- } -->

<!-- focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18) -->

<!-- top_players <- names(focal_players) -->

<!-- tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 10) %>%  -->
<!--   as_tibble() -->

<!-- init_data <- tidy_games %>%  -->
<!--   mutate(WhiteElo = as.numeric(WhiteElo),  -->
<!--          BlackElo = as.numeric(BlackElo)) %>%  -->
<!--   mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>%  -->
<!--   mutate(elo_diff = ifelse(focal_white == 1,  -->
<!--                            WhiteElo - BlackElo, BlackElo - WhiteElo)) %>%  -->
<!--   mutate(focal_id = match(focal_user, top_players)) %>%  -->
<!--   select(focal_user, focal_id, focal_white,  -->
<!--          focal_win_prop, elo_diff, focal_result) %>%  -->
<!--   group_by(focal_id) %>%  -->
<!--   mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_win_prop)) %>%  -->
<!--   filter(focal_result != 0.5) -->

<!-- init_data %>% -->
<!--   group_by(focal_white, focal_result) %>% -->
<!--   tally() %>% -->
<!--   mutate(prop = n /sum(n)) -->
<!-- #  -->

<!-- ``` -->

<!-- ### Focal Player Permutation -->

<!-- Alternatively, we could restrict our permutations to  -->
<!-- instead be at the level of each focal player. This -->
<!-- means that the total number of games won by each focal -->
<!-- player will be the same, but any temporal -->
<!-- structure to these results will be removed. -->

<!-- ## Questions -->

<!-- - Are we ignoring any covariates which could influence the model here? -->

<!-- - How do these effects change if we only look at games which were played  -->
<!-- directly before? -->

### 2024-01-10 owen/cluster_scripts/Large_Single_Prev.R
fix the plot naming convention

@@ -172,7 +172,7 @@ mcmc_hist(fit3_ave$draws(c("mu2", "mu1",  "gamma1", "gamma2")),
 
 # ggsave(filename = paste0(save_path, "/global_pars_all_rated_bullet_model_prev.png"),
                          # width = 8, height = 8, units = "in")
-ggsave(filename = paste0(save_path, "/global_pars_all_rated_blitz_model.png"),
+ggsave(filename = paste0(save_path, "/global_pars_all_rated_blitz_model_prev.png"),
        width = 8, height = 8, units = "in")
 
 mcmc_hist(fit3_ave$draws("beta"),
@@ -180,7 +180,7 @@ mcmc_hist(fit3_ave$draws("beta"),
 
 # ggsave(filename = paste0(save_path, "/winner_pars_all_rated_bullet_model_prev.png"),
 #        width = 8, height = 8, units = "in")
-ggsave(filename = paste0(save_path, "/winner_pars_all_rated_blitz_model.png"),
+ggsave(filename = paste0(save_path, "/winner_pars_all_rated_blitz_model_prev.png"),
        width = 8, height = 8, units = "in")
 
 names(players) <- paste0("alpha[", 1:length(users), "]")
@@ -190,5 +190,5 @@ mcmc_hist(fit3_ave$draws("alpha"),
 
 # ggsave(filename = paste0(save_path, "/indiv_pars_all_rated_bullet_model_prev.png"),
 #        width = 8, height = 8, units = "in")
-ggsave(filename = paste0(save_path, "/indiv_pars_all_rated_blitz_model.png"),
+ggsave(filename = paste0(save_path, "/indiv_pars_all_rated_blitz_model_prev.png"),
        width = 8, height = 8, units = "in")

### 2024-01-11 owen/cluster_scripts/Large_Single_Prev.R
fix some plot code, which wasn't run correctly, but can still get them from the saved rds

@@ -163,7 +163,7 @@ fit3_ave$save_object(file = here(save_path, "all_rated_blitz_model_prev.RDS"))
 ## create some summary plots of these results
 
 players <- users
-names(players) <- paste0("beta[", 1:length(users), "]")
+names(players) <- paste0("delta[", 1:length(users), "]")
 
 player_labels <- as_labeller(players)
 
@@ -175,7 +175,7 @@ mcmc_hist(fit3_ave$draws(c("mu2", "mu1",  "gamma1", "gamma2")),
 ggsave(filename = paste0(save_path, "/global_pars_all_rated_blitz_model_prev.png"),
        width = 8, height = 8, units = "in")
 
-mcmc_hist(fit3_ave$draws("beta"),
+mcmc_hist(fit3_ave$draws("delta"),
           facet_args = list(labeller = player_labels)) 
 
 # ggsave(filename = paste0(save_path, "/winner_pars_all_rated_bullet_model_prev.png"),

### 2024-01-11 owen/Write Ups/04_model_checking.qmd
starting to update the report with the full fits

@@
 aim is for this document to be self contained.
+
+This is being updated now to show all players in the Box Data.
 
@@
 
-### 1700-1900 Players
+### 1700-1900 Players, Bullet
 
-We fit this to games from 25 players in the 1700-1900 range. 
+We fit this to games from all players in the 1700-1900 range. 
 We looked only at standard rated bullet games (no tournaments, etc),
-which was 445k games, ranging from 500 - 50k games per player.
-Fitting this model take over 12 hours on a computing cluster.
+which 675k games, ranging from 500 - 128k games per player.
+Fitting this model takes many hours on a computing cluster.
 
@@
 ```{r}
-fit <- readRDS(here("results/", "lichess1700-1900/", 
-                    "all_rated_bullet_model3.RDS"))
+fit <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
+                    "all_rated_bullet_model.RDS"))
+
+users <- readRDS(here("results/", "Full_Fits",
+                      "lichess1700-1900/",
+                      "users_bullet.RDS"))
 
@@
 
+fit$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+
 ```
 
-- Here the mean for $\mu_2$ is -0.0651, corresponding to a 48% chance of 
+- Here the mean for $\mu_2$ is -0.0483, corresponding to a 48% chance of 
 winning as black against an equally ranked opponent. This parameter
@@
 
-- $\gamma_1$ is estimated to be about 0.115, indicating that playing 
-as white increases the win probability by 3%, holding other parameters 
+- $\gamma_1$ is estimated to be about 0.094, indicating that playing 
+as white increases the win probability by 2.5%, holding other parameters 
 fixed. Posterior SD is very small.
@@
 
-```{r get names 1700-1900}
-#| cache: true
-
-## need to get user names here, from the data again
-
-data_path <- here("box_data/lichess1700-1900/")
-
-files <- list.files(data_path)
-files <- files[1:25]
-
-read_player <- function(path, file){
-  dat <- read_csv(file = paste0(data_path, file),
-                  col_types = cols(UTCDate = col_date("%Y.%m.%d"),
-                                   WhiteTitle = col_character(),
-                                   BlackTitle = col_character(),
-                                   WhiteElo = col_character(),
-                                   BlackElo = col_character(),
-                                   FEN = col_character())) %>% 
-    select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
-           WhiteElo, BlackElo, Variant, TimeControl, Termination) %>% 
-    mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
-           BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo)))
-  dat
-}
-
-get_hist <- function(user, games, prev_n) {
-  hist_games <- games %>% 
-    filter(White == user | Black == user) %>% 
-    arrange(UTCDate, UTCTime) %>% 
-    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
-    select(White:BlackElo, focal_white) %>% 
-    mutate(focal_result = case_when(
-      (focal_white == 1 & Result == "1-0") ~ 1,
-      (focal_white == 0 & Result == "0-1") ~ 1,
-      (Result == "1/2-1/2") ~ 0.5,
-      .default = 0
-    )) %>% 
-    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
-                              roll_mean(focal_result, n = prev_n)))
-
-  hist_games
-}
-
-lichess_data <- files %>% 
-  map_dfr(~read_player(data_path, .x))
-
-## restrict to rated rapid and shorter here
-## this also removes the NAs, which makes sense
-
-small_data <- lichess_data %>% 
-  # filter(Event == "Rated Bullet game") %>% 
-  # filter(TimeControl == "60+0") %>% 
-  filter(Variant == "Standard") %>% 
-  filter(grepl("Rated Bullet game", Event)) #%>% 
-  # filter(!grepl("Classical|Correspondence", Event))
-
-users <- unique(small_data$Username)
-
-rm(lichess_data)
-rm(small_data)
-
-```
-
 ```{r winner-effects-17-19}
-#| fig.height = 7
+#| fig.height = 9
 
@@
 We have between 1200 and 89k games for each of these 10 players.
+
+<!-- Down to here in terms of updating this -->
 

### 2024-01-11 owen/cluster_scripts/Large_Fit.R

@@ -29,6 +29,10 @@ all_data_path[1] <- here("box_data/lichess1700-1900/")
 all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_save_path[1] <- here("results/lichess1700-1900/")
+
+# all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
+# ## if need to run it locally
+
 all_save_path[2] <- here("results/lichess2000-2200/")
 all_save_path[3] <- here("results/lichess2300-2500/")
 
@@ -86,19 +90,22 @@ lichess_data <- files %>%
 ## restrict to rated rapid and shorter here
 ## this also removes the NAs, which makes sense
 
+# small_data <- lichess_data %>%
+#   # filter(Event == "Rated Bullet game") %>%
+#   # filter(TimeControl == "60+0") %>%
+#   filter(Variant == "Standard") %>%
+#   filter(grepl("Rated Bullet game", Event)) 
+
 small_data <- lichess_data %>%
   # filter(Event == "Rated Bullet game") %>%
   # filter(TimeControl == "60+0") %>%
   filter(Variant == "Standard") %>%
-  filter(grepl("Rated Bullet game", Event)) 
-
-# small_data <- lichess_data %>% 
-#   # filter(Event == "Rated Bullet game") %>% 
-#   # filter(TimeControl == "60+0") %>% 
-#   filter(Variant == "Standard") %>% 
-#   filter(grepl("Rated Blitz game", Event)) 
+  filter(grepl("Rated Blitz game", Event))
 
 users <- unique(small_data$Username)
+
+# saveRDS(users, file = paste0(save_path, "users_bullet.RDS"))
+saveRDS(users, file = paste0(save_path, "users_blitz.RDS"))
 
 ## when players play less than 10 games
 ## otherwise not needed

### 2024-01-15 owen/cluster_scripts/Large_Fit.R
rerun all the bullet

@@ -90,17 +90,17 @@ lichess_data <- files %>%
 ## restrict to rated rapid and shorter here
 ## this also removes the NAs, which makes sense
 
-# small_data <- lichess_data %>%
-#   # filter(Event == "Rated Bullet game") %>%
-#   # filter(TimeControl == "60+0") %>%
-#   filter(Variant == "Standard") %>%
-#   filter(grepl("Rated Bullet game", Event)) 
-
 small_data <- lichess_data %>%
   # filter(Event == "Rated Bullet game") %>%
   # filter(TimeControl == "60+0") %>%
   filter(Variant == "Standard") %>%
-  filter(grepl("Rated Blitz game", Event))
+  filter(grepl("Rated Bullet game", Event))
+
+# small_data <- lichess_data %>%
+#   # filter(Event == "Rated Bullet game") %>%
+#   # filter(TimeControl == "60+0") %>%
+#   filter(Variant == "Standard") %>%
+#   filter(grepl("Rated Blitz game", Event))
 
 users <- unique(small_data$Username)
 
@@ -161,8 +161,8 @@ fit3_ave <- mod$sample(data = stan_data_ave,
 
 ## save the stan fit as not actually that large here
 
-# fit3_ave$save_object(file = here(save_path, "all_rated_bullet_model.RDS"))
-fit3_ave$save_object(file = here(save_path, "all_rated_blitz_model.RDS"))
+fit3_ave$save_object(file = here(save_path, "all_rated_bullet_model.RDS"))
+# fit3_ave$save_object(file = here(save_path, "all_rated_blitz_model.RDS"))
 
 ## create some summary plots of these results
 
@@ -174,25 +174,25 @@ player_labels <- as_labeller(players)
 mcmc_hist(fit3_ave$draws(c("mu2", "mu1",  "gamma1", "gamma2")),
           facet_args = list(scales = "free"))
 
-# ggsave(filename = paste0(save_path, "/global_pars_all_rated_bullet_model.png"),
-#                          width = 8, height = 8, units = "in")
-ggsave(filename = paste0(save_path, "/global_pars_all_rated_blitz_model.png"),
-       width = 8, height = 8, units = "in")
+ggsave(filename = paste0(save_path, "/global_pars_all_rated_bullet_model.png"),
+                         width = 8, height = 8, units = "in")
+# ggsave(filename = paste0(save_path, "/global_pars_all_rated_blitz_model.png"),
[…]

### 2024-01-15 owen/cluster_scripts/Large_Fit.R
run two of the blitz fits again

@@ -96,11 +96,11 @@ small_data <- lichess_data %>%
   filter(Variant == "Standard") %>%
   filter(grepl("Rated Bullet game", Event))
 
-# small_data <- lichess_data %>%
-#   # filter(Event == "Rated Bullet game") %>%
-#   # filter(TimeControl == "60+0") %>%
-#   filter(Variant == "Standard") %>%
-#   filter(grepl("Rated Blitz game", Event))
+small_data <- lichess_data %>%
+  # filter(Event == "Rated Bullet game") %>%
+  # filter(TimeControl == "60+0") %>%
+  filter(Variant == "Standard") %>%
+  filter(grepl("Rated Blitz game", Event))
 
 users <- unique(small_data$Username)
 
@@ -161,8 +161,8 @@ fit3_ave <- mod$sample(data = stan_data_ave,
 
 ## save the stan fit as not actually that large here
 
-fit3_ave$save_object(file = here(save_path, "all_rated_bullet_model.RDS"))
-# fit3_ave$save_object(file = here(save_path, "all_rated_blitz_model.RDS"))
+# fit3_ave$save_object(file = here(save_path, "all_rated_bullet_model.RDS"))
+fit3_ave$save_object(file = here(save_path, "all_rated_blitz_model.RDS"))
 
 ## create some summary plots of these results
 
@@ -174,25 +174,25 @@ player_labels <- as_labeller(players)
 mcmc_hist(fit3_ave$draws(c("mu2", "mu1",  "gamma1", "gamma2")),
           facet_args = list(scales = "free"))
 
-ggsave(filename = paste0(save_path, "/global_pars_all_rated_bullet_model.png"),
-                         width = 8, height = 8, units = "in")
-# ggsave(filename = paste0(save_path, "/global_pars_all_rated_blitz_model.png"),
-#        width = 8, height = 8, units = "in")
+# ggsave(filename = paste0(save_path, "/global_pars_all_rated_bullet_model.png"),
+#                          width = 8, height = 8, units = "in")
+ggsave(filename = paste0(save_path, "/global_pars_all_rated_blitz_model.png"),
+       width = 8, height = 8, units = "in")
 
 mcmc_hist(fit3_ave$draws("beta"),
           facet_args = list(labeller = player_labels)) 
 
-ggsave(filename = paste0(save_path, "/winner_pars_all_rated_bullet_model.png"),
-       width = 8, height = 8, units = "in")
[…]

### 2024-01-15 owen/cluster_scripts/Large_Fit.R
set everything up correctly to run the bullet games again

@@ -30,7 +30,7 @@ all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_save_path[1] <- here("results/lichess1700-1900/")
 
-# all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
+all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
 # ## if need to run it locally
 
 all_save_path[2] <- here("results/lichess2000-2200/")
@@ -95,17 +95,17 @@ small_data <- lichess_data %>%
   # filter(TimeControl == "60+0") %>%
   filter(Variant == "Standard") %>%
   filter(grepl("Rated Bullet game", Event))
-
-small_data <- lichess_data %>%
-  # filter(Event == "Rated Bullet game") %>%
-  # filter(TimeControl == "60+0") %>%
-  filter(Variant == "Standard") %>%
-  filter(grepl("Rated Blitz game", Event))
+# 
+# small_data <- lichess_data %>%
+#   # filter(Event == "Rated Bullet game") %>%
+#   # filter(TimeControl == "60+0") %>%
+#   filter(Variant == "Standard") %>%
+#   filter(grepl("Rated Blitz game", Event))
 
 users <- unique(small_data$Username)
 
-# saveRDS(users, file = paste0(save_path, "users_bullet.RDS"))
-saveRDS(users, file = paste0(save_path, "users_blitz.RDS"))
+saveRDS(users, file = paste0(save_path, "users_bullet.RDS"))
+# saveRDS(users, file = paste0(save_path, "users_blitz.RDS"))
 
 ## when players play less than 10 games
 ## otherwise not needed
@@ -161,8 +161,8 @@ fit3_ave <- mod$sample(data = stan_data_ave,
 
 ## save the stan fit as not actually that large here
 
-# fit3_ave$save_object(file = here(save_path, "all_rated_bullet_model.RDS"))
-fit3_ave$save_object(file = here(save_path, "all_rated_blitz_model.RDS"))
+fit3_ave$save_object(file = here(save_path, "all_rated_bullet_model.RDS"))
+# fit3_ave$save_object(file = here(save_path, "all_rated_blitz_model.RDS"))
 
 ## create some summary plots of these results
 
@@ -174,25 +174,25 @@ player_labels <- as_labeller(players)
[…]

### 2024-01-15 owen/cluster_scripts/Large_Fit.R
save this in the right place

@@ -30,7 +30,7 @@ all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_save_path[1] <- here("results/lichess1700-1900/")
 
-all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
+# all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
 # ## if need to run it locally
 
 all_save_path[2] <- here("results/lichess2000-2200/")

### 2024-01-15 owen/Write Ups/04_model_checking.qmd
working on updating all of this, putting in all the fits and starting some of the other things we discussed in the last meeting with James

@@
   apply(2, function(x) sum(x>0)/ndraws)
+}
+
+read_player <- function(path, file){
+  dat <- read_csv(file = paste0(data_path, file),
+                  col_types = cols(UTCDate = col_date("%Y.%m.%d"),
+                                   WhiteTitle = col_character(),
+                                   BlackTitle = col_character(),
+                                   WhiteElo = col_character(),
+                                   BlackElo = col_character(),
+                                   FEN = col_character())) %>% 
+    select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
+           WhiteElo, BlackElo, Variant, TimeControl, Termination) %>% 
+    mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
+           BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo)))
+  dat
+}
+
+get_hist <- function(user, games, prev_n) {
+  hist_games <- games %>% 
+    filter(White == user | Black == user) %>% 
+    arrange(UTCDate, UTCTime) %>% 
+    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
+    select(White:BlackElo, focal_white) %>% 
+    mutate(focal_result = case_when(
+      (focal_white == 1 & Result == "1-0") ~ 1,
+      (focal_white == 0 & Result == "0-1") ~ 1,
+      (Result == "1/2-1/2") ~ 0.5,
+      .default = 0
+    )) %>% 
+    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
+                              roll_mean(focal_result, n = prev_n)))
+
+  hist_games
 }
@@
 
-We fit this model also to 10 players rated between 2000 and 2200. 
-With the same restrictions on the games this gives us 285k games.
-We have between 1200 and 89k games for each of these 10 players.
+TO BE COMPLETED
 
-<!-- Down to here in terms of updating this -->
+We fit this model also to all players rated between 2000 and 2200. 
+With the same restrictions on the games this gives us 834k games.
+We have between 1200 and 89k games for each of these players.
 
 ```{r}
+#| eval: false
 
-fit2 <- readRDS(here("results/", "lichess2000-2200/", 
+fit2 <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/", 
                     "all_rated_bullet_model.RDS"))
@@
 
-```
-
-```{r get-player-names}
-#| cache: true
-
-data_path <- here("box_data/lichess2000-2200/")
-
-files <- list.files(data_path)
-files <- files[1:10]
-
-read_player <- function(path, file){
-  dat <- read_csv(file = paste0(data_path, file),
-                  col_types = cols(UTCDate = col_date("%Y.%m.%d"),
-                                   WhiteTitle = col_character(),
-                                   BlackTitle = col_character(),
-                                   WhiteElo = col_character(),
-                                   BlackElo = col_character(),
-                                   FEN = col_character())) %>% 
-    select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
-           WhiteElo, BlackElo, Variant, TimeControl, Termination) %>% 
-    mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
-           BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo)))
-  dat
-}
-
-get_hist <- function(user, games, prev_n) {
-  hist_games <- games %>% 
-    filter(White == user | Black == user) %>% 
-    arrange(UTCDate, UTCTime) %>% 
-    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
-    select(White:BlackElo, focal_white) %>% 
-    mutate(focal_result = case_when(
-      (focal_white == 1 & Result == "1-0") ~ 1,
-      (focal_white == 0 & Result == "0-1") ~ 1,
-      (Result == "1/2-1/2") ~ 0.5,
-      .default = 0
-    )) %>% 
-    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
-                              roll_mean(focal_result, n = prev_n)))
-
-  hist_games
-}
-
-lichess_data <- files %>% 
-  map_dfr(~read_player(data_path, .x))
-
-## restrict to rated rapid and shorter here
-## this also removes the NAs, which makes sense
-
-small_data <- lichess_data %>% 
-  # filter(Event == "Rated Bullet game") %>% 
-  # filter(TimeControl == "60+0") %>% 
-  filter(Variant == "Standard") %>% 
-  filter(grepl("Rated Bullet game", Event)) #%>% 
-  # filter(!grepl("Classical|Correspondence", Event))
-
-users_20_22 <- unique(small_data$Username)
-
-rm(lichess_data)
-rm(small_data)
+users_20_22 <- readRDS(here("results/", "Full_Fits",
+                      "lichess2000-2200/",
+                      "users_bullet.RDS"))
 
@@
 ```{r}
+#| eval: false
 
@@
 ```{r}
+#| eval: false
+
 names(players) <- paste0("alpha[", 1:length(users_20_22), "]")
@@
 
-- Here the effect of a difference in ranking appears to be slightly larger.
-This leads to a bigger change in the win probability when the difference
-in ranking between the two players is larger.
+<!-- - Here the effect of a difference in ranking appears to be slightly larger. -->
+<!-- This leads to a bigger change in the win probability when the difference -->
+<!-- in ranking between the two players is larger. -->
 
-- The posterior standard deviations of the $\mu$ parameters is slightly
-larger in this model, but this could be explained by the smaller number of 
-games.
+<!-- - The posterior standard deviations of the $\mu$ parameters is slightly -->
+<!-- larger in this model, but this could be explained by the smaller number of  -->
+<!-- games. -->
 
-- A smaller sample, but half of these posterior distributions 
-of the winner effects contain 0.
+<!-- - A smaller sample, but half of these posterior distributions  -->
+<!-- of the winner effects contain 0. -->
 
@@
 
-We look at 8 players rated between 2300 and 2500. This gives us 160k games,
-with players playing 4k-58k games.
+We look at all players rated between 2300 and 2500. This gives us 703k games,
+with players playing 400-74k games.
 
 ```{r}
-fit3 <- readRDS(here("results/", "lichess2300-2500/", 
+fit3 <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/", 
                     "all_rated_bullet_model.RDS"))
 
-mcmc_hist(fit2$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+users_23_25 <- readRDS(here("results/", "Full_Fits",
+                      "lichess2300-2500/",
+                      "users_bullet.RDS"))
+
+mcmc_hist(fit3$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
   labs(title = "Global Parameters",
@@
        Gamma2 is ELO Difference")
-```
-
-```{r get users 2300-2500}
-#| cache: true
-
-data_path <- here("box_data/lichess2300-2500/")
-
-files <- list.files(data_path)
-files <- files[1:8]
-
-lichess_data <- files %>% 
-  map_dfr(~read_player(data_path, .x))
-
-## restrict to rated rapid and shorter here
-## this also removes the NAs, which makes sense
-
-small_data <- lichess_data %>% 
-  # filter(Event == "Rated Bullet game") %>% 
-  # filter(TimeControl == "60+0") %>% 
-  filter(Variant == "Standard") %>% 
-  filter(grepl("Rated Bullet game", Event)) #%>% 
-  # filter(!grepl("Classical|Correspondence", Event))
-
-users_23_25 <- unique(small_data$Username)
-
-rm(lichess_data)
-rm(small_data)
-
 ```
@@
 
-## Other Questions
+## Repeating for Blitz Games
+
+### 1700-1900
+
+```{r}
+
+fit <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
+                    "all_rated_blitz_model.RDS"))
+
+mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+  labs(title = "Global Parameters",
+       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+       Gamma2 is ELO Difference")
+
+fit$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+
+## winner effects
+players <- users
+names(players) <- paste0("beta[", 1:length(users), "]")
+
+player_labels <- as_labeller(players)
+
+mcmc_hist(fit$draws("beta"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Player Winner Effects")
+
+## indiv effects
+names(players) <- paste0("alpha[", 1:length(users), "]")
+player_labels <- as_labeller(players)
+mcmc_hist(fit$draws("alpha"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Individual Player Effects")
+```
+
+### 2000-2200
+
+TO BE COMPLETED
+
+```{r}
+#| eval: false
+
+fit2 <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/", 
+                    "all_rated_blitz_model.RDS"))
+
+mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+  labs(title = "Global Parameters",
+       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+       Gamma2 is ELO Difference")
+
+fit$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+
+## winner effects
+players <- users
+names(players) <- paste0("delta[", 1:length(users), "]")
+
+player_labels <- as_labeller(players)
+
+mcmc_hist(fit2$draws("delta"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Player Winner Effects")
+
+## indiv effects
+names(players) <- paste0("alpha[", 1:length(users), "]")
+player_labels <- as_labeller(players)
+mcmc_hist(fit2$draws("alpha"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Individual Player Effects")
+
+```
+
+### 2300-2500
+
+TO BE COMPLETED
+
+```{r}
+#| eval: false
+
+fit3 <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/", 
+                    "all_rated_blitz_model.RDS"))
+
+```
+
+## Alternative Model
+
+The alternative model is to use only the previous game as history, 
+ignoring all other games. This is perhaps more susceptible to 
+issues with correlation, which we will need to discuss further.
+
+### 1700-1900 Bullet
+
+```{r}
+
+fit <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
+                    "all_rated_bullet_model_prev.RDS"))
+
+mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+  labs(title = "Global Parameters",
+       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+       Gamma2 is ELO Difference")
+
+fit$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+
+## winner effects
+players <- users
+names(players) <- paste0("delta[", 1:length(users), "]")
+
+player_labels <- as_labeller(players)
+
+mcmc_hist(fit$draws("delta"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Player Winner Effects")
+
+## indiv effects
+names(players) <- paste0("alpha[", 1:length(users), "]")
+player_labels <- as_labeller(players)
+mcmc_hist(fit$draws("alpha"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Individual Player Effects")
+```
+
+### 2000-2200 Bullet
+
+```{r}
+fit2 <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/", 
+                    "all_rated_bullet_model_prev.RDS"))
+
+mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+  labs(title = "Global Parameters",
+       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+       Gamma2 is ELO Difference")
+
+fit2$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+
+## winner effects
+players <- users_20_22
+names(players) <- paste0("delta[", 1:length(users_20_22), "]")
+
+player_labels <- as_labeller(players)
+
+mcmc_hist(fit2$draws("delta"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Player Winner Effects")
+
+## indiv effects
+names(players) <- paste0("alpha[", 1:length(users_20_22), "]")
+player_labels <- as_labeller(players)
+mcmc_hist(fit2$draws("alpha"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Individual Player Effects")
+```
+
+### 2300-2500 Bullet
+
+```{r}
+fit3 <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/", 
+                    "all_rated_bullet_model_prev.RDS"))
+
+mcmc_hist(fit3$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+  labs(title = "Global Parameters",
+       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+       Gamma2 is ELO Difference")
+
+fit3$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+
+## winner effects
+players <- users_23_25
+names(players) <- paste0("delta[", 1:length(users_23_25), "]")
+
+player_labels <- as_labeller(players)
+
+mcmc_hist(fit3$draws("delta"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Player Winner Effects")
+
+## indiv effects
+names(players) <- paste0("alpha[", 1:length(users_23_25), "]")
+player_labels <- as_labeller(players)
+mcmc_hist(fit3$draws("alpha"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Individual Player Effects")
+```
+
+### 1700-1900 Blitz
+
+```{r}
+fit <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
+                    "all_rated_blitz_model_prev.RDS"))
+
+mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+  labs(title = "Global Parameters",
+       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+       Gamma2 is ELO Difference")
+
+fit$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+
+## winner effects
+players <- users
+names(players) <- paste0("delta[", 1:length(users), "]")
+
+player_labels <- as_labeller(players)
+
+mcmc_hist(fit$draws("delta"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Player Winner Effects")
+
+## indiv effects
+names(players) <- paste0("alpha[", 1:length(users), "]")
+player_labels <- as_labeller(players)
+mcmc_hist(fit$draws("alpha"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Individual Player Effects")
+```
+
+### 2000-2200 Blitz
+
+```{r}
+fit2 <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/", 
+                    "all_rated_blitz_model_prev.RDS"))
+
+mcmc_hist(fit2$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+  labs(title = "Global Parameters",
+       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+       Gamma2 is ELO Difference")
+
+fit2$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+
+## winner effects
+players <- users_20_22
+names(players) <- paste0("delta[", 1:length(users_20_22), "]")
+
+player_labels <- as_labeller(players)
+
+mcmc_hist(fit2$draws("delta"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Player Winner Effects")
+
+## indiv effects
+names(players) <- paste0("alpha[", 1:length(users_20_22), "]")
+player_labels <- as_labeller(players)
+mcmc_hist(fit2$draws("alpha"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Individual Player Effects")
+```
+
+### 2300-2500 Blitz
+
+```{r}
+fit3 <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/", 
+                    "all_rated_blitz_model_prev.RDS"))
+
+mcmc_hist(fit3$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+  labs(title = "Global Parameters",
+       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+       Gamma2 is ELO Difference")
+
+fit3$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+
+## winner effects
+players <- users_23_25
+names(players) <- paste0("delta[", 1:length(users_23_25), "]")
+
+player_labels <- as_labeller(players)
+
+mcmc_hist(fit3$draws("delta"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Player Winner Effects")
+
+## indiv effects
+names(players) <- paste0("alpha[", 1:length(users_23_25), "]")
+player_labels <- as_labeller(players)
+mcmc_hist(fit3$draws("alpha"),
+          facet_args = list(labeller = player_labels)) +
+  labs(title = "Individual Player Effects")
+```
+
+## Other Questions of Interest
+
+- How much do player ratings change across their games?
+
+We can look at this for a sample of players, from the 2300-2500 category.
+
+```{r}
+## load in the data for this
+
+path_id <- 3
+
+all_data_path <- rep(NA, 3)
+all_save_path <- rep(NA, 3)
+all_data_path[1] <- here("box_data/lichess1700-1900/")
+all_data_path[2] <- here("box_data/lichess2000-2200/")
+all_data_path[3] <- here("box_data/lichess2300-2500/")
+
+data_path <- all_data_path[path_id]
+files <- list.files(data_path)
+
+lichess_data <- files %>% 
+  map_dfr(~read_player(data_path, .x))
+
+## restrict to rated rapid and shorter here
+## this also removes the NAs, which makes sense
+
+small_data <- lichess_data %>%
+  # filter(Event == "Rated Bullet game") %>%
+  # filter(TimeControl == "60+0") %>%
+  filter(Variant == "Standard") %>%
+  filter(grepl("Rated Bullet game", Event))
+
+users <- unique(small_data$Username)
+
+small_data$dttm <- as.POSIXct(paste(small_data$UTCDate, small_data$UTCTime),
+                              format="%Y-%m-%d %H:%M:%S", tz = "UTC")
+
+small_data %>% 
+  mutate(UserELO = ifelse(Username == White, WhiteElo, BlackElo)) %>% 
+  group_by(Username) %>% 
+  arrange(UTCDate, UTCTime, .by_group = TRUE) %>% 
+  select(Username, dttm, UserELO) %>% 
+  filter(Username %in% users[1:16]) %>% 
+  # slice_sample(n = 100) %>% 
+  ggplot(aes(dttm, UserELO, colour = as.factor(Username))) +
+  geom_line() +
+  labs(colour = "User", x = "Date") +
+  geom_hline(yintercept = c(2300, 2500), 
+             col = "red", linetype = "dashed", alpha = 0.5) +
+  facet_wrap(~Username, scales = "free")
+
+```
+
+Each player by default starts with a rating of 1500.
+Note that there is a lot of variability in how much time (and how many games)
+each user has played at the "Ratings Bin" they are assigned to. Given
+that current opponents are based only on your current rating, this 
+should likely be accounted for. A player who has played many 
+games where their rating is above 2300 should be different 
+to a player playing their first game after attaining that
+rating. How do we account for this in the model?
+
+```{r}
+## look at different bin
+path_id <- 1
+data_path <- all_data_path[path_id]
+files <- list.files(data_path)
+
+lichess_data <- files %>% 
+  map_dfr(~read_player(data_path, .x))
+
+## restrict to rated rapid and shorter here
+## this also removes the NAs, which makes sense
+
+small_data_1 <- lichess_data %>%
+  # filter(Event == "Rated Bullet game") %>%
+  # filter(TimeControl == "60+0") %>%
+  filter(Variant == "Standard") %>%
+  filter(grepl("Rated Bullet game", Event))
+
+users_1 <- unique(small_data_1$Username)
+
+small_data_1$dttm <- as.POSIXct(paste(small_data_1$UTCDate,
+                                      small_data_1$UTCTime),
+                              format="%Y-%m-%d %H:%M:%S", tz = "UTC")
+
+small_data_1 %>% 
+  mutate(UserELO = ifelse(Username == White, WhiteElo, BlackElo)) %>% 
+  group_by(Username) %>% 
+  arrange(UTCDate, UTCTime, .by_group = TRUE) %>% 
+  select(Username, dttm, UserELO) %>% 
+  filter(Username %in% users_1[1:16]) %>% 
+  # slice_sample(n = 100) %>% 
+  ggplot(aes(dttm, UserELO, colour = as.factor(Username))) +
+  geom_line() +
+  labs(colour = "User", x = "Date") +
+  geom_hline(yintercept = c(1700, 1900), 
+             col = "red", linetype = "dashed", alpha = 0.5) +
+  facet_wrap(~Username, scales = "free")
+```
+
+- How could we control for the experience of a player here? Would this
+impact the presence of these winner effects?
+
+## Old Questions
 

### 2024-03-06 owen/Write Ups/04_model_checking.qmd
tidying up these models

@@
 of the model and describe some of the ways we have assessed the
-suitability of this model to our data. The
-aim is for this document to be self contained.
+suitability of this model to our data. This document should be self contained.
 
@@
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
-                      fig.width = 8)
+                      fig.width = 8, dev = "png",
+                      fig.height = 5, dpi = 200)
 
@@
 We looked only at standard rated bullet games (no tournaments, etc),
-which 675k games, ranging from 500 - 128k games per player.
+which was 675k games, ranging from 500 - 128k games per player.
 Fitting this model takes many hours on a computing cluster.
@@
 
-```{r}
+```{r 17-19-bullet-global, fig.path = "figures/"}
 fit <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
@@
 mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters",
+  labs(title = "Global Parameters 17-19 Bullet",
        subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
@@
 
-```{r winner-effects-17-19}
-#| fig.height = 9
+```{r winner-effects-17-19-bullet, fig.path = "figures/"}
+#| fig-height: 9
 
@@
 mcmc_hist(fit$draws("beta"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Player Winner Effects")
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Player Winner Effects 17-19 Bullet")
 ```
@@
 
-```{r indiv-effects-17-19}
-#| fig.height = 6
+```{r indiv-effects-17-19-bullet, fig.path = "figures/"}
+#| fig-height: 9
 
@@
 mcmc_hist(fit$draws("alpha"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Individual Player Effects")
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Individual Player Effects 17-19 Bullet")
 ```
@@
 ### 2000-2200 Players
-
-TO BE COMPLETED
 
@@
 
-```{r}
-#| eval: false
+```{r global-effects-20-22-bullet, fig.path = "figures/"}
 
@@
 mcmc_hist(fit2$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters",
+  labs(title = "Global Parameters 20-22 Bullet",
        subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
@@
 
-```
-
-```{r}
-#| eval: false
+fit2$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+
+```
+
+```{r winner-effects-20-22-bullet, fig.path = "figures/"}
+#| fig-height: 9
 
@@
 mcmc_hist(fit2$draws("beta"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Player Winner Effects")
-
-```
-
-```{r}
-#| eval: false
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Player Winner Effects 20-22 Bullet")
+
+```
+
+```{r indiv-effects-20-22-bullet, fig.path = "figures/"}
+#| fig-height: 9
 
@@
 mcmc_hist(fit2$draws("alpha"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Individual Player Effects")
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Individual Player Effects, 20-22 Bullet")
 ```
@@
 
-```{r}
+```{r global-effects-23-25-bullet, fig.path = "figures/"}
 fit3 <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/", 
@@
 mcmc_hist(fit3$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters",
+  labs(title = "Global Parameters 23-25 Bullet",
        subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
        Gamma2 is ELO Difference")
-```
-
-```{r}
+
+fit3$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+```
+
+```{r winner-effects-23-25-bullet, fig.path = "figures/"}
+#| fig-height: 9
 
@@
 mcmc_hist(fit3$draws("beta"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Player Winner Effects")
-
-```
-
-```{r}
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Player Winner Effects 23-25 Bullet")
+
+```
+
+```{r indiv-effects-23-25-bullet, fig.path = "figures/"}
+#| fig-height: 9
+
 names(players) <- paste0("alpha[", 1:length(users_23_25), "]")
@@
 mcmc_hist(fit3$draws("alpha"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Individual Player Effects")
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Individual Player Effects 23-25 Bullet")
 ```
@@
 
-```{r}
-
-fit <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
+Note that now there are 187k games here, considerably fewer than 
+the number of bullet games played. This ranges from 35k games to 8 players
+who play less than 25 blitz games.
+
+```{r global-effects-17-19-blitz, fig.path = "figures/"}
+
+fit_blitz <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
                     "all_rated_blitz_model.RDS"))
 
-mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters",
+mcmc_hist(fit_blitz$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+  labs(title = "Global Parameters 17-19 Blitz",
        subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
@@
 
-fit$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+fit_blitz$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+```
+
+```{r winner-effects-17-19-blitz, fig.path = "figures/"}
+#| fig-height: 9
 
@@
 
-mcmc_hist(fit$draws("beta"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Player Winner Effects")
+mcmc_hist(fit_blitz$draws("beta"),
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Player Winner Effects, 17-19 Blitz")
+
+```
+
+```{r indiv-effects-17-19-blitz, fig.path = "figures/"}
+#| fig-height: 9
 
@@
 player_labels <- as_labeller(players)
-mcmc_hist(fit$draws("alpha"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Individual Player Effects")
+mcmc_hist(fit_blitz$draws("alpha"),
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Individual Player Effects 17-19 Blitz")
 ```
@@
 
-TO BE COMPLETED
-
-```{r}
-#| eval: false
-
-fit2 <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/", 
+Here we have 250k games, ranging from 64k to 16 per player.
+
+```{r global-effects-20-22-blitz, fig.path = "figures/"}
+#| fig-height: 9
+
+fit2_blitz <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/", 
                     "all_rated_blitz_model.RDS"))
 
-mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters",
+mcmc_hist(fit2_blitz$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+  labs(title = "Global Parameters 20-22 Blitz",
        subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
@@
 
-fit$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+fit2_blitz$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+
+```
+
+```{r winner-effects-20-22-blitz, fig.path= "figures/"}
+#| fig-height: 9
+
+## winner effects
+players <- users_20_22
+names(players) <- paste0("beta[", 1:length(users_20_22), "]")
+
+player_labels <- as_labeller(players)
+
+mcmc_hist(fit2_blitz$draws("beta"),
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Player Winner Effects 20-22 Blitz")
+
+```
+
+```{r indiv-effects-20-22-blitz, fig.path = "figures/"}
+#| fig-height: 9
+
+## indiv effects
+names(players) <- paste0("alpha[", 1:length(users_20_22), "]")
+player_labels <- as_labeller(players)
+mcmc_hist(fit2_blitz$draws("alpha"),
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Individual Player Effects 20-22 Blitz")
+
+```
+
+Only have limited comparison for these at the moment, so skipped for now.
+
+### 2300-2500
+
+Here we have only 82k total games, ranging from 15k to 16 games per player.
+
+```{r global-effects-23-25-blitz, fig.path = "figures/"}
+#| fig-height: 9
+
+fit3_blitz <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/", 
+                    "all_rated_blitz_model.RDS"))
+
+mcmc_hist(fit3_blitz$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+  labs(title = "Global Parameters 23-25 Blitz",
+       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+       Gamma2 is ELO Difference")
+
+fit3_blitz$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+
+```
+
+```{r winner-effects-23-25-blitz, fig.path = "figures/"}
+#| fig-height: 9
+
+## winner effects
+players <- users_23_25
+names(players) <- paste0("beta[", 1:length(users_23_25), "]")
+
+player_labels <- as_labeller(players)
+
+mcmc_hist(fit3_blitz$draws("beta"),
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Player Winner Effects, 23-25 Blitz")
+
+```
+
+```{r indiv-effects-23-25-blitz, fig.path = "figures/"}
+#| fig-height: 9
+
+## indiv effects
+names(players) <- paste0("alpha[", 1:length(users_23_25), "]")
+player_labels <- as_labeller(players)
+mcmc_hist(fit3_blitz$draws("alpha"),
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Individual Player Effects, 23-25 Blitz")
+
+```
+
+## Alternative Model
+
+The alternative model is to use only the previous game as history, 
+ignoring all other games. This is perhaps more susceptible to 
+issues with correlation, which we will need to discuss further.
+
+We will only consider this for the bullet games for now.
+
+### 1700-1900 Bullet
+
+```{r 17-19-bullet-prev-global, fig.path = "figures/" }
+
+fit_prev <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
+                    "all_rated_bullet_model_prev.RDS"))
+
+mcmc_hist(fit_prev$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+  labs(title = "Global Parameters, 17-19 Bullet Previous",
+       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+       Gamma2 is ELO Difference")
+
+fit_prev$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+```
+
+```{r 17-19-bullet-prev-winner, fig.path= "figures/"}
+#| fig-height: 9
 
@@
 
-mcmc_hist(fit2$draws("delta"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Player Winner Effects")
+mcmc_hist(fit_prev$draws("delta"),
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Player Winner Effects, 17-19 Bullet Previous")
+```
+
+```{r 17-19-bullet-prev-indiv, fig.path="figures/"}
+#| fig-height: 9
 
@@
 player_labels <- as_labeller(players)
-mcmc_hist(fit2$draws("alpha"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Individual Player Effects")
-
-```
-
-### 2300-2500
-
-TO BE COMPLETED
-
-```{r}
-#| eval: false
-
-fit3 <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/", 
-                    "all_rated_blitz_model.RDS"))
-
-```
-
-## Alternative Model
-
-The alternative model is to use only the previous game as history, 
-ignoring all other games. This is perhaps more susceptible to 
-issues with correlation, which we will need to discuss further.
-
-### 1700-1900 Bullet
-
-```{r}
-
-fit <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
+mcmc_hist(fit_prev$draws("alpha"),
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Individual Player Effects, 17-19 Bullet Previous")
+```
+
+### 2000-2200 Bullet
+
+```{r 20-22-bullet-prev-global, fig.path="figures/"}
+fit2_prev <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/", 
                     "all_rated_bullet_model_prev.RDS"))
 
-mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters",
+mcmc_hist(fit2_prev$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+  labs(title = "Global Parameters 20-22 Bullet Previous",
        subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
@@
 
-fit$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
-
-## winner effects
-players <- users
-names(players) <- paste0("delta[", 1:length(users), "]")
-
-player_labels <- as_labeller(players)
-
-mcmc_hist(fit$draws("delta"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Player Winner Effects")
-
-## indiv effects
-names(players) <- paste0("alpha[", 1:length(users), "]")
-player_labels <- as_labeller(players)
-mcmc_hist(fit$draws("alpha"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Individual Player Effects")
-```
-
-### 2000-2200 Bullet
-
-```{r}
-fit2 <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/", 
-                    "all_rated_bullet_model_prev.RDS"))
-
-mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters",
-       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
-       Gamma2 is ELO Difference")
-
-fit2$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+fit2_prev$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+```
+
+```{r 20-22-bullet-prev-winner, fig.path="figures/"}
+#| fig-height: 9
 
@@
 
-mcmc_hist(fit2$draws("delta"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Player Winner Effects")
+mcmc_hist(fit2_prev$draws("delta"),
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Player Winner Effects 20-22 Bullet Previous")
+```
+
+```{r 20-22-bullet-prev-indiv, fig.path="figures/"}
+#| fig-height: 9
 
@@
 player_labels <- as_labeller(players)
-mcmc_hist(fit2$draws("alpha"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Individual Player Effects")
+mcmc_hist(fit2_prev$draws("alpha"),
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Individual Player Effects 20-22 Bullet Previous")
 ```
@@
 
-```{r}
-fit3 <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/", 
+```{r 23-25-bullet-prev-global, fig.path="figures/"}
+fit3_prev <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/", 
                     "all_rated_bullet_model_prev.RDS"))
 
-mcmc_hist(fit3$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters",
+mcmc_hist(fit3_prev$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+  labs(title = "Global Parameters 23-25 Bullet Previous",
        subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
@@
 
-fit3$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+fit3_prev$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+```
+
+```{r 23-25-bullet-prev-winner, fig.path="figures/"}
+#| fig-height: 9
 
@@
 
-mcmc_hist(fit3$draws("delta"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Player Winner Effects")
+mcmc_hist(fit3_prev$draws("delta"),
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Player Winner Effects 23-25 Bullet Previous")
+
+```
+
+```{r 23-25-bullet-prev-indiv, fig.path="figures/"}
+#| fig-height: 9
 
@@
 player_labels <- as_labeller(players)
-mcmc_hist(fit3$draws("alpha"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Individual Player Effects")
-```
-
-### 1700-1900 Blitz
-
-```{r}
-fit <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
-                    "all_rated_blitz_model_prev.RDS"))
-
-mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters",
-       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
-       Gamma2 is ELO Difference")
-
-fit$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
-
-## winner effects
-players <- users
-names(players) <- paste0("delta[", 1:length(users), "]")
-
-player_labels <- as_labeller(players)
-
-mcmc_hist(fit$draws("delta"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Player Winner Effects")
-
-## indiv effects
-names(players) <- paste0("alpha[", 1:length(users), "]")
-player_labels <- as_labeller(players)
-mcmc_hist(fit$draws("alpha"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Individual Player Effects")
-```
-
-### 2000-2200 Blitz
-
-```{r}
-fit2 <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/", 
-                    "all_rated_blitz_model_prev.RDS"))
-
-mcmc_hist(fit2$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters",
-       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
-       Gamma2 is ELO Difference")
-
-fit2$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
-
-## winner effects
-players <- users_20_22
-names(players) <- paste0("delta[", 1:length(users_20_22), "]")
-
-player_labels <- as_labeller(players)
-
-mcmc_hist(fit2$draws("delta"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Player Winner Effects")
-
-## indiv effects
-names(players) <- paste0("alpha[", 1:length(users_20_22), "]")
-player_labels <- as_labeller(players)
-mcmc_hist(fit2$draws("alpha"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Individual Player Effects")
-```
-
-### 2300-2500 Blitz
-
-```{r}
-fit3 <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/", 
-                    "all_rated_blitz_model_prev.RDS"))
-
-mcmc_hist(fit3$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters",
-       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
-       Gamma2 is ELO Difference")
-
-fit3$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
-
-## winner effects
-players <- users_23_25
-names(players) <- paste0("delta[", 1:length(users_23_25), "]")
-
-player_labels <- as_labeller(players)
-
-mcmc_hist(fit3$draws("delta"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Player Winner Effects")
-
-## indiv effects
-names(players) <- paste0("alpha[", 1:length(users_23_25), "]")
-player_labels <- as_labeller(players)
-mcmc_hist(fit3$draws("alpha"),
-          facet_args = list(labeller = player_labels)) +
-  labs(title = "Individual Player Effects")
-```
+mcmc_hist(fit3_prev$draws("alpha"),
+          facet_args = list(labeller = player_labels, ncol = 5)) +
+  labs(title = "Individual Player Effects 23-25 Bullet Previous")
+```
+
+<!-- ### 1700-1900 Blitz -->
+
+<!-- ```{r} -->
+<!-- fit <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/",  -->
+<!--                     "all_rated_blitz_model_prev.RDS")) -->
+
+<!-- mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) + -->
+<!--   labs(title = "Global Parameters", -->
+<!--        subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour, -->
+<!--        Gamma2 is ELO Difference") -->
+
+<!-- fit$summary(variables = c("mu1", "mu2", "gamma1", "gamma2")) -->
+
+<!-- ## winner effects -->
+<!-- players <- users -->
+<!-- names(players) <- paste0("delta[", 1:length(users), "]") -->
+
+<!-- player_labels <- as_labeller(players) -->
+
+<!-- mcmc_hist(fit$draws("delta"), -->
+<!--           facet_args = list(labeller = player_labels)) + -->
+<!--   labs(title = "Player Winner Effects") -->
+
+<!-- ## indiv effects -->
+<!-- names(players) <- paste0("alpha[", 1:length(users), "]") -->
+<!-- player_labels <- as_labeller(players) -->
+<!-- mcmc_hist(fit$draws("alpha"), -->
+<!--           facet_args = list(labeller = player_labels)) + -->
+<!--   labs(title = "Individual Player Effects") -->
+<!-- ``` -->
+
+<!-- ### 2000-2200 Blitz -->
+
+<!-- ```{r} -->
+<!-- fit2 <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/",  -->
+<!--                     "all_rated_blitz_model_prev.RDS")) -->
+
+<!-- mcmc_hist(fit2$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) + -->
+<!--   labs(title = "Global Parameters", -->
+<!--        subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour, -->
+<!--        Gamma2 is ELO Difference") -->
+
+<!-- fit2$summary(variables = c("mu1", "mu2", "gamma1", "gamma2")) -->
+
+<!-- ## winner effects -->
+<!-- players <- users_20_22 -->
+<!-- names(players) <- paste0("delta[", 1:length(users_20_22), "]") -->
+
+<!-- player_labels <- as_labeller(players) -->
+
+<!-- mcmc_hist(fit2$draws("delta"), -->
+<!--           facet_args = list(labeller = player_labels)) + -->
+<!--   labs(title = "Player Winner Effects") -->
+
+<!-- ## indiv effects -->
+<!-- names(players) <- paste0("alpha[", 1:length(users_20_22), "]") -->
+<!-- player_labels <- as_labeller(players) -->
+<!-- mcmc_hist(fit2$draws("alpha"), -->
+<!--           facet_args = list(labeller = player_labels)) + -->
+<!--   labs(title = "Individual Player Effects") -->
+<!-- ``` -->
+
+<!-- ### 2300-2500 Blitz -->
+
+<!-- ```{r} -->
+<!-- fit3 <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/",  -->
+<!--                     "all_rated_blitz_model_prev.RDS")) -->
+
+<!-- mcmc_hist(fit3$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) + -->
+<!--   labs(title = "Global Parameters", -->
+<!--        subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour, -->
+<!--        Gamma2 is ELO Difference") -->
+
+<!-- fit3$summary(variables = c("mu1", "mu2", "gamma1", "gamma2")) -->
+
+<!-- ## winner effects -->
+<!-- players <- users_23_25 -->
+<!-- names(players) <- paste0("delta[", 1:length(users_23_25), "]") -->
+
+<!-- player_labels <- as_labeller(players) -->
+
+<!-- mcmc_hist(fit3$draws("delta"), -->
+<!--           facet_args = list(labeller = player_labels)) + -->
+<!--   labs(title = "Player Winner Effects") -->
+
+<!-- ## indiv effects -->
+<!-- names(players) <- paste0("alpha[", 1:length(users_23_25), "]") -->
+<!-- player_labels <- as_labeller(players) -->
+<!-- mcmc_hist(fit3$draws("alpha"), -->
+<!--           facet_args = list(labeller = player_labels)) + -->
+<!--   labs(title = "Individual Player Effects") -->
+<!-- ``` -->
 
@@
 
-```{r}
+```{r elo-change-23-25-bullet, fig.path = "figures/"}
+
 ## load in the data for this
@@
 
-```{r}
+```{r elo-change-17-19-bullet, fig.path= "figures/"}
 ## look at different bin
@@
 
-small_data_1 <- lichess_data %>%
+bullet_data <- lichess_data %>%
   # filter(Event == "Rated Bullet game") %>%
@@
 
-users_1 <- unique(small_data_1$Username)
-
-small_data_1$dttm <- as.POSIXct(paste(small_data_1$UTCDate,
-                                      small_data_1$UTCTime),
-                              format="%Y-%m-%d %H:%M:%S", tz = "UTC")
-
-small_data_1 %>% 
+users <- unique(bullet_data$Username)
+
+bullet_data %>% 
   mutate(UserELO = ifelse(Username == White, WhiteElo, BlackElo)) %>% 
@@
   arrange(UTCDate, UTCTime, .by_group = TRUE) %>% 
-  select(Username, dttm, UserELO) %>% 
-  filter(Username %in% users_1[1:16]) %>% 
-  # slice_sample(n = 100) %>% 
-  ggplot(aes(dttm, UserELO, colour = as.factor(Username))) +
-  geom_line() +
-  labs(colour = "User", x = "Date") +
-  geom_hline(yintercept = c(1700, 1900), 
-             col = "red", linetype = "dashed", alpha = 0.5) +
-  facet_wrap(~Username, scales = "free")
+  select(Username, UTCDate, UTCTime, UserELO) %>% 
+  slice(c(1, n()))
 ```
@@
 
-## Old Questions
-
-Do focal players play each other in this data?
+- Could control for number of games played maybe?
+
+## Change in winner effects across games
+
+- Is there evidence for different winner effects in (say)
+the first 1000 games played, vs the last 1000 games played?
+
+## Relationship between winner effects and change in rating/average rating
 
 ```{r}
-data_path <- here("box_data/lichess1700-1900/")
-
+path_id <- 1
+data_path <- all_data_path[path_id]
 files <- list.files(data_path)
-files <- files[1:25]
-
-read_player <- function(path, file){
-  dat <- read_csv(file = paste0(data_path, file),
-                  col_types = cols(UTCDate = col_date("%Y.%m.%d"),
-                                   WhiteTitle = col_character(),
-                                   BlackTitle = col_character(),
-                                   WhiteElo = col_character(),
-                                   BlackElo = col_character(),
-                                   FEN = col_character())) %>% 
-    select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
-           WhiteElo, BlackElo, Variant, TimeControl, Termination) %>% 
-    mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
-           BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo)))
-  dat
-}
 
@@
   map_dfr(~read_player(data_path, .x))
-
-users <- unique(lichess_data$Username)
-
-overlap_data <- lichess_data %>% 
-  filter(White %in% users & Black %in% users) %>% 
-  filter(Variant == "Standard") %>% 
-  filter(grepl("Rated Bullet game", Event))
-
-rm(lichess_data)
-
-overlap_data %>% group_by(Username) %>% 
-  tally() %>% 
-  arrange(-n)
-```
-
-Certain players have a large amount of overlap, looking only at the games we
-consider in the analysis. How should these games be accounted for?
-
-```{r check for 2000-2200 games also}
-#| eval: false
-
-## happens here also, but not as often for this sample
-
-data_path <- here("box_data/lichess2000-2200/")
-
-files <- list.files(data_path)
-files <- files[1:10]
-
-read_player <- function(path, file){
-  dat <- read_csv(file = paste0(data_path, file),
-                  col_types = cols(UTCDate = col_date("%Y.%m.%d"),
-                                   WhiteTitle = col_character(),
-                                   BlackTitle = col_character(),
-                                   WhiteElo = col_character(),
-                                   BlackElo = col_character(),
-                                   FEN = col_character())) %>% 
-    select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
-           WhiteElo, BlackElo, Variant, TimeControl, Termination) %>% 
-    mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
-           BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo)))
-  dat
-}
-
-lichess_data <- files %>% 
-  map_dfr(~read_player(data_path, .x))
-
-users <- unique(lichess_data$Username)
-
-overlap_data <- lichess_data %>% 
-  filter(White %in% users & Black %in% users) %>% 
-  filter(Variant == "Standard") %>% 
-  filter(grepl("Rated Bullet game", Event))
-
-rm(lichess_data)
-
-overlap_data %>% group_by(Username) %>% 
-  tally() %>% 
-  arrange(-n)
-
-```
+```
+
+<!-- ## Old Questions -->
+
+<!-- Do focal players play each other in this data? -->
+
+<!-- ```{r} -->
+<!-- data_path <- here("box_data/lichess1700-1900/") -->
+
+<!-- files <- list.files(data_path) -->
+<!-- files <- files[1:25] -->
+
+<!-- read_player <- function(path, file){ -->
+<!--   dat <- read_csv(file = paste0(data_path, file), -->
+<!--                   col_types = cols(UTCDate = col_date("%Y.%m.%d"), -->
+<!--                                    WhiteTitle = col_character(), -->
+<!--                                    BlackTitle = col_character(), -->
+<!--                                    WhiteElo = col_character(), -->
+<!--                                    BlackElo = col_character(), -->
+<!--                                    FEN = col_character())) %>%  -->
+<!--     select(Username, Event, White, Black, Result, UTCDate, UTCTime,  -->
+<!--            WhiteElo, BlackElo, Variant, TimeControl, Termination) %>%  -->
+<!--     mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)), -->
+<!--            BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo))) -->
+<!--   dat -->
+<!-- } -->
+
+<!-- lichess_data <- files %>%  -->
+<!--   map_dfr(~read_player(data_path, .x)) -->
+
+<!-- users <- unique(lichess_data$Username) -->
+
+<!-- overlap_data <- lichess_data %>%  -->
+<!--   filter(White %in% users & Black %in% users) %>%  -->
+<!--   filter(Variant == "Standard") %>%  -->
+<!--   filter(grepl("Rated Bullet game", Event)) -->
+
+<!-- rm(lichess_data) -->
+
+<!-- overlap_data %>% group_by(Username) %>%  -->
+<!--   tally() %>%  -->
+<!--   arrange(-n) -->
+<!-- ``` -->
+
+<!-- Certain players have a large amount of overlap, looking only at the games we -->
+<!-- consider in the analysis. How should these games be accounted for? -->
+
+<!-- ```{r check for 2000-2200 games also} -->
+<!-- #| eval: false -->
+
+<!-- ## happens here also, but not as often for this sample -->
+
+<!-- data_path <- here("box_data/lichess2000-2200/") -->
+
+<!-- files <- list.files(data_path) -->
+<!-- files <- files[1:10] -->
+
+<!-- read_player <- function(path, file){ -->
+<!--   dat <- read_csv(file = paste0(data_path, file), -->
+<!--                   col_types = cols(UTCDate = col_date("%Y.%m.%d"), -->
+<!--                                    WhiteTitle = col_character(), -->
+<!--                                    BlackTitle = col_character(), -->
+<!--                                    WhiteElo = col_character(), -->
+<!--                                    BlackElo = col_character(), -->
+<!--                                    FEN = col_character())) %>%  -->
+<!--     select(Username, Event, White, Black, Result, UTCDate, UTCTime,  -->
+<!--            WhiteElo, BlackElo, Variant, TimeControl, Termination) %>%  -->
+<!--     mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)), -->
+<!--            BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo))) -->
+<!--   dat -->
+<!-- } -->
+
+<!-- lichess_data <- files %>%  -->
+<!--   map_dfr(~read_player(data_path, .x)) -->
+
+<!-- users <- unique(lichess_data$Username) -->
+
+<!-- overlap_data <- lichess_data %>%  -->
+<!--   filter(White %in% users & Black %in% users) %>%  -->
+<!--   filter(Variant == "Standard") %>%  -->
+<!--   filter(grepl("Rated Bullet game", Event)) -->
+
+<!-- rm(lichess_data) -->
+
+<!-- overlap_data %>% group_by(Username) %>%  -->
+<!--   tally() %>%  -->
+<!--   arrange(-n) -->
+
+<!-- ``` -->
 

### 2024-03-06 owen/cluster_scripts/Large_Fit.R
update this to just save the user names in the right order first

@@ -97,131 +97,133 @@ small_data <- lichess_data %>%
   filter(Variant == "Standard") %>%
   filter(grepl("Rated Bullet game", Event))
 # 
-# small_data <- lichess_data %>%
-#   # filter(Event == "Rated Bullet game") %>%
-#   # filter(TimeControl == "60+0") %>%
-#   filter(Variant == "Standard") %>%
-#   filter(grepl("Rated Blitz game", Event))
+small_data <- lichess_data %>%
+  # filter(Event == "Rated Bullet game") %>%
+  # filter(TimeControl == "60+0") %>%
+  filter(Variant == "Standard") %>%
+  filter(grepl("Rated Blitz game", Event))
 
 users <- unique(small_data$Username)
+
+
+
+## when players play less than 10 games
+## otherwise not needed
+users <- small_data %>%
+  group_by(Username) %>%
+  tally() %>%
+  filter(n >= 10) %>%
+  pull(Username)
 
 saveRDS(users, file = paste0(save_path, "users_bullet.RDS"))
 # saveRDS(users, file = paste0(save_path, "users_blitz.RDS"))
 
-## when players play less than 10 games
-## otherwise not needed
-users <- small_data %>% 
-  group_by(Username) %>% 
-  tally() %>% 
-  filter(n >= 10) %>% 
-  pull(Username)
-
 tidy_games <- map_dfr(users, get_hist, small_data, prev_n = 10) %>% 
   as_tibble()
 
-
-
-init_data <- tidy_games %>% 
-  mutate(WhiteElo = as.numeric(WhiteElo), 
-         BlackElo = as.numeric(BlackElo)) %>% 
-  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
-  mutate(elo_diff = ifelse(focal_white == 1, 
-                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
-  mutate(focal_id = match(focal_user, users)) %>% 
-  select(focal_user, focal_id, focal_white, 
-         focal_win_prop, elo_diff, focal_result) %>% 
-  group_by(focal_id) %>% 
-  mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_result)) %>% 
-  filter(focal_result != 0.5)
-
-cat("----------\n")
-print(dim(init_data))
-cat("----------\n")
-
-### then fit the models
-
-stan_data_ave <- list(N = nrow(init_data),
-                      J = length(users),
[…]

### 2024-03-06 owen/cluster_scripts/Large_Fit.R
update these two scripts to run a bit tidier now, create plots that are the right size also

@@ -121,109 +121,109 @@ saveRDS(users, file = paste0(save_path, "users_bullet.RDS"))
 tidy_games <- map_dfr(users, get_hist, small_data, prev_n = 10) %>% 
   as_tibble()
 
-# 
-# 
-# init_data <- tidy_games %>% 
-#   mutate(WhiteElo = as.numeric(WhiteElo), 
-#          BlackElo = as.numeric(BlackElo)) %>% 
-#   mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
-#   mutate(elo_diff = ifelse(focal_white == 1, 
-#                            WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
-#   mutate(focal_id = match(focal_user, users)) %>% 
-#   select(focal_user, focal_id, focal_white, 
-#          focal_win_prop, elo_diff, focal_result) %>% 
-#   group_by(focal_id) %>% 
-#   mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_result)) %>% 
-#   filter(focal_result != 0.5)
-# 
-# cat("----------\n")
-# print(dim(init_data))
-# cat("----------\n")
-# 
-# ### then fit the models
-# 
-# stan_data_ave <- list(N = nrow(init_data),
-#                       J = length(users),
-#                       y = init_data$focal_result,
-#                       id = init_data$focal_id,
-#                       colour = init_data$focal_white,
-#                       elo = init_data$elo_diff,
-#                       win_prop = init_data$ave_prop)
-# 
-# 
-# stan_file <- here("owen", "cluster_scripts", "model_feb29.stan")
-# 
-# mod <- cmdstan_model(stan_file)
-# 
-# fit3_ave <- mod$sample(data = stan_data_ave,
-#                        seed = 123,
-#                        chains = 4,
-#                        parallel_chains = 4,
-#                        refresh = 100)
-# 
-# 
-# ## save the stan fit as not actually that large here, 
-# ## when no generated quantities
-# 
-# fit3_ave$save_object(file = here(save_path, "all_rated_bullet_model.RDS"))
-# # fit3_ave$save_object(file = here(save_path, "all_rated_blitz_model.RDS"))
-# 
-# ## create some summary plots of these results
-# 
-# random_effect_post <- fit3_ave$draws() %>% as_draws_df() %>% 
-#   select(starts_with("beta[")) %>% 
-#   pivot_longer(cols = everything()) %>% 
[…]

### 2024-03-06 owen/cluster_scripts/Large_Single_Prev.R

@@ -111,6 +111,8 @@ users <- small_data %>%
   filter(n >= 10) %>% 
   pull(Username)
 
+saveRDS(users, file = paste0(save_path, "users_bullet.RDS"))
+
 tidy_games <- map_dfr(users, get_hist, small_data, prev_n = 10) %>%  
   as_tibble()
 
@@ -186,8 +188,9 @@ mcmc_hist(fit3_ave$draws(c("mu_beta",  "gamma1", "gamma2",
                            "sigma_g1", "sigma_g2")),
           facet_args = list(scales = "free"))
 
-ggsave(filename = paste0(save_path, "/global_pars_all_rated_bullet_model_prev.png"),
-width = 8, height = 8, units = "in")
+ggsave(filename = paste0(save_path, 
+                         "/global_pars_all_rated_bullet_model_prev.png"),
+                         width = 8, height = 8, units = "in")
 # ggsave(filename = paste0(save_path, "/global_pars_all_rated_blitz_model_prev.png"),
 #        width = 8, height = 8, units = "in")
 
@@ -198,12 +201,13 @@ random_effect_post %>%
   ggplot(aes(value)) +
   geom_histogram(fill = "#6497b1", colour = "black", size = 0.2) +
   facet_wrap(~player_id, scales = "free",
-             labeller = player_labels) +
+             labeller = player_labels, ncol = 5) +
   labs(title = "Individual Winner Effects", y = "") +
   theme(axis.text.y = element_blank(),
         axis.ticks.y = element_blank())
 
-ggsave(filename = paste0(save_path, "/winner_pars_all_rated_bullet_model_prev.png"),
+ggsave(filename = paste0(save_path, 
+                         "/winner_pars_all_rated_bullet_model_prev.png"),
        width = 8, height = 8, units = "in")
 # ggsave(filename = paste0(save_path, "/winner_pars_all_rated_blitz_model_prev.png"),
 #        width = 8, height = 8, units = "in")
@@ -213,11 +217,12 @@ random_effect_post %>%
   ggplot(aes(value)) +
   geom_histogram(fill = "#6497b1", colour = "black", size = 0.2) +
   facet_wrap(~player_id, scales = "free",
-             labeller = player_labels) +
+             labeller = player_labels, ncol = 5) +
   labs(title = "Individual Player Effects")
 
 
-ggsave(filename = paste0(save_path, "/indiv_pars_all_rated_bullet_model_prev.png"),
[…]

### 2024-03-06 owen/cluster_scripts/Large_Single_Prev.R
save both formats here

@@ -112,6 +112,7 @@ users <- small_data %>%
   pull(Username)
 
 saveRDS(users, file = paste0(save_path, "users_bullet.RDS"))
+# saveRDS(users, file = paste0(save_path, "users_blitz.RDS"))
 
 tidy_games <- map_dfr(users, get_hist, small_data, prev_n = 10) %>%  
   as_tibble()

### 2024-04-02 owen/cluster_scripts/Large_Fit.R
run blitz versions

@@ -91,19 +91,19 @@ lichess_data <- files %>%
 ## restrict to rated rapid and shorter here
 ## this also removes the NAs, which makes sense
 
-small_data <- lichess_data %>%
-  # filter(Event == "Rated Bullet game") %>%
-  filter(TimeControl == "60+0") %>%
-  filter(Variant == "Standard") %>%
-  filter(grepl("Rated Bullet game", Event))
+# small_data <- lichess_data %>%
+#   # filter(Event == "Rated Bullet game") %>%
+#   filter(TimeControl == "60+0") %>%
+#   filter(Variant == "Standard") %>%
+#   filter(grepl("Rated Bullet game", Event))
 # 
 
 ## what time length should be
-# small_data <- lichess_data %>%
-#   # filter(Event == "Rated Bullet game") %>%
-#   filter(TimeControl == "180+0") %>%
-#   filter(Variant == "Standard") %>%
-#   filter(grepl("Rated Blitz game", Event))
+small_data <- lichess_data %>%
+  # filter(Event == "Rated Bullet game") %>%
+  filter(TimeControl == "180+0") %>%
+  filter(Variant == "Standard") %>%
+  filter(grepl("Rated Blitz game", Event))
 
 users <- unique(small_data$Username)
 
@@ -118,7 +118,7 @@ users <- small_data %>%
   pull(Username)
 
 # saveRDS(users, file = paste0(save_path, "users_bullet.RDS"))
-# saveRDS(users, file = paste0(save_path, "users_blitz.RDS"))
+saveRDS(users, file = paste0(save_path, "users_blitz.RDS"))
 
 tidy_games <- map_dfr(users, get_hist, small_data, prev_n = 10) %>% 
   as_tibble()
@@ -167,8 +167,8 @@ fit3_ave <- mod$sample(data = stan_data_ave,
 ## save the stan fit as not actually that large here,
 ## when no generated quantities
 
-fit3_ave$save_object(file = here(save_path, "all_rated_bullet_model.RDS"))
-# fit3_ave$save_object(file = here(save_path, "all_rated_blitz_model.RDS"))
+# fit3_ave$save_object(file = here(save_path, "all_rated_bullet_model.RDS"))
+fit3_ave$save_object(file = here(save_path, "all_rated_blitz_model.RDS"))
 
 ## create some summary plots of these results
 
[…]

### 2024-04-02 owen/cluster_scripts/Large_Single_Prev.R

@@ -89,17 +89,17 @@ lichess_data <- files %>%
 ## restrict to rated rapid and shorter here
 ## this also removes the NAs, which makes sense
 
-small_data <- lichess_data %>%
-  # filter(Event == "Rated Bullet game") %>%
-  filter(TimeControl == "60+0") %>%
-  filter(Variant == "Standard") %>%
-  filter(grepl("Rated Bullet game", Event))
-# 
 # small_data <- lichess_data %>%
 #   # filter(Event == "Rated Bullet game") %>%
-#   # filter(TimeControl == "180+0") %>%
+#   filter(TimeControl == "60+0") %>%
 #   filter(Variant == "Standard") %>%
-#   filter(grepl("Rated Blitz game", Event))
+#   filter(grepl("Rated Bullet game", Event))
+# # 
+small_data <- lichess_data %>%
+  # filter(Event == "Rated Bullet game") %>%
+  filter(TimeControl == "180+0") %>%
+  filter(Variant == "Standard") %>%
+  filter(grepl("Rated Blitz game", Event))
 
 users <- unique(small_data$Username)
 
@@ -111,8 +111,8 @@ users <- small_data %>%
   filter(n >= 10) %>% 
   pull(Username)
 
-saveRDS(users, file = paste0(save_path, "users_bullet.RDS"))
-# saveRDS(users, file = paste0(save_path, "users_blitz.RDS"))
+# saveRDS(users, file = paste0(save_path, "users_bullet.RDS"))
+saveRDS(users, file = paste0(save_path, "users_blitz.RDS"))
 
 tidy_games <- map_dfr(users, get_hist, small_data, prev_n = 10) %>%  
   as_tibble()
@@ -170,8 +170,8 @@ fit3_ave <- mod$sample(data = stan_data_ave,
 
 ## save the stan fit as not actually that large here
 
-fit3_ave$save_object(file = here(save_path, "all_rated_bullet_model_prev.RDS"))
-# fit3_ave$save_object(file = here(save_path, "all_rated_blitz_model_prev.RDS"))
+# fit3_ave$save_object(file = here(save_path, "all_rated_bullet_model_prev.RDS"))
+fit3_ave$save_object(file = here(save_path, "all_rated_blitz_model_prev.RDS"))
 
 ## create some summary plots of these results
 
@@ -193,12 +193,12 @@ mcmc_hist(fit3_ave$draws(c("mu_beta",  "gamma1", "gamma2",
[…]

### 2024-04-02 owen/Write Ups/04_model_checking.qmd
adding version for comparing winner effects over tinme

@@
 $$
-\alpha_j \sim \mathcal{N}(\mu_2, 1)
+\alpha_j \sim \mathcal{N}(0, \sigma_{\alpha}^2)
 $$
 $$
-\mu_2 \sim \mathcal{N}(0, 1)
+\sigma_{\alpha} \sim \mathcal{IG}(1, 1)
+$$
+
+and similarly
+
+$$
+\beta_j \sim \mathcal{N}(\mu_{\beta}, \sigma_{\beta}^2)
 $$
 $$
-\tau_2 \sim \mathcal{N}(0, 1)
-$$
-
-and similarly
-
-$$
-\beta_j \sim \mathcal{N}(\mu_1, 1)
+\mu_{\beta} \sim \mathcal{N}(0, \sigma^2)
 $$
 $$
-\mu_1 \sim \mathcal{N}(0, 1)
+\sigma_{\beta} \sim \mathcal{IG}(1, 1)
 $$
+
+and
+
 $$
-\tau_1 \sim \mathcal{N}(0, 1)
+\sigma \sim \mathcal{IG}(1,1).
 $$
@@
 
-__These predictors are not currently rescaled__.
+__These fixed effect predictors are not currently rescaled__.
 
@@
 We looked only at standard rated bullet games (no tournaments, etc),
-which was 675k games, ranging from 500 - 128k games per player.
+which was 590k games, ranging from 30 - 127k games per 36 players.
 Fitting this model takes many hours on a computing cluster.
@@
 
-```{r 17-19-bullet-global, fig.path = "figures/"}
-fit <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
+```{r 17-19-bullet-global-update, fig.path = "figures/"}
+fit <- readRDS(here("results/", "lichess1700-1900_mar/", 
                     "all_rated_bullet_model.RDS"))
 
-users <- readRDS(here("results/", "Full_Fits",
-                      "lichess1700-1900/",
+users <- readRDS(here("results/",
+                      "lichess1700-1900_mar/",
                       "users_bullet.RDS"))
 
-mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+fit$summary(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))
+
+mcmc_hist(fit$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
   labs(title = "Global Parameters 17-19 Bullet",
-       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
-       Gamma2 is ELO Difference")
-
-fit$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
-
-```
-
-- Here the mean for $\mu_2$ is -0.0483, corresponding to a 48% chance of 
-winning as black against an equally ranked opponent. This parameter
-is not clearly different from 0 here. Posterior SD is large.
+       subtitle = "mu_beta is Winner Effect, Gamma1 is colour,
+       Gamma2 is ELO Difference, others SDs")
+
+# mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+#   labs(title = "Global Parameters 17-19 Bullet",
+#        subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+#        Gamma2 is ELO Difference")
+# 
+# 
+# fit$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+
+```
+
+<!-- - Here the mean for $\mu_2$ is -0.0483, corresponding to a 48% chance of  -->
+<!-- winning as black against an equally ranked opponent. This parameter -->
+<!-- is not clearly different from 0 here. Posterior SD is large. -->
 
@@
 
-- The global estimate for $\mu_1$, the global winner effect, is 
+- The global estimate for $\mu_{\beta}$, the global winner effect, is 
 around 0. Posterior SD is large here.
 
-```{r winner-effects-17-19-bullet, fig.path = "figures/"}
-#| fig-height: 9
+```{r winner-effects-17-19-bullet-update, fig.path = "figures/"}
+#| fig-height: 9
+
+random_effect_post <- fit$draws() %>% as_draws_df() %>% 
+  select(starts_with("beta[")) %>% 
+  pivot_longer(cols = everything()) %>% 
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]"))
 
@@
 
-mcmc_hist(fit$draws("beta"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Player Winner Effects 17-19 Bullet")
+random_effect_post %>% 
+  filter(param == 2) %>% 
+  select(player_id, value) %>% 
+  group_by(player_id) %>% 
+  summarise(post_mean = mean(value)) %>% 
+  arrange(-post_mean)
+
+random_effect_post %>% 
+  filter(param == 2) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 6) +
+  labs(title = "Winner Effects", y = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
 ```
@@
 
--The largest posterior mean here is 1.32. This is saying that if on
+-The largest posterior mean here is 0.4. This is saying that if on
 average you win 50% of your games, but have instead won 10 in a row, your
-probability of winning the next would increase by ~16% over that baseline.
+probability of winning the next would increase by ~5% over that baseline.
 However most estimated effects are small, and most of these posterior
@@
 
-```{r indiv-effects-17-19-bullet, fig.path = "figures/"}
-#| fig-height: 9
-
-names(players) <- paste0("alpha[", 1:length(users), "]")
-player_labels <- as_labeller(players)
-mcmc_hist(fit$draws("alpha"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Individual Player Effects 17-19 Bullet")
+```{r indiv-effects-17-19-bullet-update, fig.path = "figures/"}
+#| fig-height: 9
+
+# names(players) <- paste0("alpha[", 1:length(users), "]")
+# player_labels <- as_labeller(players)
+# mcmc_hist(fit$draws("alpha"),
+#           facet_args = list(labeller = player_labels, ncol = 5)) +
+#   labs(title = "Individual Player Effects 17-19 Bullet")
+
+random_effect_post %>% 
+  filter(param == 1) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 6) +
+  labs(title = "Individual Effects", y = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
 ```
@@
 
-- Many are quite similar to the global parameter here.
+<!-- - Many are quite similar to the global parameter here. -->
 
@@
 We fit this model also to all players rated between 2000 and 2200. 
-With the same restrictions on the games this gives us 834k games.
-We have between 1200 and 89k games for each of these players.
-
-```{r global-effects-20-22-bullet, fig.path = "figures/"}
-
-fit2 <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/", 
+With the same restrictions on the games this gives us 722k games.
+We have between 150 and 88k games for each of these 39 players.
+
+```{r global-effects-20-22-bullet-update, fig.path = "figures/"}
+
+fit2 <- readRDS(here("results/", "lichess2000-2200_mar/", 
                     "all_rated_bullet_model.RDS"))
 
-mcmc_hist(fit2$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+# mcmc_hist(fit2$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+#   labs(title = "Global Parameters 20-22 Bullet",
+#        subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+#        Gamma2 is ELO Difference")
+mcmc_hist(fit2$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
   labs(title = "Global Parameters 20-22 Bullet",
-       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
-       Gamma2 is ELO Difference")
-
-users_20_22 <- readRDS(here("results/", "Full_Fits",
-                      "lichess2000-2200/",
+       subtitle = "mu_beta is Winner Effect, Gamma1 is colour,
+       Gamma2 is ELO Difference, others SDs")
+
+users_20_22 <- readRDS(here("results/", 
+                      "lichess2000-2200_mar/",
                       "users_bullet.RDS"))
 
-fit2$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
-
-```
-
-```{r winner-effects-20-22-bullet, fig.path = "figures/"}
+fit2$summary(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]", "tau[2]",
+                           "sigma_1", "sigma_g1", "sigma_g2"))
+
+```
+
+```{r winner-effects-20-22-bullet-update, fig.path = "figures/"}
 #| fig-height: 9
@@
 
+random_effect_post2 <- fit2$draws() %>% as_draws_df() %>% 
+  select(starts_with("beta[")) %>% 
+  pivot_longer(cols = everything()) %>% 
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]"))
+
 player_labels <- as_labeller(players)
 
-mcmc_hist(fit2$draws("beta"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Player Winner Effects 20-22 Bullet")
-
-```
-
-```{r indiv-effects-20-22-bullet, fig.path = "figures/"}
-#| fig-height: 9
-
-names(players) <- paste0("alpha[", 1:length(users_20_22), "]")
-player_labels <- as_labeller(players)
-mcmc_hist(fit2$draws("alpha"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Individual Player Effects, 20-22 Bullet")
+random_effect_post2 %>% 
+  filter(param == 2) %>% 
+  select(player_id, value) %>% 
+  group_by(player_id) %>% 
+  summarise(post_mean = mean(value)) %>% 
+  arrange(-post_mean)
+
+random_effect_post2 %>% 
+  filter(param == 2) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 5) +
+  labs(title = "Winner Effects", y = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
+
+```
+
+```{r indiv-effects-20-22-bullet-update, fig.path = "figures/"}
+#| fig-height: 9
+
+random_effect_post2 %>% 
+  filter(param == 1) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 5) +
+  labs(title = "Individual Effects", y = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
+
+# names(players) <- paste0("alpha[", 1:length(users_20_22), "]")
+# player_labels <- as_labeller(players)
+# mcmc_hist(fit2$draws("alpha"),
+#           facet_args = list(labeller = player_labels, ncol = 5)) +
+#   labs(title = "Individual Player Effects, 20-22 Bullet")
 ```
@@
 
-We look at all players rated between 2300 and 2500. This gives us 703k games,
-with players playing 400-74k games.
-
-```{r global-effects-23-25-bullet, fig.path = "figures/"}
-fit3 <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/", 
+We look at all players rated between 2300 and 2500. This gives us 599k games,
+with 38 players playing 17-74k games.
+
+Size of the colour effect increases here when the ratings are larger,
+as does the ELO difference effect.
+
+```{r global-effects-23-25-bullet-update, fig.path = "figures/"}
+fit3 <- readRDS(here("results/", "lichess2300-2500_mar/", 
                     "all_rated_bullet_model.RDS"))
 
-users_23_25 <- readRDS(here("results/", "Full_Fits",
-                      "lichess2300-2500/",
+users_23_25 <- readRDS(here("results/", 
+                      "lichess2300-2500_mar/",
                       "users_bullet.RDS"))
 
-mcmc_hist(fit3$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+mcmc_hist(fit3$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
   labs(title = "Global Parameters 23-25 Bullet",
-       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
-       Gamma2 is ELO Difference")
-
-fit3$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
-```
-
-```{r winner-effects-23-25-bullet, fig.path = "figures/"}
+       subtitle = "mu_beta is Winner Effect, Gamma1 is colour,
+       Gamma2 is ELO Difference, others SDs")
+
+fit3$summary(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))
+```
+
+```{r winner-effects-23-25-bullet-update, fig.path = "figures/"}
 #| fig-height: 9
@@
 
-mcmc_hist(fit3$draws("beta"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Player Winner Effects 23-25 Bullet")
-
-```
-
-```{r indiv-effects-23-25-bullet, fig.path = "figures/"}
-#| fig-height: 9
-
-names(players) <- paste0("alpha[", 1:length(users_23_25), "]")
+random_effect_post3 <- fit3$draws() %>% as_draws_df() %>% 
+  select(starts_with("beta[")) %>% 
+  pivot_longer(cols = everything()) %>% 
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]"))
+
+random_effect_post3 %>% 
+  filter(param == 2) %>% 
+  select(player_id, value) %>% 
+  group_by(player_id) %>% 
+  summarise(post_mean = mean(value)) %>% 
+  arrange(-post_mean)
+
+random_effect_post3 %>% 
+  filter(param == 2) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 5) +
+  labs(title = "Winner Effects", y = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
+
+```
+
+```{r indiv-effects-23-25-bullet-update, fig.path = "figures/"}
+#| fig-height: 9
+
+random_effect_post3 %>% 
+  filter(param == 1) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 5) +
+  labs(title = "Individual Effects", y = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
+
+# names(players) <- paste0("alpha[", 1:length(users_23_25), "]")
+# player_labels <- as_labeller(players)
+# mcmc_hist(fit3$draws("alpha"),
+#           facet_args = list(labeller = player_labels, ncol = 5)) +
+#   labs(title = "Individual Player Effects 23-25 Bullet")
+```
+
+## Repeating for Blitz Games
+
+NOT RERUN YET
+
+### 1700-1900
+
+Note that now there are 109k games here, considerably fewer than 
+the number of bullet games played. There are 8 players
+who play less than 25 blitz games.
+
+```{r global-effects-17-19-blitz, fig.path = "figures/"}
+
+fit_blitz <- readRDS(here("results/", "lichess1700-1900_mar/", 
+                    "all_rated_blitz_model.RDS"))
+
+users_blitz <- readRDS(here("results/",
+                      "lichess1700-1900_feb/",
+                      "users_blitz.RDS"))
+
+mcmc_hist(fit_blitz$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
+  labs(title = "Global Parameters 17-19 Blitz",
+       subtitle = "mu_beta is Winner Effect, Gamma1 is colour,
+       Gamma2 is ELO Difference")
+
+fit_blitz$summary(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))
+```
+
+```{r winner-effects-17-19-blitz, fig.path = "figures/"}
+#| fig-height: 9
+
+# ## winner effects
+# players <- users
+# names(players) <- paste0("beta[", 1:length(users), "]")
+# 
+# player_labels <- as_labeller(players)
+# 
+# mcmc_hist(fit_blitz$draws("beta"),
+#           facet_args = list(labeller = player_labels, ncol = 5)) +
+#   labs(title = "Player Winner Effects, 17-19 Blitz")
+
+random_effect_post <- fit_blitz$draws() %>% as_draws_df() %>% 
+  select(starts_with("beta[")) %>% 
+  pivot_longer(cols = everything()) %>% 
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]"))
+
+players <- users_blitz
+names(players) <- paste0("beta[", 1:length(users_blitz), "]")
+
 player_labels <- as_labeller(players)
-mcmc_hist(fit3$draws("alpha"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Individual Player Effects 23-25 Bullet")
-```
-
-## Repeating for Blitz Games
-
-### 1700-1900
-
-Note that now there are 187k games here, considerably fewer than 
-the number of bullet games played. This ranges from 35k games to 8 players
-who play less than 25 blitz games.
-
-```{r global-effects-17-19-blitz, fig.path = "figures/"}
-
-fit_blitz <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
+
+random_effect_post %>% 
+  filter(param == 2) %>% 
+  select(player_id, value) %>% 
+  group_by(player_id) %>% 
+  summarise(post_mean = mean(value)) %>% 
+  arrange(-post_mean)
+
+random_effect_post %>% 
+  filter(param == 2) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 5) +
+  labs(title = "Winner Effects, 1700-1900 Blitz", y = "", x = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
+
+```
+
+```{r indiv-effects-17-19-blitz, fig.path = "figures/"}
+#| fig-height: 9
+
+## indiv effects
+random_effect_post %>% 
+  filter(param == 1) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 5) +
+  labs(title = "Indiv Effects, 1700-1900 Blitz", y = "", x = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
+
+# names(players) <- paste0("alpha[", 1:length(users), "]")
+# player_labels <- as_labeller(players)
+# mcmc_hist(fit_blitz$draws("alpha"),
+#           facet_args = list(labeller = player_labels, ncol = 5)) +
+#   labs(title = "Individual Player Effects 17-19 Blitz")
+```
+
+### 2000-2200
+
+Here we have 190k games.
+
+```{r global-effects-20-22-blitz, fig.path = "figures/"}
+#| fig-height: 9
+
+fit2_blitz <- readRDS(here("results/", "lichess2000-2200_feb/", 
                     "all_rated_blitz_model.RDS"))
 
-mcmc_hist(fit_blitz$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters 17-19 Blitz",
-       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+users_blitz_2 <- readRDS(here("results/",
+                      "lichess2000-2200_feb/",
+                      "users_blitz.RDS"))
+
+mcmc_hist(fit2_blitz$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
+  labs(title = "Global Parameters 20-22 Blitz",
+       subtitle = "mu_beta is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
        Gamma2 is ELO Difference")
 
-fit_blitz$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
-```
-
-```{r winner-effects-17-19-blitz, fig.path = "figures/"}
+fit2_blitz$summary(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))
+
+```
+
+```{r winner-effects-20-22-blitz, fig.path= "figures/"}
+#| fig-height: 9
+
+random_effect_post2 <- fit2_blitz$draws() %>% as_draws_df() %>% 
+  select(starts_with("beta[")) %>% 
+  pivot_longer(cols = everything()) %>% 
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]"))
+
+players <- users_blitz_2
+names(players) <- paste0("beta[", 1:length(users_blitz_2), "]")
+
+player_labels <- as_labeller(players)
+
+random_effect_post2 %>% 
+  filter(param == 2) %>% 
+  select(player_id, value) %>% 
+  group_by(player_id) %>% 
+  summarise(post_mean = mean(value)) %>% 
+  arrange(-post_mean)
+
+random_effect_post2 %>% 
+  filter(param == 2) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 5) +
+  labs(title = "Winner Effects, 2000-2200 Blitz", y = "", x = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
+
+```
+
+```{r indiv-effects-20-22-blitz, fig.path = "figures/"}
+#| fig-height: 9
+
+random_effect_post2 %>% 
+  filter(param == 1) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 5) +
+  labs(title = "Indiv Effects, 2000-2200 Blitz", y = "", x = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
+
+```
+
+Only have limited comparison for these at the moment, so skipped for now.
+
+### 2300-2500
+
+Here we have only 62k total games.
+
+```{r global-effects-23-25-blitz, fig.path = "figures/"}
+#| fig-height: 9
+
+fit3_blitz <- readRDS(here("results/", "lichess2300-2500_feb/", 
+                    "all_rated_blitz_model.RDS"))
+
+users_blitz_3 <- readRDS(here("results/",
+                      "lichess2300-2500_feb/",
+                      "users_blitz.RDS"))
+
+mcmc_hist(fit3_blitz$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
+  labs(title = "Global Parameters 23-25 Blitz",
+       subtitle = "mu_beta is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+       Gamma2 is ELO Difference")
+
+fit3_blitz$summary(variables =  c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))
+
+```
+
+```{r winner-effects-23-25-blitz, fig.path = "figures/"}
 #| fig-height: 9
@@
 ## winner effects
-players <- users
-names(players) <- paste0("beta[", 1:length(users), "]")
+players <- users_blitz_3
+names(players) <- paste0("beta[", 1:length(users_blitz_3), "]")
 
@@
 
-mcmc_hist(fit_blitz$draws("beta"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Player Winner Effects, 17-19 Blitz")
-
-```
-
-```{r indiv-effects-17-19-blitz, fig.path = "figures/"}
-#| fig-height: 9
-
-## indiv effects
-names(players) <- paste0("alpha[", 1:length(users), "]")
-player_labels <- as_labeller(players)
-mcmc_hist(fit_blitz$draws("alpha"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Individual Player Effects 17-19 Blitz")
-```
-
-### 2000-2200
-
-Here we have 250k games, ranging from 64k to 16 per player.
-
-```{r global-effects-20-22-blitz, fig.path = "figures/"}
-#| fig-height: 9
-
-fit2_blitz <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/", 
-                    "all_rated_blitz_model.RDS"))
-
-mcmc_hist(fit2_blitz$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters 20-22 Blitz",
-       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
-       Gamma2 is ELO Difference")
-
-fit2_blitz$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
-
-```
-
-```{r winner-effects-20-22-blitz, fig.path= "figures/"}
-#| fig-height: 9
-
-## winner effects
-players <- users_20_22
-names(players) <- paste0("beta[", 1:length(users_20_22), "]")
-
-player_labels <- as_labeller(players)
-
-mcmc_hist(fit2_blitz$draws("beta"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Player Winner Effects 20-22 Blitz")
-
-```
-
-```{r indiv-effects-20-22-blitz, fig.path = "figures/"}
-#| fig-height: 9
-
-## indiv effects
-names(players) <- paste0("alpha[", 1:length(users_20_22), "]")
-player_labels <- as_labeller(players)
-mcmc_hist(fit2_blitz$draws("alpha"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Individual Player Effects 20-22 Blitz")
-
-```
-
-Only have limited comparison for these at the moment, so skipped for now.
-
-### 2300-2500
-
-Here we have only 82k total games, ranging from 15k to 16 games per player.
-
-```{r global-effects-23-25-blitz, fig.path = "figures/"}
-#| fig-height: 9
-
-fit3_blitz <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/", 
-                    "all_rated_blitz_model.RDS"))
-
-mcmc_hist(fit3_blitz$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters 23-25 Blitz",
-       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
-       Gamma2 is ELO Difference")
-
-fit3_blitz$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
-
-```
-
-```{r winner-effects-23-25-blitz, fig.path = "figures/"}
-#| fig-height: 9
-
-## winner effects
-players <- users_23_25
-names(players) <- paste0("beta[", 1:length(users_23_25), "]")
-
-player_labels <- as_labeller(players)
-
-mcmc_hist(fit3_blitz$draws("beta"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Player Winner Effects, 23-25 Blitz")
+random_effect_post3 <- fit3_blitz$draws() %>% as_draws_df() %>% 
+  select(starts_with("beta[")) %>% 
+  pivot_longer(cols = everything()) %>% 
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]"))
+
+random_effect_post3 %>% 
+  filter(param == 2) %>% 
+  select(player_id, value) %>% 
+  group_by(player_id) %>% 
+  summarise(post_mean = mean(value)) %>% 
+  arrange(-post_mean)
+
+random_effect_post3 %>% 
+  filter(param == 2) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 5) +
+  labs(title = "Winner Effects, 2300-2500 Blitz", y = "", x = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
 
@@
 
-## indiv effects
-names(players) <- paste0("alpha[", 1:length(users_23_25), "]")
-player_labels <- as_labeller(players)
-mcmc_hist(fit3_blitz$draws("alpha"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Individual Player Effects, 23-25 Blitz")
+random_effect_post3 %>% 
+  filter(param == 1) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 5) +
+  labs(title = "Indiv Effects, 2300-2500 Blitz", y = "", x = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
 
@@
 
-fit_prev <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
+fit_prev <- readRDS(here("results/", "lichess1700-1900_mar/", 
                     "all_rated_bullet_model_prev.RDS"))
 
-mcmc_hist(fit_prev$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
+## these will be the same users
+
+mcmc_hist(fit_prev$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
   labs(title = "Global Parameters, 17-19 Bullet Previous",
-       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+       subtitle = "mu_beta is Winner Effect, Gamma1 is colour,
        Gamma2 is ELO Difference")
 
-fit_prev$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
-```
+fit_prev$summary(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))
+```
+
+The effects when only looking at the previous game are 
+much smaller than previous 10, mostly very close to 0 here.
 
@@
 #| fig-height: 9
+
+## this needs to be updated now..
 
@@
 players <- users
-names(players) <- paste0("delta[", 1:length(users), "]")
+
+names(players) <- paste0("beta[", 1:length(users), "]")
 
@@
 
-mcmc_hist(fit_prev$draws("delta"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Player Winner Effects, 17-19 Bullet Previous")
-```
+random_effect_post4 <- fit_prev$draws() %>% as_draws_df() %>% 
+  select(starts_with("beta[")) %>% 
+  pivot_longer(cols = everything()) %>% 
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]"))
+
+random_effect_post4 %>% 
+  filter(param == 2) %>% 
+  select(player_id, value) %>% 
+  group_by(player_id) %>% 
+  summarise(post_mean = mean(value)) %>% 
+  arrange(-post_mean)
+
+random_effect_post4 %>% 
+  filter(param == 2) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 6) +
+  labs(title = "Winner Effects, 1700-1900 Bullet, Previous Game", y = "", x = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
+
+```
+
+The individual effects seem to be quite small here also.
 
@@
 ## indiv effects
-names(players) <- paste0("alpha[", 1:length(users), "]")
-player_labels <- as_labeller(players)
-mcmc_hist(fit_prev$draws("alpha"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Individual Player Effects, 17-19 Bullet Previous")
+random_effect_post4 %>% 
+  filter(param == 1) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 6) +
+  labs(title = "Indiv Effects, 1700-1900 Bullet Previous", y = "", x = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
+
 ```
@@
 ```{r 20-22-bullet-prev-global, fig.path="figures/"}
-fit2_prev <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/", 
+fit2_prev <- readRDS(here("results/", "lichess2000-2200_mar/", 
                     "all_rated_bullet_model_prev.RDS"))
 
-mcmc_hist(fit2_prev$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters 20-22 Bullet Previous",
-       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+mcmc_hist(fit2_prev$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
+  labs(title = "Global Parameters, 20-22 Bullet Previous",
+       subtitle = "mu_beta is Winner Effect, Gamma1 is colour,
        Gamma2 is ELO Difference")
 
-fit2_prev$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+fit2_prev$summary(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))
+
 ```
@@
 
+## to be updated...
 ## winner effects
 players <- users_20_22
-names(players) <- paste0("delta[", 1:length(users_20_22), "]")
+
+names(players) <- paste0("beta[", 1:length(users), "]")
 
@@
 
-mcmc_hist(fit2_prev$draws("delta"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Player Winner Effects 20-22 Bullet Previous")
+random_effect_post5 <- fit2_prev$draws() %>% as_draws_df() %>% 
+  select(starts_with("beta[")) %>% 
+  pivot_longer(cols = everything()) %>% 
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]"))
+
+random_effect_post5 %>% 
+  filter(param == 2) %>% 
+  select(player_id, value) %>% 
+  group_by(player_id) %>% 
+  summarise(post_mean = mean(value)) %>% 
+  arrange(-post_mean)
+
+random_effect_post5 %>% 
+  filter(param == 2) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 5) +
+  labs(title = "Winner Effects, 2000-2200 Bullet, Previous Game", y = "", x = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
+
 ```
@@
 ## indiv effects
-names(players) <- paste0("alpha[", 1:length(users_20_22), "]")
-player_labels <- as_labeller(players)
-mcmc_hist(fit2_prev$draws("alpha"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Individual Player Effects 20-22 Bullet Previous")
+random_effect_post5 %>% 
+  filter(param == 1) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 5) +
+  labs(title = "Indiv Effects, 2000-2200 Bullet, Previous Game", y = "", x = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
 ```
@@
 ```{r 23-25-bullet-prev-global, fig.path="figures/"}
-fit3_prev <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/", 
+fit3_prev <- readRDS(here("results/", "lichess2300-2500_mar/", 
                     "all_rated_bullet_model_prev.RDS"))
 
-mcmc_hist(fit3_prev$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
-  labs(title = "Global Parameters 23-25 Bullet Previous",
-       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
+mcmc_hist(fit3_prev$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
+  labs(title = "Global Parameters, 23-25 Bullet Previous",
+       subtitle = "mu_beta is Winner Effect, Gamma1 is colour,
        Gamma2 is ELO Difference")
 
-fit3_prev$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
+fit3_prev$summary(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
+                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))
+
 ```
@@
 players <- users_23_25
-names(players) <- paste0("delta[", 1:length(users_23_25), "]")
+
+names(players) <- paste0("beta[", 1:length(users), "]")
 
@@
 
-mcmc_hist(fit3_prev$draws("delta"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Player Winner Effects 23-25 Bullet Previous")
+random_effect_post6 <- fit3_prev$draws() %>% as_draws_df() %>% 
+  select(starts_with("beta[")) %>% 
+  pivot_longer(cols = everything()) %>% 
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]"))
+
+random_effect_post6 %>% 
+  filter(param == 2) %>% 
+  select(player_id, value) %>% 
+  group_by(player_id) %>% 
+  summarise(post_mean = mean(value)) %>% 
+  arrange(-post_mean)
+
+random_effect_post6 %>% 
+  filter(param == 2) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 5) +
+  labs(title = "Winner Effects, 2300-2500 Bullet, Previous Game", y = "", x = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
 
@@
 ## indiv effects
-names(players) <- paste0("alpha[", 1:length(users_23_25), "]")
-player_labels <- as_labeller(players)
-mcmc_hist(fit3_prev$draws("alpha"),
-          facet_args = list(labeller = player_labels, ncol = 5)) +
-  labs(title = "Individual Player Effects 23-25 Bullet Previous")
+
+random_effect_post6 %>% 
+  filter(param == 1) %>% 
+  mutate(id = as.numeric(id)) %>% 
+  ggplot(aes(value)) +
+  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
+  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
+             ncol = 5) +
+  labs(title = "Indiv Effects, 2300-2500 Bullet, Previous Game", y = "", x = "") +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
+
 ```
@@
   # filter(TimeControl == "60+0") %>%
+  ## here multiple time formats will contribute, so going to include them
+  ## for now
   filter(Variant == "Standard") %>%

### 2024-04-02 owen/cluster_scripts/Winner_Effects_Time.R

@@ -1,0 +1,350 @@
+##### January 15 2024 ######
+###########
+## Fit the current model to two sets of games from each player,
+## their first n games and their last n games, to see how potential
+## winner effects change
+##
+
+
+library(tidyverse)
+library(RcppRoll)
+library(cmdstanr)
+library(posterior)
+library(bayesplot)
+library(loo)
+library(here)
+
+options(mc.cores = parallel::detectCores())
+
+theme_set(theme_bw())
+
+path_id <- as.numeric(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+
+
+### load in the data to use
+
+## rerun for 2000-2200 next
+all_data_path <- rep(NA, 3)
+all_save_path <- rep(NA, 3)
+all_data_path[1] <- here("box_data/lichess1700-1900/")
+all_data_path[2] <- here("box_data/lichess2000-2200/")
+all_data_path[3] <- here("box_data/lichess2300-2500/")
+all_save_path[1] <- here("results/lichess1700-1900_mar/")
+
+# all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
+# ## if need to run it locally
+
+all_save_path[2] <- here("results/lichess2000-2200_mar/")
+all_save_path[3] <- here("results/lichess2300-2500_mar/")
+
+
+data_path <- all_data_path[path_id]
+save_path <- all_save_path[path_id]
+
+dir.create(save_path, showWarnings = FALSE)
+
+files <- list.files(data_path)
+# files <- files[1:8]
+
+## need to write a function to process them separately then I think
+
+read_player <- function(path, file){
+  dat <- read_csv(file = paste0(data_path, file),
+                  col_types = cols(UTCDate = col_date("%Y.%m.%d"),
+                                   WhiteTitle = col_character(),
+                                   BlackTitle = col_character(),
+                                   WhiteElo = col_character(),
+                                   BlackElo = col_character(),
+                                   FEN = col_character())) %>% 
+    select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
+           WhiteElo, BlackElo, Variant, TimeControl, Termination) %>% 
+    mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
[…]

### 2024-07-04 owen/cluster_scripts/perm_results.R
need to rerun the perm for 1700-1900

@@ -40,7 +40,7 @@ all_save_path[2] <- here("results/lichess2000-2200_mar/")
 all_save_path[3] <- here("results/lichess2300-2500_mar/")
 all_save_path[4] <- here("results/lichessGrandmasters_mar/")
 
-path_id <- 2 ## setting it for now
+path_id <- 1 ## setting it for now
 data_path <- all_data_path[path_id]
 save_path <- paste0(all_save_path[path_id], "perm/")
 
@@ -65,7 +65,7 @@ select_users <- small_data %>%
   tally() %>% 
   arrange(-n) %>% 
   ## will change this to n = 20 when run on cluster potentially
-  slice_max(order = n, n = 10) %>% 
+  slice_max(order = n, n = 20) %>% 
   pull(Username)
 
 

### 2024-07-04 owen/cluster_scripts/perm_results.R
local

@@ -40,7 +40,7 @@ all_save_path[2] <- here("results/lichess2000-2200_mar/")
 all_save_path[3] <- here("results/lichess2300-2500_mar/")
 all_save_path[4] <- here("results/lichessGrandmasters_mar/")
 
-path_id <- 2 ## setting it for now
+path_id <- 1 ## setting it for now
 data_path <- all_data_path[path_id]
 save_path <- paste0(all_save_path[path_id], "perm/")
 

### 2024-07-04 owen/cluster_scripts/perm_results.R
update the perm to correctly run

@@ -53,9 +53,10 @@ lichess_data <- files %>%
 
 
 small_data <- lichess_data %>%
+  mutate(Event = tolower(Event)) |> 
   filter(TimeControl == "60+0") %>%
   filter(Variant == "Standard") %>%
-  filter(grepl("Rated Bullet game", Event))
+  filter(grepl("rated bullet game", Event))
 
 rm(lichess_data)
 

### 2024-07-04 Paper_WriteUp/Paper_Plots.qmd
add some templates for the plots and tidy up the code to make them using this

---
title: "Paper Analysis"
author: Owen G. Ward
date: today
format: 
  # html:
  #   toc: true
  #   toc-location: left
  pdf
---

This will provide the final figures used in the paper.

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 7, dev = "png", fig.path = "output/",
                      fig.height = 5, dpi = 600)

library(tidyverse)
library(RcppRoll)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(loo)
library(scales)
library(here)

options(mc.cores = parallel::detectCores())

theme_set(theme_bw())

## to load some plot defaults
source(here("analysis", "helper.R"))
source(here("analysis", "plot_templates.R"))

all_data_path <- rep(NA, 4)
all_data_path[1] <- here("box_data/lichess1700-1900/")
all_data_path[2] <- here("box_data/lichess2000-2200/")
all_data_path[3] <- here("box_data/lichess2300-2500/")
all_data_path[4] <- here("box_data/lichessGrandmasters/")

res_data_path <- rep(NA, 4)
res_data_path[1] <- here("results/lichess1700-1900_mar/")
res_data_path[2] <- here("results/lichess2000-2200_mar/")
res_data_path[3] <- here("results/lichess2300-2500_mar/")
res_data_path[4] <- here("results/lichessGrandmasters_mar/")

```

# Summary Statistics

First get some summary statistics about the overall data. 
Construct these only for the 1700-1900 players and GMs initially.

```{r}

path_id <- 1  ## for 1700-1900

data_path <- all_data_path[path_id]

files <- list.files(data_path)

lichess_data <- files %>% 
  map_dfr(~read_player(data_path, .x))

lichess_data |> 
  group_by(Event) |> 
  count() |> 
  arrange(-n)

lichess_data |> 
  mutate(Event = tolower(Event)) |> 
  filter(grepl("rated bullet game", Event)) |> 
  group_by(TimeControl) |> 
  count() |> 
  arrange(-n) |> 
  ungroup() |> 
  mutate(prop = n/sum(n))

bullet_data <- lichess_data |> 
  mutate(Event = tolower(Event)) |>
  filter(TimeControl == "60+0") %>%
  filter(Variant == "Standard") %>%
  filter(grepl("rated bullet game", Event))

bullet_data |> nrow()

bullet_data |> 
  group_by(Username) |> 
  count() |> 
  filter(n > 3) |> 
  ungroup() |> 
  summarise(avg = mean(n))

bullet_data |> 
  group_by(Username) |> 
  count() |> 
  arrange(n)

## then repeat this sort of analysis for the gms

path_id <- 4  ## for GMs

data_path <- all_data_path[path_id]

files <- list.files(data_path)

lichess_data_gm <- files %>% 
  map_dfr(~read_player(data_path, .x))

lichess_data_gm |> 
  mutate(Event = tolower(Event)) |> 
  filter(grepl("rated bullet game", Event)) |> 
  group_by(TimeControl) |> 
  count() |> 
  arrange(-n) |> 
  ungroup() |> 
  mutate(prop = n/sum(n))

bullet_data_gm <- lichess_data_gm |> 
  mutate(Event = tolower(Event)) |>
  filter(TimeControl == "60+0") %>%
  filter(Variant == "Standard") %>%
  filter(grepl("rated bullet game", Event))

bullet_data_gm |> nrow()

bullet_data_gm |> 
  group_by(Username) |> 
  count() |> 
  arrange(n)

bullet_data_gm |> 
  group_by(Username) |> 
  count() |> 
  arrange(n) |> 
  ungroup() |> 
  summarise(mean(n))

```

# Examining Model Fit

## 1700-1900 Previous Game

### Global Parameters

```{r}

data_path <- here("results/lichess1700-1900_mar")

fit_17_19_bullet_prev <- readRDS(file = 
                                   paste0(data_path,
                                          "/all_rated_bullet_model_prev.RDS"))

bullet_17_19 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))

# color_scheme_set("blue")

## easiest way to make this grayscale would be to not use mcmc_hist at all

global_17_19_bullet_prev <- fit_17_19_bullet_prev$draws(c("mu_beta",
                                                          "gamma1", "gamma2",
                           "sigma_1", "tau[1]", "tau[2]",
                           "sigma_g1", "sigma_g2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~name, ncol = 4, scales = "free") +
  labs(x = element_blank(), y = element_blank()) +
  theme_4col() +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

global_17_19_bullet_prev

ggsave(plot = global_17_19_bullet_prev, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "17_19_prev_global.png")), dpi = 600,
       height = 5, width = 7)

```

### Winner and Player Effects

```{r}

## need to get the write names then for these parameters here
## did this somewhere before already

player_id <- as_tibble(bullet_17_19) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

player_pars <- fit_17_19_bullet_prev$draws() %>% 
  as_draws_df() %>%
  select(starts_with("beta[")) %>%
  pivot_longer(cols = everything()) %>%
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id")

## then the winner effects plot

winner_plot_17_19 <- player_pars |> 
  filter(param == 2) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~player_name, scales = "free", ncol = 6) +
  theme_6col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

winner_plot_17_19

ggsave(plot = winner_plot_17_19, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "17_19_prev_winner.png")), dpi = 600,
       height = 7, width = 7)

## then repeat for player effects

indiv_plot_17_19 <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~player_name, scales = "free", ncol = 6) +
  theme_6col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

indiv_plot_17_19

ggsave(plot = indiv_plot_17_19, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "17_19_prev_indiv.png")), dpi = 600,
       height = 7, width = 7)

```

## GM previous game

### Global Parameters

```{r}
data_path <- here("results/lichessGrandmasters_mar")

fit_gm_bullet_prev <- 
  readRDS(file =paste0(data_path,"/all_rated_bullet_model_prev.RDS"))

bullet_gm <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))

global_gm_bullet_prev <- fit_gm_bullet_prev$draws(c("mu_beta",
                                                          "gamma1", "gamma2",
                           "sigma_1", "tau[1]", "tau[2]",
                           "sigma_g1", "sigma_g2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~name, ncol = 4, scales = "free") +
  labs(x = element_blank(), y = element_blank()) +
  theme_4col() +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

global_gm_bullet_prev

ggsave(plot = global_gm_bullet_prev, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_prev_global.png")), dpi = 600,
       height = 5, width = 7)

```

### Winner and Player Effects

```{r}

player_id <- as_tibble(bullet_gm) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

player_pars <- fit_gm_bullet_prev$draws() %>% 
  as_draws_df() %>%
  select(starts_with("beta[")) %>%
  pivot_longer(cols = everything()) %>%
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id")

## then the winner effects plot

winner_plot_gm <- player_pars |> 
  filter(param == 2) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~player_name, scales = "free", ncol = 5) +
  theme_5col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

winner_plot_gm

ggsave(plot = winner_plot_gm, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_prev_winner.png")), dpi = 600,
       height = 7, width = 7)

## then repeat for player effects

indiv_plot_gm <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~player_name, scales = "free", ncol = 5) +
  theme_5col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

indiv_plot_gm

ggsave(plot = indiv_plot_gm, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_prev_indiv.png")), dpi = 600,
       height = 7, width = 7)

```

Here there are only 21 rather than 25 players at the moment, this will be
updated currently.

## Other Comparisons

## GM Prev vs Previous 10

```{r}

## global effects first

data_path <- here("results/lichessGrandmasters_mar")

fit_gm_bullet_10 <- 
  readRDS(file =paste0(data_path,"/all_rated_bullet_model.RDS"))

prev_pars <- fit_gm_bullet_prev$draws(c("mu_beta",
                                                          "gamma1", "gamma2",
                           "sigma_1", "tau[1]", "tau[2]",
                           "sigma_g1", "sigma_g2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
  mutate(model = "Prev. Game")

pre_10_pars <- fit_gm_bullet_10$draws(c("mu_beta",
                                                          "gamma1", "gamma2",
                           "sigma_1", "tau[1]", "tau[2]",
                           "sigma_g1", "sigma_g2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
  mutate(model = "Prev. 10 Games")

hist_comp_global <- bind_rows(prev_pars, pre_10_pars) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_grid(rows = vars(model), cols = vars(name), scales = "free") +
  theme_8col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 2))

hist_comp_global

ggsave(plot = hist_comp_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_comp_global.png")), dpi = 600,
       height = 5, width = 7)

## then compare the winner effects and individual effects. 
## will just show for the first 6 players for convenience

player_pars_pre <- fit_gm_bullet_prev$draws() %>% 
  as_draws_df() %>%
  select(starts_with("beta[")) %>%
  pivot_longer(cols = everything()) %>%
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id") |> 
  mutate(model = "Prev. Game")

player_pars_pre_10 <- fit_gm_bullet_10$draws() %>% 
  as_draws_df() %>%
  select(starts_with("beta[")) %>%
  pivot_longer(cols = everything()) %>%
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id") |> 
  mutate(model = "Prev. 10 Games")

hist_comp_winner <- bind_rows(player_pars_pre, player_pars_pre_10) |> 
  filter(id < 7) |> 
  filter(param == 2) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_grid(rows = vars(model), cols = vars(player_name), scales = "free") +
  theme_6col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

hist_comp_winner

ggsave(plot = hist_comp_winner, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_comp_winner.png")), dpi = 600,
       height = 5, width = 7)

## then player effects
hist_comp_player <- bind_rows(player_pars_pre, player_pars_pre_10) |> 
  filter(id < 7) |> 
  filter(param == 1) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_grid(rows = vars(model), cols = vars(player_name), scales = "free") +
  theme_6col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

hist_comp_player

ggsave(plot = hist_comp_player, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_comp_player.png")), dpi = 600,
       height = 5, width = 7)

```

## Bullet vs Blitz for 17-19

Compare the global parameters and the winner effects here, using the
previous game only.

```{r}

data_path <- here("results/lichess1700-1900_mar")

fit_17_19_bullet_prev <- readRDS(file = paste0(data_path,
                                              "/all_rated_bullet_model_prev.RDS"))

bullet_17_19 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))

fit_17_19_blitz_prev <- readRDS(file = paste0(data_path,
                                              "/all_rated_blitz_model_prev.RDS"))

blitz_17_19 <- readRDS(file = paste0(data_path, "/users_blitz.RDS"))

## compare global effects first

bullet_pars <- fit_17_19_bullet_prev$draws(c("mu_beta",
                                                          "gamma1", "gamma2",
                           "sigma_1", "tau[1]", "tau[2]",
                           "sigma_g1", "sigma_g2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
  mutate(model = "Bullet")

blitz_pars <- fit_17_19_blitz_prev$draws(c("mu_beta",
                                                          "gamma1", "gamma2",
                           "sigma_1", "tau[1]", "tau[2]",
                           "sigma_g1", "sigma_g2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
  mutate(model = "Blitz")

bul_blit_comp_global <- bind_rows(bullet_pars, blitz_pars) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_grid(rows = vars(model), cols = vars(name), scales = "free") +
  theme_8col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 2))

bul_blit_comp_global

ggsave(plot = bul_blit_comp_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "bul_blitz_comp_17_19_global.png")), dpi = 600,
       height = 5, width = 7)

## then repeat for winner and player effects

player_pars_bullet <- fit_17_19_bullet_prev$draws() %>% 
  as_draws_df() %>%
  select(starts_with("beta[")) %>%
  pivot_longer(cols = everything()) %>%
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id") |> 
  mutate(model = "Bullet")

player_pars_blitz <- fit_17_19_blitz_prev$draws() %>% 
  as_draws_df() %>%
  select(starts_with("beta[")) %>%
  pivot_longer(cols = everything()) %>%
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id") |> 
  mutate(model = "Blitz")

comp_17_19_winner <- bind_rows(player_pars_bullet, player_pars_blitz) |> 
  filter(id < 7) |> 
  filter(param == 2) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_grid(rows = vars(model), cols = vars(player_name), scales = "free") +
  theme_6col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

comp_17_19_winner

ggsave(plot = comp_17_19_winner, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "bul_blitz_comp_17_19_winner.png")), dpi = 600,
       height = 5, width = 7)

comp_17_19_player <- bind_rows(player_pars_bullet, player_pars_blitz) |> 
  filter(id < 7) |> 
  filter(param == 1) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_grid(rows = vars(model), cols = vars(player_name), scales = "free") +
  theme_6col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

comp_17_19_player

ggsave(plot = comp_17_19_player, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "bul_blitz_comp_17_19_player.png")), dpi = 600,
       height = 5, width = 7)

```

## Bullet vs Blitz for GM

```{r}
data_path <- here("results/lichessGrandmasters_mar")

fit_gm_bullet_prev <- readRDS(file = paste0(data_path,
                                              "/all_rated_bullet_model_prev.RDS"))

bullet_gm <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))

fit_gm_blitz_prev <- readRDS(file = paste0(data_path,
                                              "/all_rated_blitz_model_prev.RDS"))

blitz_gm <- readRDS(file = paste0(data_path, "/users_blitz.RDS"))

## compare global effects first

bullet_pars <- fit_gm_bullet_prev$draws(c("mu_beta",
                                                          "gamma1", "gamma2",
                           "sigma_1", "tau[1]", "tau[2]",
                           "sigma_g1", "sigma_g2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
  mutate(model = "Bullet")

blitz_pars <- fit_gm_blitz_prev$draws(c("mu_beta",
                                                          "gamma1", "gamma2",
                           "sigma_1", "tau[1]", "tau[2]",
                           "sigma_g1", "sigma_g2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
  mutate(model = "Blitz")

bul_blit_gm_comp_global <- bind_rows(bullet_pars, blitz_pars) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_grid(rows = vars(model), cols = vars(name), scales = "free") +
  theme_8col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 2))

bul_blit_gm_comp_global

ggsave(plot = bul_blit_gm_comp_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "bul_blitz_comp_gm_global.png")), dpi = 600,
       height = 5, width = 7)

## then repeat for winner and player effects

player_pars_bullet <- fit_gm_bullet_prev$draws() %>% 
  as_draws_df() %>%
  select(starts_with("beta[")) %>%
  pivot_longer(cols = everything()) %>%
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id") |> 
  mutate(model = "Bullet")

player_pars_blitz <- fit_gm_blitz_prev$draws() %>% 
  as_draws_df() %>%
  select(starts_with("beta[")) %>%
  pivot_longer(cols = everything()) %>%
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id") |> 
  mutate(model = "Blitz")

comp_gm_winner <- bind_rows(player_pars_bullet, player_pars_blitz) |> 
  filter(id < 7) |> 
  filter(param == 2) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_grid(rows = vars(model), cols = vars(player_name), scales = "free") +
  theme_6col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

comp_gm_winner

ggsave(plot = comp_gm_winner, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "bul_blitz_comp_gm_winner.png")), dpi = 600,
       height = 5, width = 7)

comp_gm_player <- bind_rows(player_pars_bullet, player_pars_blitz) |> 
  filter(id < 7) |> 
  filter(param == 1) |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_grid(rows = vars(model), cols = vars(player_name), scales = "free") +
  theme_6col() +
  labs(x = element_blank(), y = element_blank()) +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

comp_gm_player

ggsave(plot = comp_gm_player, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "bul_blitz_comp_gm_player.png")), dpi = 600,
       height = 5, width = 7)

```

# Model Checking

We currently have two initial model checking 
procedures.

1. We select 10 players from each group
and look at their first and last 1000 games.
We fit our model to each of these separately
and examine posterior predictive checks based on this.

1. Then we used the model fit to the first 1000 games
to construct posterior predictive checks for the final
1000 games and see how they compare.

## 1700-1900 Players

```{r}

## this the global and winner effect plots

path_id <- 1
data_path <- res_data_path[path_id]

fit_17_19_first <- readRDS(file =  here(data_path, "time",
                                        "select_users_first_bullet.RDS"))
fit_17_19_last <- readRDS(file =  here(data_path, "time",
                                        "select_users_last_bullet.RDS"))

global_winner <- as_draws_df(fit_17_19_first$draws(variables = "mu_beta")) %>% 
  rename("Initial Games" = mu_beta) %>% 
  bind_cols(as_draws_df(fit_17_19_last$draws(variables = "mu_beta")) %>% 
              rename("Later Games" = mu_beta)) %>% 
  select(`Initial Games`, `Later Games`) %>% 
  pivot_longer(cols = everything()) %>% 
  ggplot(aes(value, fill = name)) +
  geom_histogram(position = "identity") +
  facet_wrap(~name, ncol = 1) +
  labs(fill = element_blank(), y = element_blank(), x = element_blank()) +
  theme(legend.position = "none",
        axis.text = element_text(size = axis_text_size),
        axis.title = element_text(size = axis_title),
        legend.text = element_text(size = legend_text),
        strip.text = element_text(size = axis_title),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        plot.title = element_text(size = title_size)) +
  # labs(title = "Selected 1700-1900 Bullet Players") +
  NULL

global_winner

users <- readRDS(file = here(data_path, "time", "select_users_bullet.RDS"))

## indiv winner effects
random_effect_first <- fit_17_19_first$draws() %>% as_draws_df() %>%
  select(starts_with("beta[")) %>%
  pivot_longer(cols = everything()) %>%
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) %>% 
  mutate(when = "Initial")

random_effect_last <- fit_17_19_last$draws() %>% as_draws_df() %>%
  select(starts_with("beta[")) %>%
  pivot_longer(cols = everything()) %>%
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) %>% 
  mutate(when = "Later")

betas <- bind_rows(random_effect_first, random_effect_last) %>% 
  select(value, param, id, when)

player_id <- tibble(id = as.character(1:10), player = users)

betas %>% left_join(player_id, by = "id") %>% 
  mutate(player = factor(player, levels = users))

winner_effect_plot <- betas %>% 
  left_join(player_id, by = "id") %>% 
  mutate(player = factor(player, levels = users)) %>% 
  filter(param == 2) %>% 
  mutate(id = factor(id, levels = c(1:10))) %>% 
  ggplot(aes(value, fill = when)) +
  geom_histogram() +
  geom_vline(xintercept = 0, alpha = 0.5) +
  scale_x_continuous(breaks = c(-0.25, 0.25)) +
  # xlim(c(-1.5, 1.5)) +
  facet_grid(rows = vars(when), cols = vars(player), scales = "free") +
  labs(fill = element_blank(), y = element_blank(), 
       # title = "Individual Winner Effects, Selected 1700-1900 Bullet Players",
       x = element_blank()) +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text = element_text(size = axis_text_size - 4),
        axis.title = element_text(size = axis_title),
        legend.text = element_text(size = legend_text),
        strip.text = element_text(size = axis_title - 6),
        legend.position = "none",
        plot.title = element_text(size = title_size))

winner_effect_plot
```

```{r}
path_id <- 1
data_path <- res_data_path[path_id]

fit_17_19_first <- readRDS(file =  here(data_path, "time",
                                        "select_users_first_bullet.RDS"))
fit_17_19_last <- readRDS(file =  here(data_path, "time",
                                        "select_users_last_bullet.RDS"))

fit_samples <- as_draws_df(fit_17_19_first$draws())

fit_samp <- fit_samples %>% 
  select(!starts_with(c("log_lik", "yrep")))

y_rep <- fit_samples %>% select(starts_with("y_rep"))

rm(fit_samples)
rm(fit_samp)

stan_data_ave_first <- readRDS(file = here(data_path, "time",
                                           "stan_data_first.RDS"))

y_rep_mod <- y_rep %>% 
  rownames_to_column() %>%  
  pivot_longer(-rowname) %>% 
  pivot_wider(names_from=rowname, values_from=value) %>% 
  ## remove the draws and chains here
  mutate(focal_id = stan_data_ave_first$id)

games_won <- y_rep_mod %>% 
  pivot_longer(cols = `1`:`4000`, names_to = "draw", values_to = "y") %>% 
  group_by(focal_id, draw) %>% 
  summarise(games_won = sum(y))

orig_data <- tibble(outcome = stan_data_ave_first$y,
                    focal_id = stan_data_ave_first$id)

orig_games_won <- orig_data %>% 
  group_by(focal_id) %>% 
  summarise(games_won = sum(outcome)) %>% 
  mutate(focal_id = as.character(focal_id)) %>% 
  left_join(player_id, by = c("focal_id" = "id"))

p1 <- games_won %>% 
  mutate(focal_id = as.character(focal_id)) %>% 
  left_join(player_id, by = c("focal_id" = "id")) %>% 
  ggplot(aes(x = games_won)) +
  geom_histogram() +
  facet_wrap(~player, scales = "free", ncol = 5) +
  geom_vline(data = orig_games_won, 
             mapping = aes(xintercept = games_won), col = "red") +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  # labs(title = "Posterior Predictive Distribution, Initial Games",
  labs(y = element_blank(),
       x = "Number of Games Won") +
  theme_5col()

p1

ggsave(plot = p1, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "1700_1900_ppc_1.png")), dpi = 600,
       height = 5, width = 7)

```

```{r}
## time varying ppc for 1700-1900 players

draws <- fit_17_19_first$draws() %>% as_draws_df() %>%
  select(starts_with(c("beta[", "gamma"))) %>%
  mutate(draw = row_number()) %>% 
  pivot_longer(cols = "beta[1,1]":"beta[2,10]") %>% 
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) %>% 
  select(-name) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  rename(beta1 = `1`, beta2 = `2`)

stan_data_ave_last <- readRDS(file = here(data_path, "time",
                                          "stan_data_last.RDS"))

future_games <- tibble(focal = as.character(stan_data_ave_last$id), 
       color = stan_data_ave_last$colour, 
       elo_diff = stan_data_ave_last$elo,
       hist = stan_data_ave_last$win_prop)

ppc_data <- readRDS(file = here(data_path, "time",
                                "ppc_fit_first_last.RDS"))

future_results <- tibble(focal = as.character(stan_data_ave_last$id),
                         result = stan_data_ave_last$y) %>% 
  group_by(focal) %>%
  summarise(total = sum(result)) %>% 
  rename(id = focal) %>% 
  left_join(player_id, by = "id")

last_ppc <- ppc_data %>%  
  left_join(player_id, by = "id") %>% 
  ggplot(aes(x = games_won)) + 
  geom_histogram() + 
  geom_vline(data = future_results, 
             mapping = aes(xintercept = total), col = "red") +
  facet_wrap(~player, scales = "free_x", ncol = 5) +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  labs(y = element_blank(), x = "Number of Games Won") +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  theme_5col()

last_ppc

ggsave(plot = last_ppc, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "1700_1900_ppc_2.png")), dpi = 600,
       height = 5, width = 7)
```

### Permuting for 1700-1900

This data is complete, just need to create the plots here.
Here we use the 20 players with the most games only.

```{r}

## these don't look right just now

perm_data_path <- here("results/lichess1700-1900_mar/perm")

path_id <- 1
data_path <- res_data_path[path_id]
## read all the files in this directory
(perm_files <- list.files(perm_data_path))

perm_17_19 <- perm_files[1:1000] %>% 
  map_dfr(~read_rds(here(perm_data_path, .x)))

sim_pars <- perm_17_19 |> 
  select(variable, mean, sim_id)

fit_17_19_bullet_prev <- readRDS(file = paste0(data_path,
                                              "/all_rated_bullet_model_prev.RDS"))

bullet_17_19 <- readRDS(file = paste0(data_path,
                                      "/users_bullet.RDS"))

## then need to overlay the true values also
true_values <- fit_17_19_bullet_prev$summary() |> 
  filter(variable %in% c("mu_beta", "gamma1", "gamma2")) |> 
  select(variable, mean)

perm_17_19_global <- sim_pars |> 
  filter(variable %in% c("mu_beta",  "gamma1", "gamma2",
                           "sigma_1", "tau[1]", "tau[2]",
                           "sigma_g1", "sigma_g2")) |> 
  ggplot(aes(mean)) +
  geom_histogram() +
  facet_wrap(~variable, scales = "free") +
  geom_vline(data = true_values, mapping = aes(xintercept = mean), col = "red") +
  theme_3col() +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  labs(x = element_blank(), y = element_blank())

perm_17_19_global

ggsave(plot = perm_17_19_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_17_19_global.png")), dpi = 600,
       height = 5, width = 7)

```

```{r}

## get the selected users for this, top 20 players

path_id <- 1  ## for 1700-1900

data_path <- all_data_path[path_id]

files <- list.files(data_path)

lichess_data <- files %>% 
  map_dfr(~read_player(data_path, .x))

bullet_data <- lichess_data |> 
  mutate(Event = tolower(Event)) |>
  filter(TimeControl == "60+0") %>%
  filter(Variant == "Standard") %>%
  filter(grepl("rated bullet game", Event))

select_users <- bullet_data %>% 
  group_by(Username) %>% 
  tally() %>% 
  arrange(-n) %>% 
  ## will change this to n = 20 when run on cluster potentially
  slice_max(order = n, n = 20) %>% 
  pull(Username)

## then repeat for the winner effects
player_id <- as_tibble(select_users) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

player_pars <- sim_pars |> 
  as_draws_df() |> 
  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
         id = stringr::str_extract(variable, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  drop_na() |> 
  select(param, id, mean, sim_id) |> 
  mutate(id = as.numeric(id)) |> 
  left_join(player_id, by = "id")

perm_players <- unique(player_pars$player_name)

true_beta <- fit_17_19_bullet_prev$summary() |> 
  filter(str_starts(variable, "beta\\[")) |> 
  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
         id = stringr::str_extract(variable, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  select(param, id, mean) |> 
  mutate(id = as.numeric(id), param = as.numeric(param)) |> 
  left_join(player_id, by = "id") |> 
  filter(player_name %in% perm_players)

perm_17_19_winner <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(mean)) +
  geom_histogram() +
  facet_wrap(~player_name, ncol = 5, scales = "free") +
  labs(x = element_blank(), y = element_blank()) +
  geom_vline(data = true_beta |> filter(param == 2), aes(xintercept = mean),
             col = "red") +
  theme_5col()

perm_17_19_winner

ggsave(plot = perm_17_19_winner, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_17_19_winner.png")), dpi = 600,
       height = 5, width = 7)

```

## GMs

```{r}
## just do ppc plots for now

path_id <- 4
data_path <- res_data_path[path_id]

fit_gm_first <- readRDS(file =  here(data_path, "time",
                                        "select_users_first_bullet.RDS"))
fit_gm_last <- readRDS(file =  here(data_path, "time",
                                        "select_users_last_bullet.RDS"))

fit_samples <- as_draws_df(fit_gm_first$draws())

fit_samp <- fit_samples %>% 
  select(!starts_with(c("log_lik", "yrep")))

y_rep <- fit_samples %>% select(starts_with("y_rep"))

rm(fit_samples)
rm(fit_samp)

stan_data_ave_first <- readRDS(file = here(data_path, "time",
                                           "stan_data_first.RDS"))

y_rep_mod <- y_rep %>% 
  rownames_to_column() %>%  
  pivot_longer(-rowname) %>% 
  pivot_wider(names_from=rowname, values_from=value) %>% 
  ## remove the draws and chains here
  mutate(focal_id = stan_data_ave_first$id)

games_won <- y_rep_mod %>% 
  pivot_longer(cols = `1`:`4000`, names_to = "draw", values_to = "y") %>% 
  group_by(focal_id, draw) %>% 
  summarise(games_won = sum(y))

orig_data <- tibble(outcome = stan_data_ave_first$y,
                    focal_id = stan_data_ave_first$id)

users <- readRDS(file = here(data_path, "time", "select_users_bullet.RDS"))

## need to give these the correct names here
player_id <- tibble(id = as.character(1:10), player = users)

orig_games_won <- orig_data %>% 
  group_by(focal_id) %>% 
  summarise(games_won = sum(outcome)) %>% 
  mutate(focal_id = as.character(focal_id)) %>% 
  left_join(player_id, by = c("focal_id" = "id"))

p1 <- games_won %>% 
  mutate(focal_id = as.character(focal_id)) %>% 
  left_join(player_id, by = c("focal_id" = "id")) %>% 
  ggplot(aes(x = games_won)) +
  geom_histogram() +
  facet_wrap(~player, scales = "free", ncol = 5) +
  geom_vline(data = orig_games_won, 
             mapping = aes(xintercept = games_won), col = "red") +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  # labs(title = "Posterior Predictive Distribution, Initial Games",
  labs(y = element_blank(),
       x = "Number of Games Won") +
  theme_5col()

p1

ggsave(plot = p1, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_ppc_1.png")), dpi = 600,
       height = 5, width = 7)

draws <- fit_gm_first$draws() %>% as_draws_df() %>%
  select(starts_with(c("beta[", "gamma"))) %>%
  mutate(draw = row_number()) %>% 
  pivot_longer(cols = "beta[1,1]":"beta[2,10]") %>% 
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) %>% 
  select(-name) %>% 
  pivot_wider(names_from = param, values_from = value) %>% 
  rename(beta1 = `1`, beta2 = `2`)

stan_data_ave_last <- readRDS(file = here(data_path, "time",
                                          "stan_data_last.RDS"))

future_games <- tibble(focal = as.character(stan_data_ave_last$id), 
       color = stan_data_ave_last$colour, 
       elo_diff = stan_data_ave_last$elo,
       hist = stan_data_ave_last$win_prop)

ppc_data <- readRDS(file = here(data_path, "time",
                                "ppc_fit_first_last.RDS"))

future_results <- tibble(focal = as.character(stan_data_ave_last$id),
                         result = stan_data_ave_last$y) %>% 
  group_by(focal) %>%
  summarise(total = sum(result)) %>% 
  rename(id = focal) %>% 
  left_join(player_id, by = "id")

last_ppc <- ppc_data %>%  
  left_join(player_id, by = "id") %>% 
  ggplot(aes(x = games_won)) + 
  geom_histogram() + 
  geom_vline(data = future_results, 
             mapping = aes(xintercept = total), col = "red") +
  facet_wrap(~player, scales = "free_x", ncol = 5) +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  labs(y = element_blank(), x = "Number of Games Won") +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  theme(axis.text = element_text(size = axis_text_size - 2),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.spacing.x = unit(0.5, "lines"),
        plot.title = element_text(size = title_size),
        axis.title = element_text(size = axis_title),
        legend.text = element_text(size = legend_text),
        strip.text = element_text(size = axis_title - 6))

last_ppc

ggsave(plot = last_ppc, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_ppc_2.png")), dpi = 600,
       height = 5, width = 7)

```

What about for the other two groups?

__To be added in__

![Time-Varying PPC for 2000-2200](../results/lichess2000-2200_mar/time/ppc_fit_first_last_bullet.png)

![Time-Varying PPC for 2300-2500](../results/lichess2300-2500_mar/time/ppc_fit_first_last_bullet.png)

Maybe some very limited evidence that this becomes less
reasonable for players with a higher rating.

### Permuting for GM Data

```{r}
perm_data_path <- here("results/lichessGrandmasters_mar/perm/")

## read all the files in this directory
perm_files <- list.files(perm_data_path)

perm_gm <- perm_files[1:1000] %>% 
  map_dfr(~read_rds(here(perm_data_path, .x)))

sim_pars <- perm_gm |> 
  select(variable, mean, sim_id)

## then need to overlay the true values also
true_values <- fit_gm_bullet_prev$summary() |> 
  filter(variable %in% c("mu_beta", "gamma1", "gamma2")) |> 
  select(variable, mean)

perm_gm_global <- sim_pars |> 
  filter(variable %in% c("mu_beta",  "gamma1", "gamma2",
                           "sigma_1", "tau[1]", "tau[2]",
                           "sigma_g1", "sigma_g2")) |> 
  ggplot(aes(mean)) +
  geom_histogram() +
  facet_wrap(~variable, scales = "free") +
  geom_vline(data = true_values, mapping = aes(xintercept = mean), col = "red") +
  theme_3col() +
  labs(x = element_blank(), y = element_blank())

ggsave(plot = perm_gm_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_gm_global.png")), dpi = 600,
       height = 5, width = 7)

## then repeat for the winner effects
player_id <- as_tibble(bullet_gm) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

player_pars <- sim_pars |> 
  as_draws_df() |> 
  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
         id = stringr::str_extract(variable, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  drop_na() |> 
  select(param, id, mean, sim_id) |> 
  mutate(id = as.numeric(id)) |> 
  left_join(player_id, by = "id")

perm_players <- unique(player_pars$player_name)

true_beta <- fit_gm_bullet_prev$summary() |> 
  filter(str_starts(variable, "beta\\[")) |> 
  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
         id = stringr::str_extract(variable, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  select(param, id, mean) |> 
  mutate(id = as.numeric(id), param = as.numeric(param)) |> 
  left_join(player_id, by = "id") |> 
  filter(player_name %in% perm_players)

perm_gm_winner <- player_pars |> 
  filter(param == 2) |> 
  ggplot(aes(mean)) +
  geom_histogram() +
  facet_wrap(~player_name, ncol = 5, scales = "free") +
  labs(x = element_blank(), y = element_blank()) +
  geom_vline(data = true_beta |> filter(param == 2), aes(xintercept = mean),
             col = "red") +
  theme_5col()

perm_gm_winner

ggsave(plot = perm_gm_winner, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_gm_winner.png")), dpi = 600,
       height = 5, width = 7)

perm_gm_player <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(mean)) +
  geom_histogram() +
  facet_wrap(~player_name, ncol = 5, scales = "free") +
  labs(x = element_blank(), y = element_blank()) +
  geom_vline(data = true_beta |> filter(param == 2), aes(xintercept = mean),
             col = "red") +
  theme_5col()

perm_gm_player

ggsave(plot = perm_gm_player, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_gm_player.png")), dpi = 600,
       height = 5, width = 7)

## should make this 12 players so it looks a bit better

```

### 2024-07-04 analysis/plot_templates.R

@@ -1,0 +1,117 @@
+axis_title <- 16
+title_size <- 18
+axis_text_size <- 14
+legend_text <- 14
+line_size <- 1
+
+
+theme_3col <- function(){ 
+  theme_bw() %+replace%    #replace elements we want to change
+    
+    theme(
+      
+      #grid elements
+      panel.spacing.x = unit(1, "lines"),
+      
+      #text elements
+      plot.title = element_text(size = title_size),
+      axis.text = element_text(size = axis_text_size - 1.5),
+      axis.title = element_text(size = axis_title),
+      legend.text = element_text(size = legend_text),
+      strip.text = element_text(size = axis_title - 4, 
+                                margin = margin(1.5, 1.5, 2, 1.5)),
+      axis.text.y = element_blank(),
+      axis.ticks.y = element_blank(),
+      legend.position = "none"
+    )
+}
+
+
+theme_4col <- function(){ 
+  theme_bw() %+replace%    #replace elements we want to change
+    
+    theme(
+      
+      #grid elements
+      panel.spacing.x = unit(0.5, "lines"),
+      
+      #text elements
+      plot.title = element_text(size = title_size),
+      axis.text = element_text(size = axis_text_size - 2),
+      axis.title = element_text(size = axis_title),
+      legend.text = element_text(size = legend_text),
+      strip.text = element_text(size = axis_title - 5, 
+                                margin = margin(1.5, 1.5, 2, 1.5)),
+      axis.text.y = element_blank(),
+      axis.ticks.y = element_blank(),
+      legend.position = "none"
+    )
+}
+
+
+theme_5col <- function(){ 
+  theme_bw() %+replace%    #replace elements we want to change
+    
+    theme(
+      
+      #grid elements
+      panel.spacing.x = unit(0.5, "lines"),
+      
+      #text elements
+      plot.title = element_text(size = title_size),
+      axis.text = element_text(size = axis_text_size - 4),
+      axis.title = element_text(size = axis_title),
+      legend.text = element_text(size = legend_text),
+      strip.text = element_text(size = axis_title - 6.5, 
+                                margin = margin(1.5, 1.5, 2, 1.5)),
+      axis.text.y = element_blank(),
+      axis.ticks.y = element_blank(),
+      legend.position = "none"
+    )
+}
+
+
+
+theme_6col <- function(){ 
+  theme_bw() %+replace%    #replace elements we want to change
+    
+    theme(
+      
+      #grid elements
[…]

### 2024-07-10 Paper_WriteUp/Paper_Plots.qmd
adding some simple analyses that are used so that the qmd can reproduce all results in the paper

@@
 
+lichess_data |> 
+  mutate(Event = tolower(Event)) |> 
+  filter(grepl("rated blitz game", Event)) |> 
+  group_by(TimeControl) |> 
+  count() |> 
+  arrange(-n) |> 
+  ungroup() |> 
+  mutate(prop = n/sum(n))
+
 bullet_data <- lichess_data |> 
@@
   arrange(n)
+
+blitz_data <- lichess_data |> 
+  mutate(Event = tolower(Event)) |>
+  filter(TimeControl == "180+0") %>%
+  filter(Variant == "Standard") %>%
+  filter(grepl("rated blitz game", Event))
 
@@
 
+lichess_data_gm |> 
+  mutate(Event = tolower(Event)) |> 
+  filter(grepl("rated blitz game", Event)) |> 
+  group_by(TimeControl) |> 
+  count() |> 
+  arrange(-n) |> 
+  ungroup() |> 
+  mutate(prop = n/sum(n))
+
 bullet_data_gm <- lichess_data_gm |> 
@@
   count() |> 
-  arrange(n) |> 
   ungroup() |> 
   summarise(mean(n))
+
+blitz_data_gm <- lichess_data_gm |> 
+  mutate(Event = tolower(Event)) |>
+  filter(TimeControl == "180+0") %>%
+  filter(Variant == "Standard") %>%
+  filter(grepl("rated blitz game", Event))
+
+blitz_data_gm |> 
+  group_by(Username) |> 
+  count() |> 
+  ungroup() |> 
+  summarise(mean(n))
+
+## get the average rating of the GMs
+
+lichess_data_gm |> 
+  mutate(Event = tolower(Event)) |> 
+  filter(grepl("rated blitz game", Event)) |> 
+  group_by(Username) |> 
+  arrange(UTCDate, UTCTime) |> 
+  slice_tail(n = 1) |> 
+  mutate(user_elo = ifelse(Username == White, WhiteElo, BlackElo)) |> 
+  select(Username, White, Black, WhiteElo, BlackElo, user_elo) |>
+  ungroup() |> 
+  summarise(mean(user_elo))
 
@@
                               "gm_comp_global.png")), dpi = 600,
-       height = 5, width = 7)
+       height = 4, width = 7)
 
@@
                               "bul_blitz_comp_gm_global.png")), dpi = 600,
-       height = 5, width = 7)
+       height = 4, width = 7)
 

### 2024-07-10 analysis/helper.R

@@ -1,3 +1,8 @@
+#### July 2024 
+## Collection of helper functions for processing data
+##
+
+
 read_player <- function(path, file){
   dat <- read_csv(file = paste0(path, file),
                   col_types = cols(UTCDate = col_date("%Y.%m.%d"),

### 2024-07-15 Paper_WriteUp/Paper_Plots.qmd
add some more details looking at the proportion of games started within 3 minutes for bullet and 7 minutes for blitz

@@
 
+ bullet_data |>  group_by(Username) |> 
+  select(Username, UTCDate, UTCTime) |> 
+  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
+  group_by(Username) |> 
+  arrange(date_time) |> 
+  mutate(gap = date_time - lag(date_time),
+         gap_seconds = as.numeric(gap)) |> 
+  mutate(less_3 = ifelse(gap_seconds <= 180, TRUE, FALSE)) |> 
+  ungroup() |>   ## to get the overall proportion
+  summarise(sum(less_3, na.rm = TRUE)/ n()) |> 
+  print(n = 25)
+
 blitz_data <- lichess_data |> 
@@
   summarise(mean(user_elo))
+
+## look at the proportion of games started within 3 minutes 
+## of the previous game
+bullet_data_gm |> 
+  # group_by(Username) |> 
+  select(Username, UTCDate, UTCTime) |> 
+  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
+  group_by(Username) |> 
+  arrange(date_time) |> 
+  mutate(gap = date_time - lag(date_time),
+         gap_seconds = as.numeric(gap)) |> 
+  mutate(less_3 = ifelse(gap_seconds <= 180, TRUE, FALSE)) |> 
+  ungroup() |>   ## to get the overall proportion
+  summarise(sum(less_3, na.rm = TRUE)/ n()) |> 
+  print(n = 25)
+
+## look at the streaks, number of games played in a session
+outcomes <- bullet_data_gm |> 
+  # group_by(Username) |> 
+  select(Username, UTCDate, UTCTime) |> 
+  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
+  group_by(Username) |> 
+  arrange(date_time) |> 
+  mutate(gap = date_time - lag(date_time),
+         gap_seconds = as.numeric(gap)) |> 
+  mutate(less_3 = if_else(gap_seconds <= 180, TRUE, FALSE, FALSE))
+
+get_true_sequences <- function(x) {
+  rle_x <- rle(x)
+  true_lengths <- rle_x$lengths[rle_x$values == TRUE]
+  return(true_lengths)
+}
+
+# Apply function to each group
+result <- outcomes %>%
+  # group_by(group) %>%
+  summarise(session_length = list(get_true_sequences(less_3)))
+
+result |> 
+  unnest_longer(session_length) |> 
+  ggplot(aes(session_length)) +
+  geom_histogram()
+
+result |> 
+  unnest_longer(session_length) |> 
+  mutate(multiple = if_else(session_length > 1, TRUE, FALSE)) |> 
+  # group_by(Username) |> 
+  summarise(sum(multiple)/n())
+
+outcomes_blitz <- blitz_data_gm |> 
+  # group_by(Username) |> 
+  select(Username, UTCDate, UTCTime) |> 
+  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
+  group_by(Username) |> 
+  arrange(date_time) |> 
+  mutate(gap = date_time - lag(date_time),
+         gap_seconds = as.numeric(gap)) |> 
+  mutate(less_7_min = if_else(gap_seconds <= 420, TRUE, FALSE, FALSE))
+
+# Apply function to each group
+result_blitz <- outcomes_blitz %>%
+  # group_by(group) %>%
+  summarise(session_length = list(get_true_sequences(less_7_min)))
+
+result_blitz |> 
+  unnest_longer(session_length) |> 
+  ggplot(aes(session_length)) +
+  geom_histogram()
+
+result_blitz |> 
+  unnest_longer(session_length) |> 
+  mutate(multiple = if_else(session_length > 1, TRUE, FALSE)) |> 
+  # group_by(Username) |> 
+  summarise(sum(multiple)/n())
 

### 2024-07-22 Paper_WriteUp/Paper_Plots.qmd
adding in some more stuff

@@
 
+```
+
+## 2000-2200 Previous Game
+
+```{r}
+
+data_path <- here("results/lichess2000-2200_mar")
+
+fit_20_22_bullet_prev <- 
+  readRDS(file =paste0(data_path,"/all_rated_bullet_model_prev.RDS"))
+
+fit_20_22_bullet_prev$summary(variables = c("mu_beta",
+                                            "gamma1", "gamma2",
+                                            "sigma_1", "tau[1]", "tau[2]",
+                                            "sigma_g1", "sigma_g2"))
+```
+
+## 2300-2500 Previous Game
+
+```{r}
+
+data_path <- here("results/lichess2300-2500_mar")
+
+fit_23_25_bullet_prev <- 
+  readRDS(file =paste0(data_path,"/all_rated_bullet_model_prev.RDS"))
+
+fit_23_25_bullet_prev$summary(variables = c("mu_beta",
+                                            "gamma1", "gamma2",
+                                            "sigma_1", "tau[1]", "tau[2]",
+                                            "sigma_g1", "sigma_g2"))
 ```

### 2024-07-22 Paper_WriteUp/Paper_Plots.qmd
tidy up the data format, delete old incorrect results and update paths to reflect this

@@
 
-This will provide the final figures used in the paper.
+This document will recreate all numeric and graphical
+summaries contained in the submitted manuscript and the 
+supplementary materials.
 
@@
 res_data_path <- rep(NA, 4)
-res_data_path[1] <- here("results/lichess1700-1900_mar/")
-res_data_path[2] <- here("results/lichess2000-2200_mar/")
-res_data_path[3] <- here("results/lichess2300-2500_mar/")
-res_data_path[4] <- here("results/lichessGrandmasters_mar/")
+res_data_path[1] <- here("results/lichess1700-1900/")
+res_data_path[2] <- here("results/lichess2000-2200/")
+res_data_path[3] <- here("results/lichess2300-2500/")
+res_data_path[4] <- here("results/lichessGrandmasters/")
 
@@
 
-data_path <- here("results/lichess1700-1900_mar")
+data_path <- here("results/lichess1700-1900")
 
@@
 
-data_path <- here("results/lichess2000-2200_mar")
+data_path <- here("results/lichess2000-2200")
 
@@
 
-data_path <- here("results/lichess2300-2500_mar")
+data_path <- here("results/lichess2300-2500")
 
@@
 ```{r}
-data_path <- here("results/lichessGrandmasters_mar")
+data_path <- here("results/lichessGrandmasters")
 
@@
 
-data_path <- here("results/lichessGrandmasters_mar")
+data_path <- here("results/lichessGrandmasters")
 
@@
 
-perm_data_path <- here("results/lichess1700-1900_mar/perm")
+perm_data_path <- here("results/lichess1700-1900/perm")
 
@@
 
-![Time-Varying PPC for 2000-2200](../results/lichess2000-2200_mar/time/ppc_fit_first_last_bullet.png)
-
-![Time-Varying PPC for 2300-2500](../results/lichess2300-2500_mar/time/ppc_fit_first_last_bullet.png)
+![Time-Varying PPC for 2000-2200](../results/lichess2000-2200/time/ppc_fit_first_last_bullet.png)
+
+![Time-Varying PPC for 2300-2500](../results/lichess2300-2500/time/ppc_fit_first_last_bullet.png)
 
@@
 ```{r}
-perm_data_path <- here("results/lichessGrandmasters_mar/perm/")
+perm_data_path <- here("results/lichessGrandmasters/perm/")
 

### 2024-07-22 owen/cluster_scripts/Large_Fit.R

@@ -32,14 +32,14 @@ all_data_path[1] <- here("box_data/lichess1700-1900/")
 all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_data_path[4] <- here("box_data/lichessGrandmasters/")
-all_save_path[1] <- here("results/lichess1700-1900_mar/")
+all_save_path[1] <- here("results/lichess1700-1900/")
 
 # all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
 # ## if need to run it locally
 
-all_save_path[2] <- here("results/lichess2000-2200_mar/")
-all_save_path[3] <- here("results/lichess2300-2500_mar/")
-all_save_path[4] <- here("results/lichessGrandmasters_mar/")
+all_save_path[2] <- here("results/lichess2000-2200/")
+all_save_path[3] <- here("results/lichess2300-2500/")
+all_save_path[4] <- here("results/lichessGrandmasters/")
 
 
 data_path <- all_data_path[path_id]

### 2024-07-22 owen/cluster_scripts/Large_Single_Prev.R

@@ -32,10 +32,10 @@ all_data_path[1] <- here("box_data/lichess1700-1900/")
 all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_data_path[4] <- here("box_data/lichessGrandmasters/")
-all_save_path[1] <- here("results/lichess1700-1900_mar/")
-all_save_path[2] <- here("results/lichess2000-2200_mar/")
-all_save_path[3] <- here("results/lichess2300-2500_mar/")
-all_save_path[4] <- here("results/lichessGrandmasters_mar/")
+all_save_path[1] <- here("results/lichess1700-1900/")
+all_save_path[2] <- here("results/lichess2000-2200/")
+all_save_path[3] <- here("results/lichess2300-2500/")
+all_save_path[4] <- here("results/lichessGrandmasters/")
 
 
 data_path <- all_data_path[path_id]

### 2024-07-22 owen/cluster_scripts/Winner_Effects_Time.R

@@ -33,14 +33,14 @@ all_data_path[1] <- here("box_data/lichess1700-1900/")
 all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_data_path[4] <- here("box_data/lichessGrandmasters/")
-all_save_path[1] <- here("results/lichess1700-1900_mar/")
+all_save_path[1] <- here("results/lichess1700-1900/")
 
 # all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
 # ## if need to run it locally
 
-all_save_path[2] <- here("results/lichess2000-2200_mar/")
-all_save_path[3] <- here("results/lichess2300-2500_mar/")
-all_save_path[4] <- here("results/lichessGrandmasters_mar/")
+all_save_path[2] <- here("results/lichess2000-2200/")
+all_save_path[3] <- here("results/lichess2300-2500/")
+all_save_path[4] <- here("results/lichessGrandmasters/")
 
 
 data_path <- all_data_path[path_id]

### 2024-07-22 owen/cluster_scripts/perm_results.R

@@ -31,14 +31,14 @@ all_data_path[1] <- here("box_data/lichess1700-1900/")
 all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_data_path[4] <- here("box_data/lichessGrandmasters/")
-all_save_path[1] <- here("results/lichess1700-1900_mar/")
+all_save_path[1] <- here("results/lichess1700-1900/")
 
 # all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
 # ## if need to run it locally
 
-all_save_path[2] <- here("results/lichess2000-2200_mar/")
-all_save_path[3] <- here("results/lichess2300-2500_mar/")
-all_save_path[4] <- here("results/lichessGrandmasters_mar/")
+all_save_path[2] <- here("results/lichess2000-2200/")
+all_save_path[3] <- here("results/lichess2300-2500/")
+all_save_path[4] <- here("results/lichessGrandmasters/")
 
 path_id <- 1 ## setting it for now
 data_path <- all_data_path[path_id]

### 2024-07-23 Paper_WriteUp/Paper_Plots.qmd
updates on the plots and the summary, add a small helper function

@@
 
-# Summary Statistics
-
-First get some summary statistics about the overall data. 
-Construct these only for the 1700-1900 players and GMs initially.
-
-```{r}
+# Section 2: Data
+
+We first get some summary statistics about the overall data. 
+We construct these for the 1700-1900 players and GMs.
+We then examine the number of consecutive games played 
+across both cohorts.
+
+```{r summary statistics 1700-1900}
 
@@
 
+## overall summary of all event types
 lichess_data |> 
@@
 
+## what proportion of bullet games are of what time format
 lichess_data |> 
@@
 
+## repeat for blitz games instead
 lichess_data |> 
@@
 
+## average number, we remove the two accounts with 1 and 3 games only
+## throughout we do not consider those who play less than 10 games
+## at a given format when fitting any model
 bullet_data |> 
@@
 
+## for each player, count how many times they start
+## a game with 3 minutes of the start of a previous game
+## and get the proportion across all focal players
  bullet_data |>  
    group_by(Username) |> 
-  count() |> 
-  arrange(n)
-
- bullet_data |>  group_by(Username) |> 
    select(Username, UTCDate, UTCTime) |> 
@@
    ungroup() |>   ## to get the overall proportion
-  summarise(sum(less_3, na.rm = TRUE)/ n()) |> 
-  print(n = 25)
+   summarise(sum(less_3, na.rm = TRUE)/ n())
 
@@
 
-## then repeat this sort of analysis for the gms
-
+## repeat for blitz data using 7 minutes as the maximum gap
+ blitz_data |>  
+   group_by(Username) |> 
+   select(Username, UTCDate, UTCTime) |> 
+   mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
+   group_by(Username) |> 
+   arrange(date_time) |> 
+   mutate(gap = date_time - lag(date_time),
+         gap_seconds = as.numeric(gap)) |> 
+   mutate(less_7 = ifelse(gap_seconds <= 420, TRUE, FALSE)) |> 
+   ungroup() |>   ## to get the overall proportion
+   summarise(sum(less_7, na.rm = TRUE)/ n())
+
+## examine how many games are played in a session for the bullet data
+## look at the streaks, number of games played in a session
+outcomes <- bullet_data |> 
+  # group_by(Username) |> 
+  select(Username, UTCDate, UTCTime) |> 
+  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
+  group_by(Username) |> 
+  arrange(date_time) |> 
+  mutate(gap = date_time - lag(date_time),
+         gap_seconds = as.numeric(gap)) |> 
+  mutate(less_3 = if_else(gap_seconds <= 180, TRUE, FALSE, missing = FALSE))
+
+# Apply function to each group
+result <- outcomes |> 
+  summarise(session_length = list(get_true_sequences(less_3)))
+
+result |> 
+  unnest_longer(session_length) |> 
+  summarise(med_session = median(session_length), 
+            avg = mean(session_length))
+
+## plot the histogram of games per session
+result |> 
+  unnest_longer(session_length) |> 
+  ggplot(aes(session_length)) +
+  geom_histogram()
+
+## proportion of sessions with more than 1 game
+result |> 
+  unnest_longer(session_length) |> 
+  mutate(multiple = if_else(session_length > 1, TRUE, FALSE)) |>  
+  summarise(sum(multiple)/n())
+
+```
+
+```{r summary statistics gm}
+## then repeat this analysis for the gms
 path_id <- 4  ## for GMs
-
 data_path <- all_data_path[path_id]
-
 files <- list.files(data_path)
@@
 
+## get proportion of bullet games that are of each time format
 lichess_data_gm |> 
@@
 
+## repeat for blitz
 lichess_data_gm |> 
@@
 
-bullet_data_gm |> 
-  group_by(Username) |> 
-  count() |> 
-  arrange(n)
-
+## average across all focal players
 bullet_data_gm |> 
@@
 
-## get the average rating of the GMs
-
+## get the average rating of the GMs at the time data
+## was collected 
 lichess_data_gm |> 
@@
 bullet_data_gm |> 
-  # group_by(Username) |> 
   select(Username, UTCDate, UTCTime) |> 
@@
   ungroup() |>   ## to get the overall proportion
-  summarise(sum(less_3, na.rm = TRUE)/ n()) |> 
-  print(n = 25)
+  summarise(sum(less_3, na.rm = TRUE)/ n())
 
 ## look at the streaks, number of games played in a session
-outcomes <- bullet_data_gm |> 
+outcomes_gm <- bullet_data_gm |> 
   # group_by(Username) |> 
@@
          gap_seconds = as.numeric(gap)) |> 
-  mutate(less_3 = if_else(gap_seconds <= 180, TRUE, FALSE, FALSE))
-
+  mutate(less_3 = if_else(gap_seconds <= 180, TRUE, FALSE, missing = FALSE))
+
+## extract the sessions containing more than 1 game
 get_true_sequences <- function(x) {
@@
 # Apply function to each group
-result <- outcomes %>%
-  # group_by(group) %>%
+result_gm <- outcomes_gm |> 
   summarise(session_length = list(get_true_sequences(less_3)))
 
-result |> 
+result_gm |> 
+  unnest_longer(session_length) |> 
+  summarise(med_session = median(session_length), 
+            avg = mean(session_length))
+
+## plot the histogram of games per session
+result_gm |> 
   unnest_longer(session_length) |> 
@@
 
+## proportion of sessions with more than 1 game
 result |> 
@@
   mutate(multiple = if_else(session_length > 1, TRUE, FALSE)) |>  
-  # group_by(Username) |> 
   summarise(sum(multiple)/n())
 
+result |> 
+  unnest_longer(session_length) |> 
+  mutate(multiple = if_else(session_length > 10, TRUE, FALSE)) |>  
+  summarise(sum(multiple)/n())
+
+## repeat for blitz
 outcomes_blitz <- blitz_data_gm |> 
-  # group_by(Username) |> 
   select(Username, UTCDate, UTCTime) |> 
@@
 result_blitz <- outcomes_blitz %>%
-  # group_by(group) %>%
   summarise(session_length = list(get_true_sequences(less_7_min)))
@@
 
-# Examining Model Fit
+# Section 4: The Fitted Model
 
@@
 
-```{r}
-
+```{r plot global parameters}
 data_path <- here("results/lichess1700-1900")
@@
        height = 5, width = 7)
+
+```
+
+```{r interpret the estimated coefficients 17-19}
+
+gamma_ests <- fit_17_19_bullet_prev$summary(variables = c("gamma1",
+                                                          "gamma2")) |> 
+  pull(mean)
+## compute change in win probability for gamma1
+1/(1 + exp(-1 * (0)))
+1/(1 + exp(-1 * (gamma_ests[1])))
+
+1/(1 + exp(-1 * (0))) ## equally rated, playing as black
+1/(1 + exp(-1 * (100 * gamma_ests[2])))
 
@@
 
-## 2000-2200 Previous Game
-
-```{r}
-
-data_path <- here("results/lichess2000-2200")
-
-fit_20_22_bullet_prev <- 
-  readRDS(file =paste0(data_path,"/all_rated_bullet_model_prev.RDS"))
-
-fit_20_22_bullet_prev$summary(variables = c("mu_beta",
-                                            "gamma1", "gamma2",
-                                            "sigma_1", "tau[1]", "tau[2]",
-                                            "sigma_g1", "sigma_g2"))
-```
-
-## 2300-2500 Previous Game
-
-```{r}
-
-data_path <- here("results/lichess2300-2500")
-
-fit_23_25_bullet_prev <- 
-  readRDS(file =paste0(data_path,"/all_rated_bullet_model_prev.RDS"))
-
-fit_23_25_bullet_prev$summary(variables = c("mu_beta",
-                                            "gamma1", "gamma2",
-                                            "sigma_1", "tau[1]", "tau[2]",
-                                            "sigma_g1", "sigma_g2"))
-```
-
 ## GM previous game
@@
   theme_4col() +
-  scale_x_continuous(breaks = breaks_pretty(n = 3))
+  scale_x_continuous(breaks = breaks_pretty(n = 2))
 
@@
        height = 5, width = 7)
+
+```
+
+```{r interpret the estimated coefficients gm}
+
+gamma_ests_gm <- fit_gm_bullet_prev$summary(variables = c("gamma1",
+                                                          "gamma2")) |> 
+  pull(mean)
+
+## compute change in win probability for gamma1
+1/(1 + exp(-1 * (0)))
+1/(1 + exp(-1 * (gamma_ests_gm[1])))
+
+## change in win probability for elo difference
+
+1/(1 + exp(-1 * (0))) ## equally rated, playing as black
+1/(1 + exp(-1 * (100 * gamma_ests_gm[2])))
 
@@
 
-# Model Checking
+# Section 5: Model Validation
 
@@
 ```
+
+# Supplementary Material
+
+This contains all analysis and figures which are included in 
+the supplementary materials.
+
+## Additional Summary Information
+
+```{r}
+
+## how many times does a focal player play another focal player
+
+focal_17_19 <- unique(bullet_data$Username)
+
+bullet_data |> 
+  mutate(opponent = ifelse(Username == White, Black, White)) |> 
+  mutate(focal_opp = ifelse(opponent %in% focal_17_19, 1, 0)) |> 
+  select(Username, focal_opp) |> 
+  summarise(sum(focal_opp)/ n())
+
+focal_gm <- unique(bullet_data_gm$Username)
+
+bullet_data_gm |> 
+  mutate(opponent = ifelse(Username == White, Black, White)) |> 
+  mutate(focal_opp = ifelse(opponent %in% focal_gm, 1, 0)) |> 
+  select(Username, focal_opp) |> 
+  summarise(sum(focal_opp)/ n())
+
+```
+
+## Results for Other Cohorts
+
+## 2000-2200 Previous Game
+
+```{r}
+
+data_path <- here("results/lichess2000-2200")
+
+fit_20_22_bullet_prev <- 
+  readRDS(file =paste0(data_path,"/all_rated_bullet_model_prev.RDS"))
+
+fit_20_22_bullet_prev$summary(variables = c("mu_beta",
+                                            "gamma1", "gamma2",
+                                            "sigma_1", "tau[1]", "tau[2]",
+                                            "sigma_g1", "sigma_g2"))
+```
+
+## 2300-2500 Previous Game
+
+```{r}
+
+data_path <- here("results/lichess2300-2500")
+
+fit_23_25_bullet_prev <- 
+  readRDS(file =paste0(data_path,"/all_rated_bullet_model_prev.RDS"))
+
+fit_23_25_bullet_prev$summary(variables = c("mu_beta",
+                                            "gamma1", "gamma2",
+                                            "sigma_1", "tau[1]", "tau[2]",
+                                            "sigma_g1", "sigma_g2"))
+```

### 2024-07-23 analysis/helper.R

@@ -53,6 +53,13 @@ get_hist <- function(user, games, prev_n) {
   hist_games
 }
 
+## extract the sessions containing more than 1 game
+get_true_sequences <- function(x) {
+  rle_x <- rle(x)
+  true_lengths <- rle_x$lengths[rle_x$values == TRUE]
+  return(true_lengths)
+}
+
 
 
 ## plotting defaults

### 2024-07-23 owen/cluster_scripts/Large_Single_Prev.R
modify the stan model to use n(0,1) priors for the scale components, to check the model fit

@@ -32,10 +32,10 @@ all_data_path[1] <- here("box_data/lichess1700-1900/")
 all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_data_path[4] <- here("box_data/lichessGrandmasters/")
-all_save_path[1] <- here("results/lichess1700-1900/")
-all_save_path[2] <- here("results/lichess2000-2200/")
-all_save_path[3] <- here("results/lichess2300-2500/")
-all_save_path[4] <- here("results/lichessGrandmasters/")
+all_save_path[1] <- here("results/lichess1700-1900_test/")
+all_save_path[2] <- here("results/lichess2000-2200_test/")
+all_save_path[3] <- here("results/lichess2300-2500_test/")
+all_save_path[4] <- here("results/lichessGrandmasters_test/")
 
 
 data_path <- all_data_path[path_id]
@@ -59,18 +59,18 @@ lichess_data <- files %>%
 ## this also removes the NAs, which makes sense
 
 
-# small_data <- lichess_data %>%
-#   # filter(Event == "Rated Bullet game") %>%
-#   mutate(Event = tolower(Event)) |> 
-#   filter(TimeControl == "60+0") %>%
-#   filter(Variant == "Standard") %>%
-#   filter(grepl("rated bullet game", Event))
- 
 small_data <- lichess_data %>%
+  # filter(Event == "Rated Bullet game") %>%
   mutate(Event = tolower(Event)) |>
-  filter(TimeControl == "180+0") %>%
+  filter(TimeControl == "60+0") %>%
   filter(Variant == "Standard") %>%
-  filter(grepl("rated blitz game", Event))
+  filter(grepl("rated bullet game", Event))
+
+# small_data <- lichess_data %>%
+#   mutate(Event = tolower(Event)) |>
+#   filter(TimeControl == "180+0") %>%
+#   filter(Variant == "Standard") %>%
+#   filter(grepl("rated blitz game", Event))
 
 users <- unique(small_data$Username)
 
@@ -82,8 +82,8 @@ users <- small_data %>%
   filter(n >= 10) %>% 
   pull(Username)
 
-# saveRDS(users, file = paste0(save_path, "users_bullet.RDS"))
-saveRDS(users, file = paste0(save_path, "users_blitz.RDS"))
+saveRDS(users, file = paste0(save_path, "users_bullet.RDS"))
[…]

### 2024-07-25 Paper_WriteUp/Paper_Plots.qmd
many updates to the plots to reflect updates to the paper

@@
 
+## with more than 10 games
+result |> 
+  unnest_longer(session_length) |> 
+  mutate(multiple = if_else(session_length > 10, TRUE, FALSE)) |>  
+  summarise(sum(multiple)/n())
+
+## repeat for blitz games
+
+outcomes_blitz <- blitz_data |> 
+  # group_by(Username) |> 
+  select(Username, UTCDate, UTCTime) |> 
+  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
+  group_by(Username) |> 
+  arrange(date_time) |> 
+  mutate(gap = date_time - lag(date_time),
+         gap_seconds = as.numeric(gap)) |> 
+  mutate(less_7 = if_else(gap_seconds <= 420, TRUE, FALSE, missing = FALSE))
+
+result_blitz <- outcomes_blitz |> 
+  summarise(session_length = list(get_true_sequences(less_7)))
+
+result_blitz |> 
+  unnest_longer(session_length) |> 
+  mutate(multiple = if_else(session_length > 1, TRUE, FALSE)) |>  
+  summarise(sum(multiple)/n())
+
 ```
@@
 ## proportion of sessions with more than 1 game
-result |> 
+result_gm |> 
   unnest_longer(session_length) |> 
@@
 
-result |> 
+result_gm |> 
   unnest_longer(session_length) |> 
@@
 
-```{r plot global parameters}
+```{r plot global parameters for 17-19}
 data_path <- here("results/lichess1700-1900")
@@
 
-```{r}
+```{r player level parameters for 17-19}
 
@@
 
+## see what the smallest and largest posterior mean is 
+## for the winner effects
+player_pars |> 
+  group_by(player_name, param) |> 
+  summarise(post_mean = mean(value)) |> 
+  filter(param == 2) |> 
+  ungroup() |> 
+  slice_max(order_by = post_mean, n = 5)
+
+player_pars |> 
+  group_by(player_name, param) |> 
+  summarise(post_mean = mean(value)) |> 
+  filter(param == 2) |> 
+  ungroup() |> 
+  slice_min(order_by = post_mean, n = 5)
+
+## see how many contain 0
+player_pars |> 
+  group_by(player_name, param) |> 
+  filter(param == 2) |> 
+  summarise(lower = quantile(value, prob = 0.025),
+            upper = quantile(value, prob = 0.975)) |> 
+  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
+                         1, 0)) |> 
+  filter(contain_zero == 0)
+
 ## then the winner effects plot
-
 winner_plot_17_19 <- player_pars |> 
@@
 
-```{r}
+```{r global parameters for gm data}
 data_path <- here("results/lichessGrandmasters")
@@
 
-```{r}
-
+```{r player level parameters for gm data}
 player_id <- as_tibble(bullet_gm) |> 
@@
 
+## see what the smallest and largest posterior mean is 
+## for the winner effects
+player_pars |> 
+  group_by(player_name, param) |> 
+  summarise(post_mean = mean(value)) |> 
+  filter(param == 2) |> 
+  ungroup() |> 
+  slice_max(order_by = post_mean, n = 5)
+
+player_pars |> 
+  group_by(player_name, param) |> 
+  summarise(post_mean = mean(value)) |> 
+  filter(param == 2) |> 
+  ungroup() |> 
+  slice_min(order_by = post_mean, n = 5)
+
+## count how many intervals contain 0
+## see how many contain 0
+player_pars |> 
+  group_by(player_name, param) |> 
+  filter(param == 2) |> 
+  summarise(lower = quantile(value, prob = 0.025),
+            upper = quantile(value, prob = 0.975)) |> 
+  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
+                         1, 0)) |> 
+  filter(contain_zero == 0)
+
 ## then the winner effects plot
-
 winner_plot_gm <- player_pars |> 
@@
 
-Here there are only 21 rather than 25 players at the moment, this will be
-updated currently.
+```{r interpret changes in win probability for gm}
+## interpret the individual winner effects
+## assume wins 50% of all games
+
+1/(1 + exp(-1 * (0)))
+
+## if won previous
+1/(1 + exp(-1 * (0 + 0.216 * 0.5)))
+
+## if lost previous
+1/(1 + exp(-1 * (0 + 0.216 * - 0.5)))
+
+## what about if win probability overall was 40%
+1/(1 + exp(-1 * (0 + 0.216 * 0.6)))
+
+## 35%, pretty low
+1/(1 + exp(-1 * (0 + 0.216 * 0.65)))
+
+```
 
@@
 
-```{r}
-
+```{r compare global params n-1 vs n-10}
 ## global effects first
-
 data_path <- here("results/lichessGrandmasters")
@@
                                         "gamma1", "gamma2",
-                           "sigma_1", "tau[1]", "tau[2]",
+                                        "sigma_1", "tau[1]",
+                                        "tau[2]",
                                         "sigma_g1", "sigma_g2")) |> 
@@
 
+## check the proportion of 95% intervals which contain zero
+bind_rows(player_pars_pre, player_pars_pre_10) |> 
+  filter(param == 2) |> 
+  group_by(player_name, model) |> 
+  summarise(lower = quantile(value, prob = 0.025),
+            upper = quantile(value, prob = 0.975)) |> 
+  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
+                         1, 0)) |> 
+  group_by(model) |> 
+  summarise(prop = sum(contain_zero)/ n())
+
+## do the effects change in direction
+bind_rows(player_pars_pre, player_pars_pre_10) |> 
+  filter(param == 2) |> 
+  group_by(player_name, model) |> 
+  summarise(lower = quantile(value, prob = 0.025),
+            upper = quantile(value, prob = 0.975)) |> 
+  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
+                         1, 0)) |>
+  print(n = 50)
+
 ## then player effects
@@
 
+<!-- down to here in terms of updating the code, making sure it's complete -->
+
 ## Bullet vs Blitz for 17-19
@@
 
-```{r}
+```{r bullet v blitz for 17-19}
 
@@
 
-```{r}
+```{r bullet v blitz gm}
 data_path <- here("results/lichessGrandmasters")
@@
 
-```{r}
-
+```{r ppc 17-19}
 ## this the global and winner effect plots
-
 path_id <- 1
@@
 users <- readRDS(file = here(data_path, "time", "select_users_bullet.RDS"))
-
 ## indiv winner effects
@@
   # xlim(c(-1.5, 1.5)) +
-  facet_grid(rows = vars(when), cols = vars(player), scales = "free") +
+  facet_grid(rows = vars(when), cols = vars(player), scales = "free_x") +
   labs(fill = element_blank(), y = element_blank(), 
@@
        x = element_blank()) +
-  theme(axis.ticks.y = element_blank(),
-        axis.text.y = element_blank(),
-        axis.text = element_text(size = axis_text_size - 4),
-        axis.title = element_text(size = axis_title),
-        legend.text = element_text(size = legend_text),
-        strip.text = element_text(size = axis_title - 6),
-        legend.position = "none",
-        plot.title = element_text(size = title_size))
+  theme_10col()
+  # theme(axis.ticks.y = element_blank(),
+  #       axis.text.y = element_blank(),
+  #       axis.text = element_text(size = axis_text_size - 4),
+  #       axis.title = element_text(size = axis_title),
+  #       legend.text = element_text(size = legend_text),
+  #       strip.text = element_text(size = axis_title - 6),
+  #       legend.position = "none",
+  #       plot.title = element_text(size = title_size))
 
@@
 
-```{r}
+```{r ppc1 for 17-19}
 path_id <- 1
@@
        x = "Number of Games Won") +
-  theme_5col()
+  theme_5col() +
+  theme(axis.title.x = element_text(size = 10,
+                                    margin = margin(t = 10, b = 10)))
 
@@
 
-```{r}
+```{r ppc2 for 17-19}
 ## time varying ppc for 1700-1900 players
-
 draws <- fit_17_19_first$draws() %>% as_draws_df() %>%
@@
   scale_x_continuous(breaks = breaks_pretty(n = 3)) +
-  theme_5col()
+  theme_5col() +
+  theme(axis.title.x = element_text(size = 10,
+                                    margin = margin(t = 10, 
+                                                    b = 10,
+                                                    unit = "pt")))
 
@@
 
-This data is complete, just need to create the plots here.
-Here we use the 20 players with the most games only.
-
-```{r}
-
-## these don't look right just now
-
+```{r perm for 17-19}
 perm_data_path <- here("results/lichess1700-1900/perm")
@@
 data_path <- res_data_path[path_id]
-## read all the files in this directory
 (perm_files <- list.files(perm_data_path))
@@
   facet_wrap(~variable, scales = "free") +
-  geom_vline(data = true_values, mapping = aes(xintercept = mean), col = "red") +
+  geom_vline(data = true_values,
+             mapping = aes(xintercept = mean),
+             col = "red") +
   theme_3col() +
@@
        height = 5, width = 7)
-
-```
-
-```{r}
-
+```
+
+```{r perm winner effects}
 ## get the selected users for this, top 20 players
-
 path_id <- 1  ## for 1700-1900
-
 data_path <- all_data_path[path_id]
-
 files <- list.files(data_path)
@@
        height = 5, width = 7)
-
 ```
@@
 
-```{r}
+```{r ppc1 and ppc2 for gm}
 ## just do ppc plots for now
-
 path_id <- 4
 data_path <- res_data_path[path_id]
-
 fit_gm_first <- readRDS(file =  here(data_path, "time",
@@
        x = "Number of Games Won") +
-  theme_5col()
+  theme_5col() +
+  theme(axis.title.x = element_text(size = 10,
+                                   margin = margin(t = 10, b = 10, unit = "pt")),
+        panel.spacing.x = unit(0.7, "lines"))
 
@@
   scale_x_continuous(breaks = breaks_pretty(n = 3)) +
-  theme(axis.text = element_text(size = axis_text_size - 2),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        panel.spacing.x = unit(0.5, "lines"),
-        plot.title = element_text(size = title_size),
-        axis.title = element_text(size = axis_title),
-        legend.text = element_text(size = legend_text),
-        strip.text = element_text(size = axis_title - 6))
+  theme_5col() +
+  theme(axis.title.x = element_text(size = 10,
+                                   margin = margin(t = 10,
+                                                   b = 10,
+                                                   unit = "pt")))
 
@@
        height = 5, width = 7)
-
-```
-
-What about for the other two groups?
-
-__To be added in__
-
-![Time-Varying PPC for 2000-2200](../results/lichess2000-2200/time/ppc_fit_first_last_bullet.png)
-
-![Time-Varying PPC for 2300-2500](../results/lichess2300-2500/time/ppc_fit_first_last_bullet.png)
-
-Maybe some very limited evidence that this becomes less
-reasonable for players with a higher rating.
+```
 
@@
              col = "red") +
-  theme_5col()
+  theme_5col() +
+  scale_x_continuous(breaks = breaks_pretty(n = 3))
 
@@
 
-## should make this 12 players so it looks a bit better
-
+## should we make this 12 players so it looks a bit better?
 ```
@@
 ```
+
+## Model Checking for the Other Cohorts
+
+What about for the other two groups?
+
+__To be added in__
+
+![Time-Varying PPC for 2000-2200](../results/lichess2000-2200/time/ppc_fit_first_last_bullet.png)
+
+![Time-Varying PPC for 2300-2500](../results/lichess2300-2500/time/ppc_fit_first_last_bullet.png)
+
+Maybe some very limited evidence that this becomes less
+reasonable for players with a higher rating.

### 2024-07-25 analysis/plot_templates.R

@@ -115,3 +115,26 @@ theme_8col <- function(){
       legend.position = "none"
     )
 }
+
+
+
+theme_10col <- function(){ 
+  theme_bw() %+replace%    #replace elements we want to change
+    
+    theme(
+      
+      #grid elements
+      panel.spacing.x = unit(0.25, "lines"),
+      
+      #text elements
+      plot.title = element_text(size = title_size),
+      axis.text = element_text(size = axis_text_size - 7),
+      axis.title = element_text(size = axis_title),
+      legend.text = element_text(size = legend_text),
+      strip.text = element_text(size = axis_title - 9, 
+                                margin = margin(1.5, 1.5, 2, 1.5)),
+      axis.text.y = element_blank(),
+      axis.ticks.y = element_blank(),
+      legend.position = "none"
+    )
+}

### 2024-10-15 Paper_WriteUp/Paper_Plots.qmd
working on making these plots even nicer

@@
 library(scales)
+library(ggridges)
 library(here)
@@
        height = 7, width = 7)
+
+```
+
+```{r new boxplots}
+
+player_pars |> 
+  filter(param == 2) |> 
+  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
+  # geom_boxplot() +
+  geom_density_ridges(alpha = 0.5, 
+                      quantiles = 0.5, 
+                      quantile_lines = T, 
+                      rel_min_height = 0.005) + 
+  # geom_density_ridges2(scale = 1, rel_min_height = 0.05) +
+  # coord_flip() +
+  labs(y = "", x = "Estimated W/L-Effect") +
+  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
+  xlim(c(-0.375, 0.375))
+
+player_pars |> 
+  filter(param == 1) |> 
+  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
+  geom_boxplot(outliers = FALSE) +
+  # coord_flip() +
+  labs(y = "", x = "Estimated Player Effect") +
+  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed")
 

### 2024-10-16 Paper_WriteUp/Paper_Plots.qmd
update many of the plots and add a new template. trying to tidy up the plotting code

--- 2024-07-25 Paper_WriteUp/Paper_Plots.qmd
@@
 library(scales)
+library(ggridges)
 library(here)
-
+library(viridis)
 options(mc.cores = parallel::detectCores())
-
 theme_set(theme_bw())
@@
 all_data_path[4] <- here("box_data/lichessGrandmasters/")
-
 res_data_path <- rep(NA, 4)
@@
 
-lichess_data <- files %>% 
+lichess_data <- files |> 
   map_dfr(~read_player(data_path, .x))
@@
   mutate(Event = tolower(Event)) |>
-  filter(TimeControl == "60+0") %>%
-  filter(Variant == "Standard") %>%
+  filter(TimeControl == "60+0") |>
+  filter(Variant == "Standard") |>
   filter(grepl("rated bullet game", Event))
@@
   mutate(Event = tolower(Event)) |>
-  filter(TimeControl == "180+0") %>%
-  filter(Variant == "Standard") %>%
+  filter(TimeControl == "180+0") |>
+  filter(Variant == "Standard") |>
   filter(grepl("rated blitz game", Event))
@@
 
-lichess_data_gm <- files %>% 
+lichess_data_gm <- files |> 
   map_dfr(~read_player(data_path, .x))
@@
 # Apply function to each group
-result_blitz <- outcomes_blitz %>%
+result_blitz <- outcomes_blitz |>
   summarise(session_length = list(get_true_sequences(less_7_min)))
@@
 ## easiest way to make this grayscale would be to not use mcmc_hist at all
+
+facet_labels <- as_labeller(c(
+  mu_beta = "mu[beta]",
+  gamma1 = "gamma[1]",
+  gamma2 = "gamma[2]",
+  sigma_1 = "sigma[1]",
+  `tau[1]` = "tau[1]",
+  `tau[2]` = "tau[2]",
+  sigma_g1 = "sigma[g[1]]",
+  sigma_g2 = "sigma[g[2]]"
+), label_parsed)
 
@@
   geom_histogram() +
-  facet_wrap(~name, ncol = 4, scales = "free") +
+  facet_wrap(~name, ncol = 4, scales = "free",
+             labeller = labeller(name = facet_labels)) +
   labs(x = element_blank(), y = element_blank()) +
@@
 
-player_pars <- fit_17_19_bullet_prev$draws() %>% 
-  as_draws_df() %>%
-  select(starts_with("beta[")) %>%
-  pivot_longer(cols = everything()) %>%
+player_pars <- fit_17_19_bullet_prev$draws() |> 
+  as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
   mutate(param = stringr::str_extract(name, pattern = "\\d"),
@@
 
+```{r new boxplots}
+
+winner_plot_17_19 <- player_pars |> 
+  filter(param == 2) |> 
+  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
+  stat_histinterval() +
+  labs(x = "Estimated W/L-Effect", y = element_blank(), 
+       title = "1700-1900 Cohort") +
+  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
+  theme_single_y()
+
+ggsave(plot = winner_plot_17_19, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "17_19_prev_winner.png")), dpi = 600,
+       height = 7, width = 7)
+
+indiv_plot_17_19 <- player_pars |> 
+  filter(param == 1) |> 
+  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
+  stat_histinterval() +
+  labs(x = "Estimated Player Effect", y = element_blank()) +
+  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
+  theme_single_y()
+
+ggsave(plot = indiv_plot_17_19, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "17_19_prev_indiv.png")), dpi = 600,
+       height = 7, width = 7)
+
+```
+
 ## GM previous game
@@
   geom_histogram() +
-  facet_wrap(~name, ncol = 4, scales = "free") +
+   facet_wrap(~name, ncol = 4, scales = "free",
+             labeller = labeller(name = facet_labels)) +
   labs(x = element_blank(), y = element_blank()) +
@@
 
-player_pars <- fit_gm_bullet_prev$draws() %>% 
-  as_draws_df() %>%
-  select(starts_with("beta[")) %>%
-  pivot_longer(cols = everything()) %>%
+player_pars <- fit_gm_bullet_prev$draws() |> 
+  as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
   mutate(param = stringr::str_extract(name, pattern = "\\d"),
@@
   filter(param == 2) |> 
-  ggplot(aes(value)) +
-  geom_histogram() +
-  facet_wrap(~player_name, scales = "free", ncol = 5) +
-  theme_5col() +
-  labs(x = element_blank(), y = element_blank()) +
-  scale_x_continuous(breaks = breaks_pretty(n = 3))
+  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
+  stat_histinterval() +
+  labs(x = "Estimated W/L-Effect", y = element_blank(), 
+       title = "GM Cohort") +
+  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
+  theme_single_y()
 
@@
   filter(param == 1) |> 
-  ggplot(aes(value)) +
-  geom_histogram() +
-  facet_wrap(~player_name, scales = "free", ncol = 5) +
-  theme_5col() +
-  labs(x = element_blank(), y = element_blank()) +
-  scale_x_continuous(breaks = breaks_pretty(n = 3))
+  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
+  stat_histinterval() +
+  labs(x = "Estimated Player Effect", y = element_blank(), 
+       title = "GM Cohort") +
+  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
+  theme_single_y()
 
@@
 prev_pars <- fit_gm_bullet_prev$draws(c("mu_beta",
-                                        "gamma1", "gamma2",
-                                        "sigma_1", "tau[1]",
-                                        "tau[2]",
-                                        "sigma_g1", "sigma_g2")) |> 
+                                        "gamma1", "gamma2")) |> 
   as_draws_df() |> 
-  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |> 
   mutate(model = "Prev. Game")
@@
 pre_10_pars <- fit_gm_bullet_10$draws(c("mu_beta",
-                                        "gamma1", "gamma2",
-                                        "sigma_1", "tau[1]",
-                                        "tau[2]",
-                                        "sigma_g1", "sigma_g2")) |> 
+                                        "gamma1", "gamma2")) |> 
   as_draws_df() |> 
-  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |> 
   mutate(model = "Prev. 10 Games")
@@
 hist_comp_global <- bind_rows(prev_pars, pre_10_pars) |> 
-  ggplot(aes(value)) +
-  geom_histogram() +
-  facet_grid(rows = vars(model), cols = vars(name), scales = "free") +
-  theme_8col() +
-  labs(x = element_blank(), y = element_blank()) +
-  scale_x_continuous(breaks = breaks_pretty(n = 2))
-
-hist_comp_global
+  ggplot(aes(x = value, color = model)) +
+  stat_pointinterval(position = position_dodge(width = 0.2)) +
+  facet_wrap(~name,
+             scales = "free",
+             labeller = labeller(name = facet_labels)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_3col() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_color_viridis_d(option = "D")
 
@@
                               "gm_comp_global.png")), dpi = 600,
-       height = 4, width = 7)
+       height = 3.5, width = 7)
 
@@
 
-player_pars_pre <- fit_gm_bullet_prev$draws() %>% 
-  as_draws_df() %>%
-  select(starts_with("beta[")) %>%
-  pivot_longer(cols = everything()) %>%
+player_pars_pre <- fit_gm_bullet_prev$draws() |> 
+  as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
   mutate(param = stringr::str_extract(name, pattern = "\\d"),
@@
 
-player_pars_pre_10 <- fit_gm_bullet_10$draws() %>% 
-  as_draws_df() %>%
-  select(starts_with("beta[")) %>%
-  pivot_longer(cols = everything()) %>%
+player_pars_pre_10 <- fit_gm_bullet_10$draws() |> 
+  as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
   mutate(param = stringr::str_extract(name, pattern = "\\d"),
@@
 
-hist_comp_winner <- bind_rows(player_pars_pre, player_pars_pre_10) |> 
-  filter(id < 7) |> 
+hist_comp_winner <- player_pars_pre |> 
   filter(param == 2) |> 
-  ggplot(aes(value)) +
-  geom_histogram() +
-  facet_grid(rows = vars(model), cols = vars(player_name), scales = "free") +
-  theme_6col() +
-  labs(x = element_blank(), y = element_blank()) +
-  scale_x_continuous(breaks = breaks_pretty(n = 3))
+  mutate(y = reorder(player_name, value, FUN = median)) |> 
+  bind_rows(player_pars_pre_10 |> filter(param == 2) |> 
+              mutate(y = as.factor(player_name))) |> 
+  ggplot(aes(x = value, y = y,
+             color = model)) +
+  stat_pointinterval(position = position_dodge(width = 1)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_single_y() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_color_viridis_d(option = "D")
 
@@
 ## then player effects
+## should update this also
 hist_comp_player <- bind_rows(player_pars_pre, player_pars_pre_10) |> 
@@
 ```{r bullet v blitz for 17-19}
-
 data_path <- here("results/lichess1700-1900")
@@
   geom_histogram() +
-  facet_grid(rows = vars(model), cols = vars(name), scales = "free") +
+  facet_grid(rows = vars(model), cols = vars(name), scales = "free",
+              labeller = labeller(name = facet_labels)) +
   theme_8col() +
@@
 ## then repeat for winner and player effects
-
-player_pars_bullet <- fit_17_19_bullet_prev$draws() %>% 
-  as_draws_df() %>%
-  select(starts_with("beta[")) %>%
-  pivot_longer(cols = everything()) %>%
+player_pars_bullet <- fit_17_19_bullet_prev$draws() |> 
+  as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
   mutate(param = stringr::str_extract(name, pattern = "\\d"),
@@
 
-player_pars_blitz <- fit_17_19_blitz_prev$draws() %>% 
-  as_draws_df() %>%
-  select(starts_with("beta[")) %>%
-  pivot_longer(cols = everything()) %>%
+player_pars_blitz <- fit_17_19_blitz_prev$draws() |> 
+  as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
   mutate(param = stringr::str_extract(name, pattern = "\\d"),
@@
                                               "/all_rated_bullet_model_prev.RDS"))
-
 bullet_gm <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
-
 fit_gm_blitz_prev <- readRDS(file = paste0(data_path,
                                               "/all_rated_blitz_model_prev.RDS"))
-
 blitz_gm <- readRDS(file = paste0(data_path, "/users_blitz.RDS"))
-
 ## compare global effects first
-
 bullet_pars <- fit_gm_bullet_prev$draws(c("mu_beta",
-                                                          "gamma1", "gamma2",
-                           "sigma_1", "tau[1]", "tau[2]",
-                           "sigma_g1", "sigma_g2")) |> 
+                                          "gamma1", "gamma2")) |> 
   as_draws_df() |> 
-  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |> 
   mutate(model = "Bullet")
@@
 blitz_pars <- fit_gm_blitz_prev$draws(c("mu_beta",
-                                                          "gamma1", "gamma2",
-                           "sigma_1", "tau[1]", "tau[2]",
-                           "sigma_g1", "sigma_g2")) |> 
+                                        "gamma1", "gamma2")) |> 
   as_draws_df() |> 
-  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |> 
   mutate(model = "Blitz")
@@
 bul_blit_gm_comp_global <- bind_rows(bullet_pars, blitz_pars) |> 
-  ggplot(aes(value)) +
-  geom_histogram() +
-  facet_grid(rows = vars(model), cols = vars(name), scales = "free") +
-  theme_8col() +
-  labs(x = element_blank(), y = element_blank()) +
-  scale_x_continuous(breaks = breaks_pretty(n = 2))
+  ggplot(aes(x = value, y = name, color = model)) +
+  stat_pointinterval(position = position_dodge(width = 0.2)) +
+  facet_wrap(~name,
+             scales = "free",
+             labeller = labeller(name = facet_labels)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_3col() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_color_viridis_d(option = "D")
 
@@
                               "bul_blitz_comp_gm_global.png")), dpi = 600,
-       height = 4, width = 7)
+       height = 3.5, width = 7)
 
@@
 
-player_pars_bullet <- fit_gm_bullet_prev$draws() %>% 
-  as_draws_df() %>%
-  select(starts_with("beta[")) %>%
-  pivot_longer(cols = everything()) %>%
+player_pars_bullet <- fit_gm_bullet_prev$draws() |> 
+  as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
   mutate(param = stringr::str_extract(name, pattern = "\\d"),
@@
 
-player_pars_blitz <- fit_gm_blitz_prev$draws() %>% 
-  as_draws_df() %>%
-  select(starts_with("beta[")) %>%
-  pivot_longer(cols = everything()) %>%
+player_pars_blitz <- fit_gm_blitz_prev$draws() |> 
+  as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
   mutate(param = stringr::str_extract(name, pattern = "\\d"),
@@
 
-comp_gm_winner <- bind_rows(player_pars_bullet, player_pars_blitz) |> 
+comp_gm_winner <- player_pars_bullet |> 
+  filter(param == 2) |> 
+  mutate(y = reorder(player_name, value, FUN = median)) |> 
+  bind_rows(player_pars_blitz |> 
+              filter(param == 2) |> 
+              mutate(y = as.factor(player_name))) |> 
+  ggplot(aes(x = value, y = y,
+             color = model)) +
+  stat_pointinterval(position = position_dodge(width = 1)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_single_y() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
+  scale_color_viridis_d(option = "D")
+
+comp_gm_winner
+
+ggsave(plot = comp_gm_winner, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "bul_blitz_comp_gm_winner.png")), dpi = 600,
+       height = 5, width = 7)
+
+## TO DO - this needs to be updated
+comp_gm_player <- bind_rows(player_pars_bullet, player_pars_blitz) |> 
   filter(id < 7) |> 
-  filter(param == 2) |> 
+  filter(param == 1) |> 
   ggplot(aes(value)) +
@@
 
-comp_gm_winner
-
-ggsave(plot = comp_gm_winner, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "bul_blitz_comp_gm_winner.png")), dpi = 600,
-       height = 5, width = 7)
-
-comp_gm_player <- bind_rows(player_pars_bullet, player_pars_blitz) |> 
-  filter(id < 7) |> 
-  filter(param == 1) |> 
-  ggplot(aes(value)) +
-  geom_histogram() +
-  facet_grid(rows = vars(model), cols = vars(player_name), scales = "free") +
-  theme_6col() +
-  labs(x = element_blank(), y = element_blank()) +
-  scale_x_continuous(breaks = breaks_pretty(n = 3))
-
 comp_gm_player
@@
 
-global_winner <- as_draws_df(fit_17_19_first$draws(variables = "mu_beta")) %>% 
-  rename("Initial Games" = mu_beta) %>% 
-  bind_cols(as_draws_df(fit_17_19_last$draws(variables = "mu_beta")) %>% 
-              rename("Later Games" = mu_beta)) %>% 
-  select(`Initial Games`, `Later Games`) %>% 
-  pivot_longer(cols = everything()) %>% 
+global_winner <- as_draws_df(fit_17_19_first$draws(variables = "mu_beta")) |> 
+  rename("Initial Games" = mu_beta) |> 
+  bind_cols(as_draws_df(fit_17_19_last$draws(variables = "mu_beta")) |> 
+              rename("Later Games" = mu_beta)) |> 
+  select(`Initial Games`, `Later Games`) |> 
+  pivot_longer(cols = everything()) |> 
   ggplot(aes(value, fill = name)) +
@@
 ## indiv winner effects
-random_effect_first <- fit_17_19_first$draws() %>% as_draws_df() %>%
-  select(starts_with("beta[")) %>%
-  pivot_longer(cols = everything()) %>%
+random_effect_first <- fit_17_19_first$draws() |> as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
   mutate(param = stringr::str_extract(name, pattern = "\\d"),
@@
          id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]")) %>% 
+         player_id = paste0("beta[", id, "]")) |> 
   mutate(when = "Initial")
 
-random_effect_last <- fit_17_19_last$draws() %>% as_draws_df() %>%
-  select(starts_with("beta[")) %>%
-  pivot_longer(cols = everything()) %>%
+random_effect_last <- fit_17_19_last$draws() |> as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
   mutate(param = stringr::str_extract(name, pattern = "\\d"),
@@
          id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]")) %>% 
+         player_id = paste0("beta[", id, "]")) |> 
   mutate(when = "Later")
 
-betas <- bind_rows(random_effect_first, random_effect_last) %>% 
+betas <- bind_rows(random_effect_first, random_effect_last) |> 
   select(value, param, id, when)
@@
 
-betas %>% left_join(player_id, by = "id") %>% 
+betas |> left_join(player_id, by = "id") |> 
   mutate(player = factor(player, levels = users))
 
-winner_effect_plot <- betas %>% 
-  left_join(player_id, by = "id") %>% 
-  mutate(player = factor(player, levels = users)) %>% 
-  filter(param == 2) %>% 
-  mutate(id = factor(id, levels = c(1:10))) %>% 
+winner_effect_plot <- betas |> 
+  left_join(player_id, by = "id") |> 
+  mutate(player = factor(player, levels = users)) |> 
+  filter(param == 2) |> 
+  mutate(id = factor(id, levels = c(1:10))) |> 
   ggplot(aes(value, fill = when)) +
@@
 
-fit_samp <- fit_samples %>% 
+fit_samp <- fit_samples |> 
   select(!starts_with(c("log_lik", "yrep")))
 
-y_rep <- fit_samples %>% select(starts_with("y_rep"))
+y_rep <- fit_samples |> select(starts_with("y_rep"))
 
@@
 
-y_rep_mod <- y_rep %>% 
-  rownames_to_column() %>%  
-  pivot_longer(-rowname) %>% 
-  pivot_wider(names_from=rowname, values_from=value) %>% 
+y_rep_mod <- y_rep |> 
+  rownames_to_column() |>  
+  pivot_longer(-rowname) |> 
+  pivot_wider(names_from=rowname, values_from=value) |> 
   ## remove the draws and chains here
@@
 
-games_won <- y_rep_mod %>% 
-  pivot_longer(cols = `1`:`4000`, names_to = "draw", values_to = "y") %>% 
-  group_by(focal_id, draw) %>% 
+games_won <- y_rep_mod |> 
+  pivot_longer(cols = `1`:`4000`, names_to = "draw", values_to = "y") |> 
+  group_by(focal_id, draw) |> 
   summarise(games_won = sum(y))
@@
 
-orig_games_won <- orig_data %>% 
-  group_by(focal_id) %>% 
-  summarise(games_won = sum(outcome)) %>% 
-  mutate(focal_id = as.character(focal_id)) %>% 
+orig_games_won <- orig_data |> 
+  group_by(focal_id) |> 
+  summarise(games_won = sum(outcome)) |> 
+  mutate(focal_id = as.character(focal_id)) |> 
   left_join(player_id, by = c("focal_id" = "id"))
 
-p1 <- games_won %>% 
-  mutate(focal_id = as.character(focal_id)) %>% 
-  left_join(player_id, by = c("focal_id" = "id")) %>% 
+p1 <- games_won |> 
+  mutate(focal_id = as.character(focal_id)) |> 
+  left_join(player_id, by = c("focal_id" = "id")) |> 
   ggplot(aes(x = games_won)) +
@@
 ## time varying ppc for 1700-1900 players
-draws <- fit_17_19_first$draws() %>% as_draws_df() %>%
-  select(starts_with(c("beta[", "gamma"))) %>%
-  mutate(draw = row_number()) %>% 
-  pivot_longer(cols = "beta[1,1]":"beta[2,10]") %>% 
+draws <- fit_17_19_first$draws() |> as_draws_df() |>
+  select(starts_with(c("beta[", "gamma"))) |>
+  mutate(draw = row_number()) |> 
+  pivot_longer(cols = "beta[1,1]":"beta[2,10]") |> 
   mutate(param = stringr::str_extract(name, pattern = "\\d"),
@@
          id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]")) %>% 
-  select(-name) %>% 
-  pivot_wider(names_from = param, values_from = value) %>% 
+         player_id = paste0("beta[", id, "]")) |> 
+  select(-name) |> 
+  pivot_wider(names_from = param, values_from = value) |> 
   rename(beta1 = `1`, beta2 = `2`)
@@
 future_results <- tibble(focal = as.character(stan_data_ave_last$id),
-                         result = stan_data_ave_last$y) %>% 
-  group_by(focal) %>%
-  summarise(total = sum(result)) %>% 
-  rename(id = focal) %>% 
+                         result = stan_data_ave_last$y) |> 
+  group_by(focal) |>
+  summarise(total = sum(result)) |> 
+  rename(id = focal) |> 
   left_join(player_id, by = "id")
 
-last_ppc <- ppc_data %>%  
-  left_join(player_id, by = "id") %>% 
+last_ppc <- ppc_data |>  
+  left_join(player_id, by = "id") |> 
   ggplot(aes(x = games_won)) + 
@@
 
-perm_17_19 <- perm_files[1:1000] %>% 
+perm_17_19 <- perm_files[1:1000] |> 
   map_dfr(~read_rds(here(perm_data_path, .x)))
@@
   mutate(Event = tolower(Event)) |>
-  filter(TimeControl == "60+0") %>%
-  filter(Variant == "Standard") %>%
+  filter(TimeControl == "60+0") |>
+  filter(Variant == "Standard") |>
   filter(grepl("rated bullet game", Event))
 
-select_users <- bullet_data %>% 
-  group_by(Username) %>% 
-  tally() %>% 
-  arrange(-n) %>% 
+select_users <- bullet_data |> 
+  group_by(Username) |> 
+  tally() |> 
+  arrange(-n) |> 
   ## will change this to n = 20 when run on cluster potentially
-  slice_max(order = n, n = 20) %>% 
+  slice_max(order = n, n = 20) |> 
   pull(Username)
@@
 
-draws <- fit_gm_first$draws() %>% as_draws_df() %>%
-  select(starts_with(c("beta[", "gamma"))) %>%
-  mutate(draw = row_number()) %>% 
-  pivot_longer(cols = "beta[1,1]":"beta[2,10]") %>% 
+draws <- fit_gm_first$draws() |> as_draws_df() |>
+  select(starts_with(c("beta[", "gamma"))) |>
+  mutate(draw = row_number()) |> 
+  pivot_longer(cols = "beta[1,1]":"beta[2,10]") |> 
   mutate(param = stringr::str_extract(name, pattern = "\\d"),
@@
 
-perm_gm <- perm_files[1:1000] %>% 
+perm_gm <- perm_files[1:1000] |> 
   map_dfr(~read_rds(here(perm_data_path, .x)))

### 2024-10-16 analysis/plot_templates.R

@@ -28,6 +28,25 @@ theme_single <- function(){
 }
 
 
+theme_single_y <- function(){ 
+  theme_bw() %+replace%    #replace elements we want to change
+    
+    theme(
+      
+      #grid elements
+      panel.spacing.x = unit(1, "lines"),
+      
+      #text elements
+      plot.title = element_text(size = title_size),
+      axis.text = element_text(size = axis_text_size - 2),
+      axis.title = element_text(size = axis_title),
+      legend.text = element_text(size = legend_text),
+      strip.text = element_text(size = axis_title - 5, 
+                                margin = margin(1.5, 1.5, 2, 1.5)),
+      legend.position = "none"
+    )
+}
+
 
 theme_3col <- function(){ 
   theme_bw() %+replace%    #replace elements we want to change

### 2024-10-17 Paper_WriteUp/Paper_Plots.qmd
trying to tidy all of this up

@@
 library(here)
+library(ggdist)
 library(viridis)
@@
 ```{r summary statistics 1700-1900}
-
+#| eval: false
+## this can be run if interested
 path_id <- 1  ## for 1700-1900 
-
 data_path <- all_data_path[path_id]
-
 files <- list.files(data_path)
-
 lichess_data <- files |> 
@@
 ## for each player, count how many times they start
-## a game with 3 minutes of the start of a previous game
+## a game with 5 minutes of the start of a previous game
 ## and get the proportion across all focal players
@@
          gap_seconds = as.numeric(gap)) |> 
-   mutate(less_3 = ifelse(gap_seconds <= 180, TRUE, FALSE)) |> 
+   mutate(less_5 = ifelse(gap_seconds <= 300, TRUE, FALSE)) |> 
    ungroup() |>   ## to get the overall proportion
-   summarise(sum(less_3, na.rm = TRUE)/ n())
+   summarise(sum(less_5, na.rm = TRUE)/ n())
 
@@
 ## repeat for blitz games
-
 outcomes_blitz <- blitz_data |> 
-  # group_by(Username) |> 
   select(Username, UTCDate, UTCTime) |> 
@@
   summarise(sum(multiple)/n())
-
 ```
@@
 ```{r summary statistics gm}
+#| eval: false
+## can be run if interested
 ## then repeat this analysis for the gms
@@
 files <- list.files(data_path)
-
 lichess_data_gm <- files |> 
@@
 
-## look at the proportion of games started within 3 minutes 
+## look at the proportion of games started within 5 minutes 
 ## of the previous game
@@
          gap_seconds = as.numeric(gap)) |> 
-  mutate(less_3 = ifelse(gap_seconds <= 180, TRUE, FALSE)) |> 
+  mutate(less_5 = ifelse(gap_seconds <= 300, TRUE, FALSE)) |> 
   ungroup() |>   ## to get the overall proportion
-  summarise(sum(less_3, na.rm = TRUE)/ n())
+  summarise(sum(less_5, na.rm = TRUE)/ n())
 
@@
          gap_seconds = as.numeric(gap)) |> 
-  mutate(less_3 = if_else(gap_seconds <= 180, TRUE, FALSE, missing = FALSE))
+  mutate(less_5 = if_else(gap_seconds <= 300, TRUE, FALSE, missing = FALSE))
 
@@
 result_gm <- outcomes_gm |> 
-  summarise(session_length = list(get_true_sequences(less_3)))
+  summarise(session_length = list(get_true_sequences(less_5)))
 
@@
 
+```{r figure 1}
+
+## to be added
+
+```
+
+```{r figure 2}
+
+```
+
 # Section 4: The Fitted Model
@@
 
-```{r plot global parameters for 17-19}
+```{r figure 3 - plot global parameters for 17-19}
 data_path <- here("results/lichess1700-1900")
@@
 ```{r interpret the estimated coefficients 17-19}
-
-gamma_ests <- fit_17_19_bullet_prev$summary(variables = c("gamma1",
+#| eval: false
+gamma_ests <- fit_17_19_bullet_prev$summary(variables = 
+                                              c("gamma1",
                                                 "gamma2")) |> 
@@
 1/(1 + exp(-1 * (100 * gamma_ests[2])))
-
 ```
@@
 ```{r player level parameters for 17-19}
-
-## need to get the write names then for these parameters here
-## did this somewhere before already
-
 player_id <- as_tibble(bullet_17_19) |> 
@@
 
-## then the winner effects plot
-winner_plot_17_19 <- player_pars |> 
-  filter(param == 2) |> 
-  ggplot(aes(value)) +
-  geom_histogram() +
-  facet_wrap(~player_name, scales = "free", ncol = 6) +
-  theme_6col() +
-  labs(x = element_blank(), y = element_blank()) +
-  scale_x_continuous(breaks = breaks_pretty(n = 3))
-
-winner_plot_17_19
-
-ggsave(plot = winner_plot_17_19, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "17_19_prev_winner.png")), dpi = 600,
-       height = 7, width = 7)
-
-## then repeat for player effects
-
-indiv_plot_17_19 <- player_pars |> 
-  filter(param == 1) |> 
-  ggplot(aes(value)) +
-  geom_histogram() +
-  facet_wrap(~player_name, scales = "free", ncol = 6) +
-  theme_6col() +
-  labs(x = element_blank(), y = element_blank()) +
-  scale_x_continuous(breaks = breaks_pretty(n = 3))
-
-indiv_plot_17_19
-
-ggsave(plot = indiv_plot_17_19, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "17_19_prev_indiv.png")), dpi = 600,
-       height = 7, width = 7)
-
-```
-
-```{r new boxplots}
-
+```
+
+```{r figure 4 left}
 winner_plot_17_19 <- player_pars |> 
@@
        height = 7, width = 7)
-
 ```
@@
 
-```{r global parameters for gm data}
+```{r figure 3 - global parameters for gm data}
 data_path <- here("results/lichessGrandmasters")
@@
 ```{r interpret the estimated coefficients gm}
-
+#| eval: false
 gamma_ests_gm <- fit_gm_bullet_prev$summary(variables = c("gamma1",
@@
   pull(mean)
-
 ## compute change in win probability for gamma1
@@
 ## change in win probability for elo difference
-
 1/(1 + exp(-1 * (0))) ## equally rated, playing as black
@@
 
-```{r player level parameters for gm data}
+```{r figure 4 right - player level parameters for gm data}
 player_id <- as_tibble(bullet_gm) |> 
@@
 ## then repeat for player effects
-
 indiv_plot_gm <- player_pars |> 
@@
 ```{r interpret changes in win probability for gm}
+#| eval: false
 ## interpret the individual winner effects
 ## assume wins 50% of all games
-
 1/(1 + exp(-1 * (0)))
-
 ## if won previous
 1/(1 + exp(-1 * (0 + 0.216 * 0.5)))
-
 ## if lost previous
 1/(1 + exp(-1 * (0 + 0.216 * - 0.5)))
-
 ## what about if win probability overall was 40%
 1/(1 + exp(-1 * (0 + 0.216 * 0.6)))
-
 ## 35%, pretty low
 1/(1 + exp(-1 * (0 + 0.216 * 0.65)))
-
 ```
@@
 
-```{r compare global params n-1 vs n-10}
+```{r figure 5 - compare global params n-1 vs n-10}
 ## global effects first
 data_path <- here("results/lichessGrandmasters")
-
 fit_gm_bullet_10 <- 
@@
 
-## then compare the winner effects and individual effects. 
-## will just show for the first 6 players for convenience
-
+```
+
+```{r}
+#| eval: false
+## TO DO - move to supplement
+## not required for a paper plot
 player_pars_pre <- fit_gm_bullet_prev$draws() |> 
@@
 ## should update this also
-hist_comp_player <- bind_rows(player_pars_pre, player_pars_pre_10) |> 
-  filter(id < 7) |> 
+hist_comp_player <- player_pars_pre |> 
   filter(param == 1) |> 
-  ggplot(aes(value)) +
-  geom_histogram() +
-  facet_grid(rows = vars(model), cols = vars(player_name), scales = "free") +
-  theme_6col() +
-  labs(x = element_blank(), y = element_blank()) +
-  scale_x_continuous(breaks = breaks_pretty(n = 3))
+  mutate(y = reorder(player_name, value, FUN = median)) |> 
+  bind_rows(player_pars_pre_10 |> filter(param == 1) |> 
+              mutate(y = as.factor(player_name))) |> 
+  ggplot(aes(x = value, y = y,
+             color = model)) +
+  stat_pointinterval(position = position_dodge(width = 1)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_single_y() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_color_viridis_d(option = "D")
 
 hist_comp_player
-
 ggsave(plot = hist_comp_player, 
@@
        height = 5, width = 7)
-
-```
-
-<!-- down to here in terms of updating the code, making sure it's complete -->
+```
 
@@
 ```{r bullet v blitz for 17-19}
+#| eval: false
+
+## TO DO - not a current paper plot, not updated
 data_path <- here("results/lichess1700-1900")
-
 fit_17_19_bullet_prev <- readRDS(file = paste0(data_path,
                                               "/all_rated_bullet_model_prev.RDS"))
-
 bullet_17_19 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
-
 fit_17_19_blitz_prev <- readRDS(file = paste0(data_path,
                                               "/all_rated_blitz_model_prev.RDS"))
-
 blitz_17_19 <- readRDS(file = paste0(data_path, "/users_blitz.RDS"))
@@
 ## compare global effects first
-
 bullet_pars <- fit_17_19_bullet_prev$draws(c("mu_beta",
@@
 
-```{r bullet v blitz gm}
+```{r figure 5 - bullet v blitz gm}
 data_path <- here("results/lichessGrandmasters")
@@
 ## then repeat for winner and player effects
-
 player_pars_bullet <- fit_gm_bullet_prev$draws() |> 
@@
        height = 5, width = 7)
+```
+
+```{r}
+#| eval: false
 
@@
        height = 5, width = 7)
-
-```
+```
+
+<!-- TO DO - Update Below here -->
+<!-- Everything down as far as here up to date. -->
+<!-- setting everything below to eval: false for now -->
 
@@
 ```{r ppc 17-19}
+#| eval: false
 ## this the global and winner effect plots
@@
 ```{r ppc1 for 17-19}
+#| eval: false
 path_id <- 1
@@
 ```{r ppc2 for 17-19}
+#| eval: false
 ## time varying ppc for 1700-1900 players
@@
 ```{r perm for 17-19}
+#| eval: false
 perm_data_path <- here("results/lichess1700-1900/perm")
@@
 ```{r perm winner effects}
+#| eval: false
 ## get the selected users for this, top 20 players
@@
 ```{r ppc1 and ppc2 for gm}
+#| eval: false
 ## just do ppc plots for now
@@
 ```{r}
+#| eval: false
 perm_data_path <- here("results/lichessGrandmasters/perm/")
@@
 ```{r}
-
+#| eval: false
 data_path <- here("results/lichess2000-2200")
@@
 ```{r}
-
+#| eval: false
 data_path <- here("results/lichess2300-2500")
@@
 
-What about for the other two groups?
-
-__To be added in__
-
-![Time-Varying PPC for 2000-2200](../results/lichess2000-2200/time/ppc_fit_first_last_bullet.png)
-
-![Time-Varying PPC for 2300-2500](../results/lichess2300-2500/time/ppc_fit_first_last_bullet.png)
-
-Maybe some very limited evidence that this becomes less
-reasonable for players with a higher rating.
+<!-- What about for the other two groups? -->
+
+<!-- __To be added in__ -->
+
+<!-- ![Time-Varying PPC for 2000-2200](../results/lichess2000-2200/time/ppc_fit_first_last_bullet.png) -->
+
+<!-- ![Time-Varying PPC for 2300-2500](../results/lichess2300-2500/time/ppc_fit_first_last_bullet.png) -->
+
+<!-- Maybe some very limited evidence that this becomes less -->
+<!-- reasonable for players with a higher rating. -->

### 2024-10-17 owen/pred_glicko.R

@@ -323,6 +323,10 @@ quick_sim <- as_tibble(quick_sim) |>
   pivot_longer(cols = Sim_1:Sim_100, names_to = "Sim",
                values_to = "glicko") 
 
+
+tru_glicko <- true_elo |> 
+  rename(game = index, glicko = focal_elo)
+
 quick_sim |> 
   ggplot(aes(game, glicko)) +
   geom_line(alpha = 0.05, aes(group = Sim)) +

### 2024-10-21 Paper_WriteUp/Paper_Plots.qmd
tidying

@@
 bullet_data |> nrow()
-
 ## average number, we remove the two accounts with 1 and 3 games only

### 2024-12-10 analysis/helper.R
update the pipe but more importantly update the filter in the historic games, to only include games where the focal player was actually focal

@@ -10,9 +10,9 @@ read_player <- function(path, file){
                                    BlackTitle = col_character(),
                                    WhiteElo = col_character(),
                                    BlackElo = col_character(),
-                                   FEN = col_character())) %>% 
+                                   FEN = col_character())) |> 
     select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
-           WhiteElo, BlackElo, Variant, TimeControl, Termination) %>% 
+           WhiteElo, BlackElo, Variant, TimeControl, Termination) |> 
     mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
            BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo)))
   dat
@@ -20,32 +20,34 @@ read_player <- function(path, file){
 
 get_hist <- function(user, games, prev_n) {
   if(prev_n == 1){
-    hist_games <- games %>% 
-      filter(White == user | Black == user) %>% 
-      arrange(UTCDate, UTCTime) %>% 
-      mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
-      select(White:BlackElo, focal_white) %>% 
+    hist_games <- games |> 
+      # filter(White == user | Black == user) |>
+      filter(Username == user) |> 
+      arrange(UTCDate, UTCTime) |> 
+      mutate(focal_white = ifelse(Username == White, 1, 0)) |> 
+      select(White:BlackElo, focal_white) |> 
       mutate(focal_result = case_when(
         (focal_white == 1 & Result == "1-0") ~ 1,
         (focal_white == 0 & Result == "0-1") ~ 1,
         (Result == "1/2-1/2") ~ 0.5,
         .default = 0
-      )) %>% 
+      )) |> 
       mutate(focal_win_prop = focal_result)
     
   }
   else{
-    hist_games <- games %>% 
-      filter(White == user | Black == user) %>% 
-      arrange(UTCDate, UTCTime) %>% 
-      mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
-      select(White:BlackElo, focal_white) %>% 
+    hist_games <- games |> 
+      # filter(White == user | Black == user) |> 
+      filter(Username == user) |> 
+      arrange(UTCDate, UTCTime) |> 
+      mutate(focal_white = ifelse(Username == White, 1, 0)) |> 
[…]

### 2024-12-10 Paper_WriteUp/Paper_Plots.qmd
modify the helper in some places

@@
    group_by(Username) |> 
-   select(Username, UTCDate, UTCTime) |> 
+   dplyr::select(Username, UTCDate, UTCTime) |> 
    mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
@@
   # group_by(Username) |> 
-  select(Username, UTCDate, UTCTime) |> 
+  dplyr::select(Username, UTCDate, UTCTime) |> 
   mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
@@
 outcomes_blitz <- blitz_data |> 
-  select(Username, UTCDate, UTCTime) |> 
+  dplyr::select(Username, UTCDate, UTCTime) |> 
   mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
@@
 bullet_data_gm |> nrow()
+bullet_data_gm |> 
+  group_by(Termination) |>
+  tally() |> 
+  mutate(prop = n/sum(n))
 
@@
   mutate(user_elo = ifelse(Username == White, WhiteElo, BlackElo)) |> 
-  select(Username, White, Black, WhiteElo, BlackElo, user_elo) |>
+  dplyr::select(Username, White, Black, WhiteElo, BlackElo, user_elo) |>
   ungroup() |> 
@@
 bullet_data_gm |> 
-  select(Username, UTCDate, UTCTime) |> 
+  dplyr::select(Username, UTCDate, UTCTime) |> 
   mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
@@
 outcomes_blitz <- blitz_data_gm |> 
-  select(Username, UTCDate, UTCTime) |> 
+  dplyr::select(Username, UTCDate, UTCTime) |> 
   mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
@@
 perm_17_19_winner <- player_pars |> 
-  filter(param == 1) |> 
+  filter(param == 2) |> 
   ggplot(aes(mean)) +
@@
        height = 5, width = 7)
+
+levels <- true_beta |> 
+  filter(param == 2) |> 
+  arrange(mean) |> 
+  pull(player_name)
+
+true_vals <- true_beta |> 
+  filter(param == 2) |>  
+  mutate(y = row_number() - 0.5,
+         yend = y + 1)
+
+perm_17_19_winner <- player_pars |> 
+  filter(param == 2) |>
+  mutate(y = factor(player_name, levels = levels)) |> 
+  ggplot(aes(x = mean, y = y)) +
+  stat_histinterval(slab_fill = NA, .width = c(0.5, 0.95)) +
+  geom_segment(data = true_vals, 
+               aes(x = mean, y = y, yend = yend), col = "red") +
+  labs(y = "", x = "Estimated W/L-Effect") +
+  theme_single_y()
+
+ggsave(plot = perm_17_19_winner, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "perm_17_19_winner.png")), dpi = 600,
+       height = 5, width = 7)
+
 ```
@@
 
+perm_gm_winner <- player_pars |> 
+  filter(param == 2) |>
+  mutate(y = factor(player_name, levels = levels)) |> 
+  ggplot(aes(x = mean, y = y)) +
+  stat_histinterval(slab_fill = NA, .width = c(0.5, 0.95)) +
+  geom_segment(data = true_vals, 
+               aes(x = mean, y = y, yend = yend), col = "red") +
+  theme_single_y() +
+  labs(x = "Estimated W/L-Effect", y = "")
+
 perm_gm_winner

### 2024-12-10 analysis/helper.R

@@ -11,7 +11,7 @@ read_player <- function(path, file){
                                    WhiteElo = col_character(),
                                    BlackElo = col_character(),
                                    FEN = col_character())) |> 
-    select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
+    dplyr::select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
            WhiteElo, BlackElo, Variant, TimeControl, Termination) |> 
     mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
            BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo)))

### 2024-12-10 owen/cluster_scripts/Large_Fit.R
rerun this model with just the modification in the get_hist for previous 10 games

@@ -32,14 +32,14 @@ all_data_path[1] <- here("box_data/lichess1700-1900/")
 all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_data_path[4] <- here("box_data/lichessGrandmasters/")
-all_save_path[1] <- here("results/lichess1700-1900_check/")
+all_save_path[1] <- here("results/lichess1700-1900/")
 
 # all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
 # ## if need to run it locally
 
-all_save_path[2] <- here("results/lichess2000-2200_check/")
-all_save_path[3] <- here("results/lichess2300-2500_test/")
-all_save_path[4] <- here("results/lichessGrandmasters_check/")
+all_save_path[2] <- here("results/lichess2000-2200/")
+all_save_path[3] <- here("results/lichess2300-2500/")
+all_save_path[4] <- here("results/lichessGrandmasters/")
 
 
 data_path <- all_data_path[path_id]
@@ -49,24 +49,24 @@ dir.create(save_path, showWarnings = FALSE)
 
 files <- list.files(data_path)
 
-lichess_data <- files %>% 
+lichess_data <- files |> 
   map_dfr(~read_player(data_path, .x))
 
 ## restrict to rated rapid and shorter here
 ## this also removes the NAs, which makes sense
 
-small_data <- lichess_data %>%
+small_data <- lichess_data |>
   mutate(Event = tolower(Event)) |>
-  # filter(Event == "Rated Bullet game") %>%
-  filter(TimeControl == "60+0") %>%
-  filter(Variant == "Standard") %>%
+  # filter(Event == "Rated Bullet game") |>
+  filter(TimeControl == "60+0") |>
+  filter(Variant == "Standard") |>
   filter(grepl("rated bullet game", Event))
 
 # ## what time length should be
-# small_data <- lichess_data %>%
+# small_data <- lichess_data |>
 #   mutate(Event = tolower(Event)) |>
-#   filter(TimeControl == "180+0") %>%
-#   filter(Variant == "Standard") %>%
+#   filter(TimeControl == "180+0") |>
+#   filter(Variant == "Standard") |>
 #   filter(grepl("rated blitz game", Event))
 
 users <- unique(small_data$Username)
@@ -76,29 +76,29 @@ rm(lichess_data)
 
[…]

### 2024-12-10 owen/cluster_scripts/submit.sh
some changes to the submit file

@@ -1,8 +1,10 @@
 #! /bin/bash
 #
-#SBATCH --mem-per-cpu 12000
-#SBATCH -c 8
-#SBATCH -t 600:00
+#SBATCH --mem-per-cpu 8000
+#SBATCH -c 6
+#SBATCH -J bullet_n10
+#SBATCH -t 2000:00
+#SBATCH -a 1-4
 #SBATCH --mail-user=oward@sfu.ca
 #SBATCH --mail-type=ALL
 
@@ -12,6 +14,9 @@ echo "Launching R"
 date
 
 Rscript Large_Fit.R
+#Rscript Large_Single_Prev.R
+#Rscript Winner_Effects_Time.R
+#Rscript perm_results.R
 
 echo "Completed"
 date

### 2025-01-13 Paper_WriteUp/Paper_Plots.qmd
update the colour scheme and make it clear which players are being removed

@@
   count() |> 
-  filter(n > 3) |>
+  filter(n > 10) |> # remove 2 players with <10 games
   ungroup() |> 
@@
 
+## count how many blitz games they play as comparison
+
+blitz_data |> 
+  group_by(Username) |> 
+  count() |> 
+  filter(n > 10) |> # remove players with < 10 games
+  ungroup() |> 
+  summarise(avg = mean(n))
+
 ## repeat for blitz data using 7 minutes as the maximum gap
@@
 ```{r figure 2}
+
+## to be added
 
@@
   scale_x_continuous(breaks = breaks_pretty(n = 2)) +
-  scale_color_viridis_d(option = "D")
+  scale_color_viridis_d(option = "H")
 
@@
   scale_x_continuous(breaks = breaks_pretty(n = 3)) +
-  scale_color_viridis_d(option = "D")
+  scale_color_viridis_d(option = "H")
 

### 2025-01-13 Paper_WriteUp/Paper_Plots.qmd
add Sydney's code to the paper plots file

@@
 library(viridis)
+library(PlayerRatings)
+library(ggrepel)
+library(ggpubr)
 options(mc.cores = parallel::detectCores())
@@
 ```{r figure 1}
-
-## to be added
-
+player_cohort_1 <- "sandro2502"
+
+path_id <- 1  ## for 1700-1900 
+data_path <- all_data_path[path_id]
+files <- list.files(data_path)
+lichess_data <- files |> 
+  map_dfr(~read_player(data_path, .x))
+
+bullet_data_fig_1 <- lichess_data |> 
+  mutate(Event = tolower(Event)) |>
+  filter(Variant == "Standard") |>
+  filter(grepl("rated bullet game", Event))
+rm(lichess_data)
+
+p1 <- bullet_data |> 
+  filter(Username == player_cohort_1)
+
+p1$newELO <- ifelse(p1$White == paste0(player_cohort_1),
+                    p1$WhiteElo, p1$BlackElo)
+# p1$datetime <- as.POSIXct(paste(p1$UTCDate, p1$UTCTime),
+#                           format = "%Y.%m.%d %H:%M:%S")
+ltypes <- c(1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2, 3)
+thetitle <- paste0("Player From 1700-1900 Cohort")
+linewd <- 1
+ylim1 <- 1200
+ylim2 <- 3500
+ndays <- 1
+
+p1_sorted <- p1 |> arrange(UTCDate, UTCTime)
+p1_sorted$event <- 1:nrow(p1_sorted)
+
+tidy_plot1 <- ggplot(p1_sorted, aes(x = event, y = newELO)) +
+  # scale_colour_manual(values = getPalette(colourCount)) +
+  scale_linetype_manual(values = ltypes) +
+  xlab("Games Played") +
+  ylab("Lichess Rating") +
+  ggtitle(thetitle) +
+  ylim(ylim1, ylim2) +
+  geom_line(lwd = linewd) +
+  theme(
+    plot.title = element_text(hjust = 0, vjust = 1, size = rel(1.7)),
+    panel.background = element_blank(),
+    plot.background = element_blank(),
+    panel.grid.major.y = element_line(color = "gray75", linetype = "dotted"),
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor = element_blank(),
+    strip.background = element_blank(),
+    strip.text = element_text(size = rel(1.1)),
+    text = element_text(color = "gray20", size = 10),
+    axis.text = element_text(size = rel(1.0)),
+    axis.text.x = element_text(color = "gray20", size = rel(1.0)),
+    axis.text.y = element_text(color = "gray20", size = rel(1.0)),
+    axis.title.x = element_text(size = rel(1.0), vjust = 0),
+    axis.title.y = element_text(size = rel(1.0), vjust = 1),
+    axis.ticks.y = element_blank(),
+    axis.ticks.x = element_blank(),
+    legend.position = "none"
+  )
+
+player_cohort_gm <- "alireza2003"
+
+## load in the GM data for this
+path_id <- 4  ## for GM 
+data_path <- all_data_path[path_id]
+files <- list.files(data_path)
+lichess_data <- files |> 
+  map_dfr(~read_player(data_path, .x))
+
+bullet_data_fig_1_gm <- lichess_data |> 
+  mutate(Event = tolower(Event)) |>
+  filter(Variant == "Standard") |>
+  filter(grepl("rated bullet game", Event))
+rm(lichess_data)
+
+p2 <- bullet_data_fig_1_gm |> 
+  filter(Username == player_cohort_gm)
+
+p2$newELO <- ifelse(p2$White == paste0(player_cohort_gm),
+                    p2$WhiteElo, p2$BlackElo)
+thetitle <- paste0("Player From GM Cohort")
+p2_sorted <- p2 |> arrange(UTCDate, UTCTime)
+p2_sorted$event <- 1:nrow(p2_sorted)
+
+tidy_plot2 <- ggplot(p2_sorted, aes(x = event, y = newELO)) +
+  # scale_colour_manual(values = getPalette(colourCount)) +
+  scale_linetype_manual(values = ltypes) +
+  xlab("Games Played") +
+  ylab("Lichess Rating") +
+  ggtitle(thetitle) +
+  ylim(ylim1, ylim2) +
+  geom_line(lwd = linewd) +
+  theme(
+    plot.title = element_text(hjust = 0, vjust = 1, size = rel(1.7)),
+    panel.background = element_blank(),
+    plot.background = element_blank(),
+    panel.grid.major.y = element_line(color = "gray75", linetype = "dotted"),
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor = element_blank(),
+    strip.background = element_blank(),
+    strip.text = element_text(size = rel(1.1)),
+    text = element_text(color = "gray20", size = 10),
+    axis.text = element_text(size = rel(1.0)),
+    axis.text.x = element_text(color = "gray20", size = rel(1.0)),
+    axis.text.y = element_text(color = "gray20", size = rel(1.0)),
+    axis.title.x = element_text(size = rel(1.0), vjust = 0),
+    axis.title.y = element_text(size = rel(1.0), vjust = 1),
+    # axis.ticks.y = element_blank(),
+    axis.ticks.x = element_blank(),
+    legend.position = "none"
+  )
+
+tdy1 <- tidy_plot1 + theme_single_y()
+tdy2 <- tidy_plot2 + theme_single_y()
+
+tidy_plot <- ggarrange(tdy1, tdy2)
+
+ggsave(paste0(newfilename, ".png"), tidy_plot,
+  dpi = 600, height = 5, width = 12
+)
+ggsave(plot = tidy_plot, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "elo_plot_2_players.png")), dpi = 600,
+       height = 5, width = 12)
 ```

### 2025-01-13 Paper_WriteUp/Paper_Plots.qmd
add Adam's code for figure 2

@@
 ```{r figure 2}
-
-## to be added
-
+## Adam's code for this figure
+
+gm_players_elo <- list()
+gm_session_lengths <- list()
+gm_time_diffs <- list()
+gm_time_diffs_10 <- list()
+
+gm_players <- unique(bullet_data_gm$Username)
+
+for (i in 1:length(gm_players)) {
+  tidy_games_gms <- bullet_data_gm |>
+    filter(Username == gm_players[i]) |> # select only one player
+    mutate(UTCDateTime = ymd_hms(paste(UTCDate, UTCTime))) |> # convert date-time to lubridate type
+    arrange(UTCDateTime) |> # order date-time
+    mutate(time_diff = UTCDateTime - lag(UTCDateTime)) # time difference between games
+
+  ## A check to remove duplicated games, if they occur
+  # cat("Dupes:", sum(duplicated(tidy_games_gms[1:9])), "\n", sep = " ")
+  # # some players have all their games double recorded for some reason
+  # tidy_games_gms <- tidy_games_gms[!duplicated(tidy_games_gms[1:9]), ] # remove duplicate rows
+
+  # get players average elo across all games
+  gm_players_elo[[gm_players[i]]] <- mean(c(
+    tidy_games_gms$WhiteElo[gm_players[i] == tidy_games_gms$White],
+    tidy_games_gms$BlackElo[gm_players[i] == tidy_games_gms$Black]
+  ))
+
+  # store time diffs between games
+  gm_time_diffs[[gm_players[i]]] <- difftime(tidy_games_gms$UTCDateTime,
+    lag(tidy_games_gms$UTCDateTime),
+    units = "secs"
+  ) # time between current game and previous
+  gm_time_diffs_10[[gm_players[i]]] <- difftime(tidy_games_gms$UTCDateTime,
+    lag(tidy_games_gms$UTCDateTime, 10),
+    units = "secs"
+  ) # time between current game and 10th game ago
+}
+# store in df for plotting
+gm_time_diff_df <- data.frame("time_diff" = unlist(gm_time_diffs))
+gm_time_diff_df_10 <- data.frame("time_diff" = unlist(gm_time_diffs_10))
+
+## Same distributions as amateurs, so again proceed by assuming all 
+## games played within 300 seconds (5 min) are in the same session
+for (i in 1:length(gm_players)) {
+  tidy_games_gms <- bullet_data_gm |>
+    filter(Username == gm_players[i]) |> # select only one player
+    mutate(UTCDateTime = ymd_hms(paste(UTCDate, UTCTime))) |> # convert date-time to lubridate type
+    arrange(UTCDateTime) |> # order date-time
+    mutate(time_diff = UTCDateTime - lag(UTCDateTime)) # time difference between games
+
+  length_of_sessions <- which(tidy_games_gms$time_diff > 300) - 
+    lag(which(tidy_games_gms$time_diff > 300), default = 0)
+  # game id of first game of session minus game id of first game of previous session
+
+  gm_session_lengths[[gm_players[i]]] <- length_of_sessions # store
+}
+# store in df for plotting
+gm_session_lengths_df <- data.frame(session_length = unlist(gm_session_lengths))
+
+log_time_diff_plot_gm <- ggplot(gm_time_diff_df, aes(x = as.numeric(time_diff))) +
+  geom_histogram() +
+  scale_x_continuous(
+    trans = scales::log_trans(),
+    breaks = c(1, 60, 300, 3600, 86400, 31536000),
+    labels = c("1 sec", "1 min", "5 min", "1 hour", "1 day", "1 year")
+  ) +
+  xlab("Time Between Games (Log Scale)") +
+  geom_vline(xintercept = 300, colour = "red") +
+  ylab("Frequency") #+ ggtitle("Time Difference Between Current Game and Previous Game for Amateurs")
+fig2a <- log_time_diff_plot_gm + theme_single()
+
+ggsave(plot = fig2a, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "time_between_games_gm.png")), dpi = 600,
+       height = 5, width = 7)
+
+session_length_plot_gm <- ggplot(gm_session_lengths_df, aes(x = session_length)) +
+  geom_histogram() +
+  scale_x_continuous(
+    trans = pseudo_log_trans(sigma = 2.36),
+    breaks = c(1, 5, 10, 100),
+    labels = c("1 game", "5 games", "10 games", "100 games"),
+    limits = c(0, 389)
+  ) +
+  xlab("Number of Games in a Session (Log Scale)") +
+  ylab("Frequency") #+ ggtitle("Number of Games per Session for GMs")
+fig2b <- session_length_plot_gm + theme_single()
+
+ggsave(plot = fig2b, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "session_length_gm.png")), dpi = 600,
+       height = 5, width = 7)
 ```

### 2025-01-20 Paper_WriteUp/Paper_Plots.qmd
tidy up all the plots before section 5

@@
 ```{r figure 1}
+## Sydney's code for this figure
+
 player_cohort_1 <- "sandro2502"
@@
 
-```{r figure 3 - plot global parameters for 17-19}
+```{r figure 3 left - plot global parameters for 17-19}
 data_path <- here("results/lichess1700-1900")
@@
 bullet_17_19 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
-
-# color_scheme_set("blue")
-
-## easiest way to make this grayscale would be to not use mcmc_hist at all
 
@@
 global_17_19_bullet_prev
-
 ggsave(plot = global_17_19_bullet_prev, 
@@
 ```{r player level parameters for 17-19}
+#| eval: false
+
 player_id <- as_tibble(bullet_17_19) |> 
@@
   filter(contain_zero == 0)
-
 ```
@@
        height = 7, width = 7)
-
+```
+
+```{r plot player effects}
+#| eval: false
 indiv_plot_17_19 <- player_pars |> 
@@
 
-```{r figure 3 - global parameters for gm data}
+```{r figure 3 right - global parameters for gm data}
 data_path <- here("results/lichessGrandmasters")
@@
 fit_gm_bullet_prev <- 
-  readRDS(file =paste0(data_path,"/all_rated_bullet_model_prev.RDS"))
+  readRDS(file = paste0(data_path,"/all_rated_bullet_model_prev.RDS"))
 
@@
 global_gm_bullet_prev
-
 ggsave(plot = global_gm_bullet_prev, 
@@
 1/(1 + exp(-1 * (gamma_ests_gm[1])))
-
 ## change in win probability for elo difference
@@
 1/(1 + exp(-1 * (100 * gamma_ests_gm[2])))
-
 ```
@@
        height = 7, width = 7)
-
+```
+
+```{r plot player effects}
+#| eval: false
 ## then repeat for player effects
@@
 
-```{r figure 5 - compare global params n-1 vs n-10}
+```{r figure 5 left - compare global params n-1 vs n-10}
 ## global effects first
@@
        height = 3.5, width = 7)
-
 ```
@@
              color = model)) +
-  stat_pointinterval(position = position_dodge(width = 1)) +
+  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
   labs(x = element_blank(), y = element_blank(), color = element_blank()) +
@@
 
-## TO DO - not a current paper plot, not updated
+## Potential Supplement Plots
 data_path <- here("results/lichess1700-1900")
-fit_17_19_bullet_prev <- readRDS(file = paste0(data_path,
-                                              "/all_rated_bullet_model_prev.RDS"))
+fit_17_19_bullet_prev <- readRDS(file = paste0(
+  data_path,
+  "/all_rated_bullet_model_prev.RDS"
+))
 bullet_17_19 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
-fit_17_19_blitz_prev <- readRDS(file = paste0(data_path,
-                                              "/all_rated_blitz_model_prev.RDS"))
+fit_17_19_blitz_prev <- readRDS(file = paste0(
+  data_path,
+  "/all_rated_blitz_model_prev.RDS"
+))
 blitz_17_19 <- readRDS(file = paste0(data_path, "/users_blitz.RDS"))
 
+player_id_17_19 <- as_tibble(bullet_17_19) |> 
+  mutate(id = row_number()) |> 
+  rename(player_name = value)
+
 ## compare global effects first
-bullet_pars <- fit_17_19_bullet_prev$draws(c("mu_beta",
-                                                          "gamma1", "gamma2",
-                           "sigma_1", "tau[1]", "tau[2]",
-                           "sigma_g1", "sigma_g2")) |> 
+bullet_pars <- fit_17_19_bullet_prev$draws(c(
+  "mu_beta",
+  "gamma1", "gamma2"
+)) |>
   as_draws_df() |>
-  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |>
   mutate(model = "Bullet")
 
-blitz_pars <- fit_17_19_blitz_prev$draws(c("mu_beta",
-                                                          "gamma1", "gamma2",
-                           "sigma_1", "tau[1]", "tau[2]",
-                           "sigma_g1", "sigma_g2")) |> 
+blitz_pars <- fit_17_19_blitz_prev$draws(c(
+  "mu_beta",
+  "gamma1", "gamma2"
+)) |>
   as_draws_df() |>
-  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |>
   mutate(model = "Blitz")
@@
 bul_blit_comp_global <- bind_rows(bullet_pars, blitz_pars) |>
-  ggplot(aes(value)) +
-  geom_histogram() +
-  facet_grid(rows = vars(model), cols = vars(name), scales = "free",
+  ggplot(aes(value, color = model))  +
+  stat_pointinterval(position = position_dodge(width = 0.2)) +
+  facet_wrap(~name,
+             scales = "free",
              labeller = labeller(name = facet_labels)) +
-  theme_8col() +
-  labs(x = element_blank(), y = element_blank()) +
-  scale_x_continuous(breaks = breaks_pretty(n = 2))
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_3col() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_color_viridis_d(option = "H")
 
@@
          param = as.numeric(param)) |> 
-  left_join(player_id, by = "id") |> 
+  left_join(player_id_17_19, by = "id") |> 
   mutate(model = "Bullet")
@@
          param = as.numeric(param)) |> 
-  left_join(player_id, by = "id") |> 
+  left_join(player_id_17_19, by = "id") |> 
   mutate(model = "Blitz")
 
-comp_17_19_winner <- bind_rows(player_pars_bullet, player_pars_blitz) |> 
-  filter(id < 7) |> 
-  filter(param == 2) |> 
-  ggplot(aes(value)) +
-  geom_histogram() +
-  facet_grid(rows = vars(model), cols = vars(player_name), scales = "free") +
-  theme_6col() +
-  labs(x = element_blank(), y = element_blank()) +
-  scale_x_continuous(breaks = breaks_pretty(n = 3))
+## need to keep only players which are in both 
+## bullet and blitz
+
+common_players <- intersect(unique(player_pars_bullet$player_name),
+                            unique(player_pars_blitz$player_name))
+
+comp_17_19_winner <- player_pars_bullet |> 
+  filter(param == 2) |> 
+  filter(player_name %in% common_players) |> 
+  mutate(y = reorder(player_name, value, FUN = median)) |> 
+  bind_rows(player_pars_blitz |>
+              filter(param == 2) |>
+              filter(player_name %in% common_players) |>
+              mutate(y = as.factor(player_name))) |>
+  ggplot(aes(x = value, y = y,
+             color = model)) +
+  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_single_y() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_color_viridis_d(option = "D")
 
@@
 
-comp_17_19_player <- bind_rows(player_pars_bullet, player_pars_blitz) |> 
-  filter(id < 7) |> 
+comp_17_19_player <- player_pars_bullet |> 
   filter(param == 1) |> 
-  ggplot(aes(value)) +
-  geom_histogram() +
-  facet_grid(rows = vars(model), cols = vars(player_name), scales = "free") +
-  theme_6col() +
-  labs(x = element_blank(), y = element_blank()) +
-  scale_x_continuous(breaks = breaks_pretty(n = 3))
+  filter(player_name %in% common_players) |> 
+  mutate(y = reorder(player_name, value, FUN = median)) |> 
+  bind_rows(player_pars_blitz |>
+              filter(param == 1) |>
+              filter(player_name %in% common_players) |>
+              mutate(y = as.factor(player_name))) |>
+  ggplot(aes(x = value, y = y,
+             color = model)) +
+  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_single_y() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_color_viridis_d(option = "D")
 
@@
 
-```{r figure 5 - bullet v blitz gm}
+```{r figure 5 right - bullet v blitz gm}
 data_path <- here("results/lichessGrandmasters")
-
-fit_gm_bullet_prev <- readRDS(file = paste0(data_path,
-                                              "/all_rated_bullet_model_prev.RDS"))
+fit_gm_bullet_prev <- readRDS(file = paste0(
+  data_path,
+  "/all_rated_bullet_model_prev.RDS"
+))
 bullet_gm <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
-fit_gm_blitz_prev <- readRDS(file = paste0(data_path,
-                                              "/all_rated_blitz_model_prev.RDS"))
+fit_gm_blitz_prev <- readRDS(file = paste0(
+  data_path,
+  "/all_rated_blitz_model_prev.RDS"
+))
 blitz_gm <- readRDS(file = paste0(data_path, "/users_blitz.RDS"))
 ## compare global effects first
-bullet_pars <- fit_gm_bullet_prev$draws(c("mu_beta",
-                                          "gamma1", "gamma2")) |> 
+bullet_pars <- fit_gm_bullet_prev$draws(c(
+  "mu_beta",
+  "gamma1", "gamma2"
+)) |>
   as_draws_df() |>
@@
 
-blitz_pars <- fit_gm_blitz_prev$draws(c("mu_beta",
-                                        "gamma1", "gamma2")) |> 
+blitz_pars <- fit_gm_blitz_prev$draws(c(
+  "mu_beta",
+  "gamma1", "gamma2"
+)) |>
   as_draws_df() |>
@@
   ggplot(aes(x = value, y = name, color = model)) +
-  stat_pointinterval(position = position_dodge(width = 0.2)) +
+  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
   facet_wrap(~name,
     scales = "free",
-             labeller = labeller(name = facet_labels)) +
+    labeller = labeller(name = facet_labels)
+  ) +
   labs(x = element_blank(), y = element_blank(), color = element_blank()) +
@@
 
-ggsave(plot = bul_blit_gm_comp_global, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "bul_blitz_comp_gm_global.png")), dpi = 600,
-       height = 3.5, width = 7)
-
+ggsave(
+  plot = bul_blit_gm_comp_global,
+  filename = paste0(here(
+    "Paper_WriteUp", "paper_figures",
+    "bul_blitz_comp_gm_global.png"
+  )), dpi = 600,
+  height = 3.5, width = 7
+)
+```
+
+```{r figure 6 - comparison between bullet and blitz winner}
 ## then repeat for winner and player effects
+player_id <- as_tibble(bullet_gm) |> 
+  mutate(id = row_number()) |> 
+  rename(player_name = value)
+
 player_pars_bullet <- fit_gm_bullet_prev$draws() |> 
@@
 comp_gm_winner
-
 ggsave(plot = comp_gm_winner, 
@@
 #| eval: false
-
-## TO DO - this needs to be updated
-comp_gm_player <- bind_rows(player_pars_bullet, player_pars_blitz) |> 
-  filter(id < 7) |> 
+## player effects also
+comp_gm_player <- player_pars_bullet |> 
   filter(param == 1) |> 
-  ggplot(aes(value)) +
-  geom_histogram() +
-  facet_grid(rows = vars(model), cols = vars(player_name), scales = "free") +
-  theme_6col() +
-  labs(x = element_blank(), y = element_blank()) +
-  scale_x_continuous(breaks = breaks_pretty(n = 3))
+  mutate(y = reorder(player_name, value, FUN = median)) |> 
+  bind_rows(player_pars_blitz |> 
+              filter(param == 1) |> 
+              mutate(y = as.factor(player_name))) |> 
+  ggplot(aes(x = value, y = y,
+             color = model)) +
+  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_single_y() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
+  scale_color_viridis_d(option = "H")
 
@@
 <!-- setting everything below to eval: false for now -->
+<!-- Updating everything below here -->
 
@@
 
+## Fitting the Model with n=5 as History
+
 ## Model Checking for the Other Cohorts

### 2025-01-20 Paper_WriteUp/Paper_Plots.qmd
tidying this up a lot

@@
 
-1. We select 10 players from each group
-and look at their first and last 1000 games.
-We fit our model to each of these separately
-and examine posterior predictive checks based on this.
-
-1. Then we used the model fit to the first 1000 games
-to construct posterior predictive checks for the final
-1000 games and see how they compare.
-
-## 1700-1900 Players
-
-```{r ppc 17-19}
-#| eval: false
-## this the global and winner effect plots
-path_id <- 1
-data_path <- res_data_path[path_id]
-
-fit_17_19_first <- readRDS(file =  here(data_path, "time",
-                                        "select_users_first_bullet.RDS"))
-fit_17_19_last <- readRDS(file =  here(data_path, "time",
-                                        "select_users_last_bullet.RDS"))
-
-global_winner <- as_draws_df(fit_17_19_first$draws(variables = "mu_beta")) |> 
-  rename("Initial Games" = mu_beta) |> 
-  bind_cols(as_draws_df(fit_17_19_last$draws(variables = "mu_beta")) |> 
-              rename("Later Games" = mu_beta)) |> 
-  select(`Initial Games`, `Later Games`) |> 
-  pivot_longer(cols = everything()) |> 
-  ggplot(aes(value, fill = name)) +
-  geom_histogram(position = "identity") +
-  facet_wrap(~name, ncol = 1) +
-  labs(fill = element_blank(), y = element_blank(), x = element_blank()) +
-  theme(legend.position = "none",
-        axis.text = element_text(size = axis_text_size),
-        axis.title = element_text(size = axis_title),
-        legend.text = element_text(size = legend_text),
-        strip.text = element_text(size = axis_title),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        plot.title = element_text(size = title_size)) +
-  # labs(title = "Selected 1700-1900 Bullet Players") +
-  NULL
-
-global_winner
-
-users <- readRDS(file = here(data_path, "time", "select_users_bullet.RDS"))
-## indiv winner effects
-random_effect_first <- fit_17_19_first$draws() |> as_draws_df() |>
-  select(starts_with("beta[")) |>
-  pivot_longer(cols = everything()) |>
-  mutate(param = stringr::str_extract(name, pattern = "\\d"),
-         id = stringr::str_extract(name, pattern = "\\d+]"),
-         id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]")) |> 
-  mutate(when = "Initial")
-
-random_effect_last <- fit_17_19_last$draws() |> as_draws_df() |>
-  select(starts_with("beta[")) |>
-  pivot_longer(cols = everything()) |>
-  mutate(param = stringr::str_extract(name, pattern = "\\d"),
-         id = stringr::str_extract(name, pattern = "\\d+]"),
-         id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]")) |> 
-  mutate(when = "Later")
-
-betas <- bind_rows(random_effect_first, random_effect_last) |> 
-  select(value, param, id, when)
-
-player_id <- tibble(id = as.character(1:10), player = users)
-
-betas |> left_join(player_id, by = "id") |> 
-  mutate(player = factor(player, levels = users))
-
-winner_effect_plot <- betas |> 
-  left_join(player_id, by = "id") |> 
-  mutate(player = factor(player, levels = users)) |> 
-  filter(param == 2) |> 
-  mutate(id = factor(id, levels = c(1:10))) |> 
-  ggplot(aes(value, fill = when)) +
-  geom_histogram() +
-  geom_vline(xintercept = 0, alpha = 0.5) +
-  scale_x_continuous(breaks = c(-0.25, 0.25)) +
-  # xlim(c(-1.5, 1.5)) +
-  facet_grid(rows = vars(when), cols = vars(player), scales = "free_x") +
-  labs(fill = element_blank(), y = element_blank(), 
-       # title = "Individual Winner Effects, Selected 1700-1900 Bullet Players",
-       x = element_blank()) +
-  theme_10col()
-  # theme(axis.ticks.y = element_blank(),
-  #       axis.text.y = element_blank(),
-  #       axis.text = element_text(size = axis_text_size - 4),
-  #       axis.title = element_text(size = axis_title),
-  #       legend.text = element_text(size = legend_text),
-  #       strip.text = element_text(size = axis_title - 6),
-  #       legend.position = "none",
-  #       plot.title = element_text(size = title_size))
-
-winner_effect_plot
-```
-
-```{r ppc1 for 17-19}
-#| eval: false
-path_id <- 1
-data_path <- res_data_path[path_id]
-
-fit_17_19_first <- readRDS(file =  here(data_path, "time",
-                                        "select_users_first_bullet.RDS"))
-fit_17_19_last <- readRDS(file =  here(data_path, "time",
-                                        "select_users_last_bullet.RDS"))
-
-fit_samples <- as_draws_df(fit_17_19_first$draws())
-
-fit_samp <- fit_samples |> 
-  select(!starts_with(c("log_lik", "yrep")))
-
-y_rep <- fit_samples |> select(starts_with("y_rep"))
-
-rm(fit_samples)
-rm(fit_samp)
-
-stan_data_ave_first <- readRDS(file = here(data_path, "time",
-                                           "stan_data_first.RDS"))
-
-y_rep_mod <- y_rep |> 
-  rownames_to_column() |>  
-  pivot_longer(-rowname) |> 
-  pivot_wider(names_from=rowname, values_from=value) |> 
-  ## remove the draws and chains here
-  mutate(focal_id = stan_data_ave_first$id)
-
-games_won <- y_rep_mod |> 
-  pivot_longer(cols = `1`:`4000`, names_to = "draw", values_to = "y") |> 
-  group_by(focal_id, draw) |> 
-  summarise(games_won = sum(y))
-
-orig_data <- tibble(outcome = stan_data_ave_first$y,
-                    focal_id = stan_data_ave_first$id)
-
-orig_games_won <- orig_data |> 
-  group_by(focal_id) |> 
-  summarise(games_won = sum(outcome)) |> 
-  mutate(focal_id = as.character(focal_id)) |> 
-  left_join(player_id, by = c("focal_id" = "id"))
-
-p1 <- games_won |> 
-  mutate(focal_id = as.character(focal_id)) |> 
-  left_join(player_id, by = c("focal_id" = "id")) |> 
-  ggplot(aes(x = games_won)) +
-  geom_histogram() +
-  facet_wrap(~player, scales = "free", ncol = 5) +
-  geom_vline(data = orig_games_won, 
-             mapping = aes(xintercept = games_won), col = "red") +
-  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
-  # labs(title = "Posterior Predictive Distribution, Initial Games",
-  labs(y = element_blank(),
-       x = "Number of Games Won") +
-  theme_5col() +
-  theme(axis.title.x = element_text(size = 10,
-                                    margin = margin(t = 10, b = 10)))
-
-p1
-
-ggsave(plot = p1, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "1700_1900_ppc_1.png")), dpi = 600,
-       height = 5, width = 7)
-
-```
-
-```{r ppc2 for 17-19}
-#| eval: false
-## time varying ppc for 1700-1900 players
-draws <- fit_17_19_first$draws() |> as_draws_df() |>
-  select(starts_with(c("beta[", "gamma"))) |>
-  mutate(draw = row_number()) |> 
-  pivot_longer(cols = "beta[1,1]":"beta[2,10]") |> 
-  mutate(param = stringr::str_extract(name, pattern = "\\d"),
-         id = stringr::str_extract(name, pattern = "\\d+]"),
-         id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]")) |> 
-  select(-name) |> 
-  pivot_wider(names_from = param, values_from = value) |> 
-  rename(beta1 = `1`, beta2 = `2`)
-
-stan_data_ave_last <- readRDS(file = here(data_path, "time",
-                                          "stan_data_last.RDS"))
-
-future_games <- tibble(focal = as.character(stan_data_ave_last$id), 
-       color = stan_data_ave_last$colour, 
-       elo_diff = stan_data_ave_last$elo,
-       hist = stan_data_ave_last$win_prop)
-
-ppc_data <- readRDS(file = here(data_path, "time",
-                                "ppc_fit_first_last.RDS"))
-
-future_results <- tibble(focal = as.character(stan_data_ave_last$id),
-                         result = stan_data_ave_last$y) |> 
-  group_by(focal) |>
-  summarise(total = sum(result)) |> 
-  rename(id = focal) |> 
-  left_join(player_id, by = "id")
-
-last_ppc <- ppc_data |>  
-  left_join(player_id, by = "id") |> 
-  ggplot(aes(x = games_won)) + 
-  geom_histogram() + 
-  geom_vline(data = future_results, 
-             mapping = aes(xintercept = total), col = "red") +
-  facet_wrap(~player, scales = "free_x", ncol = 5) +
-  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
-  labs(y = element_blank(), x = "Number of Games Won") +
-  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
-  theme_5col() +
-  theme(axis.title.x = element_text(size = 10,
-                                    margin = margin(t = 10, 
-                                                    b = 10,
-                                                    unit = "pt")))
-
-last_ppc
-
-ggsave(plot = last_ppc, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "1700_1900_ppc_2.png")), dpi = 600,
-       height = 5, width = 7)
-```
-
 ### Permuting for 1700-1900
@@
 perm_17_19 <- perm_files[1:1000] |>
-  map_dfr(~read_rds(here(perm_data_path, .x)))
+  map_dfr(~ read_rds(here(perm_data_path, .x)))
 
@@
 
-fit_17_19_bullet_prev <- readRDS(file = paste0(data_path,
-                                              "/all_rated_bullet_model_prev.RDS"))
-
-bullet_17_19 <- readRDS(file = paste0(data_path,
-                                      "/users_bullet.RDS"))
+fit_17_19_bullet_prev <- readRDS(file = paste0(
+  data_path,
+  "/all_rated_bullet_model_prev.RDS"
+))
+
+bullet_17_19 <- readRDS(file = paste0(
+  data_path,
+  "/users_bullet.RDS"
+))
 
@@
 perm_17_19_global <- sim_pars |>
-  filter(variable %in% c("mu_beta",  "gamma1", "gamma2",
-                           "sigma_1", "tau[1]", "tau[2]",
-                           "sigma_g1", "sigma_g2")) |> 
+  filter(variable %in% c(
+    "mu_beta", "gamma1", "gamma2")) |>
   ggplot(aes(mean)) +
@@
   facet_wrap(~variable, scales = "free") +
-  geom_vline(data = true_values,
+  geom_vline(
+    data = true_values,
     mapping = aes(xintercept = mean),
-             col = "red") +
+    col = "red"
+  ) +
   theme_3col() +
@@
 
-ggsave(plot = perm_17_19_global, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "perm_17_19_global.png")), dpi = 600,
-       height = 5, width = 7)
-```
-
-```{r perm winner effects}
-#| eval: false
+ggsave(
+  plot = perm_17_19_global,
+  filename = paste0(here(
+    "Paper_WriteUp", "paper_figures",
+    "perm_17_19_global.png"
+  )), dpi = 600,
+  height = 5, width = 7
+)
+```
+
+```{r figure 8 left - perm winner effects}
 ## get the selected users for this, top 20 players
@@
 
-perm_17_19_winner <- player_pars |> 
-  filter(param == 2) |> 
-  ggplot(aes(mean)) +
-  geom_histogram() +
-  facet_wrap(~player_name, ncol = 5, scales = "free") +
-  labs(x = element_blank(), y = element_blank()) +
-  geom_vline(data = true_beta |> filter(param == 2), aes(xintercept = mean),
-             col = "red") +
-  theme_5col()
-
-perm_17_19_winner
-
-ggsave(plot = perm_17_19_winner, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "perm_17_19_winner.png")), dpi = 600,
-       height = 5, width = 7)
-
 levels <- true_beta |> 
@@
 
-```{r ppc1 and ppc2 for gm}
-#| eval: false
-## just do ppc plots for now
-path_id <- 4
-data_path <- res_data_path[path_id]
-fit_gm_first <- readRDS(file =  here(data_path, "time",
-                                        "select_users_first_bullet.RDS"))
-fit_gm_last <- readRDS(file =  here(data_path, "time",
-                                        "select_users_last_bullet.RDS"))
-
-fit_samples <- as_draws_df(fit_gm_first$draws())
-
-fit_samp <- fit_samples |> 
-  select(!starts_with(c("log_lik", "yrep")))
-
-y_rep <- fit_samples |> select(starts_with("y_rep"))
-
-rm(fit_samples)
-rm(fit_samp)
-
-stan_data_ave_first <- readRDS(file = here(data_path, "time",
-                                           "stan_data_first.RDS"))
-
-y_rep_mod <- y_rep |> 
-  rownames_to_column() |>  
-  pivot_longer(-rowname) |> 
-  pivot_wider(names_from=rowname, values_from=value) |> 
-  ## remove the draws and chains here
-  mutate(focal_id = stan_data_ave_first$id)
-
-games_won <- y_rep_mod |> 
-  pivot_longer(cols = `1`:`4000`, names_to = "draw", values_to = "y") |> 
-  group_by(focal_id, draw) |> 
-  summarise(games_won = sum(y))
-
-orig_data <- tibble(outcome = stan_data_ave_first$y,
-                    focal_id = stan_data_ave_first$id)
-
-users <- readRDS(file = here(data_path, "time", "select_users_bullet.RDS"))
-
-## need to give these the correct names here
-player_id <- tibble(id = as.character(1:10), player = users)
-
-orig_games_won <- orig_data |> 
-  group_by(focal_id) |> 
-  summarise(games_won = sum(outcome)) |> 
-  mutate(focal_id = as.character(focal_id)) |> 
-  left_join(player_id, by = c("focal_id" = "id"))
-
-p1 <- games_won |> 
-  mutate(focal_id = as.character(focal_id)) |> 
-  left_join(player_id, by = c("focal_id" = "id")) |> 
-  ggplot(aes(x = games_won)) +
-  geom_histogram() +
-  facet_wrap(~player, scales = "free", ncol = 5) +
-  geom_vline(data = orig_games_won, 
-             mapping = aes(xintercept = games_won), col = "red") +
-  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
-  # labs(title = "Posterior Predictive Distribution, Initial Games",
-  labs(y = element_blank(),
-       x = "Number of Games Won") +
-  theme_5col() +
-  theme(axis.title.x = element_text(size = 10,
-                                   margin = margin(t = 10, b = 10, unit = "pt")),
-        panel.spacing.x = unit(0.7, "lines"))
-
-p1
-
-ggsave(plot = p1, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "gm_ppc_1.png")), dpi = 600,
-       height = 5, width = 7)
-
-draws <- fit_gm_first$draws() |> as_draws_df() |>
-  select(starts_with(c("beta[", "gamma"))) |>
-  mutate(draw = row_number()) |> 
-  pivot_longer(cols = "beta[1,1]":"beta[2,10]") |> 
-  mutate(param = stringr::str_extract(name, pattern = "\\d"),
-         id = stringr::str_extract(name, pattern = "\\d+]"),
-         id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]")) |> 
-  select(-name) |> 
-  pivot_wider(names_from = param, values_from = value) |> 
-  rename(beta1 = `1`, beta2 = `2`)
-
-stan_data_ave_last <- readRDS(file = here(data_path, "time",
-                                          "stan_data_last.RDS"))
-
-future_games <- tibble(focal = as.character(stan_data_ave_last$id), 
-       color = stan_data_ave_last$colour, 
-       elo_diff = stan_data_ave_last$elo,
-       hist = stan_data_ave_last$win_prop)
-
-ppc_data <- readRDS(file = here(data_path, "time",
-                                "ppc_fit_first_last.RDS"))
-
-future_results <- tibble(focal = as.character(stan_data_ave_last$id),
-                         result = stan_data_ave_last$y) |> 
-  group_by(focal) |>
-  summarise(total = sum(result)) |> 
-  rename(id = focal) |> 
-  left_join(player_id, by = "id")
-
-last_ppc <- ppc_data |>  
-  left_join(player_id, by = "id") |> 
-  ggplot(aes(x = games_won)) + 
-  geom_histogram() + 
-  geom_vline(data = future_results, 
-             mapping = aes(xintercept = total), col = "red") +
-  facet_wrap(~player, scales = "free_x", ncol = 5) +
-  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
-  labs(y = element_blank(), x = "Number of Games Won") +
-  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
-  theme_5col() +
-  theme(axis.title.x = element_text(size = 10,
-                                   margin = margin(t = 10,
-                                                   b = 10,
-                                                   unit = "pt")))
-
-last_ppc
-
-ggsave(plot = last_ppc, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "gm_ppc_2.png")), dpi = 600,
-       height = 5, width = 7)
-```
-
 ### Permuting for GM Data
 
-```{r}
-#| eval: false
+```{r figure 8 right}
 perm_data_path <- here("results/lichessGrandmasters/perm/")
@@
   facet_wrap(~variable, scales = "free") +
-  geom_vline(data = true_values, mapping = aes(xintercept = mean), col = "red") +
+  geom_vline(data = true_values, mapping = aes(xintercept = mean),
+             col = "red") +
   theme_3col() +
@@
 
-perm_gm_winner <- player_pars |> 
-  filter(param == 2) |> 
-  ggplot(aes(mean)) +
-  geom_histogram() +
-  facet_wrap(~player_name, ncol = 5, scales = "free") +
-  labs(x = element_blank(), y = element_blank()) +
-  geom_vline(data = true_beta |> filter(param == 2), aes(xintercept = mean),
-             col = "red") +
-  theme_5col() +
-  scale_x_continuous(breaks = breaks_pretty(n = 3))
+levels <- true_beta |> 
+  filter(param == 2) |> 
+  arrange(mean) |> 
+  pull(player_name)
+
+true_vals <- true_beta |> 
+  filter(param == 2) |>  
+  mutate(y = row_number() - 0.5,
+         yend = y + 1)
 
@@
 
+levels <- true_beta |> 
+  filter(param == 1) |> 
+  arrange(mean) |> 
+  pull(player_name)
+
+true_vals <- true_beta |> 
+  filter(param == 1) |>  
+  mutate(y = row_number() - 0.5,
+         yend = y + 1)
+
 perm_gm_player <- player_pars |> 
   filter(param == 1) |>
-  ggplot(aes(mean)) +
-  geom_histogram() +
-  facet_wrap(~player_name, ncol = 5, scales = "free") +
-  labs(x = element_blank(), y = element_blank()) +
-  geom_vline(data = true_beta |> filter(param == 2), aes(xintercept = mean),
-             col = "red") +
-  theme_5col() +
-  scale_x_continuous(breaks = breaks_pretty(n = 3))
+  mutate(y = factor(player_name, levels = levels)) |> 
+  ggplot(aes(x = mean, y = y)) +
+  stat_histinterval(slab_fill = NA, .width = c(0.5, 0.95)) +
+  geom_segment(data = true_vals, 
+               aes(x = mean, y = y, yend = yend), col = "red") +
+  theme_single_y() +
+  labs(x = "Estimated Player Effect", y = "")
 
@@
 
-## should we make this 12 players so it looks a bit better?
+```
+
+## Posterior Predictive Plots
+
+```{r}
+
 ```

### 2025-01-20 owen/cluster_scripts/Winner_Effects_Time.R
adding the initial ppd plots for the rating here, to make sure they work

@@ -34,23 +34,19 @@ all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_data_path[4] <- here("box_data/lichessGrandmasters/")
 
-
-all_save_path[1] <- here("results/lichess1700-1900_test/")
-all_save_path[2] <- here("results/lichess2000-2200_test/")
-all_save_path[3] <- here("results/lichess2300-2500_test/")
-all_save_path[4] <- here("results/lichessGrandmasters_test/")
-
+all_save_path[1] <- here("results/lichess1700-1900_pred/")
+all_save_path[2] <- here("results/lichess2000-2200_pred/")
+all_save_path[3] <- here("results/lichess2300-2500_pred/")
+all_save_path[4] <- here("results/lichessGrandmasters_pred/")
 
 data_path <- all_data_path[path_id]
 save_path <- paste0(all_save_path[path_id], "time/")
 
-dir.create(save_path, showWarnings = FALSE)
+dir.create(save_path)
 
 files <- list.files(data_path)
-
-
-lichess_data <- files %>% 
-  map_dfr(~read_player(data_path, .x))
+lichess_data <- files |>
+  map_dfr(~ read_player(data_path, .x))
 
 print("------------")
 print(dim(lichess_data))
@@ -60,68 +56,71 @@ print("------------")
 ## this also removes the NAs, which makes sense
 
 
-small_data <- lichess_data %>%
-  mutate(Event = tolower(Event)) |> 
-  filter(TimeControl == "60+0") %>%
-  filter(Variant == "Standard") %>%
+small_data <- lichess_data |>
+  mutate(Event = tolower(Event)) |>
+  filter(TimeControl == "60+0") |>
+  filter(Variant == "Standard") |>
   filter(grepl("rated bullet game", Event))
-# 
-# small_data <- lichess_data %>%
-#   # filter(Event == "Rated Bullet game") %>%
-#   # filter(TimeControl == "60+0") %>%
-#   filter(Variant == "Standard") %>%
+#
+# small_data <- lichess_data |>
+#   # filter(Event == "Rated Bullet game") |>
+#   # filter(TimeControl == "60+0") |>
+#   filter(Variant == "Standard") |>
 #   filter(grepl("Rated Blitz game", Event))
 
 rm(lichess_data)
 
 ## then identify data for start and finish
 
-select_users <- small_data %>% 
-  group_by(Username) %>% 
-  tally() %>% 
[…]

### 2025-01-24 Paper_WriteUp/Paper_Plots.qmd
add some supplement plots

@@
 res_data_path <- rep(NA, 4)
-res_data_path[1] <- here("results/lichess1700-1900/")
-res_data_path[2] <- here("results/lichess2000-2200/")
-res_data_path[3] <- here("results/lichess2300-2500/")
-res_data_path[4] <- here("results/lichessGrandmasters/")
+res_data_path[1] <- here("results_2025/lichess1700-1900/")
+res_data_path[2] <- here("results_2025/lichess2000-2200/")
+res_data_path[3] <- here("results_2025/lichess2300-2500/")
+res_data_path[4] <- here("results_2025/lichessGrandmasters/")
 
@@
 ```{r figure 3 left - plot global parameters for 17-19}
-data_path <- here("results/lichess1700-1900")
+data_path <- here("results_2025/lichess1700-1900")
 
@@
                                    paste0(data_path,
-                                          "/all_rated_bullet_model_prev.RDS"))
+                                          "/all_rated_bullet_model_prev_n1.RDS"))
 
@@
   filter(param == 2) |> 
-  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
+  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |> 
+  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
   stat_histinterval() +
-  labs(x = "Estimated W/L-Effect", y = element_blank(), 
+  labs(x = "Change in Win Probability", y = element_blank(), 
        title = "1700-1900 Cohort") +
@@
 
-```{r plot player effects}
+```{r figure s4 left - plot player effects}
 #| eval: false
@@
        filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "17_19_prev_indiv.png")), dpi = 600,
+                              "Figure_S4_left.png")), dpi = 600,
        height = 7, width = 7)
@@
 ```{r figure 3 right - global parameters for gm data}
-data_path <- here("results/lichessGrandmasters")
+data_path <- here("results_2025/lichessGrandmasters")
 
 fit_gm_bullet_prev <- 
-  readRDS(file = paste0(data_path,"/all_rated_bullet_model_prev.RDS"))
+  readRDS(file = paste0(data_path,"/all_rated_bullet_model_prev_n1.RDS"))
 
@@
   filter(param == 2) |> 
-  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
+  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |>
+  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
   stat_histinterval() +
@@
 
-```{r plot player effects}
+```{r figure s4 right - plot player effects}
 #| eval: false
@@
        filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "gm_prev_indiv.png")), dpi = 600,
+                              "Figure_S4_right.png")), dpi = 600,
        height = 7, width = 7)
@@
 ## global effects first
-data_path <- here("results/lichessGrandmasters")
+data_path <- here("results_2025/lichessGrandmasters")
 fit_gm_bullet_10 <- 
-  readRDS(file =paste0(data_path,"/all_rated_bullet_model.RDS"))
+  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n10.RDS"))
 
@@
 ## Potential Supplement Plots
-data_path <- here("results/lichess1700-1900")
+data_path <- here("results_2025/lichess1700-1900")
 fit_17_19_bullet_prev <- readRDS(file = paste0(
   data_path,
-  "/all_rated_bullet_model_prev.RDS"
+  "/all_rated_bullet_model_prev_n1.RDS"
 ))
@@
   data_path,
-  "/all_rated_blitz_model_prev.RDS"
+  "/all_rated_blitz_model_prev_n1.RDS"
 ))
@@
 ```{r figure 5 right - bullet v blitz gm}
-data_path <- here("results/lichessGrandmasters")
+data_path <- here("results_2025/lichessGrandmasters")
 fit_gm_bullet_prev <- readRDS(file = paste0(
   data_path,
-  "/all_rated_bullet_model_prev.RDS"
+  "/all_rated_bullet_model_prev_n1.RDS"
 ))
@@
 
-<!-- TO DO - Update Below here -->
-<!-- Everything down as far as here up to date. -->
-<!-- setting everything below to eval: false for now -->
-<!-- Updating everything below here -->
-
 # Section 5: Model Validation
@@
 
+<!-- TO DO - Tidy up below here -->
+
 ## Posterior Predictive Plots
@@
 
-## 2000-2200 Previous Game
-
-```{r}
+## 2000-2200 Previous Game Winner/Loser Effect
+
+```{r figure s1-left}
 #| eval: false
-data_path <- here("results/lichess2000-2200")
-
-fit_20_22_bullet_prev <- 
-  readRDS(file =paste0(data_path,"/all_rated_bullet_model_prev.RDS"))
-
-fit_20_22_bullet_prev$summary(variables = c("mu_beta",
-                                            "gamma1", "gamma2",
-                                            "sigma_1", "tau[1]", "tau[2]",
-                                            "sigma_g1", "sigma_g2"))
-```
-
-## 2300-2500 Previous Game
-
-```{r}
+data_path <- here("results_2025/lichess2000-2200")
+
+fit_20_22_bullet_prev <- readRDS(file = 
+                                   paste0(data_path,
+                                          "/all_rated_bullet_model_prev_n1.RDS"))
+
+bullet_20_22 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
+
+player_id <- as_tibble(bullet_20_22) |> 
+  mutate(id = row_number()) |> 
+  rename(player_name = value)
+
+player_pars <- fit_20_22_bullet_prev$draws() |> 
+  as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]")) |> 
+  mutate(id = as.numeric(id),
+         param = as.numeric(param)) |> 
+  left_join(player_id, by = "id")
+
+winner_plot_20_22 <- player_pars |> 
+  filter(param == 2) |> 
+  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |> 
+  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
+  stat_histinterval() +
+  labs(x = "Change in Win Probability", y = element_blank(), 
+       title = "2000-2200 Cohort") +
+  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
+  theme_single_y()
+
+ggsave(plot = winner_plot_20_22, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "Figure_S1_left.png")), dpi = 600,
+       height = 7, width = 7)
+```
+
+## 2300-2500 Previous Game Winner/Loser Effect
+
+```{r figure-s1-right}
 #| eval: false
-data_path <- here("results/lichess2300-2500")
-
-fit_23_25_bullet_prev <- 
-  readRDS(file =paste0(data_path,"/all_rated_bullet_model_prev.RDS"))
-
-fit_23_25_bullet_prev$summary(variables = c("mu_beta",
-                                            "gamma1", "gamma2",
-                                            "sigma_1", "tau[1]", "tau[2]",
-                                            "sigma_g1", "sigma_g2"))
+data_path <- here("results_2025/lichess2300-2500")
+
+fit_23_25_bullet_prev <- readRDS(file = 
+                                   paste0(data_path,
+                                          "/all_rated_bullet_model_prev_n1.RDS"))
+
+bullet_23_25 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
+
+player_id <- as_tibble(bullet_23_25) |> 
+  mutate(id = row_number()) |> 
+  rename(player_name = value)
+
+player_pars <- fit_23_25_bullet_prev$draws() |> 
+  as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]")) |> 
+  mutate(id = as.numeric(id),
+         param = as.numeric(param)) |> 
+  left_join(player_id, by = "id")
+
+winner_plot_23_25 <- player_pars |> 
+  filter(param == 2) |> 
+  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |> 
+  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
+  stat_histinterval() +
+  labs(x = "Change in Win Probability", y = element_blank(), 
+       title = "2300-2500 Cohort") +
+  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
+  theme_single_y()
+
+ggsave(plot = winner_plot_23_25, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "Figure_S1_right.png")), dpi = 600,
+       height = 7, width = 7)
 ```
@@
 
-## Model Checking for the Other Cohorts
-
-<!-- What about for the other two groups? -->
-
-<!-- __To be added in__ -->
-
-<!-- ![Time-Varying PPC for 2000-2200](../results/lichess2000-2200/time/ppc_fit_first_last_bullet.png) -->
-
-<!-- ![Time-Varying PPC for 2300-2500](../results/lichess2300-2500/time/ppc_fit_first_last_bullet.png) -->
-
-<!-- Maybe some very limited evidence that this becomes less -->
-<!-- reasonable for players with a higher rating. -->
+To be added here for 1700-1900 and GM cohorts.
+
+```{r figure S3 left}
+## global effects first
+data_path <- here("results_2025/lichess1700-1900")
+
+fit_17_19_bullet_5 <- 
+  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n5.RDS"))
+
+fit_17_19_bullet_10 <- 
+  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n10.RDS"))
+
+prev_pars <- fit_17_19_bullet_prev$draws(c("mu_beta",
+                                        "gamma1", "gamma2")) |> 
+  as_draws_df() |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |> 
+  mutate(model = "Prev. Game")
+
+pre_5_pars <- fit_17_19_bullet_5$draws(c("mu_beta",
+                                        "gamma1", "gamma2")) |> 
+  as_draws_df() |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |> 
+  mutate(model = "Prev. 5 Games")
+
+pre_10_pars <- fit_17_19_bullet_10$draws(c("mu_beta",
+                                        "gamma1", "gamma2")) |> 
+  as_draws_df() |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |> 
+  mutate(model = "Prev. 10 Games")
+
+hist_comp_global <- bind_rows(prev_pars, pre_5_pars, pre_10_pars) |> 
+  ggplot(aes(x = value, color = model)) +
+  stat_pointinterval(position = position_dodge(width = 0.2)) +
+  facet_wrap(~name,
+             scales = "free",
+             labeller = labeller(name = facet_labels)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_3col() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_color_viridis_d(option = "H")
+
+ggsave(plot = hist_comp_global, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "Figure_S3_left.png")), dpi = 600,
+       height = 3.5, width = 7)
+
+```
+
+```{r figure S3 right}
+## global effects first
+data_path <- here("results_2025/lichessGrandmasters")
+
+fit_gm_bullet_5 <- 
+  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n5.RDS"))
+
+fit_gm_bullet_10 <- 
+  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n10.RDS"))
+
+prev_pars <- fit_gm_bullet_prev$draws(c("mu_beta",
+                                        "gamma1", "gamma2")) |> 
+  as_draws_df() |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |> 
+  mutate(model = "Prev. Game")
+
+pre_5_pars <- fit_gm_bullet_5$draws(c("mu_beta",
+                                        "gamma1", "gamma2")) |> 
+  as_draws_df() |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |> 
+  mutate(model = "Prev. 5 Games")
+
+pre_10_pars <- fit_gm_bullet_10$draws(c("mu_beta",
+                                        "gamma1", "gamma2")) |> 
+  as_draws_df() |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |> 
+  mutate(model = "Prev. 10 Games")
+
+hist_comp_global <- bind_rows(prev_pars, pre_5_pars, pre_10_pars) |> 
+  ggplot(aes(x = value, color = model)) +
+  stat_pointinterval(position = position_dodge(width = 0.2)) +
+  facet_wrap(~name,
+             scales = "free",
+             labeller = labeller(name = facet_labels)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_3col() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_color_viridis_d(option = "H")
+
+ggsave(plot = hist_comp_global, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "Figure_S3_right.png")), dpi = 600,
+       height = 3.5, width = 7)
+
+```

### 2025-02-13 owen/cluster_scripts/perm_results.R
update save path for this

@@ -31,16 +31,16 @@ all_data_path[1] <- here("box_data/lichess1700-1900/")
 all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_data_path[4] <- here("box_data/lichessGrandmasters/")
-all_save_path[1] <- here("results/lichess1700-1900_test/")
+all_save_path[1] <- here("results/lichess1700-1900_2025/")
 
 # all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
 # ## if need to run it locally
 
-all_save_path[2] <- here("results/lichess2000-2200_test/")
-all_save_path[3] <- here("results/lichess2300-2500_test/")
-all_save_path[4] <- here("results/lichessGrandmasters_test/")
+all_save_path[2] <- here("results/lichess2000-2200_2025/")
+all_save_path[3] <- here("results/lichess2300-2500_2025/")
+all_save_path[4] <- here("results/lichessGrandmasters_2025/")
 
-path_id <- 1 ## setting it for now
+path_id <- 4 ## setting it for now
 data_path <- all_data_path[path_id]
 save_path <- paste0(all_save_path[path_id], "perm/")
 

### 2025-02-13 Paper_WriteUp/Paper_Plots.qmd

@@
   stat_histinterval() +
-  labs(x = "Estimated Player Effect", y = element_blank()) +
+  labs(x = "Estimated Player Effect", y = element_blank(),
+       title = "1700-1900 Cohort") +
   geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
@@
 true_values <- fit_gm_bullet_prev$summary() |> 
-  filter(variable %in% c("mu_beta", "gamma1", "gamma2")) |> 
+  dplyr::filter(variable %in% c("mu_beta", "gamma1", "gamma2")) |> 
   select(variable, mean)
@@
 perm_gm_global <- sim_pars |> 
-  filter(variable %in% c("mu_beta",  "gamma1", "gamma2",
-                           "sigma_1", "tau[1]", "tau[2]",
-                           "sigma_g1", "sigma_g2")) |> 
+  dplyr::filter(variable %in% c("mu_beta",  "gamma1", "gamma2")) |> 
   ggplot(aes(mean)) +
@@
   filter(param == 2) |>  
-  mutate(y = row_number() - 0.5,
-         yend = y + 1)
+  mutate(y = factor(player_name, levels = levels)) |> 
+  arrange(y) |> 
+  mutate(y_fake = row_number() - 0.25,
+         yend = row_number() + 0.25)
 
@@
   geom_segment(data = true_vals,
-               aes(x = mean, y = y, yend = yend), col = "red") +
+               aes(x = mean, y = y_fake, yend = yend), col = "red") +
   theme_single_y() +

### 2025-02-13 Paper_WriteUp/Paper_Plots.qmd
update the 1700-1900 perm plot

@@
 ```{r figure 3 left - plot global parameters for 17-19}
-data_path <- here("results_2025/lichess1700-1900")
+data_path <- here("results/lichess1700-1900_2025")
 
@@
 #| eval: false
-perm_data_path <- here("results/lichess1700-1900/perm")
+perm_data_path <- here("results/lichess1700-1900_2025/perm")
 
@@
 
-fit_17_19_bullet_prev <- readRDS(file = paste0(
-  data_path,
-  "/all_rated_bullet_model_prev.RDS"
-))
+data_path <- here("results/lichess1700-1900_2025")
+
+fit_17_19_bullet_prev <- readRDS(file = 
+                                   paste0(data_path,
+                                          "/all_rated_bullet_model_prev_n1.RDS"))
 
@@
 ```{r figure 8 left - perm winner effects}
-## get the selected users for this, top 20 players
-path_id <- 1  ## for 1700-1900
-data_path <- all_data_path[path_id]
-files <- list.files(data_path)
-
-lichess_data <- files |> 
-  map_dfr(~read_player(data_path, .x))
-
-bullet_data <- lichess_data |> 
-  mutate(Event = tolower(Event)) |>
-  filter(TimeControl == "60+0") |>
-  filter(Variant == "Standard") |>
-  filter(grepl("rated bullet game", Event))
-
-select_users <- bullet_data |> 
-  group_by(Username) |> 
-  tally() |> 
-  arrange(-n) |> 
-  ## will change this to n = 20 when run on cluster potentially
-  slice_max(order = n, n = 20) |> 
-  pull(Username)
-
-## then repeat for the winner effects
-player_id <- as_tibble(select_users) |> 
-  mutate(id = row_number()) |> 
-  rename(player_name = value)
 
@@
   filter(param == 2) |>  
-  mutate(y = row_number() - 0.5,
-         yend = y + 1)
+  mutate(y = factor(player_name, levels = levels)) |> 
+  arrange(y) |> 
+  mutate(y_fake = row_number() - 0.5,
+         yend = row_number() + 0.5)
 
@@
   geom_segment(data = true_vals, 
-               aes(x = mean, y = y, yend = yend), col = "red") +
+               aes(x = mean, y = y_fake, yend = yend), col = "red") +
   labs(y = "", x = "Estimated W/L-Effect") +
@@
 ```{r figure 8 right}
-perm_data_path <- here("results/lichessGrandmasters/perm/")
+perm_data_path <- here("results/lichessGrandmasters_2025/perm/")
 
@@
   filter(param == 1) |>  
-  mutate(y = row_number() - 0.5,
-         yend = y + 1)
+  mutate(y = factor(player_name, levels = levels)) |> 
+  arrange(y) |> 
+  mutate(y_fake = row_number() - 0.5,
+         yend = row_number() + 0.5)
 
@@
   geom_segment(data = true_vals, 
-               aes(x = mean, y = y, yend = yend), col = "red") +
+               aes(x = mean, y = y_fake, yend = yend), col = "red") +
   theme_single_y() +

### 2025-02-13 owen/cluster_scripts/perm_results.R
match the priors correctly. we want N(0,1) for scale

@@ -31,14 +31,14 @@ all_data_path[1] <- here("box_data/lichess1700-1900/")
 all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_data_path[4] <- here("box_data/lichessGrandmasters/")
-all_save_path[1] <- here("results/lichess1700-1900_test/")
+all_save_path[1] <- here("results/lichess1700-1900_2025/")
 
 # all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
 # ## if need to run it locally
 
-all_save_path[2] <- here("results/lichess2000-2200_test/")
-all_save_path[3] <- here("results/lichess2300-2500_test/")
-all_save_path[4] <- here("results/lichessGrandmasters_test/")
+all_save_path[2] <- here("results/lichess2000-2200_2025/")
+all_save_path[3] <- here("results/lichess2300-2500_2025/")
+all_save_path[4] <- here("results/lichessGrandmasters_2025/")
 
 path_id <- 1 ## setting it for now
 data_path <- all_data_path[path_id]

### 2025-04-10 owen/pred_glicko.R
add some final files from previous work, before final clean up

@@ -68,8 +68,6 @@ select_users <- small_data |>
 
 
 users <- select_users
-
-
 ## will use last 1000 games each user plays
 
 last_games <- small_data |> 
@@ -102,9 +100,6 @@ init_data <- tidy_games |>
 
 
 ## get their starting scores for each player
-
-
-
 ## then just pick the player from this data
 ## need to choose the focal id and format for the glicko2 function
 
@@ -117,7 +112,6 @@ focal_init <- init_data |> group_by(focal_user) |>
   mutate(focal_elo = ifelse(focal_user == White, WhiteElo, BlackElo)) |> 
   select(focal_user, focal_elo) |> 
   rename(Player = focal_user, Rating = focal_elo)
-
 
 
 glick_data <- init_data |>
@@ -135,8 +129,6 @@ glick_part1 <- glick_data |>
          opp = ifelse(focal_user == White, Black, White)) |> 
   select(index, focal_user, opp, focal_result, focal_white) |> 
   mutate(focal_white = ifelse(focal_white == 0, -1, focal_white))
-
-
 
 
 ## second argument, giving elo of all opponents and starting of first
@@ -218,31 +210,23 @@ draws <- fit_17_19_first$draws() |> as_draws_df() |>
   pivot_wider(names_from = param, values_from = value) |> 
   rename(beta1 = `1`, beta2 = `2`)
 
-
 ## the covariates for the future games
-
 future_games <- tibble(focal = as.character(init_data$focal_id),
                        color = init_data$focal_white,
                        elo_diff = init_data$elo_diff,
                        hist = init_data$ave_prop)
 
-
 ## process for below
-
 focal_games <- future_games |> 
   filter(focal == focal_player_used)
-
 
 full_game_info <- init_data |> 
   filter(focal_id == focal_player_used)
 
 # slow for even 100 sims at the moment, due to the for loop
 num_sims <- 100
-
 num_games <- nrow(full_game_info)
-
 glicko_sim <- data.frame(NA, nrow = num_games, ncol = num_sims)
-
 game_sim <- data.frame(NA, nrow = num_games, ncol = num_sims)
 
 
@@ -310,14 +294,9 @@ for(i in 1:num_games){
   
 }
 
-
 ## quick plot to confirm
-
 quick_sim <- glicko_sim[, 1:100]  
-
 colnames(quick_sim) <- paste0("Sim_", 1:100)
-
-
[…]

### 2025-04-10 lichess_model.R
remove old original code, not used

@@ -1,324 +1,0 @@
-# lichess model
-library(tidyverse)
-setwd("/Users/mia/Desktop/lichess")
-
-ldata <- readRDS("rdata/lichess_pilot.Rdata")
-
-# is the focal player white or black?
-ldata$w <- ifelse(ldata$Username == ldata$White, 1, 0)
-
-# result column
-ldata$result <- ifelse(ldata$w == 1 & ldata$Result == "1-0" |
-                         ldata$w == 0 & ldata$Result == "0-1", 1, 0)
-
-
-
-# white/black elo adjustments
-ldata$elof <- ifelse(ldata$w == 1, ldata$WhiteElo, ldata$BlackElo)
-ldata$elonf <- ifelse(ldata$w == 0, ldata$WhiteElo, ldata$BlackElo)
-
-## focal and non focal, if focal is black then nonfocal is white
-
-
-# building a regression model
-plot_data <- ldata[c(1:500), ]
-
-ggplot(plot_data, aes(x = UTCDate, y = result)) +
-  geom_point()
-
-ldata$elof <- as.numeric(ldata$elof)
-ldata$elonf <- as.numeric(ldata$elonf)
-ldata$w <- as.numeric(ldata$w)
-ldata1 <- ldata[c(1:1000), ]
-m1 <- glm(result ~ elof + elonf + w + Event, family = binomial(), data = ldata1)
-summary(m1)
-plot(m1)
-
-
-# testing the model
-options(scipen = 999)
-probabilities <- m1 %>% predict(ldata1, type = "response")
-probabilities <- as.data.frame(probabilities)
-
-## are these used again then?
-
-# fish data WL Effect model fit to this data
-l.data <- data.frame(
-  winner = ifelse(ldata$Result == "1-0", ldata$White, ldata$Black),
-  loser = ifelse(ldata$Result == "0-1", ldata$White, ldata$Black)
-)
-
-l.data$date <- ldata$UTCDate
-l.data$type <- 0
-l.data$observation <- 0
-l.data$group <- 0
-
-
-## will fill these in based on other stuff below
-
-l.data <- l.data[, c(6, 3, 5, 1, 2, 4)] 
-l.data <- l.data[c(1:1000), ]
-
-
-## what does this function do?
-## for specific values of the coefficients, does something?
-## need to see a specific example below
-
-
-get_previous.outcome.predictions <- function(model, parameter.indexes) {
[…]

### 2025-04-10 scrape_code/scrape_lichess_clean.R
delete old initial scraping code which is not used

@@ -1,64 +1,0 @@
-
-
-## Install chessR package
-
-# # install.packages("devtools")
-# devtools::install_github("JaseZiv/chessR")
-
-library(chessR)
-library(rvest) # needed for scraping teamnames 
-
-# Note, without API token, we can only retrieve 15 games per second from lichess.
-
-
-## If we just want to get one player's data by username we do the following
-# this player only has 35 games so will be quick to get games
-lgd <- get_raw_lichess("JaseZiv")
-
-# the 18th column are game moves. This is messy, so only keep first 17
-lgd[,1:17]
-
-
-# Note: any way to do this but to not collect the game move information?
-
-
-
-### We need to grab usernames from lichess
-
-# one way is to go to the 'team' pages they have and scrape members
-# however these pages don't dynamically load all names - this only grabs approx 30 names
-# one R solution around this is using RSelenium (haven't tried yet)
-
-url <- "https://lichess.org/team/im-eric-rosen-fan-club/members"
-webpage <- read_html(url)
-player_names <- html_nodes(webpage, "td") %>% html_text()
-ids <- player_names[which(seq_along(player_names) %% 2 == 1)]
-ids
-
-
-## Another way to do this is to select from usernames from players already in our database
-# (also see scraping chess.com R file)
-
-
-
-
-### Loop to collect all games from several players at once
-# however, sometimes we get errors for certain players which stops the loop
-# unknown why these errors occur
-# so probably need to put loop into a try() function so it skips those with errors
-
-# note: only run this if prepare to wait a while to get the games.
-
-res <- vector('list',length(ids))
-ids <- 
-  for(i in 1:length(ids)){
-    res[[i]] <- get_raw_lichess(ids[i])
-  }
-res
-
-lapply(res, nrow)
-
-### To save output:
-# saveRDS(res, "rdata/name_of_file.RData")
-
-

### 2025-04-10 analysis_scripts/pred_glicko.R
tidy the file format of the scripts, place them in a higher directory

@@ -1,0 +1,376 @@
+#### Sept 18th 2024
+#### Construc a PPC of the evolution of the Lichess rating over time
+#### Using complete simulated trajectories of their final 1000 games,
+#### using parameter estimats from model fit to first 1000 games
+
+library(PlayerRatings)
+library(tidyverse)
+library(here)
+library(loo)
+library(ggplot2)
+library(posterior)
+library(RcppRoll)
+
+source(here("analysis/helper.R"))
+
+## get some initial chess data, say the final 1000 games
+## for 20 players and compute the ratings for that
+
+
+
+all_data_path <- rep(NA, 4)
+all_save_path <- rep(NA, 4)
+all_data_path[1] <- here("box_data/lichess1700-1900/")
+all_data_path[2] <- here("box_data/lichess2000-2200/")
+all_data_path[3] <- here("box_data/lichess2300-2500/")
+all_data_path[4] <- here("box_data/lichessGrandmasters/")
+all_save_path[1] <- here("results/lichess1700-1900_test/")
+
+# all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
+# ## if need to run it locally
+
+all_save_path[2] <- here("results/lichess2000-2200_test/")
+all_save_path[3] <- here("results/lichess2300-2500_test/")
+all_save_path[4] <- here("results/lichessGrandmasters_test/")
+
+path_id <- 1  ## setting it for now
+data_path <- all_data_path[path_id]
+save_path <- paste0(all_save_path[path_id], "perm/")
+
+
+res_data_path <- rep(NA, 4)
+res_data_path[1] <- here("results/lichess1700-1900/")
+res_data_path[2] <- here("results/lichess2000-2200/")
+res_data_path[3] <- here("results/lichess2300-2500/")
+res_data_path[4] <- here("results/lichessGrandmasters/")
+
+
+files <- list.files(data_path)
+
+lichess_data <- files |> 
+  map_dfr(~read_player(data_path, .x))
+
+
+small_data <- lichess_data |>
+  mutate(Event = tolower(Event)) |> 
+  filter(TimeControl == "60+0") |>
+  filter(Variant == "Standard") |>
+  filter(grepl("rated bullet game", Event))
+
+rm(lichess_data)
+
+select_users <- small_data |> 
+  group_by(Username) |> 
+  tally() |> 
+  arrange(-n) |> 
+  slice_max(order = n, n = 10) |> 
+  pull(Username)
+
+
[…]

### 2025-04-10 analysis_scripts/01_Large_Fit.R
update the scripts and report file with modified file names, ready to save revisions

@@ -19,11 +19,11 @@ options(mc.cores = parallel::detectCores())
 
 ## source helper functions for reading and transforming data
 ## along with defaults for plots, etc
-source(here("analysis/helper.R"))
+source(here("utils/helper.R"))
 
 path_id <- as.numeric(Sys.getenv("SLURM_ARRAY_TASK_ID"))
 
-n <- 5 ## number of games to use for history
+n <- 10 ## number of games to use for history
 
 ### load in the data to use
 
@@ -33,14 +33,14 @@ all_data_path[1] <- here("box_data/lichess1700-1900/")
 all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_data_path[4] <- here("box_data/lichessGrandmasters/")
-all_save_path[1] <- here("results/lichess1700-1900_2025/")
+all_save_path[1] <- here("results_revision/lichess1700-1900/")
 
 # all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
 # ## if need to run it locally
 
-all_save_path[2] <- here("results/lichess2000-2200_2025/")
-all_save_path[3] <- here("results/lichess2300-2500_2025/")
-all_save_path[4] <- here("results/lichessGrandmasters_2025/")
+all_save_path[2] <- here("results_revision/lichess2000-2200/")
+all_save_path[3] <- here("results_revision/lichess2300-2500/")
+all_save_path[4] <- here("results_revision/lichessGrandmasters/")
 
 
 data_path <- all_data_path[path_id]
@@ -117,7 +117,7 @@ stan_data_ave <- list(N = nrow(init_data),
                       win_prop = init_data$ave_prop)
 
 
-stan_file <- here("owen", "cluster_scripts", "final_model_scale_priors.stan")
+stan_file <- here("analysis_scripts", "final_model_scale_priors.stan")
 
 mod <- cmdstan_model(stan_file)
 

### 2025-04-10 analysis_scripts/02_Large_Single_Prev.R

@@ -32,10 +32,10 @@ all_data_path[1] <- here("box_data/lichess1700-1900/")
 all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_data_path[4] <- here("box_data/lichessGrandmasters/")
-all_save_path[1] <- here("results/lichess1700-1900_2025/")
-all_save_path[2] <- here("results/lichess2000-2200_2025/")
-all_save_path[3] <- here("results/lichess2300-2500_2025/")
-all_save_path[4] <- here("results/lichessGrandmasters_2025/")
+all_save_path[1] <- here("results_revision/lichess1700-1900/")
+all_save_path[2] <- here("results_revision/lichess2000-2200/")
+all_save_path[3] <- here("results_revision/lichess2300-2500/")
+all_save_path[4] <- here("results_revision/lichessGrandmasters/")
 
 
 data_path <- all_data_path[path_id]
@@ -46,23 +46,23 @@ dir.create(save_path, showWarnings = FALSE)
 files <- list.files(data_path)
 # files <- files[1:8]
 
-lichess_data <- files %>% 
+lichess_data <- files |> 
   map_dfr(~read_player(data_path, .x))
 
 ## restrict to rated rapid and shorter here
 ## this also removes the NAs, which makes sense
 
-small_data <- lichess_data %>%
-   # filter(Event == "Rated Bullet game") %>%
+small_data <- lichess_data |>
+   # filter(Event == "Rated Bullet game") |>
    mutate(Event = tolower(Event)) |>
-   filter(TimeControl == "60+0") %>%
-   filter(Variant == "Standard") %>%
+   filter(TimeControl == "60+0") |>
+   filter(Variant == "Standard") |>
    filter(grepl("rated bullet game", Event))
 
-#small_data <- lichess_data %>%
+#small_data <- lichess_data |>
 #  mutate(Event = tolower(Event)) |>
-#  filter(TimeControl == "180+0") %>%
-#  filter(Variant == "Standard") %>%
+#  filter(TimeControl == "180+0") |>
+#  filter(Variant == "Standard") |>
 #  filter(grepl("rated blitz game", Event))
 
 
@@ -70,36 +70,36 @@ users <- unique(small_data$Username)
 
 ## when players play less than 10 games
 ## otherwise not needed
-users <- small_data %>% 
-  group_by(Username) %>% 
[…]

### 2025-04-10 analysis_scripts/03_Fit_Perm.R

@@ -31,14 +31,14 @@ all_data_path[1] <- here("box_data/lichess1700-1900/")
 all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_data_path[4] <- here("box_data/lichessGrandmasters/")
-all_save_path[1] <- here("results/lichess1700-1900_2025/")
+all_save_path[1] <- here("results_revision/lichess1700-1900/")
 
 # all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
 # ## if need to run it locally
 
-all_save_path[2] <- here("results/lichess2000-2200_2025/")
-all_save_path[3] <- here("results/lichess2300-2500_2025/")
-all_save_path[4] <- here("results/lichessGrandmasters_2025/")
+all_save_path[2] <- here("results_revision/lichess2000-2200/")
+all_save_path[3] <- here("results_revision/lichess2300-2500/")
+all_save_path[4] <- here("results_revision/lichessGrandmasters/")
 
 path_id <- 4 ## setting it for now
 data_path <- all_data_path[path_id]
@@ -48,25 +48,25 @@ dir.create(save_path, showWarnings = FALSE)
 
 files <- list.files(data_path)
 
-lichess_data <- files %>% 
+lichess_data <- files |> 
   map_dfr(~read_player(data_path, .x))
 
 
-small_data <- lichess_data %>%
+small_data <- lichess_data |>
   mutate(Event = tolower(Event)) |> 
-  filter(TimeControl == "60+0") %>%
-  filter(Variant == "Standard") %>%
+  filter(TimeControl == "60+0") |>
+  filter(Variant == "Standard") |>
   filter(grepl("rated bullet game", Event))
 
 rm(lichess_data)
 
 
-select_users <- small_data %>% 
-  group_by(Username) %>% 
-  tally() %>% 
-  arrange(-n) %>% 
+select_users <- small_data |> 
+  group_by(Username) |> 
+  tally() |> 
+  arrange(-n) |> 
   ## will change this to n = 20 when run on cluster potentially
-  slice_max(order = n, n = 20) %>% 
+  slice_max(order = n, n = 20) |> 
   pull(Username)
 
 
@@ -75,35 +75,35 @@ select_users <- small_data %>%
 ## just going to use the last games for this
 ## permute within a player
 
[…]

### 2025-04-10 analysis_scripts/03_Fit_Perm.R
update path

@@ -18,7 +18,7 @@ library(ggplot2)
 library(posterior)
 library(RcppRoll)
 
-source(here("analysis/helper.R"))
+source(here("utils", "helper.R"))
 
 jobid <- Sys.getenv("SLURM_ARRAY_TASK_ID")
 jobid <- as.numeric(jobid)

### 2025-04-10 analysis_scripts/05_PPC_Pred_Glicko.R
update this path also

@@ -11,7 +11,7 @@ library(ggplot2)
 library(posterior)
 library(RcppRoll)
 
-source(here("analysis/helper.R"))
+source(here("utils", "helper.R"))
 
 ## get some initial chess data, say the final 1000 games
 ## for 20 players and compute the ratings for that

### 2025-04-10 analysis_scripts/05_PPC_Pred_Glicko.R
modify the two stan files slightly

@@ -16,6 +16,7 @@ source(here("utils", "helper.R"))
 ## get some initial chess data, say the final 1000 games
 ## for 20 players and compute the ratings for that
 
+### make some change here
 
 
 all_data_path <- rep(NA, 4)

### 2025-04-11 utils/plot_templates.R
adding theme for single plot with y-axis and legend

@@ -47,6 +47,24 @@ theme_single_y <- function(){
     )
 }
 
+theme_single_y_legend <- function(){ 
+  theme_bw() %+replace%    #replace elements we want to change
+    
+    theme(
+      
+      #grid elements
+      panel.spacing.x = unit(1, "lines"),
+      
+      #text elements
+      plot.title = element_text(size = title_size),
+      axis.text = element_text(size = axis_text_size - 2),
+      axis.title = element_text(size = axis_title),
+      legend.text = element_text(size = 12),
+      strip.text = element_text(size = axis_title - 5, 
+                                margin = margin(1.5, 1.5, 2, 1.5)),
+      legend.title = element_blank(),
+    )
+}
 
 theme_3col <- function(){ 
   theme_bw() %+replace%    #replace elements we want to change

### 2025-04-11 Paper_WriteUp/Paper_Plots.qmd
Revised figure 1 to a single figure with 3 Gms and 3 1700-1900s. Highlighting Magnus and Alireza

@@
 ## to load some plot defaults
-source(here("analysis", "helper.R"))
-source(here("analysis", "plot_templates.R"))
+source(here("utils", "helper.R"))
+source(here("utils", "plot_templates.R"))
 
@@
 ```{r figure 1}
-## Sydney's code for this figure
-
-player_cohort_1 <- "sandro2502"
-
-path_id <- 1  ## for 1700-1900 
-data_path <- all_data_path[path_id]
-files <- list.files(data_path)
-lichess_data <- files |> 
-  map_dfr(~read_player(data_path, .x))
-
-bullet_data_fig_1 <- lichess_data |> 
-  mutate(Event = tolower(Event)) |>
-  filter(Variant == "Standard") |>
-  filter(grepl("rated bullet game", Event))
-rm(lichess_data)
-
-p1 <- bullet_data |> 
-  filter(Username == player_cohort_1)
-
-p1$newELO <- ifelse(p1$White == paste0(player_cohort_1),
-                    p1$WhiteElo, p1$BlackElo)
-# p1$datetime <- as.POSIXct(paste(p1$UTCDate, p1$UTCTime),
-#                           format = "%Y.%m.%d %H:%M:%S")
-ltypes <- c(1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2, 3)
-thetitle <- paste0("Player From 1700-1900 Cohort")
-linewd <- 1
-ylim1 <- 1200
-ylim2 <- 3500
-ndays <- 1
-
-p1_sorted <- p1 |> arrange(UTCDate, UTCTime)
-p1_sorted$event <- 1:nrow(p1_sorted)
-
-tidy_plot1 <- ggplot(p1_sorted, aes(x = event, y = newELO)) +
-  # scale_colour_manual(values = getPalette(colourCount)) +
-  scale_linetype_manual(values = ltypes) +
-  xlab("Games Played") +
-  ylab("Lichess Rating") +
-  ggtitle(thetitle) +
-  ylim(ylim1, ylim2) +
-  geom_line(lwd = linewd) +
-  theme(
-    plot.title = element_text(hjust = 0, vjust = 1, size = rel(1.7)),
-    panel.background = element_blank(),
-    plot.background = element_blank(),
-    panel.grid.major.y = element_line(color = "gray75", linetype = "dotted"),
-    panel.grid.major.x = element_blank(),
-    panel.grid.minor = element_blank(),
-    strip.background = element_blank(),
-    strip.text = element_text(size = rel(1.1)),
-    text = element_text(color = "gray20", size = 10),
-    axis.text = element_text(size = rel(1.0)),
-    axis.text.x = element_text(color = "gray20", size = rel(1.0)),
-    axis.text.y = element_text(color = "gray20", size = rel(1.0)),
-    axis.title.x = element_text(size = rel(1.0), vjust = 0),
-    axis.title.y = element_text(size = rel(1.0), vjust = 1),
-    axis.ticks.y = element_blank(),
-    axis.ticks.x = element_blank(),
-    legend.position = "none"
-  )
-
-player_cohort_gm <- "alireza2003"
-
-## load in the GM data for this
-path_id <- 4  ## for GM 
-data_path <- all_data_path[path_id]
-files <- list.files(data_path)
-lichess_data <- files |> 
-  map_dfr(~read_player(data_path, .x))
-
-bullet_data_fig_1_gm <- lichess_data |> 
-  mutate(Event = tolower(Event)) |>
-  filter(Variant == "Standard") |>
-  filter(grepl("rated bullet game", Event))
-rm(lichess_data)
-
-p2 <- bullet_data_fig_1_gm |> 
-  filter(Username == player_cohort_gm)
-
-p2$newELO <- ifelse(p2$White == paste0(player_cohort_gm),
-                    p2$WhiteElo, p2$BlackElo)
-thetitle <- paste0("Player From GM Cohort")
-p2_sorted <- p2 |> arrange(UTCDate, UTCTime)
-p2_sorted$event <- 1:nrow(p2_sorted)
-
-tidy_plot2 <- ggplot(p2_sorted, aes(x = event, y = newELO)) +
-  # scale_colour_manual(values = getPalette(colourCount)) +
-  scale_linetype_manual(values = ltypes) +
-  xlab("Games Played") +
-  ylab("Lichess Rating") +
-  ggtitle(thetitle) +
-  ylim(ylim1, ylim2) +
-  geom_line(lwd = linewd) +
-  theme(
-    plot.title = element_text(hjust = 0, vjust = 1, size = rel(1.7)),
-    panel.background = element_blank(),
-    plot.background = element_blank(),
-    panel.grid.major.y = element_line(color = "gray75", linetype = "dotted"),
-    panel.grid.major.x = element_blank(),
-    panel.grid.minor = element_blank(),
-    strip.background = element_blank(),
-    strip.text = element_text(size = rel(1.1)),
-    text = element_text(color = "gray20", size = 10),
-    axis.text = element_text(size = rel(1.0)),
-    axis.text.x = element_text(color = "gray20", size = rel(1.0)),
-    axis.text.y = element_text(color = "gray20", size = rel(1.0)),
-    axis.title.x = element_text(size = rel(1.0), vjust = 0),
-    axis.title.y = element_text(size = rel(1.0), vjust = 1),
-    # axis.ticks.y = element_blank(),
-    axis.ticks.x = element_blank(),
-    legend.position = "none"
-  )
-
-tdy1 <- tidy_plot1 + theme_single_y()
-tdy2 <- tidy_plot2 + theme_single_y()
-
-tidy_plot <- ggarrange(tdy1, tdy2)
-
-ggsave(paste0(newfilename, ".png"), tidy_plot,
-  dpi = 600, height = 5, width = 12
-)
-ggsave(plot = tidy_plot, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "elo_plot_2_players.png")), dpi = 600,
-       height = 5, width = 12)
+#Adam - Revised Apr 11, 2025
+#Updated to a single figure with 3 GMs and 3 1700-1900
+#Highlighting Magnus and Alireza
+
+#import data from 1700-1900 and gm cohorts
+data_path_gm = paste0(all_data_path[4], "/")
+data_path_17_19 = paste0(all_data_path[1], "/")
+files_gm = list.files(data_path_gm, pattern = ".csv")
+files_17_19 = list.files(data_path_17_19, pattern = ".csv")
+
+lichess_data_gm = files_gm %>%
+  map_dfr(~read_player(data_path_gm, .x))
+lichess_data_17_19 = files_17_19 %>%
+  map_dfr(~read_player(data_path_17_19, .x))
+
+#combine cohorts
+lichess_data = rbind(lichess_data_17_19, lichess_data_gm)
+
+#selecting 3 users from each cohort that have around 10,000 games
+users = c("larrywheels", "SeanBambic", "dewang_007", #1700-1900s
+          "DrNykterstein", "alireza2003", "BlueGreensun") #gms
+
+#data for plot
+plot_data = lichess_data %>%
+  mutate(Event = tolower(Event)) %>%
+  filter(Event == "rated bullet game", Variant == "Standard", Username %in% users) %>%
+  mutate(focal_rating = ifelse(Username == White, WhiteElo, BlackElo)) %>%
+  group_by(Username) %>%
+  arrange(UTCDate, UTCTime, .by_group = TRUE) %>%
+  mutate(game = row_number(),
+         Username = case_when( #rename magnus and alireza usernames
+           Username == "DrNykterstein" ~ "DrNykterstein (Magnus Carlsen)",
+           Username == "alireza2003" ~ "alireza2003 (Alireza Firouzja)",
+           .default = Username
+         ),
+         Username = factor(Username, levels = c("DrNykterstein (Magnus Carlsen)", "alireza2003 (Alireza Firouzja)", 
+                                                "BlueGreensun", "SeanBambic", "dewang_007", "larrywheels"))) %>% #for order in legend
+  ungroup() %>%
+  select(Username, game, focal_rating)
+
+#plotting
+rating_plot = ggplot(data = plot_data, mapping = aes(x = game, y = focal_rating, colour = Username)) +
+  geom_line() +
+  labs(x = "Games Played", y = "Rating") + 
+  theme_single_y_legend()
+rating_plot
+
+#save
+ggsave(plot = rating_plot, 
+       filename = here("Paper_WriteUp", "paper_figures", "fig1.png"), dpi = 1000,
+       height = 4.5, width = 10)
 ```

### 2025-04-11 Paper_WriteUp/Paper_Plots.qmd
adding numbers to y-axis in fig2

@@
 ```{r figure 2}
-## Adam's code for this figure
-
-gm_players_elo <- list()
+#Adam - revised Apr 11, 2025
+#Adding numbers to y-axis
+
+#import data from gm cohort
+data_path_gm = paste0(all_data_path[4], "/")
+files_gm = list.files(data_path_gm, pattern = ".csv")
+lichess_data_gm = files_gm %>%
+  map_dfr(~read_player(data_path_gm, .x))
+
+bullet_data_gm = lichess_data_gm |> 
+  mutate(Event = tolower(Event)) |>
+  filter(TimeControl == "60+0", Variant == "Standard", Event == "rated bullet game")
+
 gm_session_lengths <- list()
@@
   # tidy_games_gms <- tidy_games_gms[!duplicated(tidy_games_gms[1:9]), ] # remove duplicate rows
-
-  # get players average elo across all games
-  gm_players_elo[[gm_players[i]]] <- mean(c(
-    tidy_games_gms$WhiteElo[gm_players[i] == tidy_games_gms$White],
-    tidy_games_gms$BlackElo[gm_players[i] == tidy_games_gms$Black]
-  ))
 
@@
   ylab("Frequency") #+ ggtitle("Time Difference Between Current Game and Previous Game for Amateurs")
-fig2a <- log_time_diff_plot_gm + theme_single()
+fig2a <- log_time_diff_plot_gm + theme_single_y()
 
@@
   ylab("Frequency") #+ ggtitle("Number of Games per Session for GMs")
-fig2b <- session_length_plot_gm + theme_single()
+fig2b <- session_length_plot_gm + theme_single_y()
 

### 2025-04-11 Paper_WriteUp/Paper_Plots.qmd
updating height/width and removing 5game in fig2b

@@
                               "time_between_games_gm.png")), dpi = 600,
-       height = 5, width = 7)
+       height = 4, width = 6.25)
 
@@
     trans = pseudo_log_trans(sigma = 2.36),
-    breaks = c(1, 5, 10, 100),
-    labels = c("1 game", "5 games", "10 games", "100 games"),
+    breaks = c(1, 10, 100),
+    labels = c("1 game", "10 games", "100 games"),
     limits = c(0, 389)
@@
                               "session_length_gm.png")), dpi = 600,
-       height = 5, width = 7)
+       height = 4, width = 6.25)
 ```

### 2025-04-11 Paper_WriteUp/Paper_Plots.qmd
making fig4 x-axis same length fr each subplot

@@
   geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
+  xlim(c(-0.1, 0.1)) +
   theme_single_y()

### 2025-04-11 analysis_scripts/06_sim_effects.R
cluter script for simulating data in parallel

@@ -1,0 +1,87 @@
+#' April 8, 2025
+#' 
+#' Script for simulating data with varying experiential effect sizes and number of games per player
+#' to test how well our model can actually capture this effects if they exist
+#' to be run on cluster in parallel
+
+library(tidyverse)
+library(here)
+library(rstanarm)
+
+source(here("utils/helper.R"))
+
+options(mc.cores = parallel::detectCores())
+
+save_path = "/home/adamgee/scratch/results/sim_data/"
+
+#' for running jobs in parallel 
+#' each player should have their own job id
+jobid = Sys.getenv("SLURM_ARRAY_TASK_ID") %>% as.numeric()
+
+
+num_players = 10 #number of players in each dataset
+player_id = jobid %% num_players #id of player in this job
+num_games = 20000 #this is roughly the same as the number of games per player in 1700-1900 cohort
+
+#' assume that the probability player j wins game i is p_ij = inv_logit(alpha_j + beta_j*x_ij + gamma_1*colour + gamma_2*rating_diff) 
+#' ie our model is correct
+#' also assume that we can't get any draws under our model
+#' also technically assuming all of these games will be played in one session...
+
+#' we vary the experiential effect size and number of games per player to see how well our model can detect true experiential effects
+#' experiential effects size of logit(0.51) represents 1% increase win percentage when coming from a win... logit(0.53) represents 3% increase...
+mu_beta_sizes = c(logit(0.51), logit(0.53), logit(0.6)) #these are the mu_betas
+
+#if we want to fit our model using smaller number of games, just take a subset of these 20,000 games
+
+
+#' we can parallelize the mu_betas and the players
+#' so each player will loop over 1 -> num_games for 3 different values of mu_beta
+#' 
+#' jobs (1 -> num_players) will be smallest mu_beta
+#' jobs num_players + (1 -> num_players) will be middle mu_beta
+#' jobs 2*num_players + (1 -> num_players) will be largest mu_beta
+#' ie for 10 players.. 1-10 have small mu_beta, 11-20 have middle mu_beta, 21-30 have largest mu_beta
+
+
+### values for other parameters
+
+#population effects values - found from posterior means of original fit
+gamma1 = 0.091 #colour effect
+gamma2 = 0.0038 #rating effect
+
[…]

### 2025-04-17 utils/plot_templates.R
adding option for legend title in theme_single_y_legend()

@@ -61,8 +61,7 @@ theme_single_y_legend <- function(){
       axis.title = element_text(size = axis_title),
       legend.text = element_text(size = 12),
       strip.text = element_text(size = axis_title - 5, 
-                                margin = margin(1.5, 1.5, 2, 1.5)),
-      legend.title = element_blank(),
+                                margin = margin(1.5, 1.5, 2, 1.5))
     )
 }
 

### 2025-04-17 Paper_WriteUp/Paper_Plots.qmd
adding plotting of simulation results

@@
 library(ggpubr)
+library(rstanarm)
 options(mc.cores = parallel::detectCores())
@@
   geom_line() +
-  labs(x = "Games Played", y = "Rating") + 
+  labs(x = "Games Played", y = "Rating", colour = "") + 
   theme_single_y_legend()
@@
 
+## Posterior Predictive Plots
+
+```{r figure 7 - PPDs for Magnus & Nihal}
+
+```
+
 ### Permuting for 1700-1900
@@
 
-<!-- TO DO - Tidy up below here -->
-
-## Posterior Predictive Plots
+## Simulation Results
 
 ```{r}
-
+#the fits and sim datasets are stored here
+sim_path = here("results_revision", "sim_data_fits")
+sim_dataset_names = list.files(sim_path, pattern = "*.csv")
+sim_fit_names = list.files(sim_path, pattern = "*.RDS")
+
+#list of all the sim datasets and corresponding fits 
+sim_datasets_list = paste0(sim_path, "/", sim_dataset_names) %>% lapply(read.csv)
+names(sim_datasets_list) = tools::file_path_sans_ext(basename(sim_dataset_names))
+
+sim_fit_list = paste0(sim_path, "/", sim_fit_names) %>% lapply(readRDS)
+names(sim_fit_list) = tools::file_path_sans_ext(basename(sim_fit_names))
+
+# Winner effects on single dataset
+
+# #the simulated dataset and corresponding fit to plot
+# sim_dataset_num = 3
+# 
+# #global effects
+# facet_labels = as_labeller(c(
+#   mu_beta = "mu[beta]",
+#   gamma1 = "gamma[1]",
+#   gamma2 = "gamma[2]",
+#   sigma_1 = "sigma[1]",
+#   `tau[1]` = "tau[1]",
+#   `tau[2]` = "tau[2]",
+#   sigma_g1 = "sigma[g[1]]",
+#   sigma_g2 = "sigma[g[2]]"
+# ), label_parsed)
+# 
+# global_plot = sim_fit_list[[sim_dataset_num]]$draws(c("mu_beta", "gamma1", "gamma2", "sigma_1", "tau[1]", "tau[2]", "sigma_g1", "sigma_g2")) |> 
+#   as_draws_df() |> 
+#   pivot_longer(cols = "mu_beta":"sigma_g2") |> 
+#   ggplot(aes(value)) +
+#   geom_histogram() +
+#    facet_wrap(~name, ncol = 4, scales = "free",
+#              labeller = labeller(name = facet_labels)) +
+#   labs(x = element_blank(), y = element_blank()) +
+#   theme_4col() +
+#   scale_x_continuous(breaks = breaks_pretty(n = 2)) 
+# global_plot
+# 
+# 
+# #experiential effects
+# mod_pars = sim_fit_list[[sim_dataset_num]]$draws() |>
+#     as_draws_df() |>
+#     select(-any_of(c("lp__", ".chain", ".iteration", ".draw"))) |>
+#     pivot_longer(cols = everything()) |>
+#     filter(startsWith(name, "beta[2")) |>
+#     mutate(id = stringr::str_extract(name, pattern = "\\d+]"),
+#            id = stringr::str_replace(id, "\\]", ""), 
+#            id = as.numeric(id))
+# 
+# #order of min to max estiamted betas - makes plot look nicer
+# beta_order = mod_pars %>% group_by(id) %>% summarise(mean = mean(value)) %>% arrange(mean) %>% pull(id)
+# 
+# #true experiential effects
+# true_betas = sim_datasets_list[[sim_dataset_num]] %>% 
+#   group_by(player_id) %>% 
+#   summarise(true_beta = unique(beta)) %>%
+#   rename(id = player_id)
+# 
+# #reordering
+# true_betas = true_betas[match(beta_order, true_betas$id),] %>% 
+#   mutate(ymin = as.numeric(factor(row_number())) - 0.2,
+#          ymax = as.numeric(factor(row_number())) + 0.2)
+# 
+# #plot
+# winner_plot = mod_pars |> 
+#   ggplot(aes(y = reorder(id, value, FUN = median), x = value)) +
+#   stat_histinterval() +
+#   geom_segment(data = true_betas,
+#                aes(x = true_beta, xend = true_beta, y = ymin, yend = ymax),
+#                color = "orange", linetype = "dashed", linewidth = 2) +
+#   labs(x = "Estimated Experiential Effect", y = "Simulated Player ID") +
+#   geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
+#   theme_single_y()
+# winner_plot
+
+#### winner plots with same mu_beta size and varying num games per player
+
+draws_list = lapply(sim_fit_list, function(x) {
+  x$draws() |>
+    as_draws_df() |>
+    select(-any_of(c("lp__", ".chain", ".iteration", ".draw"))) |>
+    pivot_longer(cols = everything()) |>
+    filter(startsWith(name, "beta[2")) |>
+    mutate(id = stringr::str_extract(name, pattern = "\\d+]"),
+           id = stringr::str_replace(id, "\\]", ""), 
+           id = as.numeric(id))
+})
+
+#combing fit results by mu_beta size
+small_effect_draws = draws_list[1:3] %>% bind_rows(.id = "column_label") %>% 
+  mutate(num_games = case_when(endsWith(column_label, "1") ~ 1000,
+                               endsWith(column_label, "2") ~ 5000,
+                               endsWith(column_label, "3") ~ 20000),
+         num_games = factor(num_games, levels = c("1000", "5000", "20000"))) #make sure the levels are messing up here
+
+med_effect_draws = draws_list[4:6] %>% bind_rows(.id = "column_label") %>% 
+  mutate(num_games = case_when(endsWith(column_label, "4") ~ 1000,
+                               endsWith(column_label, "5") ~ 5000,
+                               endsWith(column_label, "6") ~ 20000),
+         num_games = factor(num_games, levels = c("1000", "5000", "20000"))) #make sure the levels are messing up here
+
+big_effect_draws = draws_list[7:9] %>% bind_rows(.id = "column_label") %>% 
+  mutate(num_games = case_when(endsWith(column_label, "7") ~ 1000,
+                               endsWith(column_label, "8") ~ 5000,
+                               endsWith(column_label, "9") ~ 20000),
+         num_games = factor(num_games, levels = c("1000", "5000", "20000"))) #make sure the levels are messing up here
+
+## small mu_beta plot ##
+
+#true experiential effects
+true_betas = sim_datasets_list[[1]] %>% 
+  group_by(player_id) %>% 
+  summarise(true_beta = unique(beta)) %>%
+  rename(id = player_id)
+
+#reordering
+true_betas = true_betas[match(beta_order, true_betas$id),] %>% 
+  mutate(ymin = as.numeric(factor(row_number())) - 0.3,
+         ymax = as.numeric(factor(row_number())) + 0.3)
+
+#plot
+small_effect_plot = small_effect_draws  |>
+  ggplot(aes(y = reorder(id, value, FUN = median), x = value, colour = num_games)) +
+  #stat_histinterval(aes(fill = num_games), position = "dodgejust") +
+  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
+  #stat_dots(aes(colour = num_games), position = "dodgejust") +
+  geom_segment(data = true_betas,
+               aes(x = true_beta, xend = true_beta, y = ymin, yend = ymax),
+               color = "red", linewidth = 1.5) +
+  labs(x = "Estimated Experiential Effect", y = "Simulated Player ID", colour = "Number of Games\nPlayed per Player") +
+  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
+  theme_single_y_legend() +
+  scale_color_viridis_d(option = "D")
+
+#save
+ggsave(plot = small_effect_plot, 
+       filename = here("Paper_WriteUp", "paper_figures", "fig_sim_1.png"), dpi = 1000,
+       height = 6, width = 7)
+
+## med mu_beta plot ##
+
+#true experiential effects
+true_betas = sim_datasets_list[[4]] %>% 
+  group_by(player_id) %>% 
+  summarise(true_beta = unique(beta)) %>%
+  rename(id = player_id)
+
+#reordering
+true_betas = true_betas[match(beta_order, true_betas$id),] %>% 
+  mutate(ymin = as.numeric(factor(row_number())) - 0.3,
+         ymax = as.numeric(factor(row_number())) + 0.3)
+
+#plot
+med_effect_plot = med_effect_draws  |>
+  ggplot(aes(y = reorder(id, value, FUN = median), x = value, colour = num_games)) +
+  #stat_histinterval(aes(fill = num_games), position = "dodgejust") +
+  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
+  #stat_dots(aes(colour = num_games), position = "dodgejust") +
+  geom_segment(data = true_betas,
+               aes(x = true_beta, xend = true_beta, y = ymin, yend = ymax),
+               color = "red", linewidth = 1.5) +
+  labs(x = "Estimated Experiential Effect", y = "Simulated Player ID", colour = "Number of Games\nPlayed per Player") +
+  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
+  theme_single_y_legend() +
+  scale_color_viridis_d(option = "D")
+
+#save
+ggsave(plot = med_effect_plot, 
+       filename = here("Paper_WriteUp", "paper_figures", "fig_sim_2.png"), dpi = 1000,
+       height = 6, width = 7)
+
+## big mu_beta plot ##
+
+#true experiential effects
+true_betas = sim_datasets_list[[7]] %>% 
+  group_by(player_id) %>% 
+  summarise(true_beta = unique(beta)) %>%
+  rename(id = player_id)
+
+#reordering
+true_betas = true_betas[match(beta_order, true_betas$id),] %>% 
+  mutate(ymin = as.numeric(factor(row_number())) - 0.3,
+         ymax = as.numeric(factor(row_number())) + 0.3)
+
+#plot
+big_effect_plot = big_effect_draws  |>
+  ggplot(aes(y = reorder(id, value, FUN = median), x = value, colour = num_games)) +
+  #stat_histinterval(aes(fill = num_games), position = "dodgejust") +
+  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
+  #stat_dots(aes(colour = num_games), position = "dodgejust") +
+  geom_segment(data = true_betas,
+               aes(x = true_beta, xend = true_beta, y = ymin, yend = ymax),
+               color = "red", linewidth = 1.5) +
+  labs(x = "Estimated Experiential Effect", y = "Simulated Player ID", colour = "Number of Games\nPlayed per Player") +
+  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
+  theme_single_y_legend() +
+  scale_color_viridis_d(option = "D")
+
+#save
+ggsave(plot = big_effect_plot, 
+       filename = here("Paper_WriteUp", "paper_figures", "fig_sim_3.png"), dpi = 1000,
+       height = 6, width = 7)
 ```

### 2025-04-17 Paper_WriteUp/Paper_Plots.qmd
put PPD code in for Magnus & Nihal, I think it's currently bugged though

@@
 ```{r figure 7 - PPDs for Magnus & Nihal}
-
+#where models are stored
+model_path = here("results_revision", "lichessGrandmasters")
+
+#read in the model and users
+model = readRDS(file = here(model_path, "all_rated_bullet_model_prev_n1.RDS"))
+users = readRDS(file = here(model_path, "users_bullet.RDS"))
+
+#grandmaster data
+data_path = paste0(all_data_path[4], "/")
+files = list.files(data_path)
+lichess_data = files %>%
+  map_dfr(~read_player(data_path, .x))
+
+#Tidy data
+
+#selecting bullet 60+0 only
+bullet_60 = lichess_data  %>%
+  mutate(Event = tolower(Event)) %>%
+  filter(Event == "rated bullet game", Variant == "Standard", TimeControl == "60+0") %>%
+  dplyr::select(c("White", "Black", "Result", "UTCDate", "UTCTime", "Username", "WhiteElo", "BlackElo"))
+
+#remove the duplicate rows
+bullet_60_no_dup = bullet_60[!duplicated(bullet_60),]
+
+#final 1000 games each user plays
+last_games = bullet_60_no_dup %>%
+  filter(Username %in% users) %>%  
+  group_by(Username) %>%
+  arrange(UTCDate, UTCTime, .by_group = TRUE) %>%
+  slice_tail(n = 1000) %>%
+  ungroup()
+
+#only consider previous game for history
+tidy_games = map_dfr(users, get_hist, last_games, prev_n = 1) |> 
+  as_tibble()
+
+#adding performance history
+init_data = tidy_games |>
+  mutate(WhiteElo = as.numeric(WhiteElo),
+         BlackElo = as.numeric(BlackElo),
+         focal_user = ifelse(focal_white == 1, White, Black),
+         elo_diff = ifelse(focal_white == 1,
+                           WhiteElo - BlackElo, BlackElo - WhiteElo),
+         focal_id = match(focal_user, users),
+         focal_rating = ifelse(focal_white == 1, WhiteElo, BlackElo),
+         opp = ifelse(focal_white == 1, Black, White),
+         opp_rating = ifelse(focal_white == 1, BlackElo, WhiteElo),
+         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
+  dplyr::select(focal_user, focal_id, focal_white, WhiteElo, BlackElo,
+                focal_win_prop, elo_diff, focal_result, opp, opp_rating,
+                focal_rating, UTCDateTime) |>
+  group_by(focal_id) |>
+  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), 
+         ave_prop = ifelse(time_diff <= 300,  #if games played in diff sessions, make history same as draw (ie NULL)
+                           lag(focal_win_prop, default = 0.5), #removing standardization
+                           0.5)) |>
+  ungroup() |>
+  filter(focal_result != 0.5)
+
+#ids of players to plot
+magnus_id = which(users == "DrNykterstein") #magnus
+nihal_id = which(users == "nihalsarin2004") #nihal
+
+#choose who to plot
+focal_player_used = magnus_id
+
+#slightly less than 1000 games since filtered out draws
+focal_init_data = init_data |>
+  filter(focal_user == users[focal_player_used])
+
+focal_init = data.frame("Player" = users[focal_player_used],
+                        "Rating" = ifelse(focal_init_data$focal_white == 1, 
+                                          focal_init_data$WhiteElo, focal_init_data$BlackElo))
+
+num_sims = 100 #number of rating curve simulations over final games
+num_games = nrow(focal_init_data) #number of games to predict over
+
+#parameter draws for focal player
+focal_draws = model$draws() |> as_draws_df() |>
+  dplyr::select(starts_with(c("beta[", "gamma"))) |>
+  mutate(draw = row_number()) |> 
+  pivot_longer(cols = "beta[1,1]":paste0("beta[2,", as.character(length(users)), "]")) |> 
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]")) |> 
+  dplyr::select(-name) |> 
+  pivot_wider(names_from = param, values_from = value) |> 
+  rename(beta1 = `1`, beta2 = `2`) %>%
+  filter(id == focal_player_used) %>%
+  filter(draw %in% 1:num_sims)
+
+## the covariates for the future games
+future_games_cov = tibble(focal = as.character(focal_init_data$focal_id),  
+                          color = focal_init_data$focal_white,
+                          elo_diff = focal_init_data$elo_diff,
+                          hist = focal_init_data$ave_prop,
+                          focal_rating = focal_init_data$focal_rating,
+                          opponent = focal_init_data$opp,
+                          opponent_rating = focal_init_data$opp_rating)
+
+#storage
+glicko_sim = matrix(NA, nrow = num_games, ncol = num_sims)
+game_sim = matrix(NA, nrow = num_games, ncol = num_sims)
+#probs = matrix(NA, nrow = num_games, ncol = num_sims) #for debugging
+
+#sim
+set.seed(2025)
+for (i in 1:num_games) {
+  #current game info
+  curr_game_info = future_games_cov[i,]
+
+  #indicator for colour effect, 1 is white, 0 is black
+  colour_ind = curr_game_info$color
+
+  #elo diff for rating effect
+  rating_diff = curr_game_info$elo_diff
+
+  #opponent name and rating
+  opp = curr_game_info$opponent
+  opp_rating = curr_game_info$opponent_rating
+
+  #simulate wins/loss on current game
+  for (j in 1:num_sims) {
+    curr_draw = focal_draws %>% filter(draw == j)
+
+  ## if first game use default history (0), otherwise use the previous
+  if (i == 1) {
+    curr_rating = curr_game_info$focal_rating
+    curr_hist = 0.5 #assume they're current past performance is at their usual level
+  } else {
+    curr_hist = curr_game_info$hist
+    curr_rating = glicko_sim[i - 1, j]
+  }
+
+  #find probability
+  lin_comb = curr_draw$beta1 + (curr_draw$beta2 * curr_hist) +
+             (curr_draw$gamma1 * colour_ind) +
+             (curr_draw$gamma2 * rating_diff)
+  prob = invlogit(lin_comb)
+
+  #simulate result
+  sim_result = sample(c(1, 0), size = 1, prob = c(prob, 1 - prob))
+
+  #find rating
+  games = data.frame(Week = i,
+                     Player1 = users[focal_player_used],
+                     Player2 = opp,
+                     Score = sim_result)
+  initstate = tibble(Player = c(users[focal_player_used], opp),
+                     Rating = c(curr_rating, opp_elo),
+                     Deviation = 50,
+                     Volatility = 0.04)
+
+  sim_glick = glicko2(games, status = initstate, history = TRUE, gamma = colour_ind)
+  ## extract the rating of focal player and store it
+  focal_glick = sim_glick$history[users[focal_player_used], ,]
+
+  #store current result and rating
+  glicko_sim[i, j] = as.numeric(focal_glick)[1]
+  game_sim[i, j] = sim_result
+  }
+  if(i %% 10 == 0) {print(i)} #progress of number of games simulated
+}
+
+#true glicko rating of all games before last 1000
+true_glicko_first = bullet_60_no_dup |> 
+  filter(Username == users[focal_player_used]) |> 
+  arrange(UTCDate, UTCTime) |> 
+  slice(1:(n() - 1000)) |> 
+  mutate(game = row_number(), glicko = ifelse(Username == White, WhiteElo, BlackElo)) %>%
+  dplyr::select(game, glicko)
+num_prev = nrow(true_glicko_first) #number of games before final simulated games
+
+#true glicko rating of final 1000 games
+true_glicko_final = focal_init_data %>%
+  mutate(game = row_number() + num_prev, glicko = ifelse(focal_white == 1, WhiteElo, BlackElo)) %>%
+  dplyr::select(game, glicko)
+
+#glicko rating across all sims on final 100 games
+final_sims = data.frame(glicko_sim, check.names = FALSE) %>%
+  mutate(game = row_number() + num_prev) %>%
+  pivot_longer(cols = -game, names_to = "Sim", values_to = "glicko")
+
+#mean glicko rating for each game across all sims
+sim_mean_glicko = final_sims %>% group_by(game) %>% summarise(mean_glicko = mean(glicko))
+
+#plot
+ppd_rating_plot = ggplot(data = true_glicko_first, aes(x = game, y = glicko)) +
+  geom_line(linewidth = 0.5, col = "darkred") + 
+  geom_line(data = final_sims, mapping = aes(group = Sim), alpha = 0.05) +
+  geom_line(data = sim_mean_glicko, mapping = aes(x = game, y = mean_glicko), 
+            alpha = 0.7) +
+  geom_line(data = true_glicko_final, col = "red", linewidth = 0.5) +
+  ylim(c(min(glicko_sim) - 50, max(glicko_sim) + 50)) +
+  xlim(c(num_prev - 1500, num_prev + num_games)) +
+  xlab("Games Played") + ylab("Glicko2 Rating") +
+  theme_single_y()
+ppd_rating_plot
+
+# #save magnus
+# ggsave(plot = ppd_rating_plot, 
+#        filename = here("Paper_WriteUp", "paper_figures", "fig7a.png"), dpi = 1000,
+#        height = 4.5, width = 10)
+# #save nihal
+# ggsave(plot = ppd_rating_plot, 
+#        filename = here("Paper_WriteUp", "paper_figures", "fig7b.png"), dpi = 1000,
+#        height = 4.5, width = 10)
 ```

### 2025-04-17 Paper_WriteUp/Paper_Plots.qmd
putting in updated code for fig3

@@
 ### Global Parameters
+
+```{r figure 3 - Updated}
+### Adam
+#' April 17, 2025
+#' adding code to updated fig3
+#' only comparing gamma1, gamma2, & mu_beta across all 4 cohorts
+
+#read in models
+model_17_19 = readRDS(here("results_revision", "lichess1700-1900", "all_rated_bullet_model_prev_n1.RDS"))
+model_20_22 = readRDS(here("results_revision", "lichess2000-2200", "all_rated_bullet_model_prev_n1.RDS"))
+model_23_25 = readRDS(here("results_revision", "lichess2300-2500", "all_rated_bullet_model_prev_n1.RDS"))
+model_gm = readRDS(here("results_revision", "lichessGrandmasters", "all_rated_bullet_model_prev_n1.RDS"))
+
+#get global effects of each
+pars_17_19 <- model_17_19$draws(c("mu_beta", "gamma1", "gamma2")) |> 
+  as_draws_df() |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |> 
+  mutate(model = "1700-1900")
+
+pars_20_22 = model_20_22$draws(c("mu_beta", "gamma1", "gamma2")) |> 
+  as_draws_df() |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |> 
+  mutate(model = "2000-2200")
+
+pars_23_25 = model_23_25$draws(c("mu_beta", "gamma1", "gamma2")) |> 
+  as_draws_df() |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |> 
+  mutate(model = "2300-2500")
+
+pars_gm <- model_gm$draws(c("mu_beta", "gamma1", "gamma2")) |> 
+  as_draws_df() |> 
+  pivot_longer(cols = "mu_beta":"gamma2") |> 
+  mutate(model = "GM")
+
+#for plotting
+facet_labels <- as_labeller(c(
+  mu_beta = "mu[beta]",
+  gamma1 = "gamma[1]",
+  gamma2 = "gamma[2]",
+  sigma_1 = "sigma[1]",
+  `tau[1]` = "tau[1]",
+  `tau[2]` = "tau[2]",
+  sigma_g1 = "sigma[g[1]]",
+  sigma_g2 = "sigma[g[2]]"
+), label_parsed)
+
+gm_amateur_comp_global <- bind_rows(pars_17_19, pars_20_22, pars_23_25, pars_gm) |> 
+  ggplot(aes(x = value, y = name, color = factor(model, levels = c("1700-1900", "2000-2200", "2300-2500", "GM")))) +
+  stat_pointinterval(position = position_dodge(width = 0.2)) +
+  facet_wrap(~name,
+             scales = "free",
+             labeller = labeller(name = facet_labels)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_3col() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_color_viridis_d(option = "H")
+
+gm_amateur_comp_global
+
+```
 
@@
   initstate = tibble(Player = c(users[focal_player_used], opp),
-                     Rating = c(curr_rating, opp_elo),
+                     Rating = c(curr_rating, opp_rating),
                      Deviation = 50,

### 2025-04-17 analysis_scripts/06_sim_data_gen.R
changing true values for mu_beta

@@ -30,7 +30,8 @@ num_games = 20000 #this is roughly the same as the number of games per player in
 
 #' we vary the experiential effect size and number of games per player to see how well our model can detect true experiential effects
 #' experiential effects size of logit(0.51) represents 1% increase win percentage when coming from a win... logit(0.53) represents 3% increase...
-mu_beta_sizes = c(logit(0.51), logit(0.53), logit(0.6)) #these are the mu_betas
+#mu_beta_sizes = c(logit(0.51), logit(0.53), logit(0.6)) #these are the mu_betas
+mu_beta_sizes = c(-0.5, 0, 0.5)
 
 #' if we want to fit our model using smaller number of games, just take a subset of these 20,000 games
 #' 

### 2025-04-17 analysis_scripts/01_Large_Fit.R
modify the 2 main fitting scripts to only look at the win ratio, not normalized by overall mean

@@ -99,7 +99,7 @@ init_data <- tidy_games |>
   select(focal_user, focal_id, focal_white,
          focal_win_prop, elo_diff, focal_result) |>
   group_by(focal_id) |>
-  mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_result)) |>
+  mutate(ave_prop = lag(focal_win_prop, default = 0)) |> # - mean(focal_result)) |>
   filter(focal_result != 0.5)
 
 cat("----------\n")
@@ -135,63 +135,3 @@ fit3_ave$save_object(file = here(save_path,
                                  paste0("all_rated_bullet_model_n", n, ".RDS")))
 # fit3_ave$save_object(file = here(save_path,
 #                                  paste0("all_rated_blitz_model_n", n, ".RDS")))
-
-## create some summary plots of these results
-
-random_effect_post <- fit3_ave$draws() |> 
-  as_draws_df() |>
-  select(starts_with("beta[")) |>
-  pivot_longer(cols = everything()) |>
-  mutate(param = stringr::str_extract(name, pattern = "\\d"),
-         id = stringr::str_extract(name, pattern = "\\d+]"),
-         id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]"))
-
-players <- users
-names(players) <- paste0("beta[", 1:length(users), "]")
-
-player_labels <- as_labeller(players)
-
-mcmc_hist(fit3_ave$draws(c("mu_beta",  "gamma1", "gamma2",
-                           "sigma_1", "tau[1]", "tau[2]",
-                           "sigma_g1", "sigma_g2")),
-          facet_args = list(scales = "free"))
-
-ggsave(filename = paste0(save_path, "/global_pars_all_rated_bullet_model.png"),
-                         width = 8, height = 8, units = "in")
-# ggsave(filename = paste0(save_path, "/global_pars_all_rated_blitz_model.png"),
-#        width = 8, height = 8, units = "in")
-
-theme_set(bayesplot_theme_get())
-
-random_effect_post |>
-  filter(param == 2) |>
-  ggplot(aes(value)) +
-  geom_histogram(fill = "#6497b1", colour = "black", size = 0.2) +
-  facet_wrap(~player_id, scales = "free",
-             labeller = player_labels, ncol = 5) +
-  labs(title = "Individual Winner Effects", y = "") +
-  theme(axis.text.y = element_blank(),
[…]

### 2025-04-17 analysis_scripts/02_Large_Single_Prev.R

@@ -99,7 +99,7 @@ hist_data_init <- tidy_games |>
   mutate(ave_prop = lag(focal_win_prop, n = 1, default = 0) - 
            mean(focal_result), 
          ## focal_win_prop is the average of the running win proportion
-         prev_game = lag(focal_result, default = 0) - mean(focal_result)) |> 
+         prev_game = lag(focal_result, default = 0)) |> # - mean(focal_result)) |> 
   filter(focal_result != 0.5)
 
 
@@ -139,64 +139,3 @@ fit3_ave$save_object(file = here(save_path,
 #                                  paste0("all_rated_blitz_model_prev_n",
 #                                         n, ".RDS")))
 
-## create some summary plots of these results
-random_effect_post <- fit3_ave$draws() |>
-  as_draws_df() |> 
-  select(starts_with("beta[")) |> 
-  pivot_longer(cols = everything()) |> 
-  mutate(param = stringr::str_extract(name, pattern = "\\d"),
-         id = stringr::str_extract(name, pattern = "\\d+]"),
-         id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]"))
-
-players <- users
-names(players) <- paste0("beta[", 1:length(users), "]")
-
-player_labels <- as_labeller(players)
-
-mcmc_hist(fit3_ave$draws(c("mu_beta",  "gamma1", "gamma2", 
-                           "sigma_1", "tau[1]", "tau[2]",
-                           "sigma_g1", "sigma_g2")),
-          facet_args = list(scales = "free"))
-
-ggsave(filename = paste0(save_path,
-                         "/global_pars_all_rated_bullet_model_prev.png"),
-                         width = 8, height = 8, units = "in")
-# ggsave(filename = paste0(save_path,
-#                          "/global_pars_all_rated_blitz_model_prev.png"),
-#        width = 8, height = 8, units = "in")
-
-theme_set(bayesplot_theme_get())
-
-random_effect_post |> 
-  filter(param == 2) |> 
-  ggplot(aes(value)) +
-  geom_histogram(fill = "#6497b1", colour = "black", size = 0.2) +
-  facet_wrap(~player_id, scales = "free",
-             labeller = player_labels, ncol = 5) +
-  labs(title = "Individual Winner Effects", y = "") +
-  theme(axis.text.y = element_blank(),
-        axis.ticks.y = element_blank())
-
-ggsave(filename = paste0(save_path,
[…]

### 2025-04-17 Paper_WriteUp/Paper_Plots.qmd
in process of doing simulated mu_beta plots. Not finished, should probably split it all into different chunks

@@
 
-# #the simulated dataset and corresponding fit to plot
-# sim_dataset_num = 3
-# 
-# #global effects
-# facet_labels = as_labeller(c(
-#   mu_beta = "mu[beta]",
-#   gamma1 = "gamma[1]",
-#   gamma2 = "gamma[2]",
-#   sigma_1 = "sigma[1]",
-#   `tau[1]` = "tau[1]",
-#   `tau[2]` = "tau[2]",
-#   sigma_g1 = "sigma[g[1]]",
-#   sigma_g2 = "sigma[g[2]]"
-# ), label_parsed)
-# 
-# global_plot = sim_fit_list[[sim_dataset_num]]$draws(c("mu_beta", "gamma1", "gamma2", "sigma_1", "tau[1]", "tau[2]", "sigma_g1", "sigma_g2")) |> 
-#   as_draws_df() |> 
-#   pivot_longer(cols = "mu_beta":"sigma_g2") |> 
-#   ggplot(aes(value)) +
-#   geom_histogram() +
-#    facet_wrap(~name, ncol = 4, scales = "free",
-#              labeller = labeller(name = facet_labels)) +
-#   labs(x = element_blank(), y = element_blank()) +
-#   theme_4col() +
-#   scale_x_continuous(breaks = breaks_pretty(n = 2)) 
-# global_plot
-# 
+#the simulated dataset and corresponding fit to plot
+sim_dataset_num = 3
+
+#global effects
+facet_labels = as_labeller(c(
+  mu_beta = "mu[beta]",
+  gamma1 = "gamma[1]",
+  gamma2 = "gamma[2]",
+  sigma_1 = "sigma[1]",
+  `tau[1]` = "tau[1]",
+  `tau[2]` = "tau[2]",
+  sigma_g1 = "sigma[g[1]]",
+  sigma_g2 = "sigma[g[2]]"
+), label_parsed)
+
+global_plot = sim_fit_list[[sim_dataset_num]]$draws(c("mu_beta", "gamma1", "gamma2", 
+                                                      "sigma_1", "tau[1]", "tau[2]", 
+                                                      "sigma_g1", "sigma_g2")) |>
+  as_draws_df() |>
+  pivot_longer(cols = "mu_beta":"sigma_g2") |>
+  ggplot(aes(value)) +
+  geom_histogram() +
+   facet_wrap(~name, ncol = 4, scales = "free",
+             labeller = labeller(name = facet_labels)) +
+  labs(x = element_blank(), y = element_blank()) +
+  theme_4col() +
+  scale_x_continuous(breaks = breaks_pretty(n = 2))
+global_plot
+
 # 
@@
 
+#order of min to max estiamted betas - makes plot look nicer
+beta_order = sim_datasets_list[[1]] %>% group_by(player_id) %>% summarise(beta = unique(beta)) %>% arrange(beta) %>% pull(player_id)
+
 ## small mu_beta plot ##
@@
        height = 6, width = 7)
+
+## same above plot but now mu_beta
+mu_beta_draws_list = lapply(sim_fit_list, function(x) {
+  x$draws("mu_beta") |>
+  as_draws_df()
+})
+
+#combing fit results by mu_beta size
+small_effect_draws = mu_beta_draws_list[1:3] %>% bind_rows(.id = "column_label") %>% 
+  mutate(num_games = case_when(endsWith(column_label, "1") ~ 1000,
+                               endsWith(column_label, "2") ~ 5000,
+                               endsWith(column_label, "3") ~ 20000),
+         num_games = factor(num_games, levels = c("1000", "5000", "20000"))) #make sure the levels are messing up here
+
+med_effect_draws = mu_beta_draws_list[4:6] %>% bind_rows(.id = "column_label") %>% 
+  mutate(num_games = case_when(endsWith(column_label, "4") ~ 1000,
+                               endsWith(column_label, "5") ~ 5000,
+                               endsWith(column_label, "6") ~ 20000),
+         num_games = factor(num_games, levels = c("1000", "5000", "20000"))) #make sure the levels are messing up here
+
+big_effect_draws = mu_beta_draws_list[7:9] %>% bind_rows(.id = "column_label") %>% 
+  mutate(num_games = case_when(endsWith(column_label, "7") ~ 1000,
+                               endsWith(column_label, "8") ~ 5000,
+                               endsWith(column_label, "9") ~ 20000),
+         num_games = factor(num_games, levels = c("1000", "5000", "20000"))) #make sure the levels are messing up here
+
 ```

### 2025-04-18 Paper_WriteUp/Paper_Plots.qmd
added mu_beta sim plot. Also cleaned beta_sim plots chunk

--- 2025-04-17 Paper_WriteUp/Paper_Plots.qmd
@@
 
-# Winner effects on single dataset
-
-# #the simulated dataset and corresponding fit to plot
-# sim_dataset_num = 3
-# 
-# #global effects
-# facet_labels = as_labeller(c(
-#   mu_beta = "mu[beta]",
-#   gamma1 = "gamma[1]",
-#   gamma2 = "gamma[2]",
-#   sigma_1 = "sigma[1]",
-#   `tau[1]` = "tau[1]",
-#   `tau[2]` = "tau[2]",
-#   sigma_g1 = "sigma[g[1]]",
-#   sigma_g2 = "sigma[g[2]]"
-# ), label_parsed)
-# 
-# global_plot = sim_fit_list[[sim_dataset_num]]$draws(c("mu_beta", "gamma1", "gamma2", "sigma_1", "tau[1]", "tau[2]", "sigma_g1", "sigma_g2")) |> 
-#   as_draws_df() |> 
-#   pivot_longer(cols = "mu_beta":"sigma_g2") |> 
-#   ggplot(aes(value)) +
-#   geom_histogram() +
-#    facet_wrap(~name, ncol = 4, scales = "free",
-#              labeller = labeller(name = facet_labels)) +
-#   labs(x = element_blank(), y = element_blank()) +
-#   theme_4col() +
-#   scale_x_continuous(breaks = breaks_pretty(n = 2)) 
-# global_plot
-# 
-# 
-# #experiential effects
-# mod_pars = sim_fit_list[[sim_dataset_num]]$draws() |>
-#     as_draws_df() |>
-#     select(-any_of(c("lp__", ".chain", ".iteration", ".draw"))) |>
-#     pivot_longer(cols = everything()) |>
-#     filter(startsWith(name, "beta[2")) |>
-#     mutate(id = stringr::str_extract(name, pattern = "\\d+]"),
-#            id = stringr::str_replace(id, "\\]", ""), 
-#            id = as.numeric(id))
-# 
-# #order of min to max estiamted betas - makes plot look nicer
-# beta_order = mod_pars %>% group_by(id) %>% summarise(mean = mean(value)) %>% arrange(mean) %>% pull(id)
-# 
-# #true experiential effects
-# true_betas = sim_datasets_list[[sim_dataset_num]] %>% 
-#   group_by(player_id) %>% 
-#   summarise(true_beta = unique(beta)) %>%
-#   rename(id = player_id)
-# 
-# #reordering
-# true_betas = true_betas[match(beta_order, true_betas$id),] %>% 
-#   mutate(ymin = as.numeric(factor(row_number())) - 0.2,
-#          ymax = as.numeric(factor(row_number())) + 0.2)
-# 
-# #plot
-# winner_plot = mod_pars |> 
-#   ggplot(aes(y = reorder(id, value, FUN = median), x = value)) +
-#   stat_histinterval() +
-#   geom_segment(data = true_betas,
-#                aes(x = true_beta, xend = true_beta, y = ymin, yend = ymax),
-#                color = "orange", linetype = "dashed", linewidth = 2) +
-#   labs(x = "Estimated Experiential Effect", y = "Simulated Player ID") +
-#   geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
-#   theme_single_y()
-# winner_plot
-
 #### winner plots with same mu_beta size and varying num games per player
 
-draws_list = lapply(sim_fit_list, function(x) {
+beta_draws_list = lapply(sim_fit_list, function(x) {
   x$draws() |>
@@
     filter(startsWith(name, "beta[2")) |>
-    mutate(id = stringr::str_extract(name, pattern = "\\d+]"),
-           id = stringr::str_replace(id, "\\]", ""), 
-           id = as.numeric(id))
+    mutate(player_id = stringr::str_extract(name, pattern = "\\d+]"),
+           player_id = stringr::str_replace(player_id, "\\]", ""), 
+           player_id = as.numeric(player_id))
 })
 
-#combing fit results by mu_beta size
-small_effect_draws = draws_list[1:3] %>% bind_rows(.id = "column_label") %>% 
-  mutate(num_games = case_when(endsWith(column_label, "1") ~ 1000,
-                               endsWith(column_label, "2") ~ 5000,
-                               endsWith(column_label, "3") ~ 20000),
-         num_games = factor(num_games, levels = c("1000", "5000", "20000"))) #make sure the levels are messing up here
-
-med_effect_draws = draws_list[4:6] %>% bind_rows(.id = "column_label") %>% 
-  mutate(num_games = case_when(endsWith(column_label, "4") ~ 1000,
-                               endsWith(column_label, "5") ~ 5000,
-                               endsWith(column_label, "6") ~ 20000),
-         num_games = factor(num_games, levels = c("1000", "5000", "20000"))) #make sure the levels are messing up here
-
-big_effect_draws = draws_list[7:9] %>% bind_rows(.id = "column_label") %>% 
-  mutate(num_games = case_when(endsWith(column_label, "7") ~ 1000,
-                               endsWith(column_label, "8") ~ 5000,
-                               endsWith(column_label, "9") ~ 20000),
-         num_games = factor(num_games, levels = c("1000", "5000", "20000"))) #make sure the levels are messing up here
-
-## small mu_beta plot ##
-
 #true experiential effects
-true_betas = sim_datasets_list[[1]] %>% 
-  group_by(player_id) %>% 
-  summarise(true_beta = unique(beta)) %>%
-  rename(id = player_id)
-
-#reordering
-true_betas = true_betas[match(beta_order, true_betas$id),] %>% 
+true_small_betas = sim_datasets_list[[1]]$beta %>% unique()
+true_med_betas = sim_datasets_list[[4]]$beta %>% unique()
+true_big_betas = sim_datasets_list[[7]]$beta %>% unique()
+
+beta_draws = beta_draws_list %>% bind_rows(.id = "dataset_id") %>%
+  rename(draw = value) %>%
+  mutate(dataset_id = substr(dataset_id, nchar(dataset_id), nchar(dataset_id)),
+         dataset_id = as.numeric(dataset_id),
+         true_beta = case_when(dataset_id %in% 1:3 ~ true_small_betas[player_id],
+                               dataset_id %in% 4:6 ~ true_med_betas[player_id],
+                               dataset_id %in% 7:9 ~ true_big_betas[player_id]),
+         num_games = case_when(dataset_id %in% c(1, 4, 7) ~ 1000,
+                               dataset_id %in% c(2, 5, 8) ~ 5000,
+                               dataset_id %in% c(3, 6, 9) ~ 20000),
+         num_games = factor(num_games, levels = c("1000", "5000", "20000")))
+
+#for the geom_segment below
+true_betas_plot = data.frame(true_small_val = true_small_betas, 
+                             true_med_val = true_med_betas, 
+                             true_big_val = true_big_betas, 
+                             player_id = 1:length(true_small_betas)) %>%
+  arrange(true_small_val) %>% #all orders are the same
   mutate(ymin = as.numeric(factor(row_number())) - 0.3,
@@
 
-#plot
-small_effect_plot = small_effect_draws  |>
-  ggplot(aes(y = reorder(id, value, FUN = median), x = value, colour = num_games)) +
-  #stat_histinterval(aes(fill = num_games), position = "dodgejust") +
+small_effect_plot = beta_draws %>% filter(dataset_id %in% 1:3) %>%
+  ggplot(aes(y = reorder(player_id, true_beta), x = draw, colour = num_games)) +
   stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
-  #stat_dots(aes(colour = num_games), position = "dodgejust") +
-  geom_segment(data = true_betas,
-               aes(x = true_beta, xend = true_beta, y = ymin, yend = ymax),
+  geom_segment(data = true_betas_plot,
+               aes(x = true_small_val, xend = true_small_val, y = ymin, yend = ymax),
                color = "red", linewidth = 1.5) +
@@
 #save
-ggsave(plot = small_effect_plot, 
-       filename = here("Paper_WriteUp", "paper_figures", "fig_sim_1.png"), dpi = 1000,
-       height = 6, width = 7)
+# ggsave(plot = small_effect_plot,
+#        filename = here("Paper_WriteUp", "paper_figures", "fig_sim_1.png"), dpi = 1000,
+#        height = 6, width = 7)
 
@@
 
-#true experiential effects
-true_betas = sim_datasets_list[[4]] %>% 
-  group_by(player_id) %>% 
-  summarise(true_beta = unique(beta)) %>%
-  rename(id = player_id)
-
-#reordering
-true_betas = true_betas[match(beta_order, true_betas$id),] %>% 
-  mutate(ymin = as.numeric(factor(row_number())) - 0.3,
-         ymax = as.numeric(factor(row_number())) + 0.3)
-
-#plot
-med_effect_plot = med_effect_draws  |>
-  ggplot(aes(y = reorder(id, value, FUN = median), x = value, colour = num_games)) +
+med_effect_plot = beta_draws %>% filter(dataset_id %in% 4:6) %>%
+  ggplot(aes(y = reorder(player_id, true_beta), x = draw, colour = num_games)) +
   #stat_histinterval(aes(fill = num_games), position = "dodgejust") +
@@
   #stat_dots(aes(colour = num_games), position = "dodgejust") +
-  geom_segment(data = true_betas,
-               aes(x = true_beta, xend = true_beta, y = ymin, yend = ymax),
+  geom_segment(data = true_betas_plot,
+               aes(x = true_med_val, xend = true_med_val, y = ymin, yend = ymax),
                color = "red", linewidth = 1.5) +
@@
 #save
-ggsave(plot = med_effect_plot, 
-       filename = here("Paper_WriteUp", "paper_figures", "fig_sim_2.png"), dpi = 1000,
-       height = 6, width = 7)
+# ggsave(plot = med_effect_plot,
+#        filename = here("Paper_WriteUp", "paper_figures", "fig_sim_2.png"), dpi = 1000,
+#        height = 6, width = 7)
 
@@
 
-#true experiential effects
-true_betas = sim_datasets_list[[7]] %>% 
-  group_by(player_id) %>% 
-  summarise(true_beta = unique(beta)) %>%
-  rename(id = player_id)
-
-#reordering
-true_betas = true_betas[match(beta_order, true_betas$id),] %>% 
-  mutate(ymin = as.numeric(factor(row_number())) - 0.3,
-         ymax = as.numeric(factor(row_number())) + 0.3)
-
 #plot
-big_effect_plot = big_effect_draws  |>
-  ggplot(aes(y = reorder(id, value, FUN = median), x = value, colour = num_games)) +
+big_effect_plot = beta_draws %>% filter(dataset_id %in% 7:9) %>%
+  ggplot(aes(y = reorder(player_id, true_beta), x = draw, colour = num_games)) +
   #stat_histinterval(aes(fill = num_games), position = "dodgejust") +
@@
   #stat_dots(aes(colour = num_games), position = "dodgejust") +
-  geom_segment(data = true_betas,
-               aes(x = true_beta, xend = true_beta, y = ymin, yend = ymax),
+  geom_segment(data = true_betas_plot,
+               aes(x = true_big_val, xend = true_big_val, y = ymin, yend = ymax),
                color = "red", linewidth = 1.5) +
@@
 #save
-ggsave(plot = big_effect_plot, 
-       filename = here("Paper_WriteUp", "paper_figures", "fig_sim_3.png"), dpi = 1000,
-       height = 6, width = 7)
+# ggsave(plot = big_effect_plot,
+#        filename = here("Paper_WriteUp", "paper_figures", "fig_sim_3.png"), dpi = 1000,
+#        height = 6, width = 7)
+```
+
+### mu_beta simulation comparison
+
+```{r}
+## same above plot but now mu_beta
+mu_beta_draws_list = lapply(sim_fit_list, function(x) {
+  x$draws("mu_beta") |>
+  as_draws_df()
+})
+
+#true values of mu_beta
+true_mu_betas = sim_datasets_list %>% bind_rows() %>% pull("mu_beta") %>% unique()
+
+#getting the draws
+mu_beta_draws = mu_beta_draws_list %>% bind_rows(.id = "dataset_id") %>%
+  rename(draw = mu_beta) %>%
+  mutate(dataset_id = substr(dataset_id, nchar(dataset_id), nchar(dataset_id)),
+         dataset_id = as.numeric(dataset_id),
+         true_mu_beta = case_when(dataset_id %in% 1:3 ~ true_mu_betas[1],
+                                  dataset_id %in% 4:6 ~ true_mu_betas[2],
+                                  dataset_id %in% 7:9 ~ true_mu_betas[3]),
+         num_games = case_when(dataset_id %in% c(1, 4, 7) ~ 1000,
+                               dataset_id %in% c(2, 5, 8) ~ 5000,
+                               dataset_id %in% c(3, 6, 9) ~ 20000),
+         true_mu_beta = factor(true_mu_beta, levels = c(as.character(true_mu_betas))),
+         num_games = factor(num_games, levels = c("1000", "5000", "20000")))
+
+#for the geom_segment below
+true_mu_betas_plotting = data.frame(true_val = true_mu_betas) %>%
+  mutate(ymin = as.numeric(factor(true_val)) - 0.25,
+         ymax = as.numeric(factor(true_val)) + 0.25)
+
+#plotting
+mu_beta_plot = mu_beta_draws %>% ggplot(aes(y = true_mu_beta, x = draw, colour = num_games)) +
+  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
+  geom_segment(data = true_mu_betas_plotting,
+               aes(x = true_val, xend = true_val, y = ymin, yend = ymax),
+               color = "red", linewidth = 1.5) +
+  labs(x = "Estimated Experiential Effect", y = "True mu_beta Value", colour = "Number of Games\nPlayed per Player") +
+  theme_single_y_legend() +
+  scale_color_viridis_d(option = "D")
+
+#save
+# ggsave(plot = mu_beta_plot,
+#        filename = here("Paper_WriteUp", "paper_figures", "fig_sim_mu_beta.png"), dpi = 1000,
+#        height = 6, width = 7)
 ```

### 2025-04-18 Paper_WriteUp/Paper_Plots.qmd
update code for figure 3

@@
 
-## 1700-1900 Previous Game
-
 ### Global Parameters
 
-```{r figure 3 - Updated}
+```{r figure 3}
 ### Adam
@@
 
-gm_amateur_comp_global <- bind_rows(pars_17_19, pars_20_22, pars_23_25, pars_gm) |> 
+comp_global <- bind_rows(pars_17_19, pars_20_22, pars_23_25, pars_gm) |> 
   ggplot(aes(x = value, y = name, color = factor(model, levels = c("1700-1900", "2000-2200", "2300-2500", "GM")))) +
@@
 
-gm_amateur_comp_global
-
-```
-
-```{r figure 3 left - plot global parameters for 17-19}
-data_path <- here("results/lichess1700-1900_2025")
-
-fit_17_19_bullet_prev <- readRDS(file = 
-                                   paste0(data_path,
-                                          "/all_rated_bullet_model_prev_n1.RDS"))
-
-bullet_17_19 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
-
-facet_labels <- as_labeller(c(
-  mu_beta = "mu[beta]",
-  gamma1 = "gamma[1]",
-  gamma2 = "gamma[2]",
-  sigma_1 = "sigma[1]",
-  `tau[1]` = "tau[1]",
-  `tau[2]` = "tau[2]",
-  sigma_g1 = "sigma[g[1]]",
-  sigma_g2 = "sigma[g[2]]"
-), label_parsed)
-
-global_17_19_bullet_prev <- fit_17_19_bullet_prev$draws(c("mu_beta",
-                                                          "gamma1", "gamma2",
-                           "sigma_1", "tau[1]", "tau[2]",
-                           "sigma_g1", "sigma_g2")) |> 
-  as_draws_df() |> 
-  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
-  ggplot(aes(value)) +
-  geom_histogram() +
-  facet_wrap(~name, ncol = 4, scales = "free",
-             labeller = labeller(name = facet_labels)) +
-  labs(x = element_blank(), y = element_blank()) +
-  theme_4col() +
-  scale_x_continuous(breaks = breaks_pretty(n = 3))
-
-global_17_19_bullet_prev
-ggsave(plot = global_17_19_bullet_prev, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "17_19_prev_global.png")), dpi = 600,
+ggsave(plot = comp_global, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures", "fig3.png")), dpi = 600,
        height = 5, width = 7)
@@
 #| eval: false
-gamma_ests <- fit_17_19_bullet_prev$summary(variables = 
+gamma_ests <- model_17_19$summary(variables = 
                                               c("gamma1",
@@
 1/(1 + exp(-1 * (100 * gamma_ests[2])))
+```
+
+```{r interpret the estimated coefficients gm}
+#| eval: false
+gamma_ests_gm <- model_gm$summary(variables = c("gamma1",
+                                                          "gamma2")) |> 
+  pull(mean)
+## compute change in win probability for gamma1
+1/(1 + exp(-1 * (0)))
+1/(1 + exp(-1 * (gamma_ests_gm[1])))
+## change in win probability for elo difference
+1/(1 + exp(-1 * (0))) ## equally rated, playing as black
+1/(1 + exp(-1 * (100 * gamma_ests_gm[2])))
 ```
@@
        height = 7, width = 7)
-```
-
-## GM previous game
-
-### Global Parameters
-
-```{r figure 3 right - global parameters for gm data}
-data_path <- here("results_2025/lichessGrandmasters")
-
-fit_gm_bullet_prev <- 
-  readRDS(file = paste0(data_path,"/all_rated_bullet_model_prev_n1.RDS"))
-
-bullet_gm <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
-
-global_gm_bullet_prev <- fit_gm_bullet_prev$draws(c("mu_beta",
-                                                          "gamma1", "gamma2",
-                           "sigma_1", "tau[1]", "tau[2]",
-                           "sigma_g1", "sigma_g2")) |> 
-  as_draws_df() |> 
-  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
-  ggplot(aes(value)) +
-  geom_histogram() +
-   facet_wrap(~name, ncol = 4, scales = "free",
-             labeller = labeller(name = facet_labels)) +
-  labs(x = element_blank(), y = element_blank()) +
-  theme_4col() +
-  scale_x_continuous(breaks = breaks_pretty(n = 2))
-
-global_gm_bullet_prev
-ggsave(plot = global_gm_bullet_prev, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "gm_prev_global.png")), dpi = 600,
-       height = 5, width = 7)
-```
-
-```{r interpret the estimated coefficients gm}
-#| eval: false
-gamma_ests_gm <- fit_gm_bullet_prev$summary(variables = c("gamma1",
-                                                          "gamma2")) |> 
-  pull(mean)
-## compute change in win probability for gamma1
-1/(1 + exp(-1 * (0)))
-1/(1 + exp(-1 * (gamma_ests_gm[1])))
-## change in win probability for elo difference
-1/(1 + exp(-1 * (0))) ## equally rated, playing as black
-1/(1 + exp(-1 * (100 * gamma_ests_gm[2])))
 ```

### 2025-04-18 Paper_WriteUp/Paper_Plots.qmd
consistent model names and fixing xlim on fig4 right

@@
 #read in models
-model_17_19 = readRDS(here("results_revision", "lichess1700-1900", "all_rated_bullet_model_prev_n1.RDS"))
-model_20_22 = readRDS(here("results_revision", "lichess2000-2200", "all_rated_bullet_model_prev_n1.RDS"))
-model_23_25 = readRDS(here("results_revision", "lichess2300-2500", "all_rated_bullet_model_prev_n1.RDS"))
-model_gm = readRDS(here("results_revision", "lichessGrandmasters", "all_rated_bullet_model_prev_n1.RDS"))
+fit_17_19_bullet_prev  = readRDS(here("results_revision", "lichess1700-1900", "all_rated_bullet_model_prev_n1.RDS"))
+fit_20_22_bullet_prev  = readRDS(here("results_revision", "lichess2000-2200", "all_rated_bullet_model_prev_n1.RDS"))
+fit_23_25_bullet_prev = readRDS(here("results_revision", "lichess2300-2500", "all_rated_bullet_model_prev_n1.RDS"))
+fit_gm_bullet_prev  = readRDS(here("results_revision", "lichessGrandmasters", "all_rated_bullet_model_prev_n1.RDS"))
+
+#read in users
+bullet_17_19 = readRDS(here("results_revision", "lichess1700-1900", "users_bullet.RDS"))
+bullet_20_22 = readRDS(here("results_revision", "lichess2000-2200", "users_bullet.RDS"))
+bullet_23_25 = readRDS(here("results_revision", "lichess2300-2500", "users_bullet.RDS"))
+bullet_gm = readRDS(here("results_revision", "lichessGrandmasters", "users_bullet.RDS"))
 
 #get global effects of each
-pars_17_19 <- model_17_19$draws(c("mu_beta", "gamma1", "gamma2")) |> 
+pars_17_19 = fit_17_19_bullet_prev $draws(c("mu_beta", "gamma1", "gamma2")) |> 
   as_draws_df() |> 
@@
 
-pars_20_22 = model_20_22$draws(c("mu_beta", "gamma1", "gamma2")) |> 
+pars_20_22 = fit_20_22_bullet_prev$draws(c("mu_beta", "gamma1", "gamma2")) |> 
   as_draws_df() |> 
@@
 
-pars_23_25 = model_23_25$draws(c("mu_beta", "gamma1", "gamma2")) |> 
+pars_23_25 = fit_23_25_bullet_prev$draws(c("mu_beta", "gamma1", "gamma2")) |> 
   as_draws_df() |> 
@@
 
-pars_gm <- model_gm$draws(c("mu_beta", "gamma1", "gamma2")) |> 
+pars_gm = fit_gm_bullet_prev$draws(c("mu_beta", "gamma1", "gamma2")) |> 
   as_draws_df() |> 
@@
 #for plotting
-facet_labels <- as_labeller(c(
+facet_labels = as_labeller(c(
   mu_beta = "mu[beta]",
@@
 
-comp_global <- bind_rows(pars_17_19, pars_20_22, pars_23_25, pars_gm) |> 
+comp_global = bind_rows(pars_17_19, pars_20_22, pars_23_25, pars_gm) |> 
   ggplot(aes(x = value, y = name, color = factor(model, levels = c("1700-1900", "2000-2200", "2300-2500", "GM")))) +
@@
 #| eval: false
-
 player_id <- as_tibble(bullet_17_19) |> 
@@
   geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
-  xlim(c(-0.1, 0.1)) +
   theme_single_y()
@@
   geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
+  xlim(c(-0.1, 0.1)) +
   theme_single_y()
-
-winner_plot_gm
 
@@
 
+# Beta simulation comparison
+
 ```{r}
@@
 
-### mu_beta simulation comparison
+# mu_beta simulation comparison
 

### 2025-04-18 Paper_WriteUp/Paper_Plots.qmd
moving unnecessary plots to appendix section

@@
 
-```{r}
-#| eval: false
-## TO DO - move to supplement
-## not required for a paper plot
-player_pars_pre <- fit_gm_bullet_prev$draws() |> 
-  as_draws_df() |>
-  select(starts_with("beta[")) |>
-  pivot_longer(cols = everything()) |>
-  mutate(param = stringr::str_extract(name, pattern = "\\d"),
-         id = stringr::str_extract(name, pattern = "\\d+]"),
-         id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]")) |> 
-  mutate(id = as.numeric(id),
-         param = as.numeric(param)) |> 
-  left_join(player_id, by = "id") |> 
-  mutate(model = "Prev. Game")
-
-player_pars_pre_10 <- fit_gm_bullet_10$draws() |> 
-  as_draws_df() |>
-  select(starts_with("beta[")) |>
-  pivot_longer(cols = everything()) |>
-  mutate(param = stringr::str_extract(name, pattern = "\\d"),
-         id = stringr::str_extract(name, pattern = "\\d+]"),
-         id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]")) |> 
-  mutate(id = as.numeric(id),
-         param = as.numeric(param)) |> 
-  left_join(player_id, by = "id") |> 
-  mutate(model = "Prev. 10 Games")
-
-hist_comp_winner <- player_pars_pre |> 
-  filter(param == 2) |> 
-  mutate(y = reorder(player_name, value, FUN = median)) |> 
-  bind_rows(player_pars_pre_10 |> filter(param == 2) |> 
-              mutate(y = as.factor(player_name))) |> 
-  ggplot(aes(x = value, y = y,
-             color = model)) +
-  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
-  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
-  theme_single_y() +
-  theme(legend.position = "bottom") +
-  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
-  scale_color_viridis_d(option = "H")
-
-hist_comp_winner
-
-ggsave(plot = hist_comp_winner, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "gm_comp_winner.png")), dpi = 600,
-       height = 5, width = 7)
-
-## check the proportion of 95% intervals which contain zero
-bind_rows(player_pars_pre, player_pars_pre_10) |> 
-  filter(param == 2) |> 
-  group_by(player_name, model) |> 
-  summarise(lower = quantile(value, prob = 0.025),
-            upper = quantile(value, prob = 0.975)) |> 
-  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
-                         1, 0)) |> 
-  group_by(model) |> 
-  summarise(prop = sum(contain_zero)/ n())
-
-## do the effects change in direction
-bind_rows(player_pars_pre, player_pars_pre_10) |> 
-  filter(param == 2) |> 
-  group_by(player_name, model) |> 
-  summarise(lower = quantile(value, prob = 0.025),
-            upper = quantile(value, prob = 0.975)) |> 
-  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
-                         1, 0)) |>
-  print(n = 50)
-
-## then player effects
-## should update this also
-hist_comp_player <- player_pars_pre |> 
-  filter(param == 1) |> 
-  mutate(y = reorder(player_name, value, FUN = median)) |> 
-  bind_rows(player_pars_pre_10 |> filter(param == 1) |> 
-              mutate(y = as.factor(player_name))) |> 
-  ggplot(aes(x = value, y = y,
-             color = model)) +
-  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
-  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
-  theme_single_y() +
-  theme(legend.position = "bottom") +
-  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
-  scale_color_viridis_d(option = "H")
-
-hist_comp_player
-ggsave(plot = hist_comp_player, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "gm_comp_player.png")), dpi = 600,
-       height = 5, width = 7)
-```
-
-## Bullet vs Blitz for 17-19
-
 Compare the global parameters and the winner effects here, using the
 previous game only.
-
-```{r bullet v blitz for 17-19}
-#| eval: false
-
-## Potential Supplement Plots
-data_path <- here("results_2025/lichess1700-1900")
-fit_17_19_bullet_prev <- readRDS(file = paste0(
-  data_path,
-  "/all_rated_bullet_model_prev_n1.RDS"
-))
-bullet_17_19 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
-fit_17_19_blitz_prev <- readRDS(file = paste0(
-  data_path,
-  "/all_rated_blitz_model_prev_n1.RDS"
-))
-blitz_17_19 <- readRDS(file = paste0(data_path, "/users_blitz.RDS"))
-
-player_id_17_19 <- as_tibble(bullet_17_19) |> 
-  mutate(id = row_number()) |> 
-  rename(player_name = value)
-
-## compare global effects first
-bullet_pars <- fit_17_19_bullet_prev$draws(c(
-  "mu_beta",
-  "gamma1", "gamma2"
-)) |>
-  as_draws_df() |>
-  pivot_longer(cols = "mu_beta":"gamma2") |>
-  mutate(model = "Bullet")
-
-blitz_pars <- fit_17_19_blitz_prev$draws(c(
-  "mu_beta",
-  "gamma1", "gamma2"
-)) |>
-  as_draws_df() |>
-  pivot_longer(cols = "mu_beta":"gamma2") |>
-  mutate(model = "Blitz")
-
-bul_blit_comp_global <- bind_rows(bullet_pars, blitz_pars) |>
-  ggplot(aes(value, color = model))  +
-  stat_pointinterval(position = position_dodge(width = 0.2)) +
-  facet_wrap(~name,
-             scales = "free",
-             labeller = labeller(name = facet_labels)) +
-  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
-  theme_3col() +
-  theme(legend.position = "bottom") +
-  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
-  scale_color_viridis_d(option = "H")
-
-bul_blit_comp_global
-
-ggsave(plot = bul_blit_comp_global, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "bul_blitz_comp_17_19_global.png")), dpi = 600,
-       height = 5, width = 7)
-
-## then repeat for winner and player effects
-player_pars_bullet <- fit_17_19_bullet_prev$draws() |> 
-  as_draws_df() |>
-  select(starts_with("beta[")) |>
-  pivot_longer(cols = everything()) |>
-  mutate(param = stringr::str_extract(name, pattern = "\\d"),
-         id = stringr::str_extract(name, pattern = "\\d+]"),
-         id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]")) |> 
-  mutate(id = as.numeric(id),
-         param = as.numeric(param)) |> 
-  left_join(player_id_17_19, by = "id") |> 
-  mutate(model = "Bullet")
-
-player_pars_blitz <- fit_17_19_blitz_prev$draws() |> 
-  as_draws_df() |>
-  select(starts_with("beta[")) |>
-  pivot_longer(cols = everything()) |>
-  mutate(param = stringr::str_extract(name, pattern = "\\d"),
-         id = stringr::str_extract(name, pattern = "\\d+]"),
-         id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]")) |> 
-  mutate(id = as.numeric(id),
-         param = as.numeric(param)) |> 
-  left_join(player_id_17_19, by = "id") |> 
-  mutate(model = "Blitz")
-
-## need to keep only players which are in both 
-## bullet and blitz
-
-common_players <- intersect(unique(player_pars_bullet$player_name),
-                            unique(player_pars_blitz$player_name))
-
-comp_17_19_winner <- player_pars_bullet |> 
-  filter(param == 2) |> 
-  filter(player_name %in% common_players) |> 
-  mutate(y = reorder(player_name, value, FUN = median)) |> 
-  bind_rows(player_pars_blitz |>
-              filter(param == 2) |>
-              filter(player_name %in% common_players) |>
-              mutate(y = as.factor(player_name))) |>
-  ggplot(aes(x = value, y = y,
-             color = model)) +
-  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
-  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
-  theme_single_y() +
-  theme(legend.position = "bottom") +
-  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
-  scale_color_viridis_d(option = "H")
-
-comp_17_19_winner
-
-ggsave(plot = comp_17_19_winner, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "bul_blitz_comp_17_19_winner.png")), dpi = 600,
-       height = 5, width = 7)
-
-comp_17_19_player <- player_pars_bullet |> 
-  filter(param == 1) |> 
-  filter(player_name %in% common_players) |> 
-  mutate(y = reorder(player_name, value, FUN = median)) |> 
-  bind_rows(player_pars_blitz |>
-              filter(param == 1) |>
-              filter(player_name %in% common_players) |>
-              mutate(y = as.factor(player_name))) |>
-  ggplot(aes(x = value, y = y,
-             color = model)) +
-  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
-  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
-  theme_single_y() +
-  theme(legend.position = "bottom") +
-  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
-  scale_color_viridis_d(option = "H")
-
-comp_17_19_player
-
-ggsave(plot = comp_17_19_player, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "bul_blitz_comp_17_19_player.png")), dpi = 600,
-       height = 5, width = 7)
-```
 
@@
                               "bul_blitz_comp_gm_winner.png")), dpi = 600,
-       height = 5, width = 7)
-```
-
-```{r}
-#| eval: false
-## player effects also
-comp_gm_player <- player_pars_bullet |> 
-  filter(param == 1) |> 
-  mutate(y = reorder(player_name, value, FUN = median)) |> 
-  bind_rows(player_pars_blitz |> 
-              filter(param == 1) |> 
-              mutate(y = as.factor(player_name))) |> 
-  ggplot(aes(x = value, y = y,
-             color = model)) +
-  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
-  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
-  theme_single_y() +
-  theme(legend.position = "bottom") +
-  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
-  scale_color_viridis_d(option = "H")
-
-comp_gm_player
-
-ggsave(plot = comp_gm_player, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "bul_blitz_comp_gm_player.png")), dpi = 600,
        height = 5, width = 7)
@@
 ```
+
+### OTHER
+
+## Bullet vs Blitz for 17-19
+```{r bullet v blitz for 17-19}
+#| eval: false
+
+## Potential Supplement Plots
+data_path <- here("results_2025/lichess1700-1900")
+fit_17_19_bullet_prev <- readRDS(file = paste0(
+  data_path,
+  "/all_rated_bullet_model_prev_n1.RDS"
+))
+bullet_17_19 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
+fit_17_19_blitz_prev <- readRDS(file = paste0(
+  data_path,
+  "/all_rated_blitz_model_prev_n1.RDS"
+))
+blitz_17_19 <- readRDS(file = paste0(data_path, "/users_blitz.RDS"))
+
+player_id_17_19 <- as_tibble(bullet_17_19) |> 
+  mutate(id = row_number()) |> 
+  rename(player_name = value)
+
+## compare global effects first
+bullet_pars <- fit_17_19_bullet_prev$draws(c(
+  "mu_beta",
+  "gamma1", "gamma2"
+)) |>
+  as_draws_df() |>
+  pivot_longer(cols = "mu_beta":"gamma2") |>
+  mutate(model = "Bullet")
+
+blitz_pars <- fit_17_19_blitz_prev$draws(c(
+  "mu_beta",
+  "gamma1", "gamma2"
+)) |>
+  as_draws_df() |>
+  pivot_longer(cols = "mu_beta":"gamma2") |>
+  mutate(model = "Blitz")
+
+bul_blit_comp_global <- bind_rows(bullet_pars, blitz_pars) |>
+  ggplot(aes(value, color = model))  +
+  stat_pointinterval(position = position_dodge(width = 0.2)) +
+  facet_wrap(~name,
+             scales = "free",
+             labeller = labeller(name = facet_labels)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_3col() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_color_viridis_d(option = "H")
+
+bul_blit_comp_global
+
+ggsave(plot = bul_blit_comp_global, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "bul_blitz_comp_17_19_global.png")), dpi = 600,
+       height = 5, width = 7)
+
+## then repeat for winner and player effects
+player_pars_bullet <- fit_17_19_bullet_prev$draws() |> 
+  as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]")) |> 
+  mutate(id = as.numeric(id),
+         param = as.numeric(param)) |> 
+  left_join(player_id_17_19, by = "id") |> 
+  mutate(model = "Bullet")
+
+player_pars_blitz <- fit_17_19_blitz_prev$draws() |> 
+  as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]")) |> 
+  mutate(id = as.numeric(id),
+         param = as.numeric(param)) |> 
+  left_join(player_id_17_19, by = "id") |> 
+  mutate(model = "Blitz")
+
+## need to keep only players which are in both 
+## bullet and blitz
+
+common_players <- intersect(unique(player_pars_bullet$player_name),
+                            unique(player_pars_blitz$player_name))
+
+comp_17_19_winner <- player_pars_bullet |> 
+  filter(param == 2) |> 
+  filter(player_name %in% common_players) |> 
+  mutate(y = reorder(player_name, value, FUN = median)) |> 
+  bind_rows(player_pars_blitz |>
+              filter(param == 2) |>
+              filter(player_name %in% common_players) |>
+              mutate(y = as.factor(player_name))) |>
+  ggplot(aes(x = value, y = y,
+             color = model)) +
+  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_single_y() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_color_viridis_d(option = "H")
+
+comp_17_19_winner
+
+ggsave(plot = comp_17_19_winner, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "bul_blitz_comp_17_19_winner.png")), dpi = 600,
+       height = 5, width = 7)
+
+comp_17_19_player <- player_pars_bullet |> 
+  filter(param == 1) |> 
+  filter(player_name %in% common_players) |> 
+  mutate(y = reorder(player_name, value, FUN = median)) |> 
+  bind_rows(player_pars_blitz |>
+              filter(param == 1) |>
+              filter(player_name %in% common_players) |>
+              mutate(y = as.factor(player_name))) |>
+  ggplot(aes(x = value, y = y,
+             color = model)) +
+  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_single_y() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_color_viridis_d(option = "H")
+
+comp_17_19_player
+
+ggsave(plot = comp_17_19_player, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "bul_blitz_comp_17_19_player.png")), dpi = 600,
+       height = 5, width = 7)
+```
+
+#winner and player effects for bullet vs blitz
+```{r}
+#| eval: false
+## TO DO - move to supplement
+## not required for a paper plot
+player_pars_pre <- fit_gm_bullet_prev$draws() |> 
+  as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]")) |> 
+  mutate(id = as.numeric(id),
+         param = as.numeric(param)) |> 
+  left_join(player_id, by = "id") |> 
+  mutate(model = "Prev. Game")
+
+player_pars_pre_10 <- fit_gm_bullet_10$draws() |> 
+  as_draws_df() |>
+  select(starts_with("beta[")) |>
+  pivot_longer(cols = everything()) |>
+  mutate(param = stringr::str_extract(name, pattern = "\\d"),
+         id = stringr::str_extract(name, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]")) |> 
+  mutate(id = as.numeric(id),
+         param = as.numeric(param)) |> 
+  left_join(player_id, by = "id") |> 
+  mutate(model = "Prev. 10 Games")
+
+hist_comp_winner <- player_pars_pre |> 
+  filter(param == 2) |> 
+  mutate(y = reorder(player_name, value, FUN = median)) |> 
+  bind_rows(player_pars_pre_10 |> filter(param == 2) |> 
+              mutate(y = as.factor(player_name))) |> 
+  ggplot(aes(x = value, y = y,
+             color = model)) +
+  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_single_y() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_color_viridis_d(option = "H")
+
+hist_comp_winner
+
+ggsave(plot = hist_comp_winner, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "gm_comp_winner.png")), dpi = 600,
+       height = 5, width = 7)
+
+## check the proportion of 95% intervals which contain zero
+bind_rows(player_pars_pre, player_pars_pre_10) |> 
+  filter(param == 2) |> 
+  group_by(player_name, model) |> 
+  summarise(lower = quantile(value, prob = 0.025),
+            upper = quantile(value, prob = 0.975)) |> 
+  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
+                         1, 0)) |> 
+  group_by(model) |> 
+  summarise(prop = sum(contain_zero)/ n())
+
+## do the effects change in direction
+bind_rows(player_pars_pre, player_pars_pre_10) |> 
+  filter(param == 2) |> 
+  group_by(player_name, model) |> 
+  summarise(lower = quantile(value, prob = 0.025),
+            upper = quantile(value, prob = 0.975)) |> 
+  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
+                         1, 0)) |>
+  print(n = 50)
+
+## then player effects
+## should update this also
+hist_comp_player <- player_pars_pre |> 
+  filter(param == 1) |> 
+  mutate(y = reorder(player_name, value, FUN = median)) |> 
+  bind_rows(player_pars_pre_10 |> filter(param == 1) |> 
+              mutate(y = as.factor(player_name))) |> 
+  ggplot(aes(x = value, y = y,
+             color = model)) +
+  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_single_y() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_color_viridis_d(option = "H")
+
+hist_comp_player
+ggsave(plot = hist_comp_player, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "gm_comp_player.png")), dpi = 600,
+       height = 5, width = 7)
+```
+
+#bullet vs blitz player effects for GMs
+```{r}
+#| eval: false
+## player effects also
+comp_gm_player <- player_pars_bullet |> 
+  filter(param == 1) |> 
+  mutate(y = reorder(player_name, value, FUN = median)) |> 
+  bind_rows(player_pars_blitz |> 
+              filter(param == 1) |> 
+              mutate(y = as.factor(player_name))) |> 
+  ggplot(aes(x = value, y = y,
+             color = model)) +
+  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
+  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
+  theme_single_y() +
+  theme(legend.position = "bottom") +
+  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
+  scale_color_viridis_d(option = "H")
+
+comp_gm_player
+
+ggsave(plot = comp_gm_player, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "bul_blitz_comp_gm_player.png")), dpi = 600,
+       height = 5, width = 7)
+```

### 2025-04-22 Paper_WriteUp/Paper_Plots.qmd
updating figs with revised models

@@
 #read in models
-fit_17_19_bullet_prev  = readRDS(here("results_revision", "lichess1700-1900", "all_rated_bullet_model_prev_n1.RDS"))
-fit_20_22_bullet_prev  = readRDS(here("results_revision", "lichess2000-2200", "all_rated_bullet_model_prev_n1.RDS"))
-fit_23_25_bullet_prev = readRDS(here("results_revision", "lichess2300-2500", "all_rated_bullet_model_prev_n1.RDS"))
-fit_gm_bullet_prev  = readRDS(here("results_revision", "lichessGrandmasters", "all_rated_bullet_model_prev_n1.RDS"))
+fit_17_19_bullet_prev  = readRDS(here("results_revision", "lichess1700-1900", "all_rated_bullet_model_n1.RDS"))
+fit_20_22_bullet_prev  = readRDS(here("results_revision", "lichess2000-2200", "all_rated_bullet_model_n1.RDS"))
+fit_23_25_bullet_prev = readRDS(here("results_revision", "lichess2300-2500", "all_rated_bullet_model_n1.RDS"))
+fit_gm_bullet_prev  = readRDS(here("results_revision", "lichessGrandmasters", "all_rated_bullet_model_n1.RDS"))
 
@@
 
-ggsave(plot = comp_global, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures", "fig3.png")), dpi = 600,
-       height = 5, width = 7)
+# ggsave(plot = comp_global, 
+#        filename = paste0(here("Paper_WriteUp", "paper_figures", "fig3.png")), dpi = 600,
+#        height = 5, width = 7)
 ```
@@
 #| eval: false
-gamma_ests <- model_17_19$summary(variables = 
+gamma_ests <- fit_17_19_bullet_prev$summary(variables = 
                                               c("gamma1",
@@
 #| eval: false
-gamma_ests_gm <- model_gm$summary(variables = c("gamma1",
+gamma_ests_gm <- fit_gm_bullet_prev$summary(variables = c("gamma1",
                                                           "gamma2")) |> 
@@
   ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
-  stat_histinterval() +
+  stat_pointinterval() +
   labs(x = "Change in Win Probability", y = element_blank(), 
@@
   geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
-  xlim(c(-0.1, 0.1)) +
+  xlim(c(-0.125, 0.125)) +
   theme_single_y()
@@
        filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "17_19_prev_winner.png")), dpi = 600,
+                              "fig4_left.png")), dpi = 600,
        height = 7, width = 7)
@@
   ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
-  stat_histinterval() +
+  stat_pointinterval() +
   labs(x = "Estimated Player Effect", y = element_blank(),
@@
   ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
-  stat_histinterval() +
-  labs(x = "Estimated W/L-Effect", y = element_blank(), 
+  stat_pointinterval() +
+  labs(x = "Change in Win Probability", y = element_blank(), 
        title = "GM Cohort") +
   geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
-  xlim(c(-0.1, 0.1)) +
+  xlim(c(-0.125, 0.125)) +
   theme_single_y()
@@
        filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "gm_prev_winner.png")), dpi = 600,
+                              "fig4_right.png")), dpi = 600,
        height = 7, width = 7)
@@
   ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
-  stat_histinterval() +
+  stat_pointinterval() +
   labs(x = "Estimated Player Effect", y = element_blank(), 
@@
   theme_single_y()
-
-indiv_plot_gm
 
@@
                                dataset_id %in% c(3, 6, 9) ~ 20000),
-         true_mu_beta = factor(true_mu_beta, levels = c(as.character(true_mu_betas))),
+         true_mu_beta = factor(signif(true_mu_beta, 3), levels = c(as.character(signif(true_mu_betas, 3)))),
          num_games = factor(num_games, levels = c("1000", "5000", "20000")))
@@
 true_mu_betas_plotting = data.frame(true_val = true_mu_betas) %>%
-  mutate(ymin = as.numeric(factor(true_val)) - 0.25,
-         ymax = as.numeric(factor(true_val)) + 0.25)
+  mutate(ymin = as.numeric(as.factor(true_val)) - 0.3,
+         ymax = as.numeric(as.factor(true_val)) + 0.3)
 
@@
                aes(x = true_val, xend = true_val, y = ymin, yend = ymax),
-               color = "red", linewidth = 1.5) +
+               color = "red", linewidth = 1.25) +
   labs(x = "Estimated Experiential Effect", y = "True mu_beta Value", colour = "Number of Games\nPlayed per Player") +
   theme_single_y_legend() +
+  scale_x_continuous(breaks = c(signif(true_mu_betas, 3)), 
+                     labels = c(as.character(signif(true_mu_betas, 3)))) +
   scale_color_viridis_d(option = "D")
+mu_beta_plot
 

### 2025-04-23 Paper_WriteUp/Paper_Plots.qmd
updating x axis scales in fig 3

@@
 library(rstanarm)
+library(ggh4x)
 options(mc.cores = parallel::detectCores())
@@
   theme(legend.position = "bottom") +
-  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  facetted_pos_scales(x = list( #for customising scales on each facet
+    name == "mu_beta" ~ scale_x_continuous(breaks = c(-0.15, 0, 0.15), limits = c(-0.17, 0.17)),
+    name == "gamma1" ~ scale_x_continuous(breaks = breaks_pretty(n = 2)),
+    name == "gamma2" ~ scale_x_continuous(breaks = breaks_pretty(n = 2))
+  )) +
   scale_color_viridis_d(option = "H")
 
-# ggsave(plot = comp_global, 
-#        filename = paste0(here("Paper_WriteUp", "paper_figures", "fig3.png")), dpi = 600,
-#        height = 5, width = 7)
+ggsave(plot = comp_global,
+       filename = paste0(here("Paper_WriteUp", "paper_figures", "fig3.png")), dpi = 600,
+       height = 5, width = 7)
 ```
@@
   stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
-  geom_segment(data = true_mu_betas_plotting,
-               aes(x = true_val, xend = true_val, y = ymin, yend = ymax),
-               color = "red", linewidth = 1.25) +
+  # geom_segment(data = true_mu_betas_plotting,
+  #              aes(x = true_val, xend = true_val, y = ymin, yend = ymax),
+  #              color = "red", linewidth = 1.25) +
   labs(x = "Estimated Experiential Effect", y = "True mu_beta Value", colour = "Number of Games\nPlayed per Player") +
+  geom_vline(aes(xintercept = 0), col = "black", linetype = "dashed") +
   theme_single_y_legend() +
-  scale_x_continuous(breaks = c(signif(true_mu_betas, 3)), 
-                     labels = c(as.character(signif(true_mu_betas, 3)))) +
+  scale_x_continuous(breaks = c(0, signif(true_mu_betas, 3)), 
+                     labels = c(0, as.character(signif(true_mu_betas, 3)))) +
   scale_color_viridis_d(option = "D")

### 2025-04-23 Paper_WriteUp/Paper_Plots.qmd
updating mu_beta sim fig to have numerical distances on axes

@@
   theme(legend.position = "bottom") +
-  facetted_pos_scales(x = list( #for customising scales on each facet
+  facetted_pos_scales(x = list( #for customizing scales on each facet
     name == "mu_beta" ~ scale_x_continuous(breaks = c(-0.15, 0, 0.15), limits = c(-0.17, 0.17)),
@@
                                   dataset_id %in% 7:9 ~ true_mu_betas[3]),
-         num_games = case_when(dataset_id %in% c(1, 4, 7) ~ 1000,
-                               dataset_id %in% c(2, 5, 8) ~ 5000,
-                               dataset_id %in% c(3, 6, 9) ~ 20000),
-         true_mu_beta = factor(signif(true_mu_beta, 3), levels = c(as.character(signif(true_mu_betas, 3)))),
+         num_games = case_when(dataset_id %in% c(1, 4, 7) ~ "1000",
+                               dataset_id %in% c(2, 5, 8) ~ "5000",
+                               dataset_id %in% c(3, 6, 9) ~ "20000"),
+         #true_mu_beta = factor(signif(true_mu_beta, 3), levels = c(as.character(signif(true_mu_betas, 3)))),
          num_games = factor(num_games, levels = c("1000", "5000", "20000")))
@@
 true_mu_betas_plotting = data.frame(true_val = true_mu_betas) %>%
-  mutate(ymin = as.numeric(as.factor(true_val)) - 0.3,
-         ymax = as.numeric(as.factor(true_val)) + 0.3)
+  mutate(ymin = as.numeric(true_val) - 0.1,
+         ymax = as.numeric(true_val) + 0.1)
 
 #plotting
-mu_beta_plot = mu_beta_draws %>% ggplot(aes(y = true_mu_beta, x = draw, colour = num_games)) +
-  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
-  # geom_segment(data = true_mu_betas_plotting,
-  #              aes(x = true_val, xend = true_val, y = ymin, yend = ymax),
-  #              color = "red", linewidth = 1.25) +
-  labs(x = "Estimated Experiential Effect", y = "True mu_beta Value", colour = "Number of Games\nPlayed per Player") +
+mu_beta_plot = mu_beta_draws %>% ggplot(aes(x = draw, y = true_mu_beta)) +
+  stat_pointinterval(aes(group = interaction(true_mu_beta, num_games), colour = num_games), 
+                     position = ggstance::position_dodgev(height = 0.15)) +
+  geom_segment(data = true_mu_betas_plotting,
+               aes(x = true_val, xend = true_val, y = ymin, yend = ymax),
+               color = "red", linewidth = 1.1) +
+  labs(x = expression("Estimated " * mu[beta] * " Value"), 
+       y = expression("True " * mu[beta] * " Value"), 
+       colour = "Number of Games\nPlayed per Player") +
   geom_vline(aes(xintercept = 0), col = "black", linetype = "dashed") +
@@
   scale_x_continuous(breaks = c(0, signif(true_mu_betas, 3)), 
+                     labels = c(0, as.character(signif(true_mu_betas, 3)))) +
+  scale_y_continuous(breaks = c(0, signif(true_mu_betas, 3)), 
                      labels = c(0, as.character(signif(true_mu_betas, 3)))) +

### 2025-04-24 utils/helper.R
fixing focal_win_prop calculation when prev_n > 1. History only includes the games in current session, up to a max of prev_n

@@ -21,7 +21,6 @@ read_player <- function(path, file){
 get_hist <- function(user, games, prev_n) {
   if(prev_n == 1){
     hist_games <- games |> 
-      # filter(White == user | Black == user) |>
       filter(Username == user) |> 
       arrange(UTCDate, UTCTime) |> 
       mutate(focal_white = ifelse(Username == White, 1, 0)) |> 
@@ -37,21 +36,29 @@ get_hist <- function(user, games, prev_n) {
   }
   else{
     hist_games <- games |> 
-      # filter(White == user | Black == user) |> 
       filter(Username == user) |> 
-      arrange(UTCDate, UTCTime) |> 
-      mutate(focal_white = ifelse(Username == White, 1, 0)) |> 
-      select(White:BlackElo, focal_white) |> 
+      arrange(UTCDate, UTCTime) |>
+      mutate(focal_white = ifelse(Username == White, 1, 0),
+             UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime)),
+             time_diff = UTCDateTime - lag(UTCDateTime, default = NA)) |> 
+      select(White:BlackElo, focal_white, time_diff) |> 
       mutate(focal_result = case_when(
         (focal_white == 1 & Result == "1-0") ~ 1,
         (focal_white == 0 & Result == "0-1") ~ 1,
         (Result == "1/2-1/2") ~ 0.5,
-        .default = 0
-      )) |> 
-      mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
-                                roll_mean(focal_result, n = prev_n)))
+        .default = 0)
+      ) |> 
+      #updating so n = min(prev_n, number of games since first game in session)
+      mutate(n_session = (!(time_diff > 300 | is.na(time_diff))), #TRUE indicates game in a session, FALSE indicates first game in new session
+             n_session = 1 + accumulate(n_session, ~if (.y) {.x + 1} else {0}, .init = 0)[-1]) |> #the current number of games in a session
+      rowwise() |> 
+      mutate(n_hist = min(n_session, prev_n)) |> #the n to use for calculating history (ave_prop)
+      ungroup() |> #remove rowwise calculations
+      mutate(focal_win_prop = map_dbl(row_number(), function(i) { #rolling mean calculation (mean of past n_hist games)
+        start <- max(1, i - n_hist[i] + 1)
+        mean(focal_result[start:i])
+      }))
   }
-  
   hist_games
 }
 

### 2025-04-24 analysis_scripts/08_ppd_fit.R
script for fitting model on prev 1000 to final 1000 games (for PPDs)

@@ -1,0 +1,122 @@
+##### April 24th 2025 ######
+###########
+# Fit model on the previous 1000 games to the final 1000 games for each player
+# These fits will be used to generate the PPD plots
+
+library(tidyverse)
+library(RcppRoll)
+library(cmdstanr)
+library(posterior)
+library(bayesplot)
+library(loo)
+library(here)
+
+options(mc.cores = parallel::detectCores())
+
+## source helper functions for reading and transforming data
+## along with defaults for plots, etc
+source(here("utils/helper.R"))
+
+path_id <- as.numeric(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+
+n <- 1 ## number of games to use for history
+
+### load in the data to use
+
+all_data_path <- rep(NA, 4)
+all_save_path <- rep(NA, 4)
+
+all_data_path[1] <- here("box_data", "lichess1700-1900", "/")
+all_data_path[2] <- here("box_data", "lichess2000-2200", "/")
+all_data_path[3] <- here("box_data", "lichess2300-2500", "/")
+all_data_path[4] <- here("box_data", "lichessGrandmasters", "/")
+
+all_save_path[1] <- here("results_revision/lichess1700-1900/")
+all_save_path[2] <- here("results_revision/lichess2000-2200/")
+all_save_path[3] <- here("results_revision/lichess2300-2500/")
+all_save_path[4] <- here("results_revision/lichessGrandmasters/")
+
+
+data_path <- all_data_path[path_id]
+save_path <- all_save_path[path_id]
+
+dir.create(save_path, showWarnings = FALSE)
+
+files <- list.files(data_path)
+print(files)
+lichess_data <- files |> 
+  map_dfr(~read_player(data_path, .x))
+
+small_data <- lichess_data |>
+  mutate(Event = tolower(Event)) |>
+  # filter(Event == "Rated Bullet game") |>
+  filter(TimeControl == "60+0") |>
+  filter(Variant == "Standard") |>
+  filter(grepl("rated bullet game", Event)) |>
+  distinct() #remove the duplicate rows if they exist
+
+#selecting previous 1000 to the final 1000 games for each player
+mid_games <- small_data |> 
+  filter(Username %in% users) |> 
+  group_by(Username) |> 
+  arrange(UTCDate, UTCTime, .by_group = TRUE) |> 
+  slice_tail(n = 2000) |> 
+  slice_head(n = 1000) |> 
+  ungroup()
+
[…]

### 2025-04-24 analysis_scripts/08_ppd_fit.R
defining users earlier

@@ -55,6 +55,14 @@ small_data <- lichess_data |>
   filter(grepl("rated bullet game", Event)) |>
   distinct() #remove the duplicate rows if they exist
 
+## when players play less than 10 games
+## otherwise not needed
+users <- mid_games |>
+  group_by(Username) |>
+  tally() |>
+  filter(n >= 10) |>
+  pull(Username)
+
 #selecting previous 1000 to the final 1000 games for each player
 mid_games <- small_data |> 
   filter(Username %in% users) |> 
@@ -63,14 +71,6 @@ mid_games <- small_data |>
   slice_tail(n = 2000) |> 
   slice_head(n = 1000) |> 
   ungroup()
-
-## when players play less than 10 games
-## otherwise not needed
-users <- mid_games |>
-  group_by(Username) |>
-  tally() |>
-  filter(n >= 10) |>
-  pull(Username)
 
 tidy_games <- map_dfr(users, get_hist, mid_games, prev_n = n) |> 
   as_tibble() 

### 2025-04-24 analysis_scripts/08_ppd_fit.R
fixing users again

@@ -57,7 +57,7 @@ small_data <- lichess_data |>
 
 ## when players play less than 10 games
 ## otherwise not needed
-users <- mid_games |>
+users <- small_data |>
   group_by(Username) |>
   tally() |>
   filter(n >= 10) |>

### 2025-04-24 analysis_scripts/08_ppd_fit.R
fixing save path

@@ -118,5 +118,4 @@ fit <- mod$sample(data = stan_data_ave,
                   refresh = 100)
 
 #save fit
-fit$save_object(file = here(save_path,
-                            paste0("mid_fit.RDS")))
+fit$save_object(file = paste0(save_path, "mid_fit.RDS"))

### 2025-04-25 Paper_WriteUp/Paper_Plots.qmd
update fig3 height
fixing model path and using better way to remove duplicate rows

@@
        filename = paste0(here("Paper_WriteUp", "paper_figures", "fig3.png")), dpi = 600,
-       height = 5, width = 7)
+       height = 3.5, width = 7)
 ```
@@
 #read in the model and users
-model = readRDS(file = here(model_path, "all_rated_bullet_model_prev_n1.RDS"))
+model = readRDS(file = here(model_path, "mid_fit.RDS"))
 users = readRDS(file = here(model_path, "users_bullet.RDS"))
@@
   filter(Event == "rated bullet game", Variant == "Standard", TimeControl == "60+0") %>%
-  dplyr::select(c("White", "Black", "Result", "UTCDate", "UTCTime", "Username", "WhiteElo", "BlackElo"))
-
-#remove the duplicate rows
-bullet_60_no_dup = bullet_60[!duplicated(bullet_60),]
+  dplyr::select(c("White", "Black", "Result", "UTCDate", "UTCTime", "Username", "WhiteElo", "BlackElo")) %>%
+  distinct() #remove the duplicate rows if they exist
 
 #final 1000 games each user plays
-last_games = bullet_60_no_dup %>%
+last_games = bullet_60 %>%
   filter(Username %in% users) %>%  
@@
 #true glicko rating of all games before last 1000
-true_glicko_first = bullet_60_no_dup |> 
+true_glicko_first = bullet_60 |> 
   filter(Username == users[focal_player_used]) |> 

### 2025-04-25 analysis_scripts/01_Large_Fit.R
defining the time control term rather than running seperate code for bullet and blitz

@@ -34,6 +34,9 @@ path_id <- as.numeric(Sys.getenv("SLURM_ARRAY_TASK_ID"))
 
 n <- 1 ## number of games to use for history
 
+#what time control to fit model on
+time_control <- "bullet" #takes on "bullet" or "blitz" values only
+
 ### load in the data to use
 
 all_data_path <- rep(NA, 4)
@@ -42,11 +45,8 @@ all_data_path[1] <- here("box_data", "lichess1700-1900", "/")
 all_data_path[2] <- here("box_data", "lichess2000-2200", "/")
 all_data_path[3] <- here("box_data", "lichess2300-2500", "/")
 all_data_path[4] <- here("box_data", "lichessGrandmasters", "/")
+
 all_save_path[1] <- here("results_revision/lichess1700-1900/")
-
-# all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
-# ## if need to run it locally
-
 all_save_path[2] <- here("results_revision/lichess2000-2200/")
 all_save_path[3] <- here("results_revision/lichess2300-2500/")
 all_save_path[4] <- here("results_revision/lichessGrandmasters/")
@@ -62,26 +62,13 @@ print(files)
 lichess_data <- files |> 
   map_dfr(~read_player(data_path, .x))
 
-## restrict to rated rapid and shorter here
-## this also removes the NAs, which makes sense
 
 small_data <- lichess_data |>
   mutate(Event = tolower(Event)) |>
-  # filter(Event == "Rated Bullet game") |>
-  filter(TimeControl == "60+0") |>
-  filter(Variant == "Standard") |>
-  filter(grepl("rated bullet game", Event)) |>
+  filter(Event == paste("rated ", time_control, " game"),
+         Variant == "Standard") |>
+  filter(TimeControl %in% c("60+0", "180+0")) |> 
   distinct() #remove the duplicate rows if they exist
-
-
-# ## what time length should be
-# small_data <- lichess_data |>
-#   mutate(Event = tolower(Event)) |>
-#   filter(TimeControl == "180+0") |>
-#   filter(Variant == "Standard") |>
-#   filter(grepl("rated blitz game", Event))
-
-users <- unique(small_data$Username)
 
 ## when players play less than 10 games
 ## otherwise not needed
@@ -91,8 +78,7 @@ users <- small_data |>
   filter(n >= 10) |>
   pull(Username)
 
[…]

### 2025-04-25 analysis_scripts/01_Large_Fit.R
forgot to add paste0()

@@ -65,7 +65,7 @@ lichess_data <- files |>
 
 small_data <- lichess_data |>
   mutate(Event = tolower(Event)) |>
-  filter(Event == paste("rated ", time_control, " game"),
+  filter(Event == paste0("rated ", time_control, " game"),
          Variant == "Standard") |>
   filter(TimeControl %in% c("60+0", "180+0")) |> 
   distinct() #remove the duplicate rows if they exist

### 2025-04-25 Paper_WriteUp/Paper_Plots.qmd
using the rating in current simulation in ppd. Also using newer history calculation

@@
 
-#adding performance history
-init_data = tidy_games |>
+init_data <- tidy_games |>
   mutate(WhiteElo = as.numeric(WhiteElo),
@@
   group_by(focal_id) |>
-  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), 
-         ave_prop = ifelse(time_diff <= 300,  #if games played in diff sessions, make history same as draw (ie NULL)
-                           lag(focal_win_prop, default = 0.5), #removing standardization
-                           0.5)) |>
-  ungroup() |>
-  filter(focal_result != 0.5)
+  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
+         cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
+         ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  
+                           cum_win_prob, #if games played in different session, history is their mean win prob up to the current game
+                           lag(focal_win_prop))) |> #if game played in same session, rolling mean over the past n games, removing standardization 
+  filter(focal_result != 0.5) %>%
+  ungroup()
 
@@
                           elo_diff = focal_init_data$elo_diff,
+                          focal_result = focal_init_data$focal_result,
                           hist = focal_init_data$ave_prop,
@@
 
-  #elo diff for rating effect
-  rating_diff = curr_game_info$elo_diff
-
   #opponent name and rating
@@
   for (j in 1:num_sims) {
-    curr_draw = focal_draws %>% filter(draw == j)
+    curr_draw = focal_draws %>% filter(draw == j) #current draw of parameters
 
@@
   }
+
+  #rating diff in current game in jth simulation
+  rating_diff = curr_rating - opp_rating
 

### 2025-04-25 Paper_WriteUp/Paper_Plots.qmd
fixing fig7 height width

@@
 #choose who to plot
-focal_player_used = magnus_id
+focal_player_used = nihal_id
 
@@
 #        filename = here("Paper_WriteUp", "paper_figures", "fig7a.png"), dpi = 1000,
-#        height = 4.5, width = 10)
+#        height = 4, width = 6)
 # #save nihal
-# ggsave(plot = ppd_rating_plot, 
-#        filename = here("Paper_WriteUp", "paper_figures", "fig7b.png"), dpi = 1000,
-#        height = 4.5, width = 10)
+ggsave(plot = ppd_rating_plot,
+       filename = here("Paper_WriteUp", "paper_figures", "fig7b.png"), dpi = 1000,
+       height = 4, width = 6)
 ```

### 2025-04-26 Paper_WriteUp/Paper_Plots.qmd
switchin player to plot for ppd

--- 2025-04-25 Paper_WriteUp/Paper_Plots.qmd
@@
 # #save magnus
-# ggsave(plot = ppd_rating_plot, 
-#        filename = here("Paper_WriteUp", "paper_figures", "fig7a.png"), dpi = 1000,
-#        height = 4.5, width = 10)
+ggsave(plot = ppd_rating_plot,
+       filename = here("Paper_WriteUp", "paper_figures", "fig7a.png"), dpi = 1000,
+       height = 4, width = 6)
 # #save nihal
@@
 #        filename = here("Paper_WriteUp", "paper_figures", "fig7b.png"), dpi = 1000,
-#        height = 4.5, width = 10)
+#        height = 4, width = 6)
 ```

### 2025-04-28 Paper_WriteUp/Paper_Plots.qmd
updating and cleaning fig5,6 chunks

@@
 ```{r figure 5 left - compare global params n-1 vs n-10}
+model_path = here("results_revision", "lichessGrandmasters")
+
+#GM n = 1 model
+fit_gm_bullet_prev = readRDS(file = here(model_path, "all_rated_bullet_model_n1.RDS"))
+
+#GM n = 10 model
+fit_gm_bullet_10 = readRDS(file = here(model_path, "all_rated_bullet_model_n10.RDS"))
+
 ## global effects first
-data_path <- here("results_2025/lichessGrandmasters")
-fit_gm_bullet_10 <- 
-  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n10.RDS"))
-
-prev_pars <- fit_gm_bullet_prev$draws(c("mu_beta",
-                                        "gamma1", "gamma2")) |> 
+prev_pars <- fit_gm_bullet_prev$draws(c("mu_beta", "gamma1", "gamma2")) |> 
   as_draws_df() |> 
@@
 
-pre_10_pars <- fit_gm_bullet_10$draws(c("mu_beta",
-                                        "gamma1", "gamma2")) |> 
+prev_10_pars <- fit_gm_bullet_10$draws(c("mu_beta", "gamma1", "gamma2")) |> 
   as_draws_df() |> 
@@
 
-hist_comp_global <- bind_rows(prev_pars, pre_10_pars) |> 
+#plot
+hist_comp_global <- bind_rows(prev_pars, prev_10_pars) |> 
   ggplot(aes(x = value, color = model)) +
@@
 
+#save
 ggsave(plot = hist_comp_global, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "gm_comp_global.png")), dpi = 600,
+       filename = paste0(here("Paper_WriteUp", "paper_figures", "fig5_left.png")), 
+       dpi = 600,
        height = 3.5, width = 7)
@@
 ```{r figure 5 right - bullet v blitz gm}
-data_path <- here("results_2025/lichessGrandmasters")
-fit_gm_bullet_prev <- readRDS(file = paste0(
-  data_path,
-  "/all_rated_bullet_model_prev_n1.RDS"
-))
-bullet_gm <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
-fit_gm_blitz_prev <- readRDS(file = paste0(
-  data_path,
-  "/all_rated_blitz_model_prev_n1.RDS"
-))
-blitz_gm <- readRDS(file = paste0(data_path, "/users_blitz.RDS"))
+model_path = here("results_revision", "lichessGrandmasters")
+
+#GM bullet model
+fit_gm_bullet_prev = readRDS(file = here(model_path, "all_rated_bullet_model_n1.RDS"))
+
+#GM blitz model
+fit_gm_blitz_prev = readRDS(file = here(model_path, "all_rated_blitz_model_n1.RDS"))
+
+#users in each
+bullet_gm = readRDS(file = here(model_path, "users_bullet.RDS"))
+blitz_gm = readRDS(file = here(model_path, "users_blitz.RDS"))
+
 ## compare global effects first
-bullet_pars <- fit_gm_bullet_prev$draws(c(
-  "mu_beta",
-  "gamma1", "gamma2"
-)) |>
+bullet_pars <- fit_gm_bullet_prev$draws(c("mu_beta", "gamma1", "gamma2")) |>
   as_draws_df() |>
@@
 
-blitz_pars <- fit_gm_blitz_prev$draws(c(
-  "mu_beta",
-  "gamma1", "gamma2"
-)) |>
+blitz_pars <- fit_gm_blitz_prev$draws(c("mu_beta", "gamma1", "gamma2")) |>
   as_draws_df() |>
@@
 
+#plot
 bul_blit_gm_comp_global <- bind_rows(bullet_pars, blitz_pars) |>
@@
   theme(legend.position = "bottom") +
-  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  facetted_pos_scales(x = list( #for customizing scales on each facet
+    name == "mu_beta" ~ scale_x_continuous(breaks = c(0, 0.5), limits = c(0, 0.6)),
+    name == "gamma1" ~ scale_x_continuous(breaks = breaks_pretty(n = 2)),
+    name == "gamma2" ~ scale_x_continuous(breaks = breaks_pretty(n = 2), limits = c(0.005, 0.0057))
+  )) +
+  #scale_x_continuous(breaks = breaks_pretty(n = 2)) +
   scale_color_viridis_d(option = "H")
 
-bul_blit_gm_comp_global
-
-ggsave(
-  plot = bul_blit_gm_comp_global,
-  filename = paste0(here(
-    "Paper_WriteUp", "paper_figures",
-    "bul_blitz_comp_gm_global.png"
-  )), dpi = 600,
-  height = 3.5, width = 7
-)
+#save
+ggsave(plot = bul_blit_gm_comp_global,
+       filename = paste0(here("Paper_WriteUp", "paper_figures", "fig5_right.png")), 
+       dpi = 600,
+       height = 3.5, width = 7)
 ```
@@
 
+#bullet parameter draws
 player_pars_bullet <- fit_gm_bullet_prev$draws() |> 
@@
 
+#blitz parameter draws
 player_pars_blitz <- fit_gm_blitz_prev$draws() |> 
@@
 
+#plot
 comp_gm_winner <- player_pars_bullet |> 
@@
   scale_x_continuous(breaks = breaks_pretty(n = 3)) +
-  scale_color_viridis_d(option = "H")
-
-comp_gm_winner
+  scale_color_viridis_d(option = "H") + 
+  geom_vline(aes(xintercept = 0), col = "black", linetype = "dashed")
+
+#save
 ggsave(plot = comp_gm_winner, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "bul_blitz_comp_gm_winner.png")), dpi = 600,
-       height = 5, width = 7)
+       filename = paste0(here("Paper_WriteUp", "paper_figures", "fig6.png")), 
+       dpi = 600,
+       height = 7, width = 7)
 ```

### 2025-04-28 Paper_WriteUp/Paper_Plots.qmd
cleaning fig4 chunk, getting rid of stuff we don't use

@@
 
-```{r player level parameters for 17-19}
+```{r figure 4 left}
 #| eval: false
@@
 
-## see what the smallest and largest posterior mean is 
-## for the winner effects
-player_pars |> 
-  group_by(player_name, param) |> 
-  summarise(post_mean = mean(value)) |> 
-  filter(param == 2) |> 
-  ungroup() |> 
-  slice_max(order_by = post_mean, n = 5)
-
-player_pars |> 
-  group_by(player_name, param) |> 
-  summarise(post_mean = mean(value)) |> 
-  filter(param == 2) |> 
-  ungroup() |> 
-  slice_min(order_by = post_mean, n = 5)
-
-## see how many contain 0
-player_pars |> 
-  group_by(player_name, param) |> 
-  filter(param == 2) |> 
-  summarise(lower = quantile(value, prob = 0.025),
-            upper = quantile(value, prob = 0.975)) |> 
-  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
-                         1, 0)) |> 
-  filter(contain_zero == 0)
-```
-
-```{r figure 4 left}
+#plot
 winner_plot_17_19 <- player_pars |> 

### 2025-04-28 Paper_WriteUp/Paper_Plots.qmd
updating and cleaning figS1 chunks

@@
 #| eval: false
-data_path <- here("results_2025/lichess2000-2200")
-
-fit_20_22_bullet_prev <- readRDS(file = 
-                                   paste0(data_path,
-                                          "/all_rated_bullet_model_prev_n1.RDS"))
-
-bullet_20_22 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
+#read in model
+fit_20_22_bullet_prev = readRDS(here("results_revision", "lichess2000-2200", "all_rated_bullet_model_n1.RDS"))
+
+#read in users
+bullet_20_22 = readRDS(here("results_revision", "lichess2000-2200", "users_bullet.RDS"))
 
@@
 
+#draws of winner effects
 player_pars <- fit_20_22_bullet_prev$draws() |> 
@@
 
+#plot
 winner_plot_20_22 <- player_pars |> 
@@
 
+#save
 ggsave(plot = winner_plot_20_22, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "Figure_S1_left.png")), dpi = 600,
+       filename = here("Paper_WriteUp", "paper_figures", "FigS1_left.png"), dpi = 600,
        height = 7, width = 7)
@@
 #| eval: false
-data_path <- here("results_2025/lichess2300-2500")
-
-fit_23_25_bullet_prev <- readRDS(file = 
-                                   paste0(data_path,
-                                          "/all_rated_bullet_model_prev_n1.RDS"))
-
-bullet_23_25 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
+#read in model
+fit_20_22_bullet_prev = readRDS(here("results_revision", "lichess2300-2500", "all_rated_bullet_model_n1.RDS"))
+
+#read in users
+bullet_20_22 = readRDS(here("results_revision", "lichess2300-2500", "users_bullet.RDS"))
 
@@
 
+#plot
 winner_plot_23_25 <- player_pars |> 
@@
 
+#save
 ggsave(plot = winner_plot_23_25, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "Figure_S1_right.png")), dpi = 600,
+       filename = here("Paper_WriteUp", "paper_figures", "FigS1_right.png"), dpi = 600,
        height = 7, width = 7)

### 2025-04-29 Paper_WriteUp/Paper_Plots.qmd
add chunk to plot one response to reviewer

@@
 library(ggh4x)
+library(latex2exp)
 options(mc.cores = parallel::detectCores())
@@
 ```
+
+#distribution of x_{ij} - for response to reviewer comment
+```{r}
+#import GM data
+data_path = paste0(all_data_path[1], "/")
+
+files <- list.files(data_path)
+lichess_data <- files |> 
+  map_dfr(~read_player(data_path, .x))
+
+small_data <- lichess_data |>
+  mutate(Event = tolower(Event)) |>
+  filter(Event == paste0("rated ", time_control, " game"),
+         Variant == "Standard") |>
+  filter(TimeControl %in% c("60+0", "180+0")) |> 
+  distinct() #remove the duplicate rows if they exist
+
+## otherwise not needed
+users <- small_data |>
+  group_by(Username) |>
+  tally() |>
+  filter(n >= 10) |>
+  pull(Username)
+
+tidy_games <- map_dfr(users, get_hist, small_data, prev_n = n) |> 
+  as_tibble()
+
+init_data <- tidy_games |>
+  mutate(WhiteElo = as.numeric(WhiteElo),
+         BlackElo = as.numeric(BlackElo),
+         focal_user = ifelse(focal_white == 1, White, Black),
+         elo_diff = ifelse(focal_white == 1,
+                           WhiteElo - BlackElo, BlackElo - WhiteElo),
+         focal_id = match(focal_user, users),
+         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
+  dplyr::select(focal_user, focal_id, focal_white, 
+                focal_win_prop, elo_diff, focal_result,
+                UTCDateTime) |>
+  group_by(focal_id) |>
+  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
+         cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
+         ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  #current way we evaluate history
+                           0, 
+                           lag(focal_win_prop) - cum_win_prob),
+         ave_prop2 = ifelse(time_diff > 300 | is.na(time_diff), #proposed way we evaluate history
+                            0.5, 
+                            lag(focal_win_prop))) |> 
+  filter(focal_result != 0.5) %>%
+  ungroup()
+
+#just filter for one player for illustration
+init_data_plot = init_data %>% filter(focal_id == 1)
+
+#distribution of x_ij for current way we evaluate history
+curr_dist = ggplot(data = init_data_plot, mapping = aes(x = ave_prop)) +
+  geom_histogram() + 
+  labs(title = "Normalization", x = TeX("$\\tilde{x}_{ij}^n$"), y = "Count") + 
+  scale_x_continuous(breaks = c(-0.5, -0.25, 0, 0.25, 0.5), labels = c(-0.5, -0.25, 0, 0.25, 0.5)) + 
+  theme_single_y()
+
+#save
+ggsave(plot = curr_dist,
+       filename = here("Paper_WriteUp", "paper_figures", "response to comments", "curr_xij.png"), dpi = 1000,
+       height = 4, width = 6)
+
+#distribution of x_ij for proposed way to evaluate history
+prop_dist = ggplot(data = init_data_plot, mapping = aes(x = ave_prop2)) +
+  geom_histogram() + 
+  labs(title = "No Normalization", x = TeX("$\\tilde{x}_{ij}^n$"), y = "Count") + 
+  scale_x_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1), limits = c(-0.2, 1.2)) + 
+  theme_single_y()
+
+#save
+ggsave(plot = prop_dist,
+       filename = here("Paper_WriteUp", "paper_figures", "response to comments", "prop_xij.png"), dpi = 1000,
+       height = 4, width = 6)
+```

### 2025-04-29 analysis_scripts/01_Large_Fit.R
adding normalization term back

@@ -10,7 +10,6 @@
 
 ##### April 15, 2025 ######
 #' Adam revision 
-#' removing the standardization of x_ij term
 #' also assuming games played in different sessions have no relation, so make their history NULL (ie same as their historical avg past performance)
 #' removing all the plots and stuff after we save the fit, we don't use that anymore
 
@@ -98,8 +97,8 @@ init_data <- tidy_games |>
   mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
          cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
          ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  
-                           cum_win_prob, #if games played in different session, history is their mean win prob up to the current game
-                           lag(focal_win_prop))) |> #if game played in same session, rolling mean over the past n games, removing standardization 
+                           0, #if games played in different session, history is their mean win prob up to the current game
+                           lag(focal_win_prop) - cum_win_prob)) |> #if game played in same session, rolling mean over the past n games
   filter(focal_result != 0.5) %>%
   ungroup()
 

### 2025-04-29 Paper_WriteUp/Paper_Plots.qmd
updating glicko2() calculation to use RD and volatility from previous game. Not constant anymore

@@
 #storage
-glicko_sim = matrix(NA, nrow = num_games, ncol = num_sims)
-game_sim = matrix(NA, nrow = num_games, ncol = num_sims)
+glicko_sim = matrix(NA, nrow = num_games, ncol = num_sims) #storing the ratings
+rd_sim = matrix(NA, nrow = num_games, ncol = num_sims) #storing the RDs
+volatility_sim = matrix(NA, nrow = num_games, ncol = num_sims) #storing the volatilities
+game_sim = matrix(NA, nrow = num_games, ncol = num_sims) #storing the game results
 #probs = matrix(NA, nrow = num_games, ncol = num_sims) #for debugging
+
+### SET FIRST RD, VOLATILITY TO TRUE RD/VOLATILITY WE GET AFTER FINDING RDS/VOLATILITY IN RAW DATASET (IN INIT CALCULATION)
 
@@
     curr_rating = curr_game_info$focal_rating
-    curr_hist = 0.5 #assume they're current past performance is at their usual level
+    curr_rd = 50 #starting value for rating deviation
+    curr_volatility = 0.04 #starting value for volatility
+    curr_hist = 0 #assume they're current past performance is NULL
   } else {
+    #use rating info from past game
+    curr_rating = glicko_sim[i - 1, j]
+    curr_rd = rd_sim[i - 1, j]
+    curr_volatility = volatility_sim[i - 1, j]
     curr_hist = curr_game_info$hist
-    curr_rating = glicko_sim[i - 1, j]
   }
@@
                      Rating = c(curr_rating, opp_rating),
-                     Deviation = 50,
-                     Volatility = 0.04)
+                     Deviation = curr_rd,
+                     Volatility = curr_volatility)
 
@@
 
-  #store current result and rating
+  #store current result and rating info
   glicko_sim[i, j] = as.numeric(focal_glick)[1]
+  rd_sim[i, j] = as.numeric(focal_glick)[2]
+  volatility_sim[i, j] = as.numeric(focal_glick)[3]
   game_sim[i, j] = sim_result

### 2025-04-29 utils/helper.R
adding get_RDs function to helper

@@ -62,6 +62,67 @@ get_hist <- function(user, games, prev_n) {
   hist_games
 }
 
+
+#RDs/Volatility calculation
+
+#returns dataframe with RDs and volatilities for some user
+#meant to be run on init_data only
+get_RDs = function(user, games) { 
+  focal_games = games %>% filter(focal_user == user) 
+  
+  #storage
+  RDs = numeric(length = nrow(focal_games))
+  vols = numeric(length = nrow(focal_games))
+  
+  #loop through each game to use true ratings and get estimated RD/volatility
+  for (i in 1: nrow(focal_games)) {
+    curr_row = focal_games[i,]
+    
+    focal_name = curr_row$focal_user #focal players name
+    focal_rating = curr_row$focal_rating #focal players rating
+    opp_name = curr_row$opp #opponent name
+    opp_rating = curr_row$opp_rating #opponent rating
+    
+    result = curr_row$focal_result
+    colour_ind = ifelse(curr_row$focal_white == 1, 1, -1)
+    
+    #these are for the focal player
+    if (i == 1) { #guess parameters
+      curr_RD = 100
+      curr_volatility = 0.06
+    } else { #use the previously estimated parameters
+      curr_RD = RDs[i - 1]
+      curr_volatility = vols[i - 1]
+    }
+    
+    games = data.frame("Time" = i,
+                       "Player1" = focal_name,
+                       "Player2" = opp,
+                       "Result" = result)
+    initstate = data.frame("Player" = c(focal_name, opp_name),
+                           "Rating" = c(focal_rating, opp_rating),
+                           "Deviation" = c(curr_RD, 200), #doesnt matter what opponents RD/volatility is
+                           "Volatility" = c(curr_volatility, 0.05))
+    
+    #get updated rating, RDs, volatility
+    glicko2 = glicko2(games, status = initstate, history = TRUE, gamma = colour_ind)
+    
+    #extract the rating of focal player and store it
+    focal_glick = glicko2$history[focal_name, ,]
+    
+    #store them
+    RDs[i] = as.numeric(focal_glick)[2]
+    vols[i] = as.numeric(focal_glick)[3]
+  }
+  #put results back into dataframe
+  focal_games = focal_games %>% 
+    mutate(RD = RDs,
+           volatility = vols)
+}
+
+
+
+
 ## extract the sessions containing more than 1 game
 get_true_sequences <- function(x) {
   rle_x <- rle(x)

### 2025-04-29 Paper_WriteUp/Paper_Plots.qmd
adding chunk to plot RDs and volatilities for reviewers

@@
 library(latex2exp)
+library(furrr)
 options(mc.cores = parallel::detectCores())
@@
          UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
-  dplyr::select(focal_user, focal_id, focal_white, WhiteElo, BlackElo,
+  dplyr::select(focal_user, focal_id, focal_white, WhiteElo, BlackElo, White, Black,
                 focal_win_prop, elo_diff, focal_result, opp, opp_rating,
@@
 #probs = matrix(NA, nrow = num_games, ncol = num_sims) #for debugging
-
-### SET FIRST RD, VOLATILITY TO TRUE RD/VOLATILITY WE GET AFTER FINDING RDS/VOLATILITY IN RAW DATASET (IN INIT CALCULATION)
 
@@
 ```{r}
-#import GM data
+#import data
 data_path = paste0(all_data_path[1], "/")
@@
 ```
+
+#plotting RDs and volatilities for reviewer response
+```{r}
+#import data
+data_path = paste0(all_data_path[1], "/")
+
+files <- list.files(data_path)
+lichess_data <- files |> 
+  map_dfr(~read_player(data_path, .x))
+
+small_data <- lichess_data |>
+  mutate(Event = tolower(Event)) |>
+  filter(Event == paste0("rated ", time_control, " game"),
+         Variant == "Standard") |>
+  filter(TimeControl %in% c("60+0", "180+0")) |> 
+  distinct() #remove the duplicate rows if they exist
+
+## otherwise not needed
+users <- small_data |>
+  group_by(Username) |>
+  tally() |>
+  filter(n >= 10) |>
+  pull(Username)
+
+tidy_games <- map_dfr(users, get_hist, small_data, prev_n = n) |> 
+  as_tibble()
+
+start = Sys.time()
+init_data <- tidy_games |>
+  mutate(WhiteElo = as.numeric(WhiteElo),
+         BlackElo = as.numeric(BlackElo),
+         focal_user = ifelse(focal_white == 1, White, Black),
+         elo_diff = ifelse(focal_white == 1,
+                           WhiteElo - BlackElo, BlackElo - WhiteElo),
+         focal_id = match(focal_user, users),
+         focal_rating = ifelse(focal_white == 1, WhiteElo, BlackElo),
+         opp = ifelse(focal_white == 1, Black, White),
+         opp_rating = ifelse(focal_white == 1, BlackElo, WhiteElo),
+         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
+  dplyr::select(focal_user, focal_id, focal_white, WhiteElo, BlackElo, White, Black,
+                focal_win_prop, elo_diff, focal_result, opp, opp_rating,
+                focal_rating, UTCDateTime) |>
+  group_by(focal_id) |>
+  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
+         cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
+         ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  
+                           cum_win_prob, #if games played in different session, history is their mean win prob up to the current game
+                           lag(focal_win_prop))) |> #if game played in same session, rolling mean over the past n games, removing standardization 
+  filter(focal_result != 0.5) %>%
+  ungroup()
+
+#get RDs
+init_RDs = future_map_dfr(users, get_RDs, init_data) |>
+  as_tibble()
+
+end = Sys.time()
+end - start #takes like 20 minutes
+
+#RDs
+RD_plot = ggplot(data = init_RDs %>% group_by(focal_user) %>% mutate(index = row_number()), mapping = aes(x = index, y = RD)) +
+  geom_line() +
+  scale_x_continuous(n.breaks = 3) +
+  facet_wrap(~focal_user)
+
+#save
+ggsave(plot = RD_plot,
+       filename = here("Paper_WriteUp", "paper_figures", "response to comments", "RDs.png"), dpi = 1000,
+       height = 6, width = 8)
+
+#volatilities
+vol_plot = ggplot(data = init_RDs %>% group_by(focal_user) %>% mutate(index = row_number()), mapping = aes(x = index, y = volatility)) +
+  geom_line() +
+  scale_x_continuous(n.breaks = 3) +
+  facet_wrap(~focal_user)
+
+#save
+ggsave(plot = vol_plot,
+       filename = here("Paper_WriteUp", "paper_figures", "response to comments", "volatilities.png"), dpi = 1000,
+       height = 6, width = 8)
+```

### 2025-04-30 analysis_scripts/01_Large_Fit.R
small modifications, using all n(0,1) priors and modifying to correctly account for average win proportion over all previous games only

@@ -33,14 +33,10 @@ all_data_path[1] <- here("box_data/lichess1700-1900/")
 all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_data_path[4] <- here("box_data/lichessGrandmasters/")
-all_save_path[1] <- here("results_revision/lichess1700-1900/")
-
-# all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
-# ## if need to run it locally
-
-all_save_path[2] <- here("results_revision/lichess2000-2200/")
-all_save_path[3] <- here("results_revision/lichess2300-2500/")
-all_save_path[4] <- here("results_revision/lichessGrandmasters/")
+all_save_path[1] <- here("results_revision_may/lichess1700-1900/")
+all_save_path[2] <- here("results_revision_may/lichess2000-2200/")
+all_save_path[3] <- here("results_revision_may/lichess2300-2500/")
+all_save_path[4] <- here("results_revision_may/lichessGrandmasters/")
 
 
 data_path <- all_data_path[path_id]
@@ -99,7 +95,8 @@ init_data <- tidy_games |>
   select(focal_user, focal_id, focal_white,
          focal_win_prop, elo_diff, focal_result) |>
   group_by(focal_id) |>
-  mutate(ave_prop = lag(focal_win_prop, default = 0)) |> # - mean(focal_result)) |>
+  mutate(ave_prop = lag(focal_win_prop, n = 1, default = 0) - 
+           lag(cummean(focal_result), default = 0))
   filter(focal_result != 0.5)
 
 cat("----------\n")

### 2025-04-30 analysis_scripts/02_Large_Single_Prev.R

@@ -32,10 +32,10 @@ all_data_path[1] <- here("box_data/lichess1700-1900/")
 all_data_path[2] <- here("box_data/lichess2000-2200/")
 all_data_path[3] <- here("box_data/lichess2300-2500/")
 all_data_path[4] <- here("box_data/lichessGrandmasters/")
-all_save_path[1] <- here("results_revision/lichess1700-1900/")
-all_save_path[2] <- here("results_revision/lichess2000-2200/")
-all_save_path[3] <- here("results_revision/lichess2300-2500/")
-all_save_path[4] <- here("results_revision/lichessGrandmasters/")
+all_save_path[1] <- here("results_revision_may/lichess1700-1900/")
+all_save_path[2] <- here("results_revision_may/lichess2000-2200/")
+all_save_path[3] <- here("results_revision_may/lichess2300-2500/")
+all_save_path[4] <- here("results_revision_may/lichessGrandmasters/")
 
 
 data_path <- all_data_path[path_id]
@@ -97,9 +97,10 @@ hist_data_init <- tidy_games |>
          focal_win_prop, elo_diff, focal_result) |> 
   group_by(focal_id) |> 
   mutate(ave_prop = lag(focal_win_prop, n = 1, default = 0) - 
-           mean(focal_result), 
+           lag(cummean(focal_result), default = 0), 
          ## focal_win_prop is the average of the running win proportion
-         prev_game = lag(focal_result, default = 0)) |> # - mean(focal_result)) |> 
+         prev_game = lag(focal_result, default = 0) - 
+           lag(cummean(focal_result), default = 0))  |> 
   filter(focal_result != 0.5)
 
 
@@ -109,6 +110,8 @@ cat("----------\n")
 
 ### then fit the models
 
+## check only
+hist_data_init <- hist_data_init[1:5000, ]
 
 stan_data_ave <- list(N = nrow(hist_data_init),
                       J = length(users),
@@ -128,7 +131,7 @@ fit3_ave <- mod$sample(data = stan_data_ave,
                        seed = 123,
                        chains = 4,
                        parallel_chains = 4,
-                       refresh = 100)
+                       refresh = 100,iter_warmup = 100, iter_sampling = 100)
 
 ## save the stan fit as not actually that large here
 

### 2025-04-30 Paper_WriteUp/Paper_Plots.qmd
adding standardization to RDs chunk

@@
          ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  
-                           cum_win_prob, #if games played in different session, history is their mean win prob up to the current game
-                           lag(focal_win_prop))) |> #if game played in same session, rolling mean over the past n games, removing standardization 
+                           0, #if games played in different session, history is their mean win prob up to the current game
+                           lag(focal_win_prop) - cum_win_prob)) |> #if game played in same session, rolling mean over the past n games, removing standardization 
   filter(focal_result != 0.5) %>%

### 2025-04-30 analysis_scripts/06_sim_data_gen.R
adding standardization to simulation

@@ -58,13 +58,10 @@ betas = mu_beta + rnorm(n = num_players, mean = 0, sd = 0.1) #mu_beta plus some 
 
 
 #' now simulate the games for this player 
-#' 
-#' 
-#' RIGHT NOW THIS IS USING NO STANDARISATION!!! JUST USES 0, 0.5, 1 AS HISTORY FOR BETA
 
 #for storage
 sim_games = data.frame("player_id" = as.numeric(), "result" = as.numeric(), "colour" = as.numeric(), 
-                       "rating_diff" = as.numeric(), "last_result" = as.numeric(), "beta" = as.numeric(), 
+                       "rating_diff" = as.numeric(), "last_result" = as.numeric(), "ave_prop" = numeric(), "beta" = as.numeric(), 
                        "mu_beta" = as.numeric(), "alpha" = as.numeric(), "gamma1" = as.numeric(), "gamma2" = as.numeric())
 for(i in 1:num_games) {
   #current covariate (game) info
@@ -74,13 +71,16 @@ for(i in 1:num_games) {
   curr_colour = sample(c(0, 1), prob = c(0.5, 0.5), size = 1) #1 is white, 0 is black
   last_result = ifelse(i == 1, 0.5, sim_games$result[i - 1]) #assume first game history is same as draw
   
+  cum_win_prob = ifelse(i == 1, 0.5, mean(sim_games$last_result)) #the cumulative mean for focal player
+  curr_xij = last_result - cum_win_prob
+  
   #sim result
-  lin_comb = alphas[player_id] + betas[player_id]*last_result + gamma1*curr_colour + gamma2*curr_rating_diff #the predictors
+  lin_comb = alphas[player_id] + betas[player_id]*curr_xij + gamma1*curr_colour + gamma2*curr_rating_diff #the predictors
   curr_win_prob = invlogit(lin_comb) #simulate probability
   curr_result = sample(c(1, 0), prob = c(curr_win_prob, 1 - curr_win_prob), size = 1)
   
   #store everything - player number, result, colour, rating diff, last result, parameter values
-  sim_games[i,] = c(player_id, curr_result, curr_colour, curr_rating_diff, last_result, 
+  sim_games[i,] = c(player_id, curr_result, curr_colour, curr_rating_diff, last_result, curr_xij,
                     betas[player_id], mu_beta, alphas[player_id], gamma1, gamma2) #the row for this data set
 }
 

### 2025-04-30 analysis_scripts/07_sim_data_fit.R
incorporating new change from sim_data_gen

@@ -44,14 +44,14 @@ init_data = sim_data %>%
   group_by(player_id) %>%
   slice_head(n = curr_num_games) #slice the correct number of games
 
-#right now not standardizing, so history for prev_n = 1 is just 0, 0.5, 1...
+#data for stan
 stan_sim_data = list(N = nrow(init_data),
                      J = init_data$player_id %>% unique() %>% length(),
                      y = init_data$result,
                      id = init_data$player_id,
                      colour = init_data$colour,
                      elo = init_data$rating_diff,
-                     win_prop = init_data$last_result)
+                     win_prop = init_data$ave_prop)
 
 ### FITTING ###
 

### 2025-04-30 analysis_scripts/06_sim_data_gen.R
matching mu_betas used in paper

@@ -31,7 +31,7 @@ num_games = 20000 #this is roughly the same as the number of games per player in
 #' we vary the experiential effect size and number of games per player to see how well our model can detect true experiential effects
 #' experiential effects size of logit(0.51) represents 1% increase win percentage when coming from a win... logit(0.53) represents 3% increase...
 #mu_beta_sizes = c(logit(0.51), logit(0.53), logit(0.6)) #these are the mu_betas
-mu_beta_sizes = c(-0.5, 0, 0.5)
+mu_beta_sizes = c(-0.5, 0.12, 0.5)
 
 #' if we want to fit our model using smaller number of games, just take a subset of these 20,000 games
 #' 

### 2025-04-30 utils/helper.R
fix opponent in get_RDs function

@@ -97,7 +97,7 @@ get_RDs = function(user, games) {
     
     games = data.frame("Time" = i,
                        "Player1" = focal_name,
-                       "Player2" = opp,
+                       "Player2" = opp_name,
                        "Result" = result)
     initstate = data.frame("Player" = c(focal_name, opp_name),
                            "Rating" = c(focal_rating, opp_rating),

### 2025-04-30 utils/helper.R
adding tau = 0.5 and changing starting volatility to 0.05

@@ -89,7 +89,7 @@ get_RDs = function(user, games) {
     #these are for the focal player
     if (i == 1) { #guess parameters
       curr_RD = 100
-      curr_volatility = 0.06
+      curr_volatility = 0.05
     } else { #use the previously estimated parameters
       curr_RD = RDs[i - 1]
       curr_volatility = vols[i - 1]
@@ -105,7 +105,7 @@ get_RDs = function(user, games) {
                            "Volatility" = c(curr_volatility, 0.05))
     
     #get updated rating, RDs, volatility
-    glicko2 = glicko2(games, status = initstate, history = TRUE, gamma = colour_ind)
+    glicko2 = glicko2(games, status = initstate, history = TRUE, gamma = colour_ind, tau = 0.5)
     
     #extract the rating of focal player and store it
     focal_glick = glicko2$history[focal_name, ,]

### 2025-04-30 Paper_WriteUp/Paper_Plots.qmd
implementing RDs in parallel

--- 2025-04-29 Paper_WriteUp/Paper_Plots.qmd
@@
 #import data
-data_path = paste0(all_data_path[1], "/")
+data_path = paste0(all_data_path[4])
 
 files <- list.files(data_path)
+
 lichess_data <- files |> 
-  map_dfr(~read_player(data_path, .x))
+  map_dfr(~read_player(paste0(data_path, "/"), .x))
+users <- readRDS(file = here("results_revision", "lichessGrandmasters", "users_bullet.RDS"))
 
@@
   mutate(Event = tolower(Event)) |>
-  filter(Event == paste0("rated ", time_control, " game"),
+  filter(Event == paste0("rated bullet game"),
          Variant == "Standard") |>
@@
 
-tidy_games <- map_dfr(users, get_hist, small_data, prev_n = n) |> 
+tidy_games <- map_dfr(users, get_hist, small_data, prev_n = 1) |> 
   as_tibble()
 
-start = Sys.time()
 init_data <- tidy_games |>
@@
 #get RDs
+plan(multisession, workers = 12) #in parallel - dramatically speeds it up
+start = Sys.time()
 init_RDs = future_map_dfr(users, get_RDs, init_data) |>
   as_tibble()
-
 end = Sys.time()
-end - start #takes like 20 minutes
+end - start #takes 1.3 min
 
@@
   geom_line() +
-  scale_x_continuous(n.breaks = 3) +
+  scale_x_continuous(n.breaks = 3, limits = c(0, 10000)) +
+  labs(x = "Game Number", y = "Rating Deviation") +
   facet_wrap(~focal_user)

### 2025-04-30 analysis_scripts/01_Large_Fit.R
adding covariate info to init_data for RD calculation

@@ -4,13 +4,11 @@
 ## games on a cluster
 ## stan model is somewhat optimized for speed to fit to faster, using 
 ## matrix multiplication
-##
-
 
 
 ##### April 15, 2025 ######
 #' Adam revision 
-#' also assuming games played in different sessions have no relation, so make their history NULL (ie same as their historical avg past performance)
+#' Assuming games played in different sessions have no relation, so make their history NULL (0)
 #' removing all the plots and stuff after we save the fit, we don't use that anymore
 
 
@@ -88,8 +86,12 @@ init_data <- tidy_games |>
          focal_user = ifelse(focal_white == 1, White, Black),
          elo_diff = ifelse(focal_white == 1,
                            WhiteElo - BlackElo, BlackElo - WhiteElo),
-         focal_id = match(focal_user, users),
-         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
+         focal_id = match(focal_user, users), 
+         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime),
+         #adding info for RDs
+         focal_rating = ifelse(focal_white == 1, WhiteElo, BlackElo),
+         opp = ifelse(focal_white == 1, Black, White),
+         opp_rating = ifelse(focal_white == 1, BlackElo, WhiteElo))) |>
   dplyr::select(focal_user, focal_id, focal_white, 
                 focal_win_prop, elo_diff, focal_result,
                 UTCDateTime) |>

### 2025-04-30 analysis_scripts/01_Large_Fit.R
fixing brackets in init_data

@@ -87,11 +87,10 @@ init_data <- tidy_games |>
          elo_diff = ifelse(focal_white == 1,
                            WhiteElo - BlackElo, BlackElo - WhiteElo),
          focal_id = match(focal_user, users), 
-         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime),
-         #adding info for RDs
+         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime)),
          focal_rating = ifelse(focal_white == 1, WhiteElo, BlackElo),
          opp = ifelse(focal_white == 1, Black, White),
-         opp_rating = ifelse(focal_white == 1, BlackElo, WhiteElo))) |>
+         opp_rating = ifelse(focal_white == 1, BlackElo, WhiteElo)) |>
   dplyr::select(focal_user, focal_id, focal_white, 
                 focal_win_prop, elo_diff, focal_result,
                 UTCDateTime) |>

### 2025-04-30 analysis_scripts/01_Large_Fit.R
selecting the variables needed for RD calculation

@@ -86,20 +86,20 @@ init_data <- tidy_games |>
          focal_user = ifelse(focal_white == 1, White, Black),
          elo_diff = ifelse(focal_white == 1,
                            WhiteElo - BlackElo, BlackElo - WhiteElo),
-         focal_id = match(focal_user, users), 
-         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime)),
+         focal_id = match(focal_user, users),
          focal_rating = ifelse(focal_white == 1, WhiteElo, BlackElo),
          opp = ifelse(focal_white == 1, Black, White),
-         opp_rating = ifelse(focal_white == 1, BlackElo, WhiteElo)) |>
-  dplyr::select(focal_user, focal_id, focal_white, 
-                focal_win_prop, elo_diff, focal_result,
-                UTCDateTime) |>
+         opp_rating = ifelse(focal_white == 1, BlackElo, WhiteElo),
+         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
+  dplyr::select(focal_user, focal_id, focal_white, WhiteElo, BlackElo, White, Black,
+                focal_win_prop, elo_diff, focal_result, opp, opp_rating,
+                focal_rating, UTCDateTime) |>
   group_by(focal_id) |>
   mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
          cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
          ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  
                            0, #if games played in different session, history is their mean win prob up to the current game
-                           lag(focal_win_prop) - cum_win_prob)) |> #if game played in same session, rolling mean over the past n games
+                           lag(focal_win_prop) - cum_win_prob)) |> #if game played in same session, rolling mean over the past n games, removing standardization 
   filter(focal_result != 0.5) %>%
   ungroup()
 

### 2025-04-30 analysis_scripts/01_Large_Fit.R
removing RD information in this file, we don't need it

@@ -86,20 +86,17 @@ init_data <- tidy_games |>
          focal_user = ifelse(focal_white == 1, White, Black),
          elo_diff = ifelse(focal_white == 1,
                            WhiteElo - BlackElo, BlackElo - WhiteElo),
-         focal_id = match(focal_user, users),
-         focal_rating = ifelse(focal_white == 1, WhiteElo, BlackElo),
-         opp = ifelse(focal_white == 1, Black, White),
-         opp_rating = ifelse(focal_white == 1, BlackElo, WhiteElo),
+         focal_id = match(focal_user, users), 
          UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
-  dplyr::select(focal_user, focal_id, focal_white, WhiteElo, BlackElo, White, Black,
-                focal_win_prop, elo_diff, focal_result, opp, opp_rating,
-                focal_rating, UTCDateTime) |>
+  dplyr::select(focal_user, focal_id, focal_white, 
+                focal_win_prop, elo_diff, focal_result,
+                UTCDateTime) |>
   group_by(focal_id) |>
   mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
          cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
          ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  
                            0, #if games played in different session, history is their mean win prob up to the current game
-                           lag(focal_win_prop) - cum_win_prob)) |> #if game played in same session, rolling mean over the past n games, removing standardization 
+                           lag(focal_win_prop) - cum_win_prob)) |> #if game played in same session, rolling mean over the past n games
   filter(focal_result != 0.5) %>%
   ungroup()
 

### 2025-04-30 analysis_scripts/09_Large_Fit_RDs.R
script for fitting model with new rating diff covariate that factors in RDs

@@ -1,0 +1,128 @@
+##### April 30, 2025 ######
+#' Script for fitting model with new rating diff covariate
+#' rating diff now takes into account the RDs
+#' NOTE - we have to estimate the RDs, since they aren't available to us when scraping the data from lichess
+
+library(tidyverse)
+library(RcppRoll)
+library(cmdstanr)
+library(posterior)
+library(bayesplot)
+library(loo)
+library(here)
+
+options(mc.cores = parallel::detectCores())
+
+## source helper functions for reading and transforming data
+## along with defaults for plots, etc
+source(here("utils/helper.R"))
+
+path_id <- as.numeric(Sys.getenv("SLURM_ARRAY_TASK_ID"))
+
+n <- 1 ## number of games to use for history
+
+#what time control to fit model on
+time_control <- "bullet" #takes on "bullet" or "blitz" values only
+
+### load in the data to use
+
+all_data_path <- rep(NA, 4)
+all_save_path <- rep(NA, 4)
+all_data_path[1] <- here("box_data", "lichess1700-1900", "/")
+all_data_path[2] <- here("box_data", "lichess2000-2200", "/")
+all_data_path[3] <- here("box_data", "lichess2300-2500", "/")
+all_data_path[4] <- here("box_data", "lichessGrandmasters", "/")
+
+all_save_path[1] <- here("results_revision/lichess1700-1900/")
+all_save_path[2] <- here("results_revision/lichess2000-2200/")
+all_save_path[3] <- here("results_revision/lichess2300-2500/")
+all_save_path[4] <- here("results_revision/lichessGrandmasters/")
+
+
+data_path <- all_data_path[path_id]
+save_path <- all_save_path[path_id]
+
+dir.create(save_path, showWarnings = FALSE)
+
+files <- list.files(data_path)
+print(files)
+lichess_data <- files |> 
+  map_dfr(~read_player(data_path, .x))
+
+
+small_data <- lichess_data |>
+  mutate(Event = tolower(Event)) |>
+  filter(Event == paste0("rated ", time_control, " game"),
+         Variant == "Standard") |>
+  filter(TimeControl %in% c("60+0", "180+0")) |> 
+  distinct() #remove the duplicate rows if they exist
+
+## when players play less than 10 games
+## otherwise not needed
+users <- small_data |>
+  group_by(Username) |>
+  tally() |>
+  filter(n >= 10) |>
+  pull(Username)
+
+saveRDS(users, file = paste0(save_path, paste0("users_", time_control, ".RDS")))
+
[…]

### 2025-04-30 analysis_scripts/09_Large_Fit_RDs.R
adding furrr library

@@ -10,6 +10,7 @@ library(posterior)
 library(bayesplot)
 library(loo)
 library(here)
+library(furrr)
 
 options(mc.cores = parallel::detectCores())
 

### 2025-04-30 analysis_scripts/09_Large_Fit_RDs.R
adding PlayerRatings library

@@ -11,6 +11,7 @@ library(bayesplot)
 library(loo)
 library(here)
 library(furrr)
+library(PlayerRatings)
 
 options(mc.cores = parallel::detectCores())
 

### 2025-05-01 Paper_WriteUp/Paper_Plots.qmd
better x-axis limits for RD/volatility plots

@@
   geom_line() +
-  scale_x_continuous(n.breaks = 3, limits = c(0, 10000)) +
+  scale_x_continuous(n.breaks = 3, limits = c(0, 5000)) +
+  scale_y_continuous(limits = c(50, 100)) +
   labs(x = "Game Number", y = "Rating Deviation") +
@@
   geom_line() +
-  scale_x_continuous(n.breaks = 3, limits = c(0, 10000)) +
+  scale_x_continuous(n.breaks = 3, limits = c(0, 5000)) +
+  #scale_y_continuous(limits = c(0.04, 0.06)) +
   labs(x = "Game Number", y = "Rating Deviation") +

### 2025-05-01 Paper_WriteUp/Reviewer_Response.qmd
moving all code for reviewer comments to new file

---
title: "Response to Reviewer Comments"
author: Adam Gee
date: today
format: 
  pdf
---

This document will create all numeric and graphical
summaries contained in the submitted response to reviewer comments.

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 7, dev = "png", fig.path = "output/",
                      fig.height = 5, dpi = 600)

library(tidyverse)
library(RcppRoll)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(loo)
library(scales)
library(ggridges)
library(here)
library(ggdist)
library(viridis)
library(PlayerRatings)
library(ggrepel)
library(ggpubr)
library(rstanarm)
library(ggh4x)
library(latex2exp)
library(furrr)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

## to load some plot defaults
source(here("utils", "helper.R"))
source(here("utils", "plot_templates.R"))

all_data_path <- rep(NA, 4)
all_data_path[1] <- here("box_data/lichess1700-1900/")
all_data_path[2] <- here("box_data/lichess2000-2200/")
all_data_path[3] <- here("box_data/lichess2300-2500/")
all_data_path[4] <- here("box_data/lichessGrandmasters/")
res_data_path <- rep(NA, 4)
res_data_path[1] <- here("results_2025/lichess1700-1900/")
res_data_path[2] <- here("results_2025/lichess2000-2200/")
res_data_path[3] <- here("results_2025/lichess2300-2500/")
res_data_path[4] <- here("results_2025/lichessGrandmasters/")

```

#plotting RDs and volatilities - Associate Editor, Comment 5
```{r}
#import data
data_path = paste0(all_data_path[4])

files <- list.files(data_path)

lichess_data <- files |> 
  map_dfr(~read_player(paste0(data_path, "/"), .x))
users <- readRDS(file = here("results_revision", "lichessGrandmasters", "users_bullet.RDS"))

small_data <- lichess_data |>
  mutate(Event = tolower(Event)) |>
  filter(Event == paste0("rated bullet game"),
         Variant == "Standard") |>
  filter(TimeControl %in% c("60+0", "180+0")) |> 
  distinct() #remove the duplicate rows if they exist

## otherwise not needed
users <- small_data |>
  group_by(Username) |>
  tally() |>
  filter(n >= 10) |>
  pull(Username)

tidy_games <- map_dfr(users, get_hist, small_data, prev_n = 1) |> 
  as_tibble()

init_data <- tidy_games |>
  mutate(WhiteElo = as.numeric(WhiteElo),
         BlackElo = as.numeric(BlackElo),
         focal_user = ifelse(focal_white == 1, White, Black),
         elo_diff = ifelse(focal_white == 1,
                           WhiteElo - BlackElo, BlackElo - WhiteElo),
         focal_id = match(focal_user, users),
         focal_rating = ifelse(focal_white == 1, WhiteElo, BlackElo),
         opp = ifelse(focal_white == 1, Black, White),
         opp_rating = ifelse(focal_white == 1, BlackElo, WhiteElo),
         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
  dplyr::select(focal_user, focal_id, focal_white, WhiteElo, BlackElo, White, Black,
                focal_win_prop, elo_diff, focal_result, opp, opp_rating,
                focal_rating, UTCDateTime) |>
  group_by(focal_id) |>
  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
         cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
         ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  
                           0, #if games played in different session, history is their mean win prob up to the current game
                           lag(focal_win_prop) - cum_win_prob)) |> #if game played in same session, rolling mean over the past n games, removing standardization 
  filter(focal_result != 0.5) %>%
  ungroup()

#get RDs
plan(multisession, workers = 12) #in parallel - dramatically speeds it up
start = Sys.time()
init_RDs = future_map_dfr(users, get_RDs, init_data) |>
  as_tibble()
end = Sys.time()
end - start #takes 1.3 min

#RDs
RD_plot = ggplot(data = init_RDs %>% group_by(focal_user) %>% mutate(index = row_number()), mapping = aes(x = index, y = RD)) +
  geom_line() +
  scale_x_continuous(n.breaks = 3, limits = c(0, 5000)) +
  scale_y_continuous(limits = c(50, 100)) +
  labs(x = "Game Number", y = "Rating Deviation") +
  facet_wrap(~focal_user)

#save
ggsave(plot = RD_plot,
       filename = here("Paper_WriteUp", "paper_figures", "response to comments", "RDs.png"), dpi = 1000,
       height = 6, width = 8)

#volatilities
vol_plot = ggplot(data = init_RDs %>% group_by(focal_user) %>% mutate(index = row_number()), mapping = aes(x = index, y = volatility)) +
  geom_line() +
  scale_x_continuous(n.breaks = 3, limits = c(0, 5000)) +
  #scale_y_continuous(limits = c(0.04, 0.06)) +
  labs(x = "Game Number", y = "Rating Deviation") +
  facet_wrap(~focal_user)

#save
ggsave(plot = vol_plot,
       filename = here("Paper_WriteUp", "paper_figures", "response to comments", "volatilities.png"), dpi = 1000,
       height = 6, width = 8)
```

#distribution of x_{ij} - Reviewer 2, Comment 6
```{r}
#import data
data_path = paste0(all_data_path[1], "/")

files <- list.files(data_path)
lichess_data <- files |> 
  map_dfr(~read_player(data_path, .x))

small_data <- lichess_data |>
  mutate(Event = tolower(Event)) |>
  filter(Event == paste0("rated ", time_control, " game"),
         Variant == "Standard") |>
  filter(TimeControl %in% c("60+0", "180+0")) |> 
  distinct() #remove the duplicate rows if they exist

## otherwise not needed
users <- small_data |>
  group_by(Username) |>
  tally() |>
  filter(n >= 10) |>
  pull(Username)

tidy_games <- map_dfr(users, get_hist, small_data, prev_n = n) |> 
  as_tibble()

init_data <- tidy_games |>
  mutate(WhiteElo = as.numeric(WhiteElo),
         BlackElo = as.numeric(BlackElo),
         focal_user = ifelse(focal_white == 1, White, Black),
         elo_diff = ifelse(focal_white == 1,
                           WhiteElo - BlackElo, BlackElo - WhiteElo),
         focal_id = match(focal_user, users),
         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
  dplyr::select(focal_user, focal_id, focal_white, 
                focal_win_prop, elo_diff, focal_result,
                UTCDateTime) |>
  group_by(focal_id) |>
  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
         cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
         ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  #current way we evaluate history
                           0, 
                           lag(focal_win_prop) - cum_win_prob),
         ave_prop2 = ifelse(time_diff > 300 | is.na(time_diff), #proposed way we evaluate history
                            0.5, 
                            lag(focal_win_prop))) |> 
  filter(focal_result != 0.5) %>%
  ungroup()

#just filter for one player for illustration
init_data_plot = init_data %>% filter(focal_id == 1)

#distribution of x_ij for current way we evaluate history
curr_dist = ggplot(data = init_data_plot, mapping = aes(x = ave_prop)) +
  geom_histogram() + 
  labs(title = "Normalization", x = TeX("$\\tilde{x}_{ij}^n$"), y = "Count") + 
  scale_x_continuous(breaks = c(-0.5, -0.25, 0, 0.25, 0.5), labels = c(-0.5, -0.25, 0, 0.25, 0.5)) + 
  theme_single_y()

#save
ggsave(plot = curr_dist,
       filename = here("Paper_WriteUp", "paper_figures", "response to comments", "curr_xij.png"), dpi = 1000,
       height = 4, width = 6)

#distribution of x_ij for proposed way to evaluate history
prop_dist = ggplot(data = init_data_plot, mapping = aes(x = ave_prop2)) +
  geom_histogram() + 
  labs(title = "No Normalization", x = TeX("$\\tilde{x}_{ij}^n$"), y = "Count") + 
  scale_x_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1), limits = c(-0.2, 1.2)) + 
  theme_single_y()

#save
ggsave(plot = prop_dist,
       filename = here("Paper_WriteUp", "paper_figures", "response to comments", "prop_xij.png"), dpi = 1000,
       height = 4, width = 6)
```

### 2025-05-01 Paper_WriteUp/Paper_Plots.qmd
moving all code for reviewer comments to new file

@@
   facetted_pos_scales(x = list( #for customizing scales on each facet
-    name == "mu_beta" ~ scale_x_continuous(breaks = c(0, 0.5), limits = c(0, 0.6)),
+    name == "mu_beta" ~ scale_x_continuous(breaks = breaks_pretty(n = 2)),
     name == "gamma1" ~ scale_x_continuous(breaks = breaks_pretty(n = 2)),
-    name == "gamma2" ~ scale_x_continuous(breaks = breaks_pretty(n = 2), limits = c(0.005, 0.0057))
+    name == "gamma2" ~ scale_x_continuous(breaks = breaks_pretty(n = 2))
   )) +

### 2025-05-01 Paper_WriteUp/Paper_Plots.qmd
moving code for reviewer comments to other file

@@
 ```
-
-#distribution of x_{ij} - for response to reviewer comment
-```{r}
-#import data
-data_path = paste0(all_data_path[1], "/")
-
-files <- list.files(data_path)
-lichess_data <- files |> 
-  map_dfr(~read_player(data_path, .x))
-
-small_data <- lichess_data |>
-  mutate(Event = tolower(Event)) |>
-  filter(Event == paste0("rated ", time_control, " game"),
-         Variant == "Standard") |>
-  filter(TimeControl %in% c("60+0", "180+0")) |> 
-  distinct() #remove the duplicate rows if they exist
-
-## otherwise not needed
-users <- small_data |>
-  group_by(Username) |>
-  tally() |>
-  filter(n >= 10) |>
-  pull(Username)
-
-tidy_games <- map_dfr(users, get_hist, small_data, prev_n = n) |> 
-  as_tibble()
-
-init_data <- tidy_games |>
-  mutate(WhiteElo = as.numeric(WhiteElo),
-         BlackElo = as.numeric(BlackElo),
-         focal_user = ifelse(focal_white == 1, White, Black),
-         elo_diff = ifelse(focal_white == 1,
-                           WhiteElo - BlackElo, BlackElo - WhiteElo),
-         focal_id = match(focal_user, users),
-         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
-  dplyr::select(focal_user, focal_id, focal_white, 
-                focal_win_prop, elo_diff, focal_result,
-                UTCDateTime) |>
-  group_by(focal_id) |>
-  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
-         cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
-         ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  #current way we evaluate history
-                           0, 
-                           lag(focal_win_prop) - cum_win_prob),
-         ave_prop2 = ifelse(time_diff > 300 | is.na(time_diff), #proposed way we evaluate history
-                            0.5, 
-                            lag(focal_win_prop))) |> 
-  filter(focal_result != 0.5) %>%
-  ungroup()
-
-#just filter for one player for illustration
-init_data_plot = init_data %>% filter(focal_id == 1)
-
-#distribution of x_ij for current way we evaluate history
-curr_dist = ggplot(data = init_data_plot, mapping = aes(x = ave_prop)) +
-  geom_histogram() + 
-  labs(title = "Normalization", x = TeX("$\\tilde{x}_{ij}^n$"), y = "Count") + 
-  scale_x_continuous(breaks = c(-0.5, -0.25, 0, 0.25, 0.5), labels = c(-0.5, -0.25, 0, 0.25, 0.5)) + 
-  theme_single_y()
-
-#save
-ggsave(plot = curr_dist,
-       filename = here("Paper_WriteUp", "paper_figures", "response to comments", "curr_xij.png"), dpi = 1000,
-       height = 4, width = 6)
-
-#distribution of x_ij for proposed way to evaluate history
-prop_dist = ggplot(data = init_data_plot, mapping = aes(x = ave_prop2)) +
-  geom_histogram() + 
-  labs(title = "No Normalization", x = TeX("$\\tilde{x}_{ij}^n$"), y = "Count") + 
-  scale_x_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1), limits = c(-0.2, 1.2)) + 
-  theme_single_y()
-
-#save
-ggsave(plot = prop_dist,
-       filename = here("Paper_WriteUp", "paper_figures", "response to comments", "prop_xij.png"), dpi = 1000,
-       height = 4, width = 6)
-```
-
-#plotting RDs and volatilities for reviewer response
-```{r}
-#import data
-data_path = paste0(all_data_path[4])
-
-files <- list.files(data_path)
-
-lichess_data <- files |> 
-  map_dfr(~read_player(paste0(data_path, "/"), .x))
-users <- readRDS(file = here("results_revision", "lichessGrandmasters", "users_bullet.RDS"))
-
-small_data <- lichess_data |>
-  mutate(Event = tolower(Event)) |>
-  filter(Event == paste0("rated bullet game"),
-         Variant == "Standard") |>
-  filter(TimeControl %in% c("60+0", "180+0")) |> 
-  distinct() #remove the duplicate rows if they exist
-
-## otherwise not needed
-users <- small_data |>
-  group_by(Username) |>
-  tally() |>
-  filter(n >= 10) |>
-  pull(Username)
-
-tidy_games <- map_dfr(users, get_hist, small_data, prev_n = 1) |> 
-  as_tibble()
-
-init_data <- tidy_games |>
-  mutate(WhiteElo = as.numeric(WhiteElo),
-         BlackElo = as.numeric(BlackElo),
-         focal_user = ifelse(focal_white == 1, White, Black),
-         elo_diff = ifelse(focal_white == 1,
-                           WhiteElo - BlackElo, BlackElo - WhiteElo),
-         focal_id = match(focal_user, users),
-         focal_rating = ifelse(focal_white == 1, WhiteElo, BlackElo),
-         opp = ifelse(focal_white == 1, Black, White),
-         opp_rating = ifelse(focal_white == 1, BlackElo, WhiteElo),
-         UTCDateTime = ymd_hms(paste0(UTCDate, "_", UTCTime))) |>
-  dplyr::select(focal_user, focal_id, focal_white, WhiteElo, BlackElo, White, Black,
-                focal_win_prop, elo_diff, focal_result, opp, opp_rating,
-                focal_rating, UTCDateTime) |>
-  group_by(focal_id) |>
-  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
-         cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
-         ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  
-                           0, #if games played in different session, history is their mean win prob up to the current game
-                           lag(focal_win_prop) - cum_win_prob)) |> #if game played in same session, rolling mean over the past n games, removing standardization 
-  filter(focal_result != 0.5) %>%
-  ungroup()
-
-#get RDs
-plan(multisession, workers = 12) #in parallel - dramatically speeds it up
-start = Sys.time()
-init_RDs = future_map_dfr(users, get_RDs, init_data) |>
-  as_tibble()
-end = Sys.time()
-end - start #takes 1.3 min
-
-#RDs
-RD_plot = ggplot(data = init_RDs %>% group_by(focal_user) %>% mutate(index = row_number()), mapping = aes(x = index, y = RD)) +
-  geom_line() +
-  scale_x_continuous(n.breaks = 3, limits = c(0, 5000)) +
-  scale_y_continuous(limits = c(50, 100)) +
-  labs(x = "Game Number", y = "Rating Deviation") +
-  facet_wrap(~focal_user)
-
-#save
-ggsave(plot = RD_plot,
-       filename = here("Paper_WriteUp", "paper_figures", "response to comments", "RDs.png"), dpi = 1000,
-       height = 6, width = 8)
-
-#volatilities
-vol_plot = ggplot(data = init_RDs %>% group_by(focal_user) %>% mutate(index = row_number()), mapping = aes(x = index, y = volatility)) +
-  geom_line() +
-  scale_x_continuous(n.breaks = 3, limits = c(0, 5000)) +
-  #scale_y_continuous(limits = c(0.04, 0.06)) +
-  labs(x = "Game Number", y = "Rating Deviation") +
-  facet_wrap(~focal_user)
-
-#save
-ggsave(plot = vol_plot,
-       filename = here("Paper_WriteUp", "paper_figures", "response to comments", "volatilities.png"), dpi = 1000,
-       height = 6, width = 8)
-```

### 2025-05-01 Paper_WriteUp/Paper_Plots.qmd
moving permutation section to appendix

@@
 # #save magnus
-ggsave(plot = ppd_rating_plot,
-       filename = here("Paper_WriteUp", "paper_figures", "fig7a.png"), dpi = 1000,
-       height = 4, width = 6)
+# ggsave(plot = ppd_rating_plot,
+#        filename = here("Paper_WriteUp", "paper_figures", "fig7a.png"), dpi = 1000,
+#        height = 4, width = 6)
 # #save nihal
@@
 #        height = 4, width = 6)
-```
-
-### Permuting for 1700-1900
-
-```{r perm for 17-19}
-#| eval: false
-perm_data_path <- here("results/lichess1700-1900_2025/perm")
-
-path_id <- 1
-data_path <- res_data_path[path_id]
-(perm_files <- list.files(perm_data_path))
-
-perm_17_19 <- perm_files[1:1000] |>
-  map_dfr(~ read_rds(here(perm_data_path, .x)))
-
-sim_pars <- perm_17_19 |>
-  select(variable, mean, sim_id)
-
-data_path <- here("results/lichess1700-1900_2025")
-
-fit_17_19_bullet_prev <- readRDS(file = 
-                                   paste0(data_path,
-                                          "/all_rated_bullet_model_prev_n1.RDS"))
-
-bullet_17_19 <- readRDS(file = paste0(
-  data_path,
-  "/users_bullet.RDS"
-))
-
-## then need to overlay the true values also
-true_values <- fit_17_19_bullet_prev$summary() |>
-  filter(variable %in% c("mu_beta", "gamma1", "gamma2")) |>
-  select(variable, mean)
-
-perm_17_19_global <- sim_pars |>
-  filter(variable %in% c(
-    "mu_beta", "gamma1", "gamma2")) |>
-  ggplot(aes(mean)) +
-  geom_histogram() +
-  facet_wrap(~variable, scales = "free") +
-  geom_vline(
-    data = true_values,
-    mapping = aes(xintercept = mean),
-    col = "red"
-  ) +
-  theme_3col() +
-  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
-  labs(x = element_blank(), y = element_blank())
-
-perm_17_19_global
-
-ggsave(
-  plot = perm_17_19_global,
-  filename = paste0(here(
-    "Paper_WriteUp", "paper_figures",
-    "perm_17_19_global.png"
-  )), dpi = 600,
-  height = 5, width = 7
-)
-```
-
-```{r figure 8 left - perm winner effects}
-
-player_pars <- sim_pars |> 
-  as_draws_df() |> 
-  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
-         id = stringr::str_extract(variable, pattern = "\\d+]"),
-         id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]")) |> 
-  drop_na() |> 
-  select(param, id, mean, sim_id) |> 
-  mutate(id = as.numeric(id)) |> 
-  left_join(player_id, by = "id")
-
-perm_players <- unique(player_pars$player_name)
-
-true_beta <- fit_17_19_bullet_prev$summary() |> 
-  filter(str_starts(variable, "beta\\[")) |> 
-  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
-         id = stringr::str_extract(variable, pattern = "\\d+]"),
-         id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]")) |> 
-  select(param, id, mean) |> 
-  mutate(id = as.numeric(id), param = as.numeric(param)) |> 
-  left_join(player_id, by = "id") |> 
-  filter(player_name %in% perm_players)
-
-levels <- true_beta |> 
-  filter(param == 2) |> 
-  arrange(mean) |> 
-  pull(player_name)
-
-true_vals <- true_beta |> 
-  filter(param == 2) |>  
-  mutate(y = factor(player_name, levels = levels)) |> 
-  arrange(y) |> 
-  mutate(y_fake = row_number() - 0.5,
-         yend = row_number() + 0.5)
-
-perm_17_19_winner <- player_pars |> 
-  filter(param == 2) |>
-  mutate(y = factor(player_name, levels = levels)) |> 
-  ggplot(aes(x = mean, y = y)) +
-  stat_histinterval(slab_fill = NA, .width = c(0.5, 0.95)) +
-  geom_segment(data = true_vals, 
-               aes(x = mean, y = y_fake, yend = yend), col = "red") +
-  labs(y = "", x = "Estimated W/L-Effect") +
-  theme_single_y()
-
-ggsave(plot = perm_17_19_winner, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "perm_17_19_winner.png")), dpi = 600,
-       height = 5, width = 7)
-
-```
-
-## GMs
-
-### Permuting for GM Data
-
-```{r figure 8 right}
-perm_data_path <- here("results/lichessGrandmasters_2025/perm/")
-
-## read all the files in this directory
-perm_files <- list.files(perm_data_path)
-
-perm_gm <- perm_files[1:1000] |> 
-  map_dfr(~read_rds(here(perm_data_path, .x)))
-
-sim_pars <- perm_gm |> 
-  select(variable, mean, sim_id)
-
-## then need to overlay the true values also
-true_values <- fit_gm_bullet_prev$summary() |> 
-  dplyr::filter(variable %in% c("mu_beta", "gamma1", "gamma2")) |> 
-  select(variable, mean)
-
-perm_gm_global <- sim_pars |> 
-  dplyr::filter(variable %in% c("mu_beta",  "gamma1", "gamma2")) |> 
-  ggplot(aes(mean)) +
-  geom_histogram() +
-  facet_wrap(~variable, scales = "free") +
-  geom_vline(data = true_values, mapping = aes(xintercept = mean),
-             col = "red") +
-  theme_3col() +
-  labs(x = element_blank(), y = element_blank())
-
-ggsave(plot = perm_gm_global, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "perm_gm_global.png")), dpi = 600,
-       height = 5, width = 7)
-
-## then repeat for the winner effects
-player_id <- as_tibble(bullet_gm) |> 
-  mutate(id = row_number()) |> 
-  rename(player_name = value)
-
-player_pars <- sim_pars |> 
-  as_draws_df() |> 
-  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
-         id = stringr::str_extract(variable, pattern = "\\d+]"),
-         id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]")) |> 
-  drop_na() |> 
-  select(param, id, mean, sim_id) |> 
-  mutate(id = as.numeric(id)) |> 
-  left_join(player_id, by = "id")
-
-perm_players <- unique(player_pars$player_name)
-
-true_beta <- fit_gm_bullet_prev$summary() |> 
-  filter(str_starts(variable, "beta\\[")) |> 
-  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
-         id = stringr::str_extract(variable, pattern = "\\d+]"),
-         id = stringr::str_replace(id, "\\]", ""),
-         player_id = paste0("beta[", id, "]")) |> 
-  select(param, id, mean) |> 
-  mutate(id = as.numeric(id), param = as.numeric(param)) |> 
-  left_join(player_id, by = "id") |> 
-  filter(player_name %in% perm_players)
-
-levels <- true_beta |> 
-  filter(param == 2) |> 
-  arrange(mean) |> 
-  pull(player_name)
-
-true_vals <- true_beta |> 
-  filter(param == 2) |>  
-  mutate(y = factor(player_name, levels = levels)) |> 
-  arrange(y) |> 
-  mutate(y_fake = row_number() - 0.25,
-         yend = row_number() + 0.25)
-
-perm_gm_winner <- player_pars |> 
-  filter(param == 2) |>
-  mutate(y = factor(player_name, levels = levels)) |> 
-  ggplot(aes(x = mean, y = y)) +
-  stat_histinterval(slab_fill = NA, .width = c(0.5, 0.95)) +
-  geom_segment(data = true_vals,
-               aes(x = mean, y = y_fake, yend = yend), col = "red") +
-  theme_single_y() +
-  labs(x = "Estimated W/L-Effect", y = "")
-
-perm_gm_winner
-
-ggsave(plot = perm_gm_winner, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "perm_gm_winner.png")), dpi = 600,
-       height = 5, width = 7)
-
-levels <- true_beta |> 
-  filter(param == 1) |> 
-  arrange(mean) |> 
-  pull(player_name)
-
-true_vals <- true_beta |> 
-  filter(param == 1) |>  
-  mutate(y = factor(player_name, levels = levels)) |> 
-  arrange(y) |> 
-  mutate(y_fake = row_number() - 0.5,
-         yend = row_number() + 0.5)
-
-perm_gm_player <- player_pars |> 
-  filter(param == 1) |>
-  mutate(y = factor(player_name, levels = levels)) |> 
-  ggplot(aes(x = mean, y = y)) +
-  stat_histinterval(slab_fill = NA, .width = c(0.5, 0.95)) +
-  geom_segment(data = true_vals, 
-               aes(x = mean, y = y_fake, yend = yend), col = "red") +
-  theme_single_y() +
-  labs(x = "Estimated Player Effect", y = "")
-
-perm_gm_player
-
-ggsave(plot = perm_gm_player, 
-       filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "perm_gm_player.png")), dpi = 600,
-       height = 5, width = 7)
-
 ```
@@
 
+### Permuting for 1700-1900
+
+```{r perm for 17-19}
+#| eval: false
+perm_data_path <- here("results/lichess1700-1900_2025/perm")
+
+path_id <- 1
+data_path <- res_data_path[path_id]
+(perm_files <- list.files(perm_data_path))
+
+perm_17_19 <- perm_files[1:1000] |>
+  map_dfr(~ read_rds(here(perm_data_path, .x)))
+
+sim_pars <- perm_17_19 |>
+  select(variable, mean, sim_id)
+
+data_path <- here("results/lichess1700-1900_2025")
+
+fit_17_19_bullet_prev <- readRDS(file = 
+                                   paste0(data_path,
+                                          "/all_rated_bullet_model_prev_n1.RDS"))
+
+bullet_17_19 <- readRDS(file = paste0(
+  data_path,
+  "/users_bullet.RDS"
+))
+
+## then need to overlay the true values also
+true_values <- fit_17_19_bullet_prev$summary() |>
+  filter(variable %in% c("mu_beta", "gamma1", "gamma2")) |>
+  select(variable, mean)
+
+perm_17_19_global <- sim_pars |>
+  filter(variable %in% c(
+    "mu_beta", "gamma1", "gamma2")) |>
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  facet_wrap(~variable, scales = "free") +
+  geom_vline(
+    data = true_values,
+    mapping = aes(xintercept = mean),
+    col = "red"
+  ) +
+  theme_3col() +
+  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
+  labs(x = element_blank(), y = element_blank())
+
+perm_17_19_global
+
+ggsave(
+  plot = perm_17_19_global,
+  filename = paste0(here(
+    "Paper_WriteUp", "paper_figures",
+    "perm_17_19_global.png"
+  )), dpi = 600,
+  height = 5, width = 7
+)
+```
+
+```{r figure 8 left - perm winner effects}
+
+player_pars <- sim_pars |> 
+  as_draws_df() |> 
+  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
+         id = stringr::str_extract(variable, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]")) |> 
+  drop_na() |> 
+  select(param, id, mean, sim_id) |> 
+  mutate(id = as.numeric(id)) |> 
+  left_join(player_id, by = "id")
+
+perm_players <- unique(player_pars$player_name)
+
+true_beta <- fit_17_19_bullet_prev$summary() |> 
+  filter(str_starts(variable, "beta\\[")) |> 
+  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
+         id = stringr::str_extract(variable, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]")) |> 
+  select(param, id, mean) |> 
+  mutate(id = as.numeric(id), param = as.numeric(param)) |> 
+  left_join(player_id, by = "id") |> 
+  filter(player_name %in% perm_players)
+
+levels <- true_beta |> 
+  filter(param == 2) |> 
+  arrange(mean) |> 
+  pull(player_name)
+
+true_vals <- true_beta |> 
+  filter(param == 2) |>  
+  mutate(y = factor(player_name, levels = levels)) |> 
+  arrange(y) |> 
+  mutate(y_fake = row_number() - 0.5,
+         yend = row_number() + 0.5)
+
+perm_17_19_winner <- player_pars |> 
+  filter(param == 2) |>
+  mutate(y = factor(player_name, levels = levels)) |> 
+  ggplot(aes(x = mean, y = y)) +
+  stat_histinterval(slab_fill = NA, .width = c(0.5, 0.95)) +
+  geom_segment(data = true_vals, 
+               aes(x = mean, y = y_fake, yend = yend), col = "red") +
+  labs(y = "", x = "Estimated W/L-Effect") +
+  theme_single_y()
+
+ggsave(plot = perm_17_19_winner, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "perm_17_19_winner.png")), dpi = 600,
+       height = 5, width = 7)
+
+```
+
+## GMs
+
+### Permuting for GM Data
+
+```{r figure 8 right}
+perm_data_path <- here("results/lichessGrandmasters_2025/perm/")
+
+## read all the files in this directory
+perm_files <- list.files(perm_data_path)
+
+perm_gm <- perm_files[1:1000] |> 
+  map_dfr(~read_rds(here(perm_data_path, .x)))
+
+sim_pars <- perm_gm |> 
+  select(variable, mean, sim_id)
+
+## then need to overlay the true values also
+true_values <- fit_gm_bullet_prev$summary() |> 
+  dplyr::filter(variable %in% c("mu_beta", "gamma1", "gamma2")) |> 
+  select(variable, mean)
+
+perm_gm_global <- sim_pars |> 
+  dplyr::filter(variable %in% c("mu_beta",  "gamma1", "gamma2")) |> 
+  ggplot(aes(mean)) +
+  geom_histogram() +
+  facet_wrap(~variable, scales = "free") +
+  geom_vline(data = true_values, mapping = aes(xintercept = mean),
+             col = "red") +
+  theme_3col() +
+  labs(x = element_blank(), y = element_blank())
+
+ggsave(plot = perm_gm_global, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "perm_gm_global.png")), dpi = 600,
+       height = 5, width = 7)
+
+## then repeat for the winner effects
+player_id <- as_tibble(bullet_gm) |> 
+  mutate(id = row_number()) |> 
+  rename(player_name = value)
+
+player_pars <- sim_pars |> 
+  as_draws_df() |> 
+  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
+         id = stringr::str_extract(variable, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]")) |> 
+  drop_na() |> 
+  select(param, id, mean, sim_id) |> 
+  mutate(id = as.numeric(id)) |> 
+  left_join(player_id, by = "id")
+
+perm_players <- unique(player_pars$player_name)
+
+true_beta <- fit_gm_bullet_prev$summary() |> 
+  filter(str_starts(variable, "beta\\[")) |> 
+  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
+         id = stringr::str_extract(variable, pattern = "\\d+]"),
+         id = stringr::str_replace(id, "\\]", ""),
+         player_id = paste0("beta[", id, "]")) |> 
+  select(param, id, mean) |> 
+  mutate(id = as.numeric(id), param = as.numeric(param)) |> 
+  left_join(player_id, by = "id") |> 
+  filter(player_name %in% perm_players)
+
+levels <- true_beta |> 
+  filter(param == 2) |> 
+  arrange(mean) |> 
+  pull(player_name)
+
+true_vals <- true_beta |> 
+  filter(param == 2) |>  
+  mutate(y = factor(player_name, levels = levels)) |> 
+  arrange(y) |> 
+  mutate(y_fake = row_number() - 0.25,
+         yend = row_number() + 0.25)
+
+perm_gm_winner <- player_pars |> 
+  filter(param == 2) |>
+  mutate(y = factor(player_name, levels = levels)) |> 
+  ggplot(aes(x = mean, y = y)) +
+  stat_histinterval(slab_fill = NA, .width = c(0.5, 0.95)) +
+  geom_segment(data = true_vals,
+               aes(x = mean, y = y_fake, yend = yend), col = "red") +
+  theme_single_y() +
+  labs(x = "Estimated W/L-Effect", y = "")
+
+perm_gm_winner
+
+ggsave(plot = perm_gm_winner, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "perm_gm_winner.png")), dpi = 600,
+       height = 5, width = 7)
+
+levels <- true_beta |> 
+  filter(param == 1) |> 
+  arrange(mean) |> 
+  pull(player_name)
+
+true_vals <- true_beta |> 
+  filter(param == 1) |>  
+  mutate(y = factor(player_name, levels = levels)) |> 
+  arrange(y) |> 
+  mutate(y_fake = row_number() - 0.5,
+         yend = row_number() + 0.5)
+
+perm_gm_player <- player_pars |> 
+  filter(param == 1) |>
+  mutate(y = factor(player_name, levels = levels)) |> 
+  ggplot(aes(x = mean, y = y)) +
+  stat_histinterval(slab_fill = NA, .width = c(0.5, 0.95)) +
+  geom_segment(data = true_vals, 
+               aes(x = mean, y = y_fake, yend = yend), col = "red") +
+  theme_single_y() +
+  labs(x = "Estimated Player Effect", y = "")
+
+perm_gm_player
+
+ggsave(plot = perm_gm_player, 
+       filename = paste0(here("Paper_WriteUp", "paper_figures",
+                              "perm_gm_player.png")), dpi = 600,
+       height = 5, width = 7)
+
+```
+
 ### OTHER

### 2025-05-01 analysis_scripts/09_Large_Fit_RDs.R
setting all opponent RDs to mean of focal players

@@ -98,10 +98,24 @@ init_data <- tidy_games |>
 #get RDs
 plan(multisession, workers = 8) #in parallel - dramatically speeds it up
 start = Sys.time()
-init_RDs = future_map_dfr(users, get_RDs, init_data) |>
-  as_tibble()
+init_RDs = future_map_dfr(users, get_RDs, init_data) |> #get RDs
+  as_tibble() 
 end = Sys.time()
 end - start 
+
+#add updated rating diff covariate
+
+g_RD = function(s) { #function for RD calculation
+  q_constant = log(10)/400 #constant for formula
+  1/sqrt((1+3*q_constant^2*s^2)/pi^2)
+}
+mean_RD = mean(init_RDs$RD) #the avg RD across all games for all focal players
+#use this to fill in opponents RD
+
+init_RDs = init_RDs |>
+  mutate(opp_RD = mean_RD,
+         new_rating_diff = elo_diff*g_RD(sqrt(RD^2 + opp_RD^2)))
+
 
 
 ### then fit the models
@@ -111,7 +125,7 @@ stan_data_ave <- list(N = nrow(init_RDs),
                       y = init_RDs$focal_result,
                       id = init_RDs$focal_id,
                       colour = init_RDs$focal_white,
-                      elo = init_RDs$elo_diff,
+                      elo = init_RDs$new_rating_diff,
                       win_prop = init_RDs$ave_prop)
 
 

### 2025-05-05 analysis_scripts/03_a_Fit_Perm_Session.R
add version of the perm code for dealing with sessions

@@ -1,0 +1,146 @@
+#### September 5th 2023
+#### Updated May 13th 2024
+
+## Considering a permutation test for evaluating the robustness of
+## the results to serial correlations in the data
+
+
+## We consider this first for the subset of data consisting of 18 focal players
+
+library(cmdstanr)
+
+options(mc.cores = parallel::detectCores())
+
+library(tidyverse)
+library(here)
+library(loo)
+library(ggplot2)
+library(posterior)
+library(RcppRoll)
+
+source(here("utils", "helper.R"))
+
+jobid <- Sys.getenv("SLURM_ARRAY_TASK_ID")
+jobid <- as.numeric(jobid)
+sim_id <- jobid
+
+
+all_data_path <- rep(NA, 4)
+all_save_path <- rep(NA, 4)
+all_data_path[1] <- here("box_data/lichess1700-1900/")
+all_data_path[2] <- here("box_data/lichess2000-2200/")
+all_data_path[3] <- here("box_data/lichess2300-2500/")
+all_data_path[4] <- here("box_data/lichessGrandmasters/")
+all_save_path[1] <- here("results_revision_may/lichess1700-1900/")
+
+# all_save_path[1] <- here("results/Full_Fits/lichess1700-1900/")
+# ## if need to run it locally
+
+all_save_path[2] <- here("results_revision_may/lichess2000-2200/")
+all_save_path[3] <- here("results_revision_may/lichess2300-2500/")
+all_save_path[4] <- here("results_revision_may/lichessGrandmasters/")
+
+path_id <- 4 ## setting it for now
+data_path <- all_data_path[path_id]
+save_path <- paste0(all_save_path[path_id], "perm_session/")
+
+dir.create(save_path, showWarnings = FALSE)
+
+files <- list.files(data_path)
+
+lichess_data <- files |> 
+  map_dfr(~read_player(data_path, .x))
+
+
+small_data <- lichess_data |>
+  mutate(Event = tolower(Event)) |> 
+  filter(TimeControl == "60+0") |>
+  filter(Variant == "Standard") |>
+  filter(grepl("rated bullet game", Event))
+
+rm(lichess_data)
+
+
+select_users <- small_data |> 
+  group_by(Username) |> 
+  tally() |> 
+  arrange(-n) |> 
+  ## will change this to n = 20 when run on cluster potentially
+  slice_max(order = n, n = 20) |> 
+  pull(Username)
+
+
+
+
+## just going to use the last games for this
[…]

### 2025-05-05 analysis_scripts/01_Large_Fit.R
better comment

@@ -95,8 +95,8 @@ init_data <- tidy_games |>
   mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
          cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
          ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  
-                           0, #if games played in different session, history is their mean win prob up to the current game
-                           lag(focal_win_prop) - cum_win_prob)) |> #if game played in same session, rolling mean over the past n games
+                           0, #if games played in different session, history is NULL (0)
+                           lag(focal_win_prop) - cum_win_prob)) |> #if game played in same session, rolling mean over the past min(n, num games in curr session) games
   filter(focal_result != 0.5) %>%
   ungroup()
 

### 2025-05-05 analysis_scripts/03_a_Fit_Perm_Session.R
tidy this file a little bit

@@ -104,11 +104,16 @@ init_data <- tidy_games |>
   select(focal_user, focal_id, focal_white, 
          focal_win_prop, elo_diff, focal_result, UTCDateTime) |> 
   group_by(focal_id) |>
-  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), #default is to ensure first game is always start of a new session
-         cum_win_prob = cummean(focal_result), #the mean win probability for the focal player up to the ith (current) game 
+  mutate(time_diff = UTCDateTime - lag(UTCDateTime, default = NA), 
+         #default is to ensure first game is always start of a new session
+         cum_win_prob = cummean(focal_result), 
+         #the mean win probability for the focal player up to the ith (current) game 
          ave_prop = ifelse(time_diff > 300 | is.na(time_diff),  
-                           0, #if games played in different session, history is their mean win prob up to the current game
-                           lag(focal_win_prop) - cum_win_prob)) |> #if game played in same session, rolling mean over the past n games
+                           0, 
+                           #if games played in different session, 
+                           # history is their mean win prob up to the current game
+                           lag(focal_result) - cum_win_prob)) |> 
+  #if game played in same session, rolling mean over the past n games
   filter(focal_result != 0.5) %>%
   ungroup()
 

### 2025-05-05 Paper_WriteUp/Paper_Plots.qmd
fixing breaks and limits for various plot axes

@@
   ggplot(aes(x = value, y = name, color = factor(model, levels = c("1700-1900", "2000-2200", "2300-2500", "GM")))) +
-  stat_pointinterval(position = position_dodge(width = 0.2)) +
+  stat_pointinterval(position = position_dodge(width = 0.3)) +
   facet_wrap(~name,
@@
   facetted_pos_scales(x = list( #for customizing scales on each facet
-    name == "mu_beta" ~ scale_x_continuous(breaks = c(-0.15, 0, 0.15), limits = c(-0.17, 0.17)),
+    name == "mu_beta" ~ scale_x_continuous(breaks = c(-0.05, 0, 0.05), limits = c(-0.07, 0.07)),
     name == "gamma1" ~ scale_x_continuous(breaks = breaks_pretty(n = 2)),
@@
   geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
-  xlim(c(-0.125, 0.125)) +
+  xlim(c(-0.075, 0.075)) +
   theme_single_y()
@@
 
-## count how many intervals contain 0
-## see how many contain 0
-player_pars |> 
-  group_by(player_name, param) |> 
-  filter(param == 2) |> 
-  summarise(lower = quantile(value, prob = 0.025),
-            upper = quantile(value, prob = 0.975)) |> 
-  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
-                         1, 0)) |> 
-  filter(contain_zero == 0)
-
 ## then the winner effects plot
@@
   facetted_pos_scales(x = list( #for customizing scales on each facet
-    name == "mu_beta" ~ scale_x_continuous(breaks = breaks_pretty(n = 2)),
     name == "gamma1" ~ scale_x_continuous(breaks = breaks_pretty(n = 2)),
-    name == "gamma2" ~ scale_x_continuous(breaks = breaks_pretty(n = 2))
+    name == "gamma2" ~ scale_x_continuous(breaks = c(0.0052, 0.0056), limits = c(0.0051, 0.0058)),
+    name == "mu_beta" ~ scale_x_continuous(breaks = c(-0.1, 0, 0.1))
   )) +

### 2025-05-06 Paper_WriteUp/Paper_Plots.qmd
fixing figS3 chunk

@@
 ```{r figure S3 left}
-## global effects first
-data_path <- here("results_2025/lichess1700-1900")
-
-fit_17_19_bullet_5 <- 
-  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n5.RDS"))
-
-fit_17_19_bullet_10 <- 
-  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n10.RDS"))
+model_path = here("results_revision", "lichess1700-1900")
+
+#GM n = 1 model
+fit_17_19_bullet_prev = readRDS(file = here(model_path, "all_rated_bullet_model_n1.RDS"))
+
+#GM n = 5 model
+fit_17_19_bullet_5 = readRDS(file = here(model_path, "all_rated_bullet_model_n5.RDS"))
+
+#GM n = 10 model
+fit_17_19_bullet_10 = readRDS(file = here(model_path, "all_rated_bullet_model_n10.RDS"))
 
@@
   theme(legend.position = "bottom") +
-  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
+  scale_x_continuous(n.breaks = 3) +
   scale_color_viridis_d(option = "H")
@@
 ```{r figure S3 right}
-## global effects first
-data_path <- here("results_2025/lichessGrandmasters")
-
-fit_gm_bullet_5 <- 
-  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n5.RDS"))
-
-fit_gm_bullet_10 <- 
-  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n10.RDS"))
+model_path = here("results_revision", "lichessGrandmasters")
+
+#GM n = 1 model
+fit_gm_bullet_prev = readRDS(file = here(model_path, "all_rated_bullet_model_n1.RDS"))
+
+#GM n = 5 model
+fit_gm_bullet_5 = readRDS(file = here(model_path, "all_rated_bullet_model_n5.RDS"))
+
+#GM n = 10 model
+fit_gm_bullet_10 = readRDS(file = here(model_path, "all_rated_bullet_model_n10.RDS"))
 

### 2025-05-07 Paper_WriteUp/Paper_Plots.qmd
remove duplicate games for GMs in fig2 generation

@@
   mutate(Event = tolower(Event)) |>
-  filter(TimeControl == "60+0", Variant == "Standard", Event == "rated bullet game")
+  filter(TimeControl == "60+0", Variant == "Standard", Event == "rated bullet game") %>%
+  distinct()
 
@@
     mutate(time_diff = UTCDateTime - lag(UTCDateTime)) # time difference between games
-
-  ## A check to remove duplicated games, if they occur
-  # cat("Dupes:", sum(duplicated(tidy_games_gms[1:9])), "\n", sep = " ")
-  # # some players have all their games double recorded for some reason
-  # tidy_games_gms <- tidy_games_gms[!duplicated(tidy_games_gms[1:9]), ] # remove duplicate rows
 

### 2025-05-07 Paper_WriteUp/Paper_Plots.qmd
updating filenames for fig2

@@
        filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "time_between_games_gm.png")), dpi = 600,
+                              "fig2a.png")), dpi = 600,
        height = 4, width = 6.25)
@@
        filename = paste0(here("Paper_WriteUp", "paper_figures",
-                              "session_length_gm.png")), dpi = 600,
+                              "fig2b.png")), dpi = 600,
        height = 4, width = 6.25)

### 2025-05-08 Paper_WriteUp/Paper_Plots.qmd
removing parts we dont use in fig4 chunk

@@
   left_join(player_id, by = "id")
-
-## see what the smallest and largest posterior mean is 
-## for the winner effects
-player_pars |> 
-  group_by(player_name, param) |> 
-  summarise(post_mean = mean(value)) |> 
-  filter(param == 2) |> 
-  ungroup() |> 
-  slice_max(order_by = post_mean, n = 5)
-
-player_pars |> 
-  group_by(player_name, param) |> 
-  summarise(post_mean = mean(value)) |> 
-  filter(param == 2) |> 
-  ungroup() |> 
-  slice_min(order_by = post_mean, n = 5)
 

### 2025-05-23 Paper_WriteUp/Paper_Plots.qmd
all experiential plots to have same x axis width

@@
   geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
-  xlim(c(-0.075, 0.075)) +
+  scale_x_continuous(breaks = c(-0.06, -0.03, 0, 0.03, 0.06), 
+                     labels = c(-0.06, -0.03, 0, 0.03, 0.06),
+                     limits = c(-0.062, 0.062)) +
   theme_single_y()
@@
   geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
+  scale_x_continuous(breaks = c(-0.06, -0.03, 0, 0.03, 0.06), 
+                     labels = c(-0.06, -0.03, 0, 0.03, 0.06),
+                     limits = c(-0.062, 0.062)) +
   theme_single_y()
@@
 #read in model
-fit_20_22_bullet_prev = readRDS(here("results_revision", "lichess2300-2500", "all_rated_bullet_model_n1.RDS"))
+fit_23_25_bullet_prev = readRDS(here("results_revision", "lichess2300-2500", "all_rated_bullet_model_n1.RDS"))
 
 #read in users
-bullet_20_22 = readRDS(here("results_revision", "lichess2300-2500", "users_bullet.RDS"))
+bullet_23_25 = readRDS(here("results_revision", "lichess2300-2500", "users_bullet.RDS"))
 

### 2025-05-29 Paper_WriteUp/Reviewer_Response.qmd

@@
 
-#plotting RDs and volatilities - Associate Editor, Comment 5
+#Associate Editor, Comment 5 - plotting RDs and volatilities
 ```{r}
@@
 
-#distribution of x_{ij} - Reviewer 2, Comment 6
+#Reviewer 2, Comment 6 - distribution of x_{ij}
 ```{r}
